### 2025-04-04 analysis/analysis.qmd
update

---
title: "BAIT"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
execute: 
  cache: false
---

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
library(ggdist)
library(brms)
```

## BAIT Questionnaire

### Data Preparation

```{r}
df1 <- read.csv("https://raw.githubusercontent.com/RealityBending/FakeFace/refs/heads/main/data/data.csv") |> 
  select(Participant, Age, Sex, Date, starts_with("AI_")) |> 
  rename(
    # I think current Artificial Intelligence algorithms can generate very realistic images
    BAIT_ImagesRealistic = AI_1_RealisticImages,
    # Organisations use Artificial Intelligence unethically
    BAIT_Unethical = AI_2_Unethical,
    # I think videos generated by Artificial Intelligence are impossible to distinguish from real videos
    BAIT_VideosReal = AI_3_VideosReal,
    # I am interested in using artificially intelligent systems in my daily life
    BAIT_DailyLife = AI_4_DailyLife,
    # I think Computer-Generated Imagery (CGI) software is capable of perfectly imitating reality
    BAIT_ImitatingReality = AI_5_ImitatingReality,
    # I think Artificial Intelligence is dangerous
    BAIT_Dangerous = AI_6_Dangerous,
    # I think current Artificial Intelligence algorithms can generate very realistic videos
    BAIT_VideosRealistic = AI_7_RealisticVideos,
    # Artificial Intelligence is exciting
    BAIT_Exciting = AI_8_Exciting,
    # There are many beneficial applications of Artificial Intelligence
    BAIT_Applications = AI_9_Applications,
    # Images of faces or people generated by Artificial Intelligence will contain errors
    BAIT_FaceErrors = AI_10_FaceErrors
    )
df1 <- df1[!duplicated(df1), ]
```

```{r}
df2 <- read.csv("https://raw.githubusercontent.com/RealityBending/FictionEro/refs/heads/main/study1/data/data_participants.csv") |> 
  select(Participant, Age, Sex, Date, starts_with("BAIT_"), starts_with("GAAIS_"), -ends_with("Duration"),
         -BAIT_Text, -BAIT_Visual, -BAIT_Videos) |> 
  rename(
    # Artificial Intelligence might take control of people
    BAIT_Control = GAAIS_Negative_9,
    # Artificial Intelligence is dangerous
    BAIT_Dangerous = GAAIS_Negative_10,
    # I am worried about future uses of Artificial Intelligence
    BAIT_Worry = GAAIS_Negative_15,
    # I am interested in using artificially intelligent systems in my daily life
    BAIT_DailyLife = GAAIS_Positive_7,
    # Artificial Intelligence is exciting
    BAIT_Exciting = GAAIS_Positive_12,
    # Much of society will benefit from a future full of Artificial Intelligence
    BAIT_Benefit = GAAIS_Positive_17,
    # Current Artificial Intelligence algorithms can generate very realistic images
    BAIT_ImagesRealistic = BAIT_1_ImagesRealistic,
    # Images of faces or people generated by Artificial Intelligence always contain errors and artifacts
    BAIT_ImagesIssues = BAIT_2_ImagesIssues,
    # Videos generated by Artificial Intelligence have obvious problems that make them easy to spot as fake
    BAIT_VideosIssues = BAIT_3_VideosRealistic, # NOTE: WRONG LABEL IN DATA
    # Current Artificial Intelligence algorithms can generate very realistic videos
    BAIT_VideosRealistic = BAIT_4_VideosIssues,
    # Computer-Generated Images (CGI) are capable of perfectly imitating reality
    BAIT_ImitatingReality = BAIT_5_ImitatingReality,
    # Technology allows the creation of environments that seem just as real as reality
    BAIT_EnvironmentReal = BAIT_6_EnvironmentReal, 
    # Artificial Intelligence assistants can write texts that are indistinguishable from those written by humans
    BAIT_TextRealistic = BAIT_7_TextRealistic,
    # Documents and paragraphs written by Artificial Intelligence usually read differently compared to Human productions
    BAIT_TextIssues = BAIT_8_TextIssues
    )
```

```{r}
df3 <- read.csv("https://raw.githubusercontent.com/RealityBending/FakeFace2/refs/heads/main/data/data_participants.csv") |> 
  select(Participant, Age, Sex=Gender, Date, starts_with("BAIT_"), -ends_with("Duration")) |> 
  rename(
    # Current AI algorithms can generate very realistic images
    BAIT_ImagesRealistic = BAIT_1_ImagesRealistic,
    # Images of faces or people generated by AI always contain errors and artifacts
    BAIT_ImagesIssues = BAIT_2_ImagesIssues,
    # Videos generated by AI have obvious problems that make them easy to spot as fake
    BAIT_VideosIssues = BAIT_3_VideosRealistic,  # NOTE: WRONG LABEL IN DATA
    # Current AI algorithms can generate very realistic videos
    BAIT_VideosRealistic = BAIT_4_VideosIssues,
    # Computer-Generated Images (CGI) are capable of perfectly imitating reality
    BAIT_ImitatingReality = BAIT_5_ImitatingReality,
    # Technology allows the creation of environments that seem just as real as reality
    BAIT_EnvironmentReal = BAIT_6_EnvironmentReal,
    # AI assistants can write texts that are indistinguishable from those written by humans
    BAIT_TextRealistic = BAIT_7_TextRealistic,
    # Documents and paragraphs written by AI usually read differently compared to Human productions
    BAIT_TextIssues = BAIT_8_TextIssues,
    # AI is dangerous
    BAIT_Dangerous = BAIT_9_NegativeAttitutes,
    # I am worried about future uses of AI
    BAIT_Worry = BAIT_10_NegativeAttitutes,
    # AI is exciting
    BAIT_Exciting = BAIT_11_PositiveAttitutes,
    # Much of society will benefit from a future full of AI
    BAIT_Benefit = BAIT_12_PositiveAttitutes,
    # How knowledgeable do you consider yourself about Artificial Intelligence (AI) technology?
    BAIT_Expertise = BAIT_AI_Knowledge
    )
```

### EGA

::: {.panel-tabset}

#### Sample 1

```{r}
uva <- EGAnet::UVA(data = select(df1, starts_with("BAIT")), cut.off = 0.4)
uva
uva$keep_remove
```

```{r}
ega <- EGAnet::bootEGA(
  data = select(df1, starts_with("BAIT")),
  seed=123,
  model="glasso",
  algorithm="leiden",
  EGA.type="riEGA",
  type="resampling",
  plot.typicalStructure=FALSE,
  verbose=FALSE)

plot(ega$EGA)

make_table <- function(ega) {
  EGAnet::net.loads(ega$EGA)$std |> 
  as.data.frame() |> 
  rownames_as_column("Item") |> 
  gt::gt() |> 
  gt::data_color(
    method = "numeric",
    palette = c("red", "white", "green"),
    domain = c(-1, 1)
  ) |> 
  gt::fmt_number(decimals = 2)
}
make_table(ega)
```

#### Sample 2

```{r}
uva <- EGAnet::UVA(data = select(df2, starts_with("BAIT")), cut.off = 0.4)
uva
uva$keep_remove
```

```{r}
ega <- EGAnet::bootEGA(
  data = select(df2, starts_with("BAIT")),
  seed=123,
  model="glasso",
  algorithm="leiden",
  EGA.type="riEGA",
  type="resampling",
  plot.typicalStructure=FALSE,
  verbose=FALSE)

plot(ega$EGA)

make_table(ega)
```

#### Sample 3

```{r}
uva <- EGAnet::UVA(data = select(df3, starts_with("BAIT")), cut.off = 0.4)
uva
uva$keep_remove
```

```{r}
ega <- EGAnet::bootEGA(
  data = select(df3, starts_with("BAIT")),
  seed=123,
  model="glasso",
  algorithm="leiden",
  EGA.type="riEGA",
  type="resampling",
  plot.typicalStructure=FALSE,
  verbose=FALSE)

plot(ega$EGA)

make_table(ega)
```

:::

## CHOCO Model

### Data Preparation

```{r}
#| message: false
#| warning: false

options(mc.cores = parallel::detectCores() - 2)

library(tidyverse)
library(easystats)
library(patchwork)
library(brms)

# source("https://raw.githubusercontent.com/RealityBending/BAIT/refs/heads/main/functions/choco.R")
source("../functions/choco.R")
```

### Analysis

::: {.panel-tabset}

#### Sample 1

```{r}
df <- read.csv("https://raw.githubusercontent.com/RealityBending/FakeFace/refs/heads/main/data/data.csv")
df$Real <- (df$Belief_Answer + 1) / 2  # Rescale
df$Item <- gsub(".jpg", "", df$Stimulus)
df <- df[c("Participant", "Item", "Real")]

# df <- df[df$Participant %in% unique(df$Participant)[1:20],]
```

```{r}
p1 <- df |> 
  ggplot(aes(x = Real)) +
  geom_histogram(fill = "blue", bins = 100) +
  scale_y_sqrt()

# p1 + geom_histogram(
#   data=data.frame(Real = rchoco(nrow(df), mu = plogis(0.4), muleft = plogis(0), phileft = 3, kleft = plogis(1.4), kd=0.3)),
#   alpha = 0.3, position = "identity", fill = "black", bins = 100)
```

##### Model

```{r}
#| eval: false

# Define formula and priors
f <- bf(
  Real ~ 0 + Intercept + (0 + Intercept | Participant),
  muleft ~ 0 + Intercept,
  phileft ~ 0 + Intercept,
  kleft ~ 0 + Intercept + (0 + Intercept | Participant),
  mud = 0,
  phid = 0,
  kd = 0,
  family = choco7d(link_mu = "logit", link_muleft = "logit", link_mud = "identity",
                   link_phileft = "softplus", link_phid = "identity",
                   link_kleft = "logit", link_kd = "identity")
)

# Define priors (brms::default_prior(f, data=df))
priors <- c(
  # prior("normal(0, 1)", class = "sd", lb = 0),
  prior("normal(0, 1)", class = "sd", coef = "", dpar = "", group = "Participant", lb = 0),
  prior("normal(0, 1)", class = "sd", coef = "", dpar = "kleft", group = "Participant", lb = 0),
  prior("normal(0, 0.5)", class = "b", coef = "Intercept"),
  prior("normal(0, 0.3)", class = "b", coef = "Intercept", dpar = "muleft"),
  prior("normal(3, 0.3)", class = "b", coef = "Intercept", dpar = "phileft"),
  prior("normal(1, 0.5)", class = "b", coef = "Intercept", dpar = "kleft")
) |> brms::validate_prior(formula = f, data = df)

# plot(seq(0, 10, length.out = 100), dgamma(seq(0, 10, length.out = 100), shape = 3, scale = 1))

# Fit the model
m1 <- brm(
  formula = f,
  data = df,
  prior = priors,
  family = choco7d(),
  stanvars = choco_stanvars("choco7d"),
  algorithm = "sampling",
  chains = 4,
  iter = 500,
  backend = "cmdstanr"
)

save(m1, file = "models/m1.RData")
```

```{r}
load("models/m1.RData")

# Summary and posterior predictive check
summary(m1)

insight::get_predicted(m1, predict = "prediction", iterations = 100) |>
  as.data.frame() |>
  cbind(df) |>
  bayestestR::reshape_iterations() |>
  ggplot(aes(x = iter_value)) +
  geom_histogram(data = df, aes(x = Real), fill = "blue", bins = 100) +
  geom_histogram(aes(group = iter_group), alpha = 0.02, fill = "black", bins = 100, position = "identity") +
  scale_y_sqrt()

g1 <- modelbased::estimate_grouplevel(m1)
plot(g1)
# performance::performance_dvour(g1)
```

##### Empirical

```{r}
features <- df |> 
  summarize(Empirical_pright = sum(Real > 0.5) / n(), 
            Empirical_conf = mean(abs(Real - 0.5) * 2),
            Empirical_pextreme = sum(Real %in% c(0, 1)) / n(),
            .by = "Participant") |> 
  full_join(reshape_grouplevel(g1, indices = "Median"), by = "Participant") |> 
  mutate(Sample = "Sample1")
```

##### Correlation

```{r}
correlation::correlation(features) |> 
  summary() |> 
  plot()
```

:::

### 2025-04-04 analysis/make_models.R

@@ -1,0 +1,82 @@
+# Fit model with 16 chains per task
+
+library(brms)
+library(cmdstanr)
+
+# Get array task ID from Slurm
+task_id <- as.numeric(commandArgs(trailingOnly = TRUE)[1])
+
+# Number of chains per task from Slurm (fixed at 16)
+chains_per_task <- as.numeric(Sys.getenv("SLURM_CPUS_PER_TASK"))
+
+# Calculate starting chain ID
+start_chain <- (task_id - 1) * chains_per_task + 1
+
+
+
+# Test --------------------------------------------------------------------
+
+# model <- brm(
+#   formula = mpg ~ wt,
+#   data = mtcars,
+#   family = gaussian(),
+#   chains = chains_per_task,    # 16 chains per task
+#   cores = chains_per_task,     # Use all 16 CPUs
+#   iter = 2000,
+#   warmup = 1000,
+#   seed = 1234 + start_chain,   # Unique seed per task
+#   backend = "cmdstanr",
+#   file = paste0("./modeltoy_task_", task_id, ".rds")
+# )
+
+
+
+# Sample 1 ----------------------------------------------------------------
+
+df <- read.csv("https://raw.githubusercontent.com/RealityBending/FakeFace/refs/heads/main/data/data.csv")
+df$Real <- (df$Belief_Answer + 1) / 2  # Rescale
+df$Item <- gsub(".jpg", "", df$Stimulus)
+df <- df[c("Participant", "Item", "Real")]
+
+
+# Define formula and priors
+f <- brms::bf(
+  Real ~ 0 + Intercept + (0 + Intercept | Participant),
+  muleft ~ 0 + Intercept,
+  phileft ~ 0 + Intercept,
+  kleft ~ 0 + Intercept + (0 + Intercept | Participant),
+  mud = 0,
+  phid = 0,
+  kd = 0,
+  family = choco7d(link_mu = "logit", link_muleft = "logit", link_mud = "identity",
+                   link_phileft = "softplus", link_phid = "identity",
+                   link_kleft = "logit", link_kd = "identity")
+)
+
+
+# Define priors (brms::default_prior(f, data=df))
+priors <- c(
+  # prior("normal(0, 1)", class = "sd", lb = 0),
+  prior("normal(0, 1)", class = "sd", coef = "", dpar = "", group = "Participant", lb = 0),
[…]

### 2025-04-10 analysis/1_bait.qmd
data

---
title: "Analysis 1 - BAIT Questionnaire"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
execute: 
  cache: false
---

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
library(ggdist)
```

## Data Preparation

```{r}
df1 <- read.csv("https://raw.githubusercontent.com/RealityBending/FakeFace/refs/heads/main/data/data.csv") |> 
  select(Participant, Age, Sex, Date, starts_with("AI_")) |> 
  rename(
    # I think current Artificial Intelligence algorithms can generate very realistic images
    BAIT_ImagesRealistic = AI_1_RealisticImages,
    # Organisations use Artificial Intelligence unethically
    BAIT_Unethical = AI_2_Unethical,
    # I think videos generated by Artificial Intelligence are impossible to distinguish from real videos
    BAIT_VideosReal = AI_3_VideosReal,
    # I am interested in using artificially intelligent systems in my daily life
    BAIT_DailyLife = AI_4_DailyLife,
    # I think Computer-Generated Imagery (CGI) software is capable of perfectly imitating reality
    BAIT_ImitatingReality = AI_5_ImitatingReality,
    # I think Artificial Intelligence is dangerous
    BAIT_Dangerous = AI_6_Dangerous,
    # I think current Artificial Intelligence algorithms can generate very realistic videos
    BAIT_VideosRealistic = AI_7_RealisticVideos,
    # Artificial Intelligence is exciting
    BAIT_Exciting = AI_8_Exciting,
    # There are many beneficial applications of Artificial Intelligence
    BAIT_Applications = AI_9_Applications,
    # Images of faces or people generated by Artificial Intelligence will contain errors
    BAIT_FaceErrors = AI_10_FaceErrors
    )
df1 <- df1[!duplicated(df1), ]
row.names(df1) <- NULL
```

```{r}
df2 <- read.csv("https://raw.githubusercontent.com/RealityBending/FakeFace2/refs/heads/main/data/data_participants.csv") |> 
  select(Participant, Age, Sex=Gender, Date, starts_with("BAIT_"), -ends_with("Duration")) |> 
  rename(
    # Current AI algorithms can generate very realistic images
    BAIT_ImagesRealistic = BAIT_1_ImagesRealistic,
    # Images of faces or people generated by AI always contain errors and artifacts
    BAIT_ImagesIssues = BAIT_2_ImagesIssues,
    # Videos generated by AI have obvious problems that make them easy to spot as fake
    BAIT_VideosIssues = BAIT_3_VideosRealistic,  # NOTE: WRONG LABEL IN DATA
    # Current AI algorithms can generate very realistic videos
    BAIT_VideosRealistic = BAIT_4_VideosIssues,
    # Computer-Generated Images (CGI) are capable of perfectly imitating reality
    BAIT_ImitatingReality = BAIT_5_ImitatingReality,
    # Technology allows the creation of environments that seem just as real as reality
    BAIT_EnvironmentReal = BAIT_6_EnvironmentReal,
    # AI assistants can write texts that are indistinguishable from those written by humans
    BAIT_TextRealistic = BAIT_7_TextRealistic,
    # Documents and paragraphs written by AI usually read differently compared to Human productions
    BAIT_TextIssues = BAIT_8_TextIssues,
    # AI is dangerous
    BAIT_Dangerous = BAIT_9_NegativeAttitutes,
    # I am worried about future uses of AI
    BAIT_Worry = BAIT_10_NegativeAttitutes,
    # AI is exciting
    BAIT_Exciting = BAIT_11_PositiveAttitutes,
    # Much of society will benefit from a future full of AI
    BAIT_Benefit = BAIT_12_PositiveAttitutes,
    # How knowledgeable do you consider yourself about Artificial Intelligence (AI) technology?
    BAIT_Expertise = BAIT_AI_Knowledge
    )
```

```{r}
df3 <- read.csv("https://raw.githubusercontent.com/RealityBending/FictionEro/refs/heads/main/study1/data/data_participants.csv") |> 
  select(Participant, Age, Sex, Date, starts_with("BAIT_"), starts_with("GAAIS_"), -ends_with("Duration"),
         -BAIT_Text, -BAIT_Visual, -BAIT_Videos) |> 
  rename(
    # Artificial Intelligence might take control of people
    BAIT_Control = GAAIS_Negative_9,
    # Artificial Intelligence is dangerous
    BAIT_Dangerous = GAAIS_Negative_10,
    # I am worried about future uses of Artificial Intelligence
    BAIT_Worry = GAAIS_Negative_15,
    # I am interested in using artificially intelligent systems in my daily life
    BAIT_DailyLife = GAAIS_Positive_7,
    # Artificial Intelligence is exciting
    BAIT_Exciting = GAAIS_Positive_12,
    # Much of society will benefit from a future full of Artificial Intelligence
    BAIT_Benefit = GAAIS_Positive_17,
    # Current Artificial Intelligence algorithms can generate very realistic images
    BAIT_ImagesRealistic = BAIT_1_ImagesRealistic,
    # Images of faces or people generated by Artificial Intelligence always contain errors and artifacts
    BAIT_ImagesIssues = BAIT_2_ImagesIssues,
    # Videos generated by Artificial Intelligence have obvious problems that make them easy to spot as fake
    BAIT_VideosIssues = BAIT_3_VideosRealistic, # NOTE: WRONG LABEL IN DATA
    # Current Artificial Intelligence algorithms can generate very realistic videos
    BAIT_VideosRealistic = BAIT_4_VideosIssues,
    # Computer-Generated Images (CGI) are capable of perfectly imitating reality
    BAIT_ImitatingReality = BAIT_5_ImitatingReality,
    # Technology allows the creation of environments that seem just as real as reality
    BAIT_EnvironmentReal = BAIT_6_EnvironmentReal, 
    # Artificial Intelligence assistants can write texts that are indistinguishable from those written by humans
    BAIT_TextRealistic = BAIT_7_TextRealistic,
    # Documents and paragraphs written by Artificial Intelligence usually read differently compared to Human productions
    BAIT_TextIssues = BAIT_8_TextIssues
    )
```

## EGA

::: {.panel-tabset}

### Sample 1

```{r}
# EGAnet::itemDiagnostics(select(df1, starts_with("BAIT")), seed = 1)
```

```{r}
uva <- EGAnet::UVA(data = select(df1, starts_with("BAIT")), cut.off = 0.4)
uva
uva$keep_remove
```

```{r}
ega1 <- EGAnet::bootEGA(
  data = select(df1, starts_with("BAIT")),
  seed=123,
  model="glasso",
  algorithm="leiden",
  EGA.type="riEGA",
  type="resampling",
  plot.typicalStructure=FALSE,
  verbose=FALSE)

plot(ega1$EGA)

make_table <- function(ega) {
  EGAnet::net.loads(ega$EGA)$std |> 
  as.data.frame() |> 
  rownames_as_column("Item") |> 
  gt::gt() |> 
  gt::data_color(
    method = "numeric",
    palette = c("red", "white", "green"),
    domain = c(-1, 1)
  ) |> 
  gt::fmt_number(decimals = 2)
}
make_table(ega1)
```

```{r}
features1 <- data.frame(
  Participant = df1$Participant,
  Sample = "Sample 1"
)
features1$BAIT_Realism <- rowMeans(df1[c("BAIT_VideosRealistic", "BAIT_ImagesRealistic", "BAIT_ImitatingReality")])
features1$BAIT_Positive <- df1 |> 
  datawizard::reverse_scale(select = c("BAIT_Dangerous")) |> 
  datawizard::data_select(select = c("BAIT_Exciting", "BAIT_DailyLife", "BAIT_Applications", "BAIT_Dangerous")) |> 
  rowMeans()
```

### Sample 3

```{r}
uva <- EGAnet::UVA(data = select(df3, starts_with("BAIT")), cut.off = 0.4)
uva
uva$keep_remove
```

```{r}
ega3 <- EGAnet::bootEGA(
  data = select(df3, starts_with("BAIT")),
  seed=123,
  model="glasso",
  algorithm="leiden",
  EGA.type="riEGA",
  type="resampling",
  plot.typicalStructure=FALSE,
  verbose=FALSE)

plot(ega3$EGA)

make_table(ega3)
```

```{r}
features3 <- data.frame(
  Participant = df3$Participant,
  Sample = "Sample 3"
)

features3$BAIT_Realism <- df3 |> 
  datawizard::reverse_scale(select = c("BAIT_ImagesIssues", "BAIT_VideosIssues")) |> 
  datawizard::data_select(select = c("BAIT_VideosRealistic", "BAIT_ImitatingReality", "BAIT_ImagesRealistic", "BAIT_EnvironmentReal", "BAIT_ImagesIssues", "BAIT_VideosIssues")) |> 
  rowMeans()

features3$BAIT_Positive <- df3 |> 
  datawizard::reverse_scale(select = c("BAIT_Worry", "BAIT_Dangerous", "BAIT_Control")) |> 
  datawizard::data_select(select = c("BAIT_DailyLife", "BAIT_Benefit", "BAIT_Exciting", "BAIT_Control", "BAIT_Worry", "BAIT_Dangerous")) |> 
  rowMeans()

features3$BAIT_TextRealism <- df3 |> 
  datawizard::reverse_scale(select = c("BAIT_TextIssues")) |> 
  datawizard::data_select(select = c("BAIT_TextRealistic", "BAIT_TextIssues")) |> 
  rowMeans()
```

### Sample 2

```{r}
uva <- EGAnet::UVA(data = select(df2, starts_with("BAIT")), cut.off = 0.4)
uva
uva$keep_remove
```

```{r}
ega2 <- EGAnet::bootEGA(
  data = select(df2, starts_with("BAIT"), -BAIT_Expertise),
  seed=123,
  model="glasso",
  algorithm="leiden",
  EGA.type="riEGA",
  type="resampling",
  plot.typicalStructure=FALSE,
  verbose=FALSE)

plot(ega2$EGA)

make_table(ega2)
```

```{r}
features2 <- data.frame(
  Participant = df2$Participant,
  Sample = "Sample 2"
)

features2$BAIT_Realism <- df2 |> 
  datawizard::reverse_scale(select = c("BAIT_ImagesIssues", "BAIT_VideosIssues")) |> 
  datawizard::data_select(select = c("BAIT_VideosRealistic", "BAIT_ImagesRealistic", "BAIT_ImitatingReality", "BAIT_VideosIssues")) |> 
  rowMeans() / 6

features2$BAIT_Positive <- df2 |> 
  datawizard::reverse_scale(select = c("BAIT_Worry", "BAIT_Dangerous")) |> 
  datawizard::data_select(select = c("BAIT_Benefit", "BAIT_Exciting", "BAIT_Worry", "BAIT_Dangerous")) |> 
  rowMeans() / 6

features2$BAIT_Simulate <- df2 |> 
  datawizard::reverse_scale(select = c("BAIT_TextIssues")) |> 
  datawizard::data_select(select = c("BAIT_ImitatingReality", "BAIT_TextRealistic", "BAIT_EnvironmentReal", "BAIT_TextIssues")) |> 
  rowMeans() / 6
```

:::

## Demographics

### Date

```{r}
dat <- rbind(
  df1 |> 
    select(Participant, Date, starts_with("BAIT")) |> 
    full_join(features1, by = "Participant") |>
    tidyr::pivot_longer(starts_with("BAIT")) |>
    mutate(Sample = "Sample 1"),
  df3 |> 
    select(Participant, Date, starts_with("BAIT")) |> 
    full_join(features3, by = "Participant") |>
    tidyr::pivot_longer(starts_with("BAIT")) |>
    mutate(Sample = "Sample 3", value = value),
  df2 |> 
    select(Participant, Date, starts_with("BAIT")) |> 
    full_join(features2, by = "Participant") |>
    tidyr::pivot_longer(starts_with("BAIT")) |>
    mutate(Sample = "Sample 2", value = value / 6)
) |> 
  # summarize(
  #   value = median(value, na.rm = TRUE),
  #   .by = c("Date", "Sample", "name")
  # ) |> 
  filter(name %in% c("BAIT_Realism", "BAIT_Positive", "BAIT_Simulate"))

dat |> 
  mutate(Date = lubridate::dmy(Date)) |> 
  ggplot(aes(x = Date, y = value, color = name, group = interaction(Sample, name))) +
  geom_point() +
  geom_smooth(method="lm", se = FALSE)
```

### Gender

```{r}
dat <- rbind(
  full_join(df1, features1, by = "Participant") |> 
  select(Sex, Age, Sample, BAIT_Realism, BAIT_Positive) |> 
    mutate(BAIT_Simulate = NA, BAIT_TextRealism = NA),
  full_join(df2, features2, by = "Participant") |> 
  select(Sex, Age, Sample, BAIT_Realism, BAIT_Positive, BAIT_Simulate) |> 
    mutate(BAIT_TextRealism = NA),
  full_join(df3, features3, by = "Participant") |> 
  select(Sex, Age, Sample, BAIT_Realism, BAIT_Positive, BAIT_TextRealism) |> 
    mutate(BAIT_Simulate = NA)) |> 
  mutate(Sample = as.factor(Sample))

m1 <- lm(BAIT_Realism ~ Sample / Sex, data = filter(dat, Sex %in% c("Male", "Female")))
m2 <- lm(BAIT_Positive ~ Sample / Sex, data = filter(dat, Sex %in% c("Male", "Female")))

rbind(
  mutate(estimate_relation(m1), Outcome = "Realism"),
  mutate(estimate_relation(m2), Outcome = "Positive")
) |> 
  ggplot(aes(x = Sample, y = Predicted, color = Sex, group = Sex)) +
  geom_pointrange(aes(ymin = CI_low, ymax = CI_high), position = position_dodge(width = 0.1))  +
  facet_wrap(~Outcome)

estimate_contrasts(m1, contrast = "Sex", by = "Sample")
estimate_contrasts(m2, contrast = "Sex", by = "Sample")
```

### Age

```{r}
library(mgcv)

m1 <- gam(BAIT_Realism ~ s(Age, by = Sample, k=5), data = filter(dat, Sex %in% c("Male", "Female")))
m2 <- gam(BAIT_Positive ~ s(Age, by = Sample, k=5), data = filter(dat, Sex %in% c("Male", "Female")))

rbind(
  mutate(estimate_relation(m1, length = 100), Outcome = "Realism"),
  mutate(estimate_relation(m2, length = 100), Outcome = "Positive")
) |> 
  ggplot(aes(x = Age, y = Predicted)) +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = Sample), alpha = 0.2) +
  geom_line(aes(color = Sample)) +
  facet_wrap(~Outcome)
```

## Save

```{r}
features <- rbind(features1)

write.csv(features, "../data/features_ega.csv", row.names = FALSE)
```

### 2025-04-10 analysis/2_choco.qmd
data

---
title: "Analysis 2 - Models"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
execute: 
  cache: false
---

```{r}
#| message: false
#| warning: false

options(mc.cores = parallel::detectCores() - 2)

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
library(ggdist)
library(brms)
```

## Sample 1

### Data Preprocessing

```{r}
df <- read.csv("https://raw.githubusercontent.com/RealityBending/FakeFace/refs/heads/main/data/data.csv") |> 
  mutate(Real = (Belief_Answer + 1) / 2,  # Rescale
         Orientation = ifelse(Sexual_Orientation == "Straight ", "Heterosexual", Sexual_Orientation),
         Item = gsub(".jpg", "", Stimulus)) |> 
  filter(Sex %in% c("Male", "Female"), Orientation %in% c("Heterosexual", "Bisexual", "Homosexual")) |> 
  mutate(Relevance = case_when(
    Sex == "Male" & Orientation == "Heterosexual" & grepl("NF", Stimulus) ~ "Relevant",
    Sex == "Female" & Orientation == "Heterosexual" & grepl("NM", Stimulus) ~ "Relevant",
    Sex == "Male" & Orientation == "Bisexual" & grepl("NF", Stimulus) ~ "Relevant",
    Sex == "Female" & Orientation == "Bisexual" & grepl("NM", Stimulus) ~ "Relevant",
    Sex == "Male" & Orientation == "Homosexual" & grepl("NM", Stimulus) ~ "Relevant",
    Sex == "Female" & Orientation == "Homosexual" & grepl("NF", Stimulus) ~ "Relevant",
    .default = "Irrelevant"
  )) |> 
  filter(Relevance == "Relevant") |> 
  select(Participant, Sex, Item, Real, Attractive, Beauty)

# Nudge mid-values to avoid model fitting issues
df$Real[df$Real == 0.5] <- sample(c(0.4999, 0.5001), sum(df$Real == 0.5), replace = TRUE)  # Avoid 0.5

write.csv(df, "../data/sample1.csv", row.names = FALSE)
```

### Models

::: {.panel-tabset}

```{r}
p1 <- df |> 
  ggplot(aes(x = Real)) +
  geom_histogram(fill = "blue", bins = 100) +
  scale_y_sqrt()

# p1 + geom_histogram(
#   data=data.frame(Real = rchoco(nrow(df), mu = plogis(0.4), muleft = plogis(0), phileft = 3, kleft = plogis(1.4), kd=0.3)),
#   alpha = 0.3, position = "identity", fill = "black", bins = 100)
```

#### Empirical

```{r}
features1 <- df |> 
  summarize(Sample = "Sample1", 
            Empirical_pright = sum(Real > 0.5) / n(), 
            Empirical_conf = mean(abs(Real - 0.5) * 2),
            Empirical_pex = sum(Real %in% c(0, 1)) / n(),
            Empirical_mean = mean(Real),
            .by = "Participant") 
```

#### ZOIB

```{r}
m1_zoib <- readRDS("models/sample1_zoib.rds")

# performance::r2(m1_zoib)

make_ppcheck  <- function(m, fill = "#FF9800") {
  pred <- m |> 
    insight::get_predicted(data = df[sample(nrow(df), 1000, replace = FALSE), ], 
                           predict="prediction", iterations = 100) |>
    bayestestR::reshape_iterations()

  pred |>
    ggplot(aes(x = iter_value)) +
    stat_bin(aes(group = iter_group, y = after_stat(count / max(count))), position = "identity",
             geom="bar", breaks = seq(-0.005, 1.005, by = 0.01), pad = TRUE,
             alpha = 0.01, fill = fill) + 
    stat_bin(data = df, aes(x = Real, y = after_stat(count / max(count))), 
             geom="step", breaks = seq(-0.01, 1.0, by = 0.01),  pad = TRUE, 
             linewidth = 1, color = "#9C27B0") + 
    scale_y_sqrt() +
    scale_x_continuous(expand = c(0, 0)) +
    theme_minimal() 
}

p1_zoib <- make_ppcheck(m1_zoib, fill = "#FF9800")
p1_zoib
```

```{r}
g1_zoib <- modelbased::estimate_grouplevel(m1_zoib)
plot(g1_zoib)

features1 <- reshape_grouplevel(g1_zoib, indices = "Median", group = "Participant") |> 
  datawizard::data_addprefix("ZOIB_", exclude = "Participant") |> 
  full_join(features1, by = "Participant")

# modelbased::estimate_grouplevel(insight::download_model("brms_chocomini_1"))
# modelbased::estimate_grouplevel(insight::download_model("brms_chocomini_1"), type = "total")
# performance::performance_dvour(g1)
```

#### BEXT

```{r}
m1_bext <- readRDS("models/sample1_bext.rds")

p1_bext <- make_ppcheck(m1_bext, fill = "#FF5722")
p1_bext
```

```{r}
g1_bext <- modelbased::estimate_grouplevel(m1_bext)
plot(g1_bext)

features1 <- reshape_grouplevel(g1_bext, indices = "Median", group = "Participant") |> 
  datawizard::data_addprefix("BEXT_", exclude = "Participant") |> 
  full_join(features1, by = "Participant") 
```

#### CHOCO

```{r}
m1_choco <- readRDS("models/sample1_choco.rds")

p1_choco <- make_ppcheck(m1_choco, fill = "#795548")
p1_choco
```

```{r}
g1_choco <- modelbased::estimate_grouplevel(m1_choco)
plot(g1_choco)

performance::performance_dvour(g1_choco) |> 
  arrange(desc(D_vour))

features1 <- reshape_grouplevel(g1_choco, indices = "Median", group = "Participant") |> 
  datawizard::data_addprefix("CHOCO_", exclude = "Participant") |> 
  full_join(features1, by = "Participant") 
```

:::

### Model Comparison

#### Correlation

```{r}
p1_zoib | p1_bext | p1_choco
```

```{r}
correlation::correlation(select(features1, -starts_with("Empirical_"), -Participant, -Sample), 
                         select(features1, starts_with("Empirical_")), 
                         p_adjust = "none") |> 
  summary() |> 
  plot()
```

#### Clusters

```{r}
# features1 |> 
#   ggplot(aes(x = muleft_Intercept, y = mud_Intercept)) +
#   geom_point(size = 2)

selected_features <- select(features1, contains("Intercept"))

parameters::n_clusters(selected_features)

rez <- parameters::cluster_analysis(selected_features, 
                             method = "hkmeans", 
                             n = 3) 
plot(rez)
features1$Cluster <- paste0("C", predict(rez))

summary(lm(Realism ~ Cluster, data = features1))
summary(lm(Exciting ~ Cluster, data = features1))
```

#### Moderators

```{r}
dfsub <- read.csv("../data/features_ega.csv") |> 
  full_join(features1, by = c("Participant", "Sample"))

dfsub1 <- select(filter(dfsub, Sample == "Sample1"), -Participant, -Sample)
correlation::correlation(select(dfsub1, -Realism, -Exciting), 
                         select(dfsub1, Realism, Exciting), 
                         p_adjust = "none") |> 
  summary() |> 
  plot()
```

```{r}
m1 <- lm(Realism ~ ., data = select(dfsub1, Realism, starts_with("Empirical_")))
m2 <- lm(Realism ~ ., data = select(dfsub1, Realism, starts_with("ZOIB_"))) 
m3 <- lm(Realism ~ ., data = select(dfsub1, Realism, starts_with("BEXT_")))
m4 <- lm(Realism ~ ., data = select(dfsub1, Realism, starts_with("CHOCO_")))

performance::test_performance(m1, m2, m3, m4)
performance::test_vuong(m1, m2, m3, m4)
r2(m1)
r2(m2)
r2(m3)
r2(m4)
summary(m4)
```

### Attractiveness

```{r}
m1_chocoat <- readRDS("models/sample1_chocoattractiveness.rds")

p1_chocoat <- make_ppcheck(m1_chocoat, fill = "#795548")
p1_chocoat
```

```{r}
performance::performance_dvour(m1_chocoat) |> 
  arrange(desc(D_vour))
```

```{r}
m1_chocoat <- readRDS(url("https://raw.github.com/DominiqueMakowski/cogmod/main/man/figures/m_zoib.rds"))
f <- bf(
  Real ~ 0 + Intercept + Attractive,
  phi ~ 0 + Intercept + Attractive,
  zoi ~ 0 + Intercept + Attractive,
  coi ~ 0 + Intercept + Attractive
)

m1_chocoat <- brm(f,
  data = df[1:1000,], family = zero_one_inflated_beta(), init = 0,
  chains = 4, iter = 500, backend = "cmdstanr"
)

estimate_relation(m1_chocoat, predict = "phi")
# modelbased::estimate_slopes(m1_chocoat, trend = c("Attractive"))

```

## Sample 2

### Models

```{r}
df <- read.csv("https://raw.githubusercontent.com/RealityBending/FakeFace2/refs/heads/main/data/data_task.csv")
df$Real <- (df$Realness + 1) / 2  
df <- df[c("Participant", "Item", "Real", "Attractive", "Beauty")]
```

```{r}
p1 <- df |> 
  ggplot(aes(x = Real)) +
  geom_histogram(fill = "blue", bins = 100) +
  scale_y_sqrt()

# p1 + geom_histogram(
#   data=data.frame(Real = rchoco(nrow(df), mu = plogis(0.4), muleft = plogis(0), phileft = 3, kleft = plogis(1.4), kd=0.3)),
#   alpha = 0.3, position = "identity", fill = "black", bins = 100)
```

### 2025-04-10 analysis/combine_models.R

@@ -1,0 +1,23 @@
+library(brms)
+library(cmdstanr)
+library(rstan)
+
+combine_and_save <- function(name = "sample1_zoib") {
+  files <- list.files('./models/', pattern = paste0(name, '_.*rds$'), full.names = TRUE)
+  m <- brms::combine_models(mlist = lapply(files, readRDS))
+
+  # m <- brms::add_criterion(m, criterion="waic", model_name = name)
+  # loo1 <- loo::loo_subsample(m, observations = 10, loo_approximation_draws = 100, cores = 4)
+  # fit_laplace <- rstan::optimizing(m, data = insight::get_data(m), draws = 1000, importance_resampling = TRUE)
+  # parameter_draws_laplace <- fit_laplace$theta_tilde # draws from approximate posterior
+  # log_p <- fit_laplace$log_p # log density of the posterior
+  # log_g <- fit_laplace$log_g # log density of the approximation
+
+
+  saveRDS(m, paste0("models/", name, ".rds"))
+}
+
+combine_and_save("sample1_zoib")
+combine_and_save("sample1_bext")
+combine_and_save("sample1_choco")
+combine_and_save("sample1_chocoattractiveness")

### 2025-04-10 analysis/make_models.R

@@ -2,6 +2,9 @@
 
 library(brms)
 library(cmdstanr)
+library(cogmod)
+
+options(brms.backend = "cmdstanr")
 
 # Get array task ID from Slurm
 task_id <- as.numeric(commandArgs(trailingOnly = TRUE)[1])
@@ -12,7 +15,7 @@ chains_per_task <- as.numeric(Sys.getenv("SLURM_CPUS_PER_TASK"))
 # Calculate starting chain ID
 start_chain <- (task_id - 1) * chains_per_task + 1
 
-
+iter <- 1000 # 50% of that is warmup
 
 # Test --------------------------------------------------------------------
 
@@ -22,10 +25,8 @@ start_chain <- (task_id - 1) * chains_per_task + 1
 #   family = gaussian(),
 #   chains = chains_per_task,    # 16 chains per task
 #   cores = chains_per_task,     # Use all 16 CPUs
-#   iter = 2000,
-#   warmup = 1000,
+#   iter = 500,
 #   seed = 1234 + start_chain,   # Unique seed per task
-#   backend = "cmdstanr",
 #   file = paste0("./modeltoy_task_", task_id, ".rds")
 # )
 
@@ -35,34 +36,190 @@ start_chain <- (task_id - 1) * chains_per_task + 1
 
 df <- read.csv("https://raw.githubusercontent.com/RealityBending/FakeFace/refs/heads/main/data/data.csv")
 df$Real <- (df$Belief_Answer + 1) / 2  # Rescale
+df$Real[df$Real == 0.5] <- sample(c(0.4999, 0.5001), sum(df$Real == 0.5), replace = TRUE)  # Avoid 0.5
 df$Item <- gsub(".jpg", "", df$Stimulus)
-df <- df[c("Participant", "Item", "Real")]
+df <- df[c("Participant", "Item", "Real", "Attractive", "Beauty")]
+
+# df <- df[df$Participant %in% unique(df$Participant)[1:10],]
+
+# # ZOIB
+# print("===== ZOIB =====")
+# f <- bf(
+#   Real ~ 0 + Intercept + (0 + Intercept | Participant) + (0 + Intercept | Item),
+#   phi ~ 0 + Intercept + (0 + Intercept | Participant) + (0 + Intercept | Item),
+#   zoi ~ 0 + Intercept + (0 + Intercept | Participant) + (0 + Intercept | Item),
[…]

### 2025-04-11 analysis/2_choco.qmd
sample2

@@
 execute: 
-  cache: false
+  cache: true
 ---
@@
 library(brms)
+library(cmdtsanr)
 ```
@@
 ```{r}
+#| eval: false
+
 df <- read.csv("https://raw.githubusercontent.com/RealityBending/FakeFace/refs/heads/main/data/data.csv") |> 
@@
 ```
+```{r}
+df <- read.csv("https://raw.githubusercontent.com/RealityBending/FictionChoco/refs/heads/main/data/sample1.csv")
+df$Sex <- factor(df$Sex, levels = c("Male", "Female"))
+```
 
@@
 ::: {.panel-tabset}
-
-```{r}
-p1 <- df |> 
-  ggplot(aes(x = Real)) +
-  geom_histogram(fill = "blue", bins = 100) +
-  scale_y_sqrt()
-
-# p1 + geom_histogram(
-#   data=data.frame(Real = rchoco(nrow(df), mu = plogis(0.4), muleft = plogis(0), phileft = 3, kleft = plogis(1.4), kd=0.3)),
-#   alpha = 0.3, position = "identity", fill = "black", bins = 100)
-```
 
@@
   pred <- m |> 
-    insight::get_predicted(data = df[sample(nrow(df), 1000, replace = FALSE), ], 
-                           predict="prediction", iterations = 100) |>
+    modelbased::estimate_prediction(df[sample(nrow(df), 4000, replace = FALSE), ], keep_iterations = 200) |>
+    select(Sex, starts_with("iter_")) |> 
+    # insight::get_predicted(data = df[sample(nrow(df), 1000, replace = FALSE), ],
+    #                        predict="prediction", iterations = 5) |>
+    # as.data.frame() |> 
     bayestestR::reshape_iterations()
@@
     scale_x_continuous(expand = c(0, 0)) +
-    theme_minimal() 
+    theme_minimal() +
+    facet_wrap(~Sex, scales = "free_y")
 }
@@
 plot(g1_zoib)
+performance::performance_dvour(g1_zoib)
 
@@
 # modelbased::estimate_grouplevel(insight::download_model("brms_chocomini_1"), type = "total")
-# performance::performance_dvour(g1)
+
 ```
@@
 
+```{r}
+#| eval: false
+
+df <- read.csv("https://raw.githubusercontent.com/RealityBending/FakeFace2/refs/heads/main/data/data_task.csv") |> 
+  filter(Relevance == "Relevant") |> 
+  mutate(Real = (Realness + 3) / 6,
+         Attractive = Attractiveness / 6,
+         Beauty = Beauty / 6) |> 
+  select(Participant, Sex=Gender, Condition, Item=Stimulus, Real, Attractive, Beauty)
+
+write.csv(df, "../data/sample2.csv", row.names = FALSE)
+```
+```{r}
+df <- read.csv("https://raw.githubusercontent.com/RealityBending/FictionChoco/refs/heads/main/data/sample2.csv")
+df$Sex <- factor(df$Sex, levels = c("Male", "Female"))
+```
+
 ### Models
-
-```{r}
-df <- read.csv("https://raw.githubusercontent.com/RealityBending/FakeFace2/refs/heads/main/data/data_task.csv")
-df$Real <- (df$Realness + 1) / 2  
-df <- df[c("Participant", "Item", "Real", "Attractive", "Beauty")]
-```
 

### 2025-04-11 analysis/make_models.R

@@ -34,94 +34,100 @@ iter <- 1000 # 50% of that is warmup
 
 # Sample 1 ----------------------------------------------------------------
 
-df <- read.csv("https://raw.githubusercontent.com/RealityBending/FakeFace/refs/heads/main/data/data.csv")
-df$Real <- (df$Belief_Answer + 1) / 2  # Rescale
-df$Real[df$Real == 0.5] <- sample(c(0.4999, 0.5001), sum(df$Real == 0.5), replace = TRUE)  # Avoid 0.5
-df$Item <- gsub(".jpg", "", df$Stimulus)
-df <- df[c("Participant", "Item", "Real", "Attractive", "Beauty")]
-
+df <- read.csv("https://raw.githubusercontent.com/RealityBending/FictionChoco/refs/heads/main/data/sample1.csv")
+df$Sex <- factor(df$Sex, levels = c("Male", "Female"))
 # df <- df[df$Participant %in% unique(df$Participant)[1:10],]
 
-# # ZOIB
-# print("===== ZOIB =====")
-# f <- bf(
-#   Real ~ 0 + Intercept + (0 + Intercept | Participant) + (0 + Intercept | Item),
-#   phi ~ 0 + Intercept + (0 + Intercept | Participant) + (0 + Intercept | Item),
-#   zoi ~ 0 + Intercept + (0 + Intercept | Participant) + (0 + Intercept | Item),
-#   coi ~ 0 + Intercept + (0 + Intercept | Participant) + (0 + Intercept | Item)
-# )
-#
-# model <- brm(
-#   formula = f,
-#   data = df,
-#   family = zero_one_inflated_beta(),
-#   init = 0,
-#   chains = chains_per_task,    # 16 chains per task
-#   cores = chains_per_task,     # Use all 16 CPUs
-#   iter = iter,  # Number of iterations and warmup must be equal (warmup is 50% iter)
-#   seed = 1234 + start_chain,   # Unique seed per task
-#   file = paste0("./sample1_zoib_task_", task_id, ".rds")
-# )
+# ZOIB
+print("===== ZOIB =====")
+print(Sys.time())
+
+f <- bf(
+  Real ~ 0 + Intercept + Sex + (0 + Intercept | Participant) + (0 + Intercept | Item),
+  phi ~ 0 + Intercept + Sex + (0 + Intercept | Participant) + (0 + Intercept | Item),
[…]

### 2025-04-27 analysis/1_study1.qmd
update

---
title: "Analysis - Study 1"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    toc: true
    code-fold: true
    self-contained: false
execute: 
  cache: true
---

```{r}
#| message: false
#| warning: false

options(mc.cores = parallel::detectCores() - 2)

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
library(ggdist)
library(brms)
library(cmdstanr)
library(rstan)

set.seed(123)
```

## Data Preprocessing

```{r}
#| eval: false

dfraw <- read.csv("https://raw.githubusercontent.com/RealityBending/FakeFace/refs/heads/main/data/data.csv") |> 
  mutate(Real = (Belief_Answer + 1) / 2,  # Rescale
         Orientation = ifelse(Sexual_Orientation == "Straight ", "Heterosexual", Sexual_Orientation),
         Item = gsub(".jpg", "", Stimulus)) |> 
  filter(Sex %in% c("Male", "Female"), Orientation %in% c("Heterosexual", "Bisexual", "Homosexual")) |> 
  mutate(Relevance = case_when(
    Sex == "Male" & Orientation == "Heterosexual" & grepl("NF", Stimulus) ~ "Relevant",
    Sex == "Female" & Orientation == "Heterosexual" & grepl("NM", Stimulus) ~ "Relevant",
    Sex == "Male" & Orientation == "Bisexual" & grepl("NF", Stimulus) ~ "Relevant",
    Sex == "Female" & Orientation == "Bisexual" & grepl("NM", Stimulus) ~ "Relevant",
    Sex == "Male" & Orientation == "Homosexual" & grepl("NM", Stimulus) ~ "Relevant",
    Sex == "Female" & Orientation == "Homosexual" & grepl("NF", Stimulus) ~ "Relevant",
    .default = "Irrelevant"
  )) |> 
  mutate(SexOrientation = paste(Sex, "-", Orientation)) |>
  filter(Relevance == "Relevant", Orientation == "Heterosexual")

# Nudge mid-values to avoid model fitting issues
dfraw$Real[dfraw$Real == 0.5] <- sample(c(0.49, 0.51), sum(dfraw$Real == 0.5), replace = TRUE)  # Avoid 0.5

dfraw |> 
  select(Participant, Sex, Item, Real, Attractive, Beauty) |> 
  write.csv("../data/sample1.csv", row.names = FALSE)
```

```{r}
#| eval: false

# For demographic section
dfraw |> 
  mutate(n_stims = n(), .by = "Participant") |> 
  select(Participant, Age, Sex, n_stims) |> 
  slice(1, .by = "Participant") |> 
  report::report_participants()

report::report(dfraw$Sex)
```

```{r}
#| code-fold: false

df <- read.csv("https://raw.githubusercontent.com/RealityBending/FictionChoco/refs/heads/main/data/sample1.csv")
df$Sex <- factor(df$Sex, levels = c("Male", "Female"))
```

## Models

::: {.panel-tabset}

### ZOIB

```{r}
m1_zoib <- readRDS("models/sample1_zoib.rds")
params_table <- function(m) {
  d <- as.data.frame(m, variable = insight::find_parameters(m, effects = "fixed", flatten  = TRUE))
  t <- bayestestR::describe_posterior(d, effects = "fixed", component = "all", ci_method = "hdi", test = "p_direction")
  out <- insight::format_table(t)
  out |> 
    mutate(col = t$pd) |> 
    gt::gt() |> 
    gt::data_color(columns = "col", method = "numeric", target_columns = "pd", 
                   row = col > 0.95, palette = "yellow") |> 
    gt::data_color(columns = "col", method = "numeric", target_columns = "pd", 
                   row = col > 0.97, palette = "green") |> 
    gt::cols_hide("col")
 }
params_table(m1_zoib)
summary(m1_zoib, robust = TRUE)

make_ppcheck  <- function(m, fill = "#FF9800") {
  pred <- m |> 
    modelbased::estimate_prediction(df, keep_iterations = 100, iterations = 100) |>
    select(Sex, starts_with("iter_")) |> 
    bayestestR::reshape_iterations()

  pred |>
    ggplot(aes(x = iter_value)) +
    stat_bin(aes(group = iter_group, y = after_stat(count / max(count))), position = "identity",
             geom="bar", breaks = seq(0, 1.0, by = 0.02), pad = TRUE,
             alpha = 0.01, fill = fill) +
    stat_bin(data = df, aes(x = Real, y = after_stat(count / max(count))), 
             geom="step", breaks = seq(0, 1.0, by = 0.02), 
             linewidth = 0.5, color = "#9C27B0", position = position_nudge(x=-0.01), pad = TRUE) + 
    scale_y_sqrt(expand = c(0, 0.01)) +
    scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c("      0%", "50%", "100%      ")) +
    theme_minimal() +
    theme(axis.title.x = element_text(face = "bold"), 
          axis.title.y = element_blank(), 
          axis.text.y = element_blank(), 
          strip.background = element_rect(fill = "lightgrey", color = NA)) +
    facet_wrap(~Sex, scales = "free_y") +
    coord_cartesian(xlim = c(-0.003, 1.005), ylim=c(0, 1), expand = TRUE)
}

p1_zoib <- make_ppcheck(m1_zoib, fill = "#FF9800") +
  labs(x = "Zero-Inflated Beta (ZOIB)")
p1_zoib
```

- Women: higher real mean
- Women: no difference in precision
- Women: less extreme values BUT more ones

### Beta-Gate

```{r}
m1_betagate <- readRDS("models/sample1_betagate.rds")
params_table(m1_betagate)
summary(m1_betagate, robust = TRUE)

p1_betagate <- make_ppcheck(m1_betagate, fill = "#FF5722") +
  labs(x = "Beta-Gate")
p1_betagate

# d <- brms::as_draws(m1_betagate, draw=1:10, variable = "^b_", regex = TRUE) |> 
#   bayestestR::describe_posterior(centrality = "median", test = "p_direction")
```

- Women: higher real mean
- Women: no difference in precision
- Women: less extreme values BUT more ones

### CHOCO

```{r}
m1_choco <- readRDS("models/sample1_choco.rds")
params_table(m1_choco)
summary(m1_choco, robust = TRUE)

p1_choco <- make_ppcheck(m1_choco, fill = "#7B3F00") +
  labs(x = "CHOCO")
p1_choco
```

- Women: higher probability of "Real"
- Women: Less confidence when "Fake"
- Women: no difference in precision
- Women: less extreme values (BUT possibly more ones)

```{r}
visualize_dchoco <- function(m, newdata) {

  p <- modelbased::estimate_prediction(m, data = newdata, predict = "mu", iterations = 200)
  confright <- modelbased::estimate_prediction(m, data = newdata, predict = "confright", iterations = 200)
  confleft <- modelbased::estimate_prediction(m, data = newdata, predict = "confleft", iterations = 200)
  precright <- modelbased::estimate_prediction(m, data = newdata, predict = "precright", iterations = 200)
  precleft <- modelbased::estimate_prediction(m, data = newdata, predict = "precleft", iterations = 200)
  pex <- modelbased::estimate_prediction(m, data = newdata, predict = "pex", iterations = 200)
  bex <- modelbased::estimate_prediction(m, data = newdata, predict = "bex", iterations = 200)

  data <- data.frame()
  x <- seq(0, 1, length.out = 1001)
  for(i in 1:nrow(newdata)) {
     data <- data.frame(
       x = c(-0.005, x, 1.005),
       y = c(0, cogmod::dchoco(x, p=p$Predicted[i], confright=confright$Predicted[i], confleft = confleft$Predicted[i],
                          precright = precright$Predicted[i], precleft = precleft$Predicted[i],
                          pex = pex$Predicted[i], bex = bex$Predicted[i], pmid = 0), 0)
     ) |> 
       merge(newdata[i, ]) |> 
       rbind(data)
  }
  data
}

pred <- visualize_dchoco(m1_choco, newdata=insight::get_datagrid(m1_choco))
p1_density <- pred |> 
  mutate(y = ifelse(x %in% c(0, 1), y * 3, y)) |>  # Inflate extremes for visibility
  ggplot(aes(x=x, y=y, color=Sex)) +
  geom_line(linewidth = 1.5) +
  scale_color_manual(values = c("Male"="#2196F3", "Female"="#E91E63")) +
  scale_x_continuous(breaks = c(0, 0.25, 0.5, 0.75, 1), 
                     labels = c("0%\nAI-Generated", "25%", "50%", "75%", "100%\nPhotograph")) +
  theme_minimal() +
  labs(y = "Distribution", x = "Do you think this image is real?") +
  theme(axis.text.y = element_blank(),
        axis.title.x = element_text(face = "bold"))
p1_density
```

:::

## Model Comparison

### Fitting Duration

```{r}
rbind(
  mutate(attributes(m1_zoib$fit)$metadata$time$chain, Model="ZOIB"),
  mutate(attributes(m1_betagate$fit)$metadata$time$chain, Model="BEXT"),
  mutate(attributes(m1_choco$fit)$metadata$time$chain, Model="CHOCO")
) |> 
  mutate(Model = fct_relevel(Model, "ZOIB", "BEXT", "CHOCO"),
         total = total / 60) |>
  ggplot(aes(x = Model, y = total, fill = Model)) +
  geom_boxplot() +
  labs(y = "Sampling Duration (min)") +
  scale_y_log10() +
  scale_fill_manual(values = c("#FF9800", "#FF5722", "#795548")) +
  theme_minimal() 
```

### Fit Quality

```{r}
loocomp <- loo::loo_compare(m1_zoib, m1_betagate, m1_choco, criterion = "waic")

parameters(loocomp, include_ENP = TRUE)

report::report(loocomp)

p1_zoib | p1_betagate | p1_choco
```

## Individual-Level Parameters

### Reliability

::: {.panel-tabset}

#### ZOIB

```{r}
table_dvour <- function(g) {
  performance::performance_dvour(g) |> 
    arrange(desc(D_vour)) |> 
    gt::gt() |> 
    gt::data_color("D_vour", palette = c("red", "red", "orange", "yellow", "green"), domain = c(0, 1)) |> 
    gt::fmt_auto()
}

g1_zoib <- modelbased::estimate_grouplevel(m1_zoib)
table_dvour(g1_zoib)
```

#### Beta-Gate

```{r}
g1_betagate <- modelbased::estimate_grouplevel(m1_betagate)
table_dvour(g1_betagate)
```

#### CHOCO

```{r}
g1_choco <- modelbased::estimate_grouplevel(m1_choco)
table_dvour(g1_choco)
```

```{r}
colors <- c("p" = "#9C27B0", "confright" = "#795548", "confleft" = "#795548",
            "precright" = "#607D8B","precleft" = "#607D8B", "pex" = "#FF9800", "bex" = "#F44336")

p2_dvour <- g1_choco |> 
  mutate(Component = ifelse(Component == "conditional", "p",  Component),
         Component = fct_relevel(Component, "p", "confright", "confleft", "precright", "precleft", "pex", "bex"),
         Group = fct_relevel(Group, "Participant", "Item"),
         .group = paste0(Group, " (", Component, ")")) |>
  ggplot(aes(x = Median, y = Level)) +
  geom_pointrange(aes(xmin = CI_low, xmax = CI_high, color= Component), linewidth = 0.3, size = 0.05) +
  facet_grid(Group~Component, scales = "free", switch="both") +
  ggside::geom_xsidedensity(aes(group = .group, fill = Component), color = NA) +
  ggside::ggside(scales  = "free") +
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.placement.x = "outside",
        strip.background = element_rect(fill = NA, color = NA),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position = "none")
p2_dvour
```

:::

### Correlation

```{r}
features1 <- df |> 
  summarize(Sample = "Sample1", 
            Empirical_p = sum(Real > 0.5) / n(), 
            Empirical_conf = mean(abs(Real - 0.5) * 2),
            Empirical_pex = sum(Real %in% c(0, 1)) / n(),
            Empirical_bex = sum(Real == 1) / sum(Real %in% c(0, 1)),
            Empirical_mu = mean(Real),
            .by = "Participant")

features1 <- reshape_grouplevel(g1_zoib, indices = "Median", group = "Participant") |> 
  datawizard::data_rename("Intercept", "mu") |> 
  select(-coi_Intercept) |> 
  datawizard::data_addprefix("ZOIB_", exclude = "Participant") |> 
  full_join(features1, by = "Participant")

features1 <- reshape_grouplevel(g1_betagate, indices = "Median", group = "Participant") |> 
  datawizard::data_rename("Intercept", "mu") |> 
  select(-bex_Intercept) |> 
  datawizard::data_addprefix("BGATE_", exclude = "Participant") |> 
  full_join(features1, by = "Participant")

features1 <- reshape_grouplevel(g1_choco, indices = "Median", group = "Participant") |> 
  datawizard::data_rename("Intercept", "p") |> 
  select(-bex_Intercept) |> 
  datawizard::data_addprefix("CHOCO_", exclude = "Participant") |> 
  full_join(features1, by = "Participant")
```

```{r}
cordata <- correlation::correlation(select(features1, -Participant, -Sample), p_adjust = "none", redundant = TRUE) |> 
  correlation::cor_sort() |> 
  correlation::cor_lower() |>
  as.data.frame() |> 
  mutate(Parameter2 = fct_rev(Parameter2),
         Parameter1 = fct_relabel(Parameter1, \(x) gsub("_Intercept", "", x)),
         Parameter2 = fct_relabel(Parameter2, \(x) gsub("_Intercept", "", x)),
         Parameter1 = fct_relabel(Parameter1, \(x) gsub("_", " - ", x)),
         Parameter2 = fct_relabel(Parameter2, \(x) gsub("_", " - ", x)))

p2_cor <- cordata |> 
  ggplot(aes(x=Parameter2, y = Parameter1)) +
  geom_tile(aes(fill = r)) +
  geom_text(aes(label = insight::format_value(r, 2, zap_small = TRUE, lead_zero = FALSE)), color = ifelse(cordata$p < 0.01, "black", "grey")) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "none"
  )
p2_cor
```

Choco parameters are the most interpretable (relate to empirical indices).

## Figures

```{r}
#| eval: true

p1_a <- wrap_elements((p1_zoib | p1_betagate | p1_choco) + 
         patchwork::plot_annotation(title = "Posterior Predictive Checks", theme = theme(plot.title = element_text(face = "bold"))))

p1_b <- wrap_elements(p1_density + 
         patchwork::plot_annotation(title = "Effect of Sex", theme = theme(plot.title = element_text(face = "bold"))))

p1 <- p1_a / p1_b

p2_a <- wrap_elements(p2_dvour + 
         patchwork::plot_annotation(title = "Reliability of CHOCO Participant-Level Estimates", theme = theme(plot.title = element_text(face = "bold"))))

p2_b <- wrap_elements(p2_cor + 
         patchwork::plot_annotation(title = "Correlation of Participant-Level Estimates", theme = theme(plot.title = element_text(face = "bold"))))
p2 <- p2_a / p2_b + plot_layout(heights = c(1.5, 1))

ggsave("../paper/figures/fig3.png", p1, width = 8, height = 8, dpi = 300)
ggsave("../paper/figures/fig4.png", p2, width = 8, height = 12, dpi = 300)
```

### 2025-04-27 analysis/2_study2.qmd
update

---
title: "Analysis - Study 2"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    toc: true
    code-fold: true
    self-contained: false
execute: 
  cache: true
  cache-lazy: false
---

```{r}
#| message: false
#| warning: false

options(mc.cores = parallel::detectCores() - 2)

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
library(ggdist)
library(brms)
library(cmdstanr)
library(rstan)
```

## Data Preprocessing

```{r}
df <- read.csv("https://raw.githubusercontent.com/RealityBending/FictionChoco/refs/heads/main/data/sample1.csv")
df$Sex <- factor(df$Sex, levels = c("Male", "Female"))
```

## Sample 1

### Attractiveness

```{r}
m <- lme4::lmer(Real ~ Sex / poly(Attractive, 2) + (1|Participant) + (1|Item), data = df)
display(parameters::parameters(m, effects = "fixed"))

estimate_relation(m, length = 30) |> 
  ggplot(aes(x=Attractive, y = Predicted)) + 
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high,, fill = Sex), alpha = 0.2) +
  geom_line(aes(color = Sex), linewidth = 1) +
  geom_smooth(data=df, aes(color = Sex, y = Real), method = "loess", se = FALSE, linetype = "dotted", formula = 'y ~ x') +
  scale_color_manual(values = c("Male"="#2196F3", "Female"="#E91E63")) +
  scale_fill_manual(values = c("Male"="#2196F3", "Female"="#E91E63")) +
  theme_minimal()
```

#### Model Description

::: {.panel-tabset}

##### Effect of Attractiveness

```{r}
m1_chocoat <- readRDS("models/sample1_chocoattractiveness.rds")

params_table <- function(m, keep = "polyAttractive") {
  params <- insight::find_parameters(m, effects = "fixed", flatten  = TRUE)
  d <- as.data.frame(m, variable = params[str_detect(params, keep)])
  tab <- bayestestR::describe_posterior(d, effects = "fixed", component = "all", 
                                        ci_method = "hdi", test = "p_direction")
  tab |> 
    mutate(Parameter = str_replace(Parameter, "b_Sex", "b_p_Sex"),
           Parameter = str_replace(Parameter, "b_Condition", "b_p_Condition")) |> 
    separate(Parameter, into = c("temp", "Parameter", "Effect"), sep = "_") |> 
    insight::format_table() |> 
    select(-temp) |> 
    mutate(col = tab$pd) |> 
    gt::gt() |> 
    gt::data_color(columns = "Parameter") |> 
    gt::data_color(columns = "col", method = "numeric", target_columns = "pd", 
                   row = col > 0.95, palette = "yellow") |> 
    gt::data_color(columns = "col", method = "numeric", target_columns = "pd", 
                   row = col > 0.97, palette = "green") |> 
    gt::cols_hide("col")
}
params_table(m1_chocoat, keep = "polyAttractive")
```

##### Random Effects

```{r}
table_dvour <- function(g) {
  performance::performance_dvour(g) |> 
    arrange(desc(D_vour)) |> 
    gt::gt() |> 
    gt::data_color("D_vour", palette = c("red", "red", "orange", "yellow", "green"), domain = c(0, 1)) |> 
    gt::fmt_auto()
}

g1 <- modelbased::estimate_grouplevel(m1_chocoat)
table_dvour(g1)
```

##### Full

```{r}
summary(m1_chocoat, robust = TRUE)
```

:::

- p: poly1 male
- confright: poly2 male
- confleft: poly1 male (neg)
- bex: poly1 male
- Suggests that (related to the original FakeFace1 hypothesis) reality would be driven by extremes degrees of attractiveness.

#### Model Visualization

```{r}
visualize_dchoco <- function(m, newdata, x = c(-0.005, seq(0, 1, length.out = 1001), 1.005)) {

  p <- modelbased::estimate_prediction(m, data = newdata, predict = "mu", iterations = 50)
  confright <- modelbased::estimate_prediction(m, data = newdata, predict = "confright", iterations = 50)
  confleft <- modelbased::estimate_prediction(m, data = newdata, predict = "confleft", iterations = 50)
  precright <- modelbased::estimate_prediction(m, data = newdata, predict = "precright", iterations = 50)
  precleft <- modelbased::estimate_prediction(m, data = newdata, predict = "precleft", iterations = 50)
  pex <- modelbased::estimate_prediction(m, data = newdata, predict = "pex", iterations = 50)
  bex <- modelbased::estimate_prediction(m, data = newdata, predict = "bex", iterations = 50)
  if ("pmid" %in% names(insight::find_parameters(m))) {
    pmid <- modelbased::estimate_prediction(m, data = newdata, predict = "pmid", iterations = 50)
  } else {
    pmid <- list(Predicted = rep(0, nrow(newdata)))
  }

  densities <- data.frame()
  mus <- data.frame()
  for(i in 1:nrow(newdata)) {
    # Get distribution that will be shown
    d <- data.frame(
       x = x,
       y = cogmod::dchoco(x, p=p$Predicted[i], confright=confright$Predicted[i], confleft = confleft$Predicted[i],
                          precright = precright$Predicted[i], precleft = precleft$Predicted[i],
                          pex = pex$Predicted[i], bex = bex$Predicted[i], pmid = pmid$Predicted[i])
     )

    densities <- rbind(densities,  merge(d, newdata[i, ]))

    # Get mode
    x_mus <- c(-0.005, seq(0, 1, length.out = 1001), 1.005)
    d <- data.frame(
       x = x_mus,
       y = cogmod::dchoco(x_mus, p=p$Predicted[i], confright=confright$Predicted[i], confleft = confleft$Predicted[i],
                          precright = precright$Predicted[i], precleft = precleft$Predicted[i],
                          pex = pex$Predicted[i], bex = bex$Predicted[i], pmid = pmid$Predicted[i])
     )
    y <- c(max(d[d$x > 0.5, "y"]), max(d[d$x < 0.5, "y"]))
    mu <- data.frame(
     y = y ,
     x = c(d[d$x > 0.5 & d$y == y[1], "x"], 
           d[d$x < 0.5 & d$y == y[2], "x"]),
     side = c("confright", "confleft")
    )
    mus <- rbind(mus,  merge(mu, newdata[i, ]))
  }

  list(densities=densities, mu=mus)
}

pred <- visualize_dchoco(m1_chocoat,
                         newdata=insight::get_datagrid(m1_chocoat, length = 5, preserve_range = FALSE))

p3_density <- pred$densities |> 
  mutate(Attractive = as.factor(Attractive),
         y = ifelse(x %in% c(0, 1), y * 3, y)) |>  # Inflate extremes for visibility
  ggplot(aes(x=x, y=y)) +
  geom_line(aes(color=Attractive), linewidth = 1.25) +
  geom_path(data = pred$mu, aes(group = side), linewidth=0, lineend = 'round', 
            arrow = arrow(length = unit(0.025, "npc"), type = "closed")) +
  facet_wrap(~Sex) +
  labs(y = "Distribution", x = "Do you think this image is real?", color = "Attractiveness") +
  scale_colour_manual(values=scales::seq_gradient_pal("#FFCDD2", "#B71C1C", "Lab")(seq(0,1,length.out=5))) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.25, 0.5, 0.75, 1), labels = c("      0%", "25%", "50%", "75%", "100%      ")) +
  scale_linewidth_continuous(range = c(0.5, 2), guide = "none") +
  theme_minimal() +
  theme(axis.title.x = element_text(face = "bold"), 
        axis.text.y = element_blank(), 
        # axis.title.y = element_blank(), 
        strip.background = element_rect(fill = "lightgrey", color = NA)) 
p3_density
```

```{r}
#| cache-lazy: false

colors <- c("p" = "#9C27B0", "confright" = "#795548", "confleft" = "#795548",
            "precright" = "#607D8B","precleft" = "#607D8B", "pex" = "#FF9800", "bex" = "#F44336", "pmid" = "black")

plot_params1 <- function(m, xlabel = "Attractiveness") {
  pred_params <- data.frame()
  for(param in c("mu", "confright", "confleft", "precright", "precleft", "pex", "bex")) {
    pred_params <- m |> 
      modelbased::estimate_prediction(data = "grid", length = 20, preserve_range = FALSE, predict = param, iterations = 10) |>
      as.data.frame() |> 
      dplyr::mutate(Parameter = param) |> 
      rbind(pred_params)
  }

  if(xlabel == "Attractiveness") {
    pred_params$x <- pred_params$Attractive
  } else {
    pred_params$x <- pred_params$Beauty
  }

  pred_params |> 
    mutate(Parameter = ifelse(Parameter == "mu", "p", Parameter),
           Parameter = fct_relevel(Parameter, "p", "confright", "precright", "pex", "confleft", "precleft", "bex")) |> 
    filter(Parameter != "bex") |>
    ggplot(aes(x = x, y = Predicted)) +
    geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = Parameter, group = Sex), alpha = 0.2) +
    geom_line(aes(color = Parameter, linetype = Sex), linewidth = 1) +
    facet_wrap(~Parameter, scales = "free_y", ncol=3, strip.position = "left") +
    labs(x = xlabel) +
    scale_fill_manual(values = colors, guide = "none") +
    scale_color_manual(values = colors, guide = "none") +
    # scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c("      0%", "50%", "100%      ")) +
    theme_minimal() +
    theme(strip.placement = "outside",
          axis.title.y = element_blank(),
          axis.title.x = element_text(face = "bold"),
          strip.text = element_text(size = rel(1), face = "italic"),
          panel.grid.minor.x = element_blank())
}

plot_params1(m1_chocoat, xlabel = "Attractiveness")
```

### Beauty

```{r}
m <- lme4::lmer(Real ~ Sex / poly(Beauty, 2) + (1|Participant) + (1|Item), data = df)
display(parameters::parameters(m, effects = "fixed"))

estimate_relation(m, length = 30) |> 
  ggplot(aes(x=Beauty, y = Predicted)) + 
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high,, fill = Sex), alpha = 0.2) +
  geom_line(aes(color = Sex), linewidth = 1) +
  geom_smooth(data=df, aes(color = Sex, y = Real), method = "loess", se = FALSE, linetype = "dotted", formula = 'y ~ x') +
  scale_color_manual(values = c("Male"="#2196F3", "Female"="#E91E63")) +
  scale_fill_manual(values = c("Male"="#2196F3", "Female"="#E91E63")) +
  theme_minimal()
```

#### Model Description

::: {.panel-tabset}

##### Effect of Beauty

```{r}
m1_chocobeauty <- readRDS("models/sample1_chocobeauty.rds")
params_table(m1_chocobeauty, keep = "polyBeauty")
```

##### Random Effects

```{r}
table_dvour(modelbased::estimate_grouplevel(m1_chocobeauty))
```

##### Full

```{r}
summary(m1_chocobeauty, robust = TRUE)
```

:::

#### Model Visualization

```{r}
pred <- visualize_dchoco(m1_chocobeauty,
                         newdata=insight::get_datagrid(m1_chocobeauty, length = 5, preserve_range = FALSE))

p4_density <- pred$densities |> 
  mutate(Attractive = as.factor(Beauty),
         y = ifelse(x %in% c(0, 1), y * 3, y)) |>  # Inflate extremes for visibility
  ggplot(aes(x=x, y=y)) +
  geom_line(aes(color=Attractive), linewidth = 1.25) +
  # ggforce::geom_link2(data = pred$mu, aes(group = side, linewidth = Attractive, color = as.factor(Attractive)), 
  #                     lineend = 'round', n = 500) +
  geom_path(data = pred$mu, aes(group = side), linewidth=0, lineend = 'round', 
            arrow = arrow(length = unit(0.025, "npc"), type = "closed")) +
  facet_wrap(~Sex) +
  labs(y = "Distribution", x = "Do you think this image is real?", color = "Attractiveness") +
  scale_colour_manual(values=scales::seq_gradient_pal("#C5CAE9", "#283593", "Lab")(seq(0,1,length.out=5))) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.25, 0.5, 0.75, 1), labels = c("      0%", "25%", "50%", "75%", "100%      ")) +
  scale_linewidth_continuous(range = c(0.5, 2), guide = "none") +
  theme_minimal() +
  theme(axis.title.x = element_text(face = "bold"), 
        axis.text.y = element_blank(), 
        # axis.title.y = element_blank(), 
        strip.background = element_rect(fill = "lightgrey", color = NA)) 
p4_density
```

```{r}
#| cache-lazy: false

plot_params1(m1_chocobeauty, xlabel = "Beauty")
```

## Sample 2

```{r}
#| eval: false

df <- read.csv("https://raw.githubusercontent.com/RealityBending/FakeFace2/refs/heads/main/data/data_task.csv") |> 
  filter(Relevance == "Relevant") |> 
  mutate(Real = (Realness + 3) / 6,
         Attractive = Attractiveness / 6,
         Beauty = Beauty / 6,
         RealNoex = datawizard::change_scale(Real, to = c(0.125, 0.875))) |> 
  select(Participant, Sex=Gender, Condition, Item=Stimulus, Real, RealNoex, Attractive, Beauty)

write.csv(df, "../data/sample2.csv", row.names = FALSE)
```

```{r}
df <- read.csv("https://raw.githubusercontent.com/RealityBending/FictionChoco/refs/heads/main/data/sample2.csv")
df$Sex <- factor(df$Sex, levels = c("Male", "Female"))
df$Condition <- factor(df$Condition, levels = c("Photograph", "AI-Generated"))
```

### Attractiveness

```{r}
m <- lme4::lmer(Real ~ Sex / poly(Attractive, 2) * Condition + (1|Participant) + (1|Item), data = df)
display(parameters::parameters(m, effects = "fixed"))

estimate_relation(m, length = 30) |> 
  ggplot(aes(x=Attractive, y = Predicted)) + 
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high,, fill = Condition), alpha = 0.2) +
  geom_line(aes(color = Condition), linewidth = 1) +
  geom_smooth(data=df, aes(color = Condition, y = Real), method = "loess", se = FALSE, linetype = "dotted", formula = 'y ~ x') +
  scale_color_manual(values = c("AI-Generated"="#3F51B5", "Photograph"="#FF5722")) +
  scale_fill_manual(values = c("AI-Generated"="#3F51B5", "Photograph"="#FF5722")) +
  facet_wrap(~Sex) +
  theme_minimal()
```

#### Posterior Predictive Check

```{r}
m2_chocoat <- readRDS("models/sample2_chocoattractiveness.rds")

m <- m2_chocoat

make_ppcheck  <- function(m, fill = "#FF9800") {
  pred <- m |> 
    modelbased::estimate_prediction(df, keep_iterations = 100, iterations = 100) |>
    select(Sex, starts_with("iter_")) |> 
    bayestestR::reshape_iterations()

  pred |>
    ggplot(aes(x = iter_value)) +
    stat_bin(data = df, aes(x = Real, y = after_stat(density)),
             geom="bar", breaks = seq(0, 1.0, by = 0.04), fill = "#424242") +
    stat_bin(aes(group = iter_group, y = after_stat(density)), position = "identity",
             geom="bar", breaks = seq(0, 1.0, by = 0.04), pad = TRUE,
             alpha = 0.01, fill = fill) +

    scale_y_sqrt(expand = c(0, 0.01)) +
    # scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c("      0%", "50%", "100%      ")) +
    theme_minimal() +
    theme(axis.title.x = element_text(face = "bold"), 
          axis.title.y = element_blank(), 
          axis.text.y = element_blank(), 
          strip.background = element_rect(fill = "lightgrey", color = NA)) +
    facet_wrap(~Sex, scales = "free_y") +
    coord_cartesian(xlim = c(-0.003, 1.005), expand = TRUE)
}

make_ppcheck(m2_chocoat) +
  labs("Real")
```

#### Model Description

::: {.panel-tabset}

##### Effect of Attractiveness

```{r}
params_table(m2_chocoat, keep = "polyAttractive|Condition")
```

##### Effect of Condition

```{r}
# c <- estimate_contrasts(m2_chocoat, contrast = "Condition", by = "Sex", predict = "mu")
# c <- estimate_contrasts(m2_chocoat, contrast = "Condition", by = "Sex", predict = "mu", backend = "emmeans")
# c2 <- estimate_contrasts(m2_chocoat, contrast = "Condition", by = "Sex", predict = "confright", backend = "emmeans")
```

##### Full

```{r}
summary(m2_chocoat, robust = TRUE)
```

:::

#### Model Visualization

```{r}
pred <- visualize_dchoco(m2_chocoat,
                         newdata=insight::get_datagrid(m2_chocoat, length = 5, preserve_range = FALSE),
                         x = seq(0, 1, length.out=7))

p4_density <- pred$densities |> 
  mutate(Attractive = as.factor(Attractive),
         y = ifelse(x %in% c(0, 1), y * 3, y),  # Inflate extremes for visibility
         x = ifelse(Condition == "AI-Generated", x + 0.1, x)) |>  
  filter(Condition == "Photograph") |> 
  ggplot(aes(x=x, y=y)) +
  geom_line(aes(color=Attractive), linewidth = 1.25) +
  geom_path(data = filter(pred$mu, Condition == "Photograph"), aes(group = side), linewidth=0, lineend = 'round', 
            arrow = arrow(length = unit(0.025, "npc"), type = "closed")) +
  facet_wrap(~Sex) +
  labs(y = "Distribution", x = "Do you think this image is real?", color = "Attractiveness") +
  scale_colour_manual(values=scales::seq_gradient_pal("#FFCDD2", "#B71C1C", "Lab")(seq(0,1,length.out=5))) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.25, 0.5, 0.75, 1), labels = c("      0%", "25%", "50%", "75%", "100%      ")) +
  scale_linewidth_continuous(range = c(0.5, 2), guide = "none") +
  theme_minimal() +
  theme(axis.title.x = element_text(face = "bold"), 
        axis.text.y = element_blank(), 
        # axis.title.y = element_blank(), 
        strip.background = element_rect(fill = "lightgrey", color = NA)) 
p4_density
```

```{r}
# pred <- visualize_dchoco(m2_chocoat,
#                          newdata=insight::get_datagrid(m2_chocoat, length = 5, preserve_range = FALSE),
#                          x = seq(0, 1, length.out=7))
```

### Beauty

```{r}
m <- lme4::lmer(Real ~ Sex / poly(Beauty, 2) * Condition + (1|Participant) + (1|Item), data = df)
display(parameters::parameters(m, effects = "fixed"))

estimate_relation(m, length = 30) |> 
  ggplot(aes(x=Beauty, y = Predicted)) + 
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high,, fill = Condition), alpha = 0.2) +
  geom_line(aes(color = Condition), linewidth = 1) +
  geom_smooth(data=df, aes(color = Condition, y = Real), method = "loess", se = FALSE, linetype = "dotted", formula = 'y ~ x') +
  scale_color_manual(values = c("AI-Generated"="#3F51B5", "Photograph"="#FF5722")) +
  scale_fill_manual(values = c("AI-Generated"="#3F51B5", "Photograph"="#FF5722")) +
  facet_wrap(~Sex) +
  theme_minimal()
```

## Figures

```{r}
#| eval: false

p3_a <- wrap_elements(p3_density / p3_params + 
         patchwork::plot_annotation(title = "Effect of Attractiveness", theme = theme(plot.title = element_text(face = "bold"))))
```

### 2025-04-27 paper/manuscript.qmd
update

---
title: "**Introducing the Choice-Confidence (CHOCO) Model for Bimodal Scales Data: Application to the Relationship between Reality Beliefs about AI-Generated Faces and Beauty**"
shorttitle: "Choice-Confidence (CHOCO) Model"
author:
  - name: Dominique Makowski
    corresponding: true
    orcid: 0000-0001-5375-9967
    email: D.Makowski@sussex.ac.uk
    url: https://realitybending.github.io/
    roles:
      - Conceptualization
      - Data curation
      - Formal Analysis
      - Funding acquisition
      - Investigation
      - Methodology
      - Project administration
      - Resources
      - Software
      - Supervision
      - Validation
      - Visualization
      - Writing – original draft
    affiliations:
      - id: "id1"
        name: "University of Sussex"
        department: "School of Psychology"
        city: "Brighton"
        country: "UK"
        postal-code: "BN1 9RH"
      - name: "University of Sussex"
        department:  "Sussex Centre for Consciousness Science"
  - name: Ana Neves  
    orcid: 0009-0006-0020-7599
    role:
      - Data curation
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - ref: "id1"
  - name: Andy Field 
    orcid: ""
    role:
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - ref: "id1"
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    gratitude: |
      ::: {.callout-note icon=false appearance="simple"}
      This preprint is a non-peer-reviewed work from the [**Reality Bending Lab**](https://realitybending.github.io/).
      ![](https://realitybending.github.io/media/ReBeL_LogoOnly_hu11484441381606756729.png){width=20% fig-align="center"}
      :::
    study-registration: ~
    data-sharing: ~
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    authorship-agreements: ~
abstract: |
  TO DO.
#   **Significance Statement**. TODO.
keywords: "subjective scales, choice confidence model, reality beliefs, AI-generated faces, attractiveness"
numbered-lines: false
floatsintext: true
mask: false
bibliography: bibliography.bib
format:
  apaquarto-pdf:
    documentmode: jou
    keep-tex: true
  apaquarto-docx: default
---

# Introduction

<!-- Problem statement -->
Despite significant advancements in psychological science following the replication crisis [@open2015estimating], its progress is still hindered by its sub-optimal (or inappropriate) usage of statistical tools [@makowski2023we]. 
A prevalent issue is the continued reliance on linear models that assume normally distributed (Gaussian) data - as this assumption often does not hold true for many types of psychological outcomes. 
For instance, reaction times typically exhibit skewed distributions, choices can be represented as binary variables, and count data consists of strictly positive integers. 
Applying models that presume normality and model the "mean" of the outcome variable can lead to misinterpretation and potentially misleading conclusions when applied indiscriminately. It is thus important that psychologists use models that can best describe (or generate) the data they collect, to fully exploit them and bring more nuance and accuracy to their conclusions.

<!-- Bounded data, Beta-models and ZOIB -->
Among the most commonly collected data in psychology are responses on subjective scales, such as Likert-type items or visual analog scales, which exhibit some fundamental properties: these responses are bounded (and can be rescaled to a 0-1 range) and frequently display clustering at the extremes. 
Traditional linear models being ill-suited for such data, researchers have turned to using Beta distributions to model this data (instead of Gaussian), suited for continuous data within the (0,1) interval (i.e., excluding extreme responses). 
To address the frequent occurrence of exact zeros and ones (i.e., extreme values), zero-one inflated beta (ZOIB) models have been developed [@ospina2012general] to accommodate the excess of boundary values by incorporating additional components that model the probabilities of responses at 0 and 1 as a separate, independent process.

## The Beta-Gate Model

The Beta-Gate model is a reparametrized Ordered Beta model [@kubinec2023ordered]^[In the Ordered Beta model, the cutpoints on the log-scale are directly used as parameters, instead of being derived from *pex* and *bex*.] available in the *cogmod* package in R (https://github.com/DominiqueMakowski/cogmod), in which participants' answers on bounded scales are conceptualized as latent responses that can fall past a pair of probabilistic "gates" (or cutpoints) that control whether the response is recorded as an extreme (0 or 1) or as a nuanced, continuous value in between (@fig-one). 
These distance of these gates from the edges of the scale varies based on two interpretable parameters: *pex* (the propensity by which people are likely to answer extreme values), and *bex*, a bias toward the upper extreme (1) versus the lower (0). 
A person's internal response that lies close to the edge might be "caught" by a gate and recorded as an extreme, while others pass through to express a continuous response (Beta-distributed with $\mu$ (*mu*) and $\phi$ (*phi*) as its mean and precision parameters).
The Beta-Gate model is based on the idea that extreme values can emerge not just from a fundamentally different underlying processes - as assumed in ZOIB models - but from a common process governed by thresholds of decisiveness and confidence.

```{r}
#| label: fig-one
#| fig-cap: "The Beta-Gate Distribution is a reparametrized ordered Beta model (Kubinec, 2023) that is governed by 4 parameters. 'Mu' and 'phi' correspond to the mean and precision of the continuous part of the distribution (between 0 and 1), and 'pex' (propensity of extremes) and 'bex' (balance of extremes) indirectly control the proportion of zeros and ones by specifying the location of the \"gates\", past which the latent response process is likely to generate extreme values. Specifically, 'pex' defines the total distance of both gates from the extremes (in yellow), and 'bex' determines the proportion of the right gate distance relative to the left. In this example, the total distance from the extremes is 'pex' = 0.1, with 40% ('bex' = 0.4) of that distance being on the right (and 60% on the left). The left gate is thus located at 0.6, and the right at 1-0.04 = 0.96."
#| apa-twocolumn: false  # A Figure Spanning Two Columns When in Journal Mode
#| out-width: "50%"

knitr::include_graphics("./figures/fig1.png")
```

Mathematically, the Beta-Gate distribution defines the observed outcome $x \in [0, 1]$ as a mixture of three components; a point mass at 0, a point mass at 1, and a continuous Beta density over $x \in (0,1)$, scaled by the remaining probability mass. 
The probability of these components are:

- $P(x = 0) = \text{logistic}\left( \text{logit}(pex \cdot (1 - bex)) - \text{logit}(\mu) \right)$
- $P(x = 1) = 1 - \text{logistic}\left( \text{logit}(1 - pex \cdot bex) - \text{logit}(\mu) \right)$
- $P(x \in (0,1)) = 1 - P(x = 0) - P(x = 1)$

The continuous part follows a Beta distribution with parameters^[Note that *phi* is scaled in Beta-Gate models relative to the traditional mu/phi Beta particularization so that a phi of 1 corresponds to a uniform distribution - to facilitate setting priors on this parameter]:

$$
Beta(\alpha = \mu \cdot 2 \phi, \quad \beta = (1 - \mu) \cdot 2 \phi)
$$

## The Choice-Confidence (CHOCO) Model

Decision-making is often conceptualized as involving distinct processes: the choice itself, and the confidence associated with that choice. 
In experimental paradigms, these can be somewhat disentangled by prompting participants to make a discrete choice selection (e.g., "True" vs. "False"), followed by a separate confidence rating. 
However, this artificial separation makes its joint analysis difficulty, and may not reflect real-world decision-making, where individuals often express both choice and confidence simultaneously using a single, continuous scale. In such scales, each side can represent a distinct latent category, and the distance from the midpoint can indicates the level of confidence or certainty.
This integrated response format typically results in bimodal distributions, with peaks corresponding to the mean confidence on either side.
Traditional beta regression models, which assume unimodal distributions within the (0,1) interval, are ill-suited for such data. One alternative is to transform the data into two variables a posteriori: binarizing the side to represent choice and calculating the absolute distance from the midpoint to represent confidence. 
These can then be modeled separately, for instance, using logistic regression for choice and beta regression for confidence [see @makowski2025too for an example]. 
While this approach can provide additional insights into underlying mechanisms compared to a unique model, it assumes psychological and statistical independence between choice and confidence, which may not hold true in practice.

```{r}
#| label: fig-two
#| fig-cap: "The CHOCO Model uses a mixture of Beta-Gate distributions to model separately the right and left sides of the scale (e.g., a rating of whether a statement was 'Truth' vs. 'Lie'), as well as their relative proportion. In this example, the participants are more likely overall (p = 60%) to select the right side of the scale ('Lie') than the left ('Truth'). They are also more confident in their choice (confright = 0.6 vs. confleft = 0.3). Extreme values (zeros and ones) are governed by the same mechanism as for Beta-Gate models."
#| apa-twocolumn: false  # A Figure Spanning Two Columns When in Journal Mode
#| out-width: "50%"

knitr::include_graphics("./figures/fig2.png")
```

To model data of subjective scales in which the left and right sides can be conceptualized as two different choices (e.g., True/False, Agree/Disagree, etc.) and the magnitude of the response (how much the cursor is set away from the mid-point) as the confidence, we introduce the Choice-Confidence (CHOCO) model (@fig-two).
It consists of a three‐part mixture on $x \in [0, 1]$:

- An (optional) point-mass at the midpoint *mid* (typically 0.5) of weight $p_{\text{mid}}$ for undecided or neutral responses.  
- A left-choice component governed by a Beta-Gate density on the rescaled variable $x/mid$ with mean *1 - confleft*, precision *precleft*, and boundary-excess parameter *pex(1 - bex)*.  
- A right-choice component governed by a Beta-Gate density on the rescaled variable $(x - mid)/(1 - mid)$ with mean *confright*, precision *precright*, and boundary-excess parameter *pex x bex*.

The overall probability of the right choice (relative to the left choice) is controlled by a main parameter *p*.
The full CHOCO density is:

$$
\small
\begin{cases}
pmid, & x = mid, \\[6pt]
(1 - pmid)(1 - p)\,\dfrac{1}{mid} \cdot \mathrm{BetaGate}\left(\dfrac{x}{mid}\right), & 0 < x < mid, \\[6pt]
(1 - pmid)\,p\,\dfrac{1}{1 - mid} \cdot \mathrm{BetaGate}\left(\dfrac{x - mid}{1 - mid}\right), & mid < x < 1. \\
\end{cases}
$$

By coupling choice probability $p$, midpoint mass $p_{\text{mid}}$, and side‐specific Beta-Gate parameters ($conf$, $prec$, $pex$, $bex$), CHOCO flexibly captures both bimodality and confidence intensity in a single unified model.
Despite this theoretical appeal, it is unclear whether this heavily parametrized model can be estimated reliably from data, and whether it can provide more useful insights than simpler alternatives.

## Aim of the Present Study

**Study 1** aims to evaluate the CHOCO model’s ability to better capture subjective scale responses that (potentially) reflect an underlying discrete choice, in comparison to existing models such as the ZOIB and Beta-Gate. Specifically, we will assess whether 1) CHOCO provides improved model fit, 2) yields deeper insights into population-level effects than traditional approaches (gender differences in reality beliefs), and 3) allows for the reliable estimation of interpretable individual-level parameters through random effects.

To this end, we analyze data from two separate studies in which participants judged whether a face image was AI-generated ("fake") or a real photograph. **Study 2** will apply this model to more subtle effects, such as the effect perceived facial attractiveness and beauty on reality judgments.

# Study 1

**TODO**: Write some introduction about reality beliefs in the context of ambiguity. Underline that what drives judgments and beliefs of reality remains unclear. And this is important in today's post-truth era.

Another unknown is whether these beliefs are fully stimuli-driven or whether there are some stable individual tendencies that are at stake.

## Methods

### Participants

The sample is the same as in [@makowski2025too]. We included all heterosexual male and female participants, for a final sample of 125 participants (Mean age = 28.6, SD = 9.2, range: [19, 66]; Sex: 44.8% females). We included only sexually relevant stimuli for men and women (i.e., all 89 female faces for men and 20 male faces for women).

### Procedure

In the first phase, participants viewed 109 neutral-expression photographs of faces (random order, display time of 500 ms) from the American Multiracial Face Database [@chen2021broadening]. After each image, participants rated the face on trustworthiness, familiarity, attractiveness, and beauty using visual analog scales. In the second phase, participants were informed that "about half of the previously seen images were AI-generated". The same faces were presented again in a new random order (same display time), followed by ratings of "reality" (whether they believed the image was AI-generated - left anchor - or a real photograph - right anchor). @makowski2025too showed that most stimuli yielded highly variable reality ratings, suggesting that idiosyncratic cognitive processes were at play.

### Data Analysis

We fitted 3 models to predict the reality ratings: a ZOIB model, a Beta-Gate model, and the CHOCO model. For all models and each parameter, the full formula was entered: $Real\sim Sex + (1 | Participant) + (1 | Item)$ (with Sex as the main predictor and participants and items entered as random intercepts). The models were run using *brms* [@burkner2017brms] R package, and analyzed using the *easystats* collection of packages [**REFS**]. To maximize the comparability across models We used the default priors (uniform) for all models, and we ran 16 chains of 1400 iterations each on the University of Sussex High-Performance Computing (HPC) cluster.

Model comparisons were performed using the *loo* R package [@vehtari2017practical], which computes the Widely Applicable Information Criterion (WAIC) and estimates the Expected Log Predictive Density (ELPD) and penalizes the number of parameters. We assessed model performance by examining ELPD differences and their standard errors (SE), reporting corresponding *p*-values to determine significant differences in predictive accuracy.

For the population-level effects, we will consider significant and report (using the median of the posterior distribution) effects for which the 95% Credible Interval (CI) does not include zero [and when the probability of direction *pd* is > \~97\%; @makowski2019indices]. For the individual-level parameters (i.e., the random intercepts of each parameter for each participant and each item), we will first analyze their reliability using the Variance-Over-Uncertainty Ratio index (*D-vour*). This index, implemented in the *performance* package [@ludecke2021performance], is inspired by recent work on mixed models reliability [@williams2021beneath; @rouder2024hierarchical], and corresponds to the normalized ratio of observed variability to uncertainty in random effect estimates, defined as:

$$
D_{\text{vour}} = \frac{\sigma_B^2}{\sigma_B^2 + \mu_{\text{SE}}^2}
$$

Where $\sigma_B^2$ is the between-group variability (computed as the SD of the random effect point-estimates) and $\mu_{\text{SE}}^2$ is the mean squared uncertainty in random effect estimates (i.e., the average uncertainty).
We use as *D-vour* = 0.666 as the threshold for moderately reliable random effect estimates, which corresponds to a 2:1 ratio of between-group variance to uncertainty.

Finally, we will run a correlation analysis of the models' individual-level estimates against "empirical" (indices computed directly on the observed data), including the empirical *p* ($P(y > 0.5)$), the overall *conf* ($mean(|y - 0.5|)$), *pex* ($P(y \in [0, 1])$) and *bex* ($P(y == 1) / P(y \in [0, 1])$), assessing whether the model's estimate are in-line with easily interpretable indices.

## Results

The reproducible code and full result report are available at https://github.com/RealityBending/FictionChoco.

### Model Comparison

```{r}
#| label: fig-three
#| fig-cap: "TODO."
#| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
#| out-width: "100%"

knitr::include_graphics("./figures/fig3.png")
```

The models did converge without divergent transitions, and the effective sample size was sufficient for all parameters (all $n_{\text{eff}} > 1000$).
The difference in predictive accuracy, as indexed by Expected Log Predictive Density (ELPD-WAIC), suggests that the *CHOCO* is the best
model ($ELPD = -138.45$), followed by *Beta-Gate* ($\Delta_{ELPD} = -1875.31 \pm 61.18, p < .001)$ and *ZOIB* ($\Delta_{ELPD} = -1916.11 \pm 61.75, p < .001$). See @fig-three for the posterior predictive checks, showing that only the CHOCO model managed to capture the bimodal distribution of data.

### Effect of Sex

The ZOIB model suggested that women had higher mean (*mu* parameter) scores of reality beliefs ($\Delta_{\text{Female}} = 0.16,~95\%~ CI~[0.01, 0.32], pd = 98.10\%$), less extreme values (*zoi* parameter; STATS) but more ones relative to zeros (*coi* parameter; STATS).

The Beta-Gate model similarly suggested that women had higher mean (*mu* parameter) scores of reality beliefs (STATS), less extreme values (*pex* parameter; STATS), and a greater tendency to answer one relative to zero (*bex* parameter; STATS).

The CHOCO model shows that women had a higher probability *p* of judging faces as real (STATS), but are not more confident when doing so (*confright* parameter; STATS). However, they were less confident when answering that an image was AI-generated (*confleft* parameter; STATS). There were also less likely to produce extreme answers (*pex* parameter; STATS), but no strong evidence supporting a directional bias was observed (*bex* parameter; STATS).

Across all these models, no effect of Sex on the precision parameter was observed.

### Individual-Level Parameters

```{r}
#| label: fig-four
#| fig-cap: "TODO."
#| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
#| out-width: "100%"

knitr::include_graphics("./figures/fig4.png")
```

The ZOIB model estimated reliable variability in the participant's *phi* parameter (D-vour = 0.88) and *zoi* parameter (D-vour = 0.84), as well as in the *mu* parameter related to individual items (D-vour = 0.82). Moderate reliability was also observed for items in the *coi* parameter (D-vour = 0.71) and for participants in the *mu* parameter (D-vour = 0.69). The Beta-Gate model yielded similar results: a high reliability of participant's *phi* parameter (D-vour = 0.88), *pex* parameter (D-vour = 0.84). The *mu* parameter's variability was reliably captured for items (D-vour = 0.85) and moderately for participants (D-vour = 0.72). 
Of note that the parameters of Beta-Gate and ZOIB correlate almost perfectly, underlining their empirical similarity despite a different parametrization.

The CHOCO model yielded reliable estimates for all parameters except *bex* for participants (*confright* D-vour = 0.94, *confleft* D-vour = 0.91, *pex* D-vour = 0.80, *p* D-vour = 0.73, *precright* D-vour = 0.79, *precleft* D-vour = 0.73). Item's variability was primarily reflected through the *p* parameter (D-vour = 0.86).

Finally, the empirical average correlated the most strongly with CHOCO's *p* (r = .94), rather than ZOIB's *mu* (r = .74) or
Beta-Gate's *mu* (r = .82). The empirical overall confidence was the strongest correlated with CHOCO's *confright* (r = .91), followed by ZOIB's *phi* (r = -.86), Beta-Gate's phi (r = -.83), and other CHOCO's parameters. The empirical proportion of "right" answer *p* correlated the strongest with CHOCO's *p* (r = 0.94), followed by Beta-Gate's *mu* (r = .82) and ZOIB's *mu* (r = .77). The empirical *pex* correlated the strongest with ZOIB's *zoi* (r = .90), and Beta-Gate's *pex* (r = .90), and CHOCO's *pex* (r = .88).

## Discussion

Study 1 revealed that the Choice-Confidence (CHOCO) model was a much better fit for bimodal bounded data, compared to other alternatives like the Zero-and-One Inflated Beta (ZOIB) and Beta-Gate (Ordered Beta) models. It also allowed for a deeper understanding through its interpretable parameters, offering insights into possibly distinct cognitive mechanisms, such as the probability of answering real vs. fake, and the associated confidence in these two choices. This was illustrated by modelling the effect of sex on all the CHOCO parameters.

Note that the observed gender differences are primarily presented as a proof-of-principle, to showcase the model's ability to capture group-level effects and to provide deeper insights compared to other models. However, given that they were based on different items (female and male faces), these differences might just be a reflection of stimuli characteristics rather than true sex dymorphism in the formation of reality beliefs.

Finally, we also show that the CHOCO model was able to capture reliable and interpretable individual-level parameters, supporting its value to measure inter-individual differences. An interesting dissociation emerged between participant- and item-level variability: the latter seemed mostly to be represented in the *p* parameter, while participants reliably varied in most of the components (aside from *bex*). This could be interpreted as that external item characteristics primarily influence the probability of being judged as real vs. fake, while the expressed confidence is first and foremost an individual characteristic.

# Study 2

The explosion of accessibility of state-of-the-art AI tools has made it effortless to generate realistic images, including human 
faces that are often indistinguishable from real ones **[REF]**. These synthetic visuals are flooding the cyberspace across various domains, such as art, advertising and entertainment, to education and information. This technological advancement carries an important potential for misuse, such as in disinformation campaigns, scams (e.g., AI-bots, identity theft), and abuse. The democratization of such technology raises pressing concerns about the value of authenticity and the potential erosion of media trust in our increasingly *post-truth* society.

In this evolving landscape, understanding the cognitive mechanisms that underpin our judgments of reality becomes paramount. Despite the increasing prevalence of digitally altered or AI-generated content, humans still rely on certain heuristics to assess authenticity. One such heuristic might be facial attractiveness. Attractiveness appraisals are known to be automatic and unconscious [@luo2019robust; @hung2016unconscious; @hou2023neural], and carry strong real-life consequences, as demonstrated by the large body of literature on the "beauty premium" [@little2021facial; @gulati2024beautiful, **MORE**]^[As an eloquent example, @monk2021beholding reported in a large representative US sample that the magnitude of the earnings disparities among white women along the perceived attractiveness continuum exceeds in magnitude the canonical black-white race gap.]

**TODO:** However, it is important to allow for the possibility of a U-shaped relationship. For instance, Peng et al. (2020) reported that both attractive and unattractive people sold more (underlining a "plainness penalty").
<!-- The Faces of Success: Beauty and Ugliness Premiums in e-Commerce Platforms -->

Although its role as a potential modulator of reality beliefs remains underexplored, Makowski et al... **TODO: detail findings of that study + other studies**.

In this study, we will apply the CHOCO model to the data of MAKOWSKI, and see whether different analysis approaches lead to different conclusions. We will then try to replicate these findings in a new sample of participants.

## Methods

### Participants

The first sample includes the same participants as study 1 (see above). 
**TODO** for sample 2.

### Procedure

For the second sample, the procedure was relatively similar with a few key differences.
**TODO** Describe  fake Face 2. 
The main difference is that the subjective ratings were done on a X-point Likert scale, rather than analog, with a mid-point value.

### Data Analysis

To compare the benefits of CHOCO models to a "traditional" analytic approach, we started by fitting a frequentist linear mixed model to predict reality beliefs with the formula $Real\sim Sex / poly(Attractive, 2) + (poly(Attractive, 2)|Participant) + (1|Item)$. The 2nd degree raw polynomial term was included to allow for potentially non-linear relationships (note that the first and second degree effects of *raw* polynomials can be interpreted independently as the linear part and the "curvy" part of the relationship). For the CHOCO model, mildly informative and effect-agnostic (i.e., centered at zero) priors were used. The same formula was used for all parameters, except that (based on the reliability analysis of Study 1), items were only included as random effects for *p*, and participants were not included for *bex*.

**TODO**. For sample 2, the analysis was based on that of Sample 1. The main differences are 1) the inclusion of the "Condition" (whether the picture was presented as "Real" of "Fake" in the first phase of the experiment) as an additional predictor (it was entered as the only random slope for all participant random effects); 2) the inclusion of an additional parameters, *pmid*, modelling the probability of answering the middle-point of the Likert scale (representing an "undecided" option).

The same analysis was applied for Beauty.

## Results

As the complete model parameters tables are available at https://github.com/RealityBending/FictionChoco, we will focus on reporting noteworthy findings below.

In sample 1, the traditional approach suggested a significant linear relationship between the mean level of "Reality" and attractiveness ($\beta_{poly1}=3.63, ~95\%~CI [2.71, 4.56], pd = ...\%$) for males only. The CHOCO model revealed that attractiveness had a significantly positive linear relationship with the probability *p* of judging faces as real (STATS), but a quadratic relationship with the confidence in real judgments (*confright*; STATS). Attractiveness was also associated with less confidence in fake judgments (*confleft*; STATS), and related linearly with a stronger bias towards extreme Real responses (*bex*; STATS). These effects were present for males only, and no relationship was found for females. The reliability of the effect of attractiveness (the random slopes) was very low for all parameters (D-vour < 0.1).

The effect of Beauty was the same as above, with some additional evidence in favour of an effect in women, in particular for a linear relationship with *p* for females as well ($\beta_{poly1}=7.02, ~95\%~CI [-0.90, 14.97], pd = 95.91\%$), as well as a quadradic relationship with the confidence in fake judgments (*confleft*; STATS).

In Sample 2, the traditional approach suggested a significant linear relationship between the mean level of "Reality" and attractiveness for males (STATS) and women (STATS), with no effect of the Condition. The CHOCO model revealed that attractiveness had a significantly positive linear relationship with the probability *p* of judging faces as real for males (STATS) and females (STATS). It also had a linear relationship with the confidence in real judgments (*confright*) for males (STATS) and females (STATS), as well as a significant quadratic curvature for males only (STATS). Attractiveness also linearly decreased the confidence in fake judgments only for males (*confleft*; STATS), but increased the probability of an undecided response (*pmid*) in women (STATS), and displayed in females only a quadratic relationship with the amount of extreme answers (*pex*; STATS) and its bias in favour of right-side extreme response (*bex*; STATS). No consistent and reliable effect of condition was found.

The effect of Beauty was...

## Discussion

The fact that the effect was less obvious in women in the sample 1 might be explained with power (less relevant items were included for women).

# General Discussion

Little difference between Beauty and Attractiveness, aside that Beauty revealed more effects in women. Why?

Likert-scales (with a limited number of discrete options) are commonly used in psychology, and often modeled as continuous.
While this fact has spurred its own polarized debate (REFS), our position is that the model used should ideally reflect the data *generating* mechanism rather than necessarily focus on the *observed* data^[Ideally, the data recording method should be aligned with the assumed generating mechanism, which is an experiment design issue rather than a statistical one.] - although the former cannot always be easily inferred or known. 
For the data of Sample 2, we held the assumption that the underlying latent distribution was CHOCO distributed (as there is no reason to assume it would be different from Sample 1), treated as continuous CHOCO-distributed data, despite the fact that it was collected on a 7-point Likert scale. 
The motivation was to see if the model could extrapolate and reliably estimate all the parameters based on a limited number of data points but means in our case that, effectively, each Beta distribution was extrapolated from the frequency of only two unique values (or 3 for the no-extreme version). 
While this is far from ideal, the CHOCO model still provided a better fit to the data than alternative approaches. 
**TODO** However, it might have lead to a lesser ability to detect finer grain changes, including U-shaped relationships. Having 7-points (or 6 with no mid-value) is the strict minimum to use the CHOCO model.

Future studies should assess the ability of CHOCO models to also fit unimodal distributions (so it can be used more generally), as well as the benefits of more parsimonious parametrization (for instance, a unique precision parameter controlling both the left and right sides, or a linked confidence parameter e.g. where *confleft* is expressed as a function of *confright*).

# Data Availability

Data, code and everything is available at https://github.com/RealityBending/FictionChoco.
The CHOCO model is implemented in the *cogmod* R package (https://github.com/DominiqueMakowski/cogmod).

# Acknowledgements

We would like to thank the dissertation students from the University of Sussex for their help in data collection.
DM would also like to thank ... for the motivation provided to write this paper.

# References

<!-- References will auto-populate in the refs div below -->

::: {#refs}
:::

### 2025-04-27 paper/manuscript.tex
update

\documentclass[
  jou,
  floatsintext,
  longtable,
  nolmodern,
  notxfonts,
  notimes,
  colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{apa7}

\usepackage{amsmath}
\usepackage{amssymb}

\RequirePackage{longtable}
\RequirePackage{threeparttablex}

\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-.5em}%
	{\normalfont\normalsize\bfseries\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{0.5em}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-\z@\relax}%
	{\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\usepackage{longtable, booktabs, multirow, multicol, colortbl, hhline, caption, array, float, xpatch}
\usepackage{subcaption}
\renewcommand\thesubfigure{\Alph{subfigure}}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.7}

\usepackage{tcolorbox}
\tcbuselibrary{listings,theorems, breakable, skins}
\usepackage{fontawesome5}

\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{ACACAC}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582EC}
\definecolor{quarto-callout-important-color-frame}{HTML}{D9534F}
\definecolor{quarto-callout-warning-color-frame}{HTML}{F0AD4E}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02B875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{FD7E14}

%\newlength\Oldarrayrulewidth
%\newlength\Oldtabcolsep

\usepackage{hyperref}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{newtx}

\defaultfontfeatures{Scale=MatchLowercase}
\defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}

\title{\textbf{Introducing the Choice-Confidence (CHOCO) Model for
Bimodal Scales Data: Application to the Relationship between Reality
Beliefs about AI-Generated Faces and Beauty}}

\shorttitle{Choice-Confidence (CHOCO) Model}

\usepackage{etoolbox}

\authorsnames[{1,2},{1},{1}]{Dominique Makowski,Ana Neves,Andy Field}

\authorsaffiliations{
{School of Psychology, University of Sussex},{Sussex Centre for
Consciousness Science, University of Sussex}}

\leftheader{Makowski, Neves and Field}

\abstract{TO DO. }

\keywords{subjective scales, choice confidence model, reality beliefs,
AI-generated faces, attractiveness}

\authornote{\par{\addORCIDlink{Dominique
Makowski}{0000-0001-5375-9967}}\par{\addORCIDlink{Ana
Neves}{0009-0006-0020-7599}}\par{\addORCIDlink{Andy Field}{true}} 
\par{ }
\par{     \begin{tcolorbox}[enhanced jigsaw, left=2mm, arc=.35mm, bottomrule=.15mm, colframe=quarto-callout-note-color-frame, leftrule=.75mm, breakable, rightrule=.15mm, toprule=.15mm, colback=white, opacityback=0]

This preprint is a non-peer-reviewed work from the
\href{https://realitybending.github.io/}{\textbf{Reality Bending Lab}}.
\begin{center}
\includegraphics[width=0.2\textwidth,height=\textheight]{manuscript_files/mediabag/ReBeL_LogoOnly_hu114.png}
\end{center}

\end{tcolorbox}  Author roles were classified using the Contributor Role
Taxonomy (CRediT; https://credit.niso.org/) as follows:  Dominique
Makowski:   Conceptualization, Data curation, Formal Analysis, Funding
acquisition, Investigation, Methodology, Project
administration, Resources, Software, Supervision, Validation, Visualization, Writing
-- original draft; Ana Neves:   Data curation, Writing -- original
draft, Writing -- review \& editing; Andy Field:   Writing -- original
draft, Writing -- review \& editing}
\par{Correspondence concerning this article should be addressed
to Dominique
Makowski, Email: \href{mailto:D.Makowski@sussex.ac.uk}{D.Makowski@sussex.ac.uk}}
}

\usepackage{pbalance} 
\usepackage{float}
\makeatletter
\let\oldtpt\ThreePartTable
\let\endoldtpt\endThreePartTable
\def\ThreePartTable{\@ifnextchar[\ThreePartTable@i \ThreePartTable@ii}
\def\ThreePartTable@i[#1]{\begin{figure}[!htbp]
\onecolumn
\begin{minipage}{0.5\textwidth}
\oldtpt[#1]
}
\def\ThreePartTable@ii{\begin{figure}[!htbp]
\onecolumn
\begin{minipage}{0.5\textwidth}
\oldtpt
}
\def\endThreePartTable{
\endoldtpt
\end{minipage}
\twocolumn
\end{figure}}
\makeatother

\makeatletter
\let\endoldlt\endlongtable		
\def\endlongtable{
\hline
\endoldlt}
\makeatother

\newenvironment{twocolumntable}% environment name
{% begin code
\begin{table*}[!htbp]%
\onecolumn%
}%
{%
\twocolumn%
\end{table*}%
}% end code

\urlstyle{same}

\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

% From https://tex.stackexchange.com/a/645996/211326
%%% apa7 doesn't want to add appendix section titles in the toc
%%% let's make it do it
\makeatletter
\xpatchcmd{\appendix}
  {\par}
  {\addcontentsline{toc}{section}{\@currentlabelname}\par}
  {}{}
\makeatother

%% Disable longtable counter
%% https://tex.stackexchange.com/a/248395/211326

\usepackage{etoolbox}

\makeatletter
\patchcmd{\LT@caption}
  {\bgroup}
  {\bgroup\global\LTpatch@captiontrue}
  {}{}
\patchcmd{\longtable}
  {\par}
  {\par\global\LTpatch@captionfalse}
  {}{}
\apptocmd{\endlongtable}
  {\ifLTpatch@caption\else\addtocounter{table}{-1}\fi}
  {}{}
\newif\ifLTpatch@caption
\makeatother

\begin{document}

\maketitle

\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\setlength\LTleft{0pt}

\section{Introduction}\label{introduction}

Despite significant advancements in psychological science following the
replication crisis (\citeproc{ref-open2015estimating}{Collaboration,
2015}), its progress is still hindered by its sub-optimal (or
inappropriate) usage of statistical tools
(\citeproc{ref-makowski2023we}{Makowski \& Waggoner, 2023}). A prevalent
issue is the continued reliance on linear models that assume normally
distributed (Gaussian) data - as this assumption often does not hold
true for many types of psychological outcomes. For instance, reaction
times typically exhibit skewed distributions, choices can be represented
as binary variables, and count data consists of strictly positive
integers. Applying models that presume normality and model the ``mean''
of the outcome variable can lead to misinterpretation and potentially
misleading conclusions when applied indiscriminately. It is thus
important that psychologists use models that can best describe the data
they collect, to fully exploit them and bring more nuance and accuracy
to their conclusions.

Among the most commonly collected data in psychology are responses on
subjective scales, such as Likert-type items or visual analog scales,
which exhibit some fundamental properties: these responses are bounded
(and can be rescaled to a 0-1 range) and frequently display clustering
at the extremes. Traditional linear models being ill-suited for such
data, researchers have turned to using Beta distributions to model this
data (instead of Gaussian), suited for continuous data within the (0,1)
interval (i.e., excluding extreme responses). To address the frequent
occurrence of exact zeros and ones (i.e., extreme values), zero-one
inflated beta (ZOIB) models have been developed
(\citeproc{ref-ospina2012general}{Ospina \& Ferrari, 2012}) to
accommodate the excess of boundary values by incorporating additional
components that model the probabilities of responses at 0 and 1 as a
separate, independent process.

\subsection{The Beta-Gate Model}\label{the-beta-gate-model}

The Beta-Gate model is a reparametrized Ordered Beta model
(\citeproc{ref-kubinec2023ordered}{Kubinec, 2023})\footnote{In the
  Ordered Beta model, the cutpoints on the log-scale are directly used
  as parameters, instead of being derived from \emph{pex} and
  \emph{bex}.} available in the \emph{cogmod} package in R
(https://github.com/DominiqueMakowski/cogmod), in which participants'
answers on bounded scales are conceptualized as latent responses that
can fall past a pair of probabilistic ``gates'' (or cutpoints) that
control whether the response is recorded as an extreme (0 or 1) or as a
nuanced, continuous value in between (Figure~\ref{fig-one}). These
distance of these gates from the edges of the scale varies based on two
interpretable parameters: \emph{pex} (the propensity by which people are
likely to answer extreme values), and \emph{bex}, a bias toward the
upper extreme (1) versus the lower (0). A person's internal response
that lies close to the edge might be ``caught'' by a gate and recorded
as an extreme, while others pass through to express a continuous
response (Beta-distributed with \(\mu\) (\emph{mu}) and \(\phi\)
(\emph{phi}) as its mean and precision parameters). The Beta-Gate model
is based on the idea that extreme values can emerge not just from a
fundamentally different underlying processes - as assumed in ZOIB models
- but from a common process governed by thresholds of decisiveness and
confidence.

\begin{figure}

\caption{\label{fig-one}The Beta-Gate Distribution is a reparametrized
ordered Beta model (Kubinec, 2023) that is governed by 4 parameters.
`Mu' and `phi' correspond to the mean and precision of the continuous
part of the distribution (between 0 and 1), and `pex' (propensity of
extremes) and `bex' (balance of extremes) indirectly control the
proportion of zeros and ones by specifying the location of the
``gates'', past which the latent response process is likely to generate
extreme values. Specifically, `pex' defines the total distance of both
gates from the extremes (in yellow), and `bex' determines the proportion
of the right gate distance relative to the left. In this example, the
total distance from the extremes is `pex' = 0.1, with 40\% (`bex' = 0.4)
of that distance being on the right (and 60\% on the left). The left
gate is thus located at 0.6, and the right at 1-0.04 = 0.96.}

\centering{

\includegraphics[width=0.5\textwidth,height=\textheight]{./figures/fig1.png}

}

\end{figure}%

Mathematically, the Beta-Gate distribution defines the observed outcome
\(x \in [0, 1]\) as a mixture of three components; a point mass at 0, a
point mass at 1, and a continuous Beta density over \(x \in (0,1)\),
scaled by the remaining probability mass. The probability of these
components are:

\begin{itemize}
\tightlist
\item
  \(P(x = 0) = \text{logistic}\left( \text{logit}(pex \cdot (1 - bex)) - \text{logit}(\mu) \right)\)
\item
  \(P(x = 1) = 1 - \text{logistic}\left( \text{logit}(1 - pex \cdot bex) - \text{logit}(\mu) \right)\)
\item
  \(P(x \in (0,1)) = 1 - P(x = 0) - P(x = 1)\)
\end{itemize}

The continuous part follows a Beta distribution with
parameters\footnote{Note that \emph{phi} is scaled in Beta-Gate models
  relative to the traditional mu/phi Beta particularization so that a
  phi of 1 corresponds to a uniform distribution - to facilitate setting
  priors on this parameter}:

\[
Beta(\alpha = \mu \cdot 2 \phi, \quad \beta = (1 - \mu) \cdot 2 \phi)
\]

\subsection{The Choice-Confidence (CHOCO)
Model}\label{the-choice-confidence-choco-model}

Decision-making is often conceptualized as involving distinct processes:
the choice itself, and the confidence associated with that choice. In
experimental paradigms, these can be somewhat disentangled by prompting
participants to make a discrete choice selection (e.g., ``True''
vs.~``False''), followed by a separate confidence rating. However, this
artificial separation makes its joint analysis difficulty, and may not
reflect real-world decision-making, where individuals often express both
choice and confidence simultaneously using a single, continuous scale.
In such scales, each side can represent a distinct latent category, and
the distance from the midpoint can indicates the level of confidence or
certainty. This integrated response format typically results in bimodal
distributions, with peaks corresponding to the mean confidence on either
side. Traditional beta regression models, which assume unimodal
distributions within the (0,1) interval, are ill-suited for such data.
One alternative is to transform the data into two variables a
posteriori: binarizing the side to represent choice and calculating the
absolute distance from the midpoint to represent confidence. These can
then be modeled separately, for instance, using logistic regression for
choice and beta regression for confidence (see
\citeproc{ref-makowski2025too}{Makowski et al., 2025} for an example).
While this approach can provide additional insights into underlying
mechanisms compared to a unique model, it assumes psychological and
statistical independence between choice and confidence, which may not
hold true in practice.

\begin{figure}

\caption{\label{fig-two}The CHOCO Model uses a mixture of Beta-Gate
distributions to model separately the right and left sides of the scale
(e.g., a rating of whether a statement was `Truth' vs.~`Lie'), as well
as their relative proportion. In this example, the participants are more
likely overall (p = 60\%) to select the right side of the scale (`Lie')
than the left (`Truth'). They are also more confident in their choice
(confright = 0.6 vs.~confleft = 0.3). Extreme values (zeros and ones)
are governed by the same mechanism as for Beta-Gate models.}

\centering{

\includegraphics[width=0.5\textwidth,height=\textheight]{./figures/fig2.png}

}

\end{figure}%

To model data of subjective scales in which the left and right sides can
be conceptualized as two different choices (e.g., True/False,
Agree/Disagree, etc.) and the magnitude of the response (how much the
cursor is set away from the mid-point) as the confidence, we introduce
the Choice-Confidence (CHOCO) model (Figure~\ref{fig-two}). It consists
of a three‐part mixture on \(x \in [0, 1]\):

\begin{itemize}
\tightlist
\item
  An (optional) point-mass at the midpoint \emph{mid} (typically 0.5) of
  weight \(p_{\text{mid}}\) for undecided or neutral responses.\\
\item
  A left-choice component governed by a Beta-Gate density on the
  rescaled variable \(x/mid\) with mean \emph{1 - confleft}, precision
  \emph{precleft}, and boundary-excess parameter \emph{pex(1 - bex)}.\\
\item
  A right-choice component governed by a Beta-Gate density on the
  rescaled variable \((x - mid)/(1 - mid)\) with mean \emph{confright},
  precision \emph{precright}, and boundary-excess parameter \emph{pex x
  bex}.
\end{itemize}

The overall probability of the right choice (relative to the left
choice) is controlled by a main parameter \emph{p}. The full CHOCO
density is:

\[
\small
\begin{cases}
pmid, & x = mid, \\[6pt]
(1 - pmid)(1 - p)\,\dfrac{1}{mid} \cdot \mathrm{BetaGate}\left(\dfrac{x}{mid}\right), & 0 < x < mid, \\[6pt]
(1 - pmid)\,p\,\dfrac{1}{1 - mid} \cdot \mathrm{BetaGate}\left(\dfrac{x - mid}{1 - mid}\right), & mid < x < 1. \\
\end{cases}
\]

By coupling choice probability \(p\), midpoint mass \(p_{\text{mid}}\),
and side‐specific Beta-Gate parameters (\(conf\), \(prec\), \(pex\),
\(bex\)), CHOCO flexibly captures both bimodality and confidence
intensity in a single unified model. Despite this theoretical appeal, it
is unclear whether this heavily parametrized model can be estimated
reliably from data, and whether it can provide more useful insights than
simpler alternatives.

\subsection{Aim of the Present Study}\label{aim-of-the-present-study}

\textbf{Study 1} aims to evaluate the CHOCO model's ability to better
capture subjective scale responses that (potentially) reflect an
underlying discrete choice, in comparison to existing models such as the
ZOIB and Beta-Gate. Specifically, we will assess whether 1) CHOCO
provides improved model fit, 2) yields deeper insights into
population-level effects than traditional approaches (gender differences
in reality beliefs), and 3) allows for the reliable estimation of
interpretable individual-level parameters through random effects.

To this end, we analyze data from two separate studies in which
participants judged whether a face image was AI-generated (``fake'') or
a real photograph. \textbf{Study 2} will apply this model to more subtle
effects, such as the effect perceived facial attractiveness and beauty
on reality judgments.

\section{Study 1}\label{study-1}

\textbf{TODO}: Write some introduction about reality beliefs in the
context of ambiguity. Underline that what drives judgments and beliefs
of reality remains unclear. And this is important in today's post-truth
era.

Another unknown is whether these beliefs are fully stimuli-driven or
whether there are some stable individual tendencies that are at stake.

\subsection{Methods}\label{methods}

\subsubsection{Participants}\label{participants}

The sample is the same as in (\citeproc{ref-makowski2025too}{Makowski et
al., 2025}). We included all heterosexual male and female participants,
for a final sample of 125 participants (Mean age = 28.6, SD = 9.2,
range: {[}19, 66{]}; Sex: 44.8\% females). We included only sexually
relevant stimuli for men and women (i.e., all 89 female faces for men
and 20 male faces for women).

\subsubsection{Procedure}\label{procedure}

In the first phase, participants viewed 109 neutral-expression
photographs of faces (random order, display time of 500 ms) from the
American Multiracial Face Database
(\citeproc{ref-chen2021broadening}{Chen et al., 2021}). After each
image, participants rated the face on trustworthiness, familiarity,
attractiveness, and beauty using visual analog scales. In the second
phase, participants were informed that ``about half of the previously
seen images were AI-generated''. The same faces were presented again in
a new random order (same display time), followed by ratings of
``reality'' (whether they believed the image was AI-generated - left
anchor - or a real photograph - right anchor). Makowski et al.
(\citeproc{ref-makowski2025too}{2025}) showed that most stimuli yielded
highly variable reality ratings, suggesting that idiosyncratic cognitive
processes were at play.

\subsubsection{Data Analysis}\label{data-analysis}

We fitted 3 models to predict the reality ratings: a ZOIB model, a
Beta-Gate model, and the CHOCO model. For all models and each parameter,
the full formula was entered:
\(Real\sim Sex + (1 | Participant) + (1 | Item)\) (with Sex as the main
predictor and participants and items entered as random intercepts). The
models were run using \emph{brms}
(\citeproc{ref-burkner2017brms}{Bürkner, 2017}) R package, and analyzed
using the \emph{easystats} collection of packages {[}\textbf{REFS}{]}.
To maximize the comparability across models We used the default priors
(uniform) for all models, and we ran 16 chains of 1400 iterations each
on the University of Sussex High-Performance Computing (HPC) cluster.

Model comparisons were performed using the \emph{loo} R package
(\citeproc{ref-vehtari2017practical}{Vehtari et al., 2017}), which
computes the Widely Applicable Information Criterion (WAIC) and
estimates the Expected Log Predictive Density (ELPD) and penalizes the
number of parameters. We assessed model performance by examining ELPD
differences and their standard errors (SE), reporting corresponding
\emph{p}-values to determine significant differences in predictive
accuracy.

For the population-level effects, we will consider significant and
report (using the median of the posterior distribution) effects for
which the 95\% Credible Interval (CI) does not include zero {[}and when
the probability of direction \emph{pd} is \textgreater{}
\textasciitilde97\%; Makowski et al.
(\citeproc{ref-makowski2019indices}{2019}){]}. For the individual-level
parameters (i.e., the random intercepts of each parameter for each
participant and each item), we will first analyze their reliability
using the Variance-Over-Uncertainty Ratio index (\emph{D-vour}). This
index, implemented in the \emph{performance} package
(\citeproc{ref-ludecke2021performance}{Lüdecke et al., 2021}), is
inspired by recent work on mixed models reliability
(\citeproc{ref-rouder2024hierarchical}{Rouder \& Mehrvarz, 2024};
\citeproc{ref-williams2021beneath}{Williams et al., 2021}), and
corresponds to the normalized ratio of observed variability to
uncertainty in random effect estimates, defined as:

\[
D_{\text{vour}} = \frac{\sigma_B^2}{\sigma_B^2 + \mu_{\text{SE}}^2}
\]

Where \(\sigma_B^2\) is the between-group variability (computed as the
SD of the random effect point-estimates) and \(\mu_{\text{SE}}^2\) is
the mean squared uncertainty in random effect estimates (i.e., the
average uncertainty). We use as \emph{D-vour} = 0.666 as the threshold
for moderately reliable random effect estimates, which corresponds to a
2:1 ratio of between-group variance to uncertainty.

Finally, we will run a correlation analysis of the models'
individual-level estimates against ``empirical'' (indices computed
directly on the observed data), including the empirical \emph{p}
(\(P(y > 0.5)\)), the overall \emph{conf} (\(mean(|y - 0.5|)\)),
\emph{pex} (\(P(y \in [0, 1])\)) and \emph{bex}
(\(P(y == 1) / P(y \in [0, 1])\)), assessing whether the model's
estimate are in-line with easily interpretable indices.

\subsection{Results}\label{results}

The reproducible code and full result report are available at
https://github.com/RealityBending/FictionChoco.

\subsubsection{Model Comparison}\label{model-comparison}

\begin{figure*}[!htbp]

{\caption{{TODO.}{\label{fig-three}}}}

\includegraphics[width=1\textwidth,height=\textheight]{./figures/fig3.png}

\end{figure*}

The models did converge without divergent transitions, and the effective
sample size was sufficient for all parameters (all
\(n_{\text{eff}} > 1000\)). The difference in predictive accuracy, as
indexed by Expected Log Predictive Density (ELPD-WAIC), suggests that
the \emph{CHOCO} is the best model (\(ELPD = -138.45\)), followed by
\emph{Beta-Gate} (\(\Delta_{ELPD} = -1875.31 \pm 61.18, p < .001)\) and
\emph{ZOIB} (\(\Delta_{ELPD} = -1916.11 \pm 61.75, p < .001\)). See
Figure~\ref{fig-three} for the posterior predictive checks, showing that
only the CHOCO model managed to capture the bimodal distribution of
data.

\subsubsection{Effect of Sex}\label{effect-of-sex}

The ZOIB model suggested that women had higher mean (\emph{mu}
parameter) scores of reality beliefs
(\(\Delta_{\text{Female}} = 0.16,~95\%~ CI~[0.01, 0.32], pd = 98.10\%\)),
less extreme values (\emph{zoi} parameter; STATS) but more ones relative
to zeros (\emph{coi} parameter; STATS).

The Beta-Gate model similarly suggested that women had higher mean
(\emph{mu} parameter) scores of reality beliefs (STATS), less extreme
values (\emph{pex} parameter; STATS), and a greater tendency to answer
one relative to zero (\emph{bex} parameter; STATS).

The CHOCO model shows that women had a higher probability \emph{p} of
judging faces as real (STATS), but are not more confident when doing so
(\emph{confright} parameter; STATS). However, they were less confident
when answering that an image was AI-generated (\emph{confleft}
parameter; STATS). There were also less likely to produce extreme
answers (\emph{pex} parameter; STATS), but no strong evidence supporting
a directional bias was observed (\emph{bex} parameter; STATS).

Across all these models, no effect of Sex on the prediction parameter
was observed.

\subsubsection{Individual-Level
Parameters}\label{individual-level-parameters}

\begin{figure*}[!htbp]

{\caption{{TODO.}{\label{fig-four}}}}

\includegraphics[width=1\textwidth,height=\textheight]{./figures/fig4.png}

\end{figure*}

The ZOIB model estimated reliable variability in the participant's
\emph{phi} parameter (D-vour = 0.88) and \emph{zoi} parameter (D-vour =
0.84), as well as in the \emph{mu} parameter related to individual items
(D-vour = 0.82). Moderate reliability was also observed for items in the
\emph{coi} parameter (D-vour = 0.71) and for participants in the
\emph{mu} parameter (D-vour = 0.69). The Beta-Gate model yielded similar
results: a high reliability of participant's \emph{phi} parameter
(D-vour = 0.88), \emph{pex} parameter (D-vour = 0.84). The \emph{mu}
parameter's variability was reliably captured for items (D-vour = 0.85)
and moderately for participants (D-vour = 0.72). Of note that the
parameters of Beta-Gate and ZOIB correlate almost perfectly, underlining
their empirical similarity despite a different parametrization.

The CHOCO model yielded reliable estimates for all parameters except
\emph{bex} for participants (\emph{confright} D-vour = 0.94,
\emph{confleft} D-vour = 0.91, \emph{pex} D-vour = 0.80, \emph{p} D-vour
= 0.73, \emph{precright} D-vour = 0.79, \emph{precleft} D-vour = 0.73).
Item's variability was primarily reflected through the \emph{p}
parameter (D-vour = 0.86).

Finally, the empirical average correlated the most strongly with CHOCO's
\emph{p} (r = .94), rather than ZOIB's \emph{mu} (r = .74) or
Beta-Gate's \emph{mu} (r = .82). The empirical overall confidence was
the strongest correlated with CHOCO's \emph{confright} (r = .91),
followed by ZOIB's \emph{phi} (r = -.86), Beta-Gate's phi (r = -.83),
and other CHOCO's parameters. The empirical proportion of ``right''
answer \emph{p} correlated the strongest with CHOCO's \emph{p} (r =
0.94), followed by Beta-Gate's \emph{mu} (r = .82) and ZOIB's \emph{mu}
(r = .77). The empirical \emph{pex} correlated the strongest with ZOIB's
\emph{zoi} (r = .90), and Beta-Gate's \emph{pex} (r = .90), and CHOCO's
\emph{pex} (r = .88).

\subsection{Discussion}\label{discussion}

Study 1 revealed that the Choice-Confidence (CHOCO) model was a much
better fit for bimodal bounded data, compared to other alternatives like
the Zero-and-One Inflated Beta (ZOIB) and Beta-Gate (Ordered Beta)
models. It also allowed for a deeper understanding through its
interpretable parameters, offering insights into possibly distinct
cognitive mechanisms, such as the probability of answering real
vs.~fake, and the associated confidence in these two choices. This was
illustrated by modelling the effect of sex on all the CHOCO parameters.

Note that the observed gender differences are primarily presented as a
proof-of-principle, to showcase the model's ability to capture
group-level effects and to provide deeper insights compared to other
models. However, given that they were based on different items (female
and male faces), these differences might just be a reflection of stimuli
characteristics rather than true sex dymorphism in the formation of
reality beliefs.

Finally, we also show that the CHOCO model was able to capture reliable
and interpretable individual-level parameters, supporting its value to
measure inter-individual differences. An interesting dissociation
emerged between participant- and item-level variability: the latter
seemed mostly to be represented in the \emph{p} parameter, while
participants reliably varied in most of the components (aside from
\emph{bex}). This could be interpreted as that external item
characteristics primarily influence the probability of being judged as
real vs.~fake, while the expressed confidence is first and foremost an
individual characteristic.

\section{Study 2}\label{study-2}

The explosion of accessibility of state-of-the-art AI tools has made it
effortless to generate realistic images, including human faces that are
often indistinguishable from real ones \textbf{{[}REF{]}}. These
synthetic visuals are flooding the cyberspace across various domains,
such as art, advertising and entertainment, to education and
information. This technological advancement carries an important
potential for misuse, such as in disinformation campaigns, scams (e.g.,
AI-bots, identity theft), and abuse. The democratization of such
technology raises pressing concerns about the value of authenticity and
the potential erosion of media trust in our increasingly
\emph{post-truth} society.

In this evolving landscape, understanding the cognitive mechanisms that
underpin our judgments of reality becomes paramount. Despite the
increasing prevalence of digitally altered or AI-generated content,
humans still rely on certain heuristics to assess authenticity. One such
heuristic might be facial attractiveness. Attractiveness appraisals are
known to be automatic and unconscious (\citeproc{ref-hou2023neural}{Hou
et al., 2023}; \citeproc{ref-hung2016unconscious}{Hung et al., 2016};
\citeproc{ref-luo2019robust}{Luo et al., 2019}), and carry strong
real-life consequences, as demonstrated by the large body of literature
on the ``beauty premium'' (\citeproc{ref-gulati2024beautiful}{Gulati et
al., 2024}, \textbf{MORE}; \citeproc{ref-little2021facial}{Little,
2021})\footnote{As an eloquent example, Monk Jr et al.
  (\citeproc{ref-monk2021beholding}{2021}) reported that the the
  magnitude of the earnings disparities among white women along the
  perceived attractiveness continuum exceeds in magnitude the canonical
  black-white race gap in a large representative US sample.}

\textbf{TODO:} However, it is important to allow for the possibility of
a U-shaped relationship. For instance, Peng et al.~(2020) reported that
both attractive and unattractive people sold more (underlining a
``plainness penalty'').

Although its role as a potential modulator of reality beliefs remains
underexplored, Makowski et al\ldots{} \textbf{TODO: detail findings of
that study + other studies}.

In this study, we will apply the CHOCO model to the data of MAKOWSKI,
and see whether different analysis approaches lead to different
conclusions. We will then try to replicate these findings in a new
sample of participants.

\subsection{Methods}\label{methods-1}

\subsubsection{Participants}\label{participants-1}

The first sample includes the same participants as study 1 (see above).
\textbf{TODO} for sample 2.

\subsubsection{Procedure}\label{procedure-1}

For the second sample, the procedure was relatively similar with a few
key differences. \textbf{TODO} Describe fake Face 2. The main difference
is that the subjective ratings were done on a X-point Likert scale,
rather than analog, with a mid-point value.

\subsubsection{Data Analysis}\label{data-analysis-1}

To compare the benefits of CHOCO models to a ``traditional'' analytic
approach, we started by fitting a frequentist linear mixed model to
predict reality beliefs with the formula
\(Real\sim Sex / poly(Attractive, 2) + (poly(Attractive, 2)|Participant) + (1|Item)\).
The 2nd degree raw polynomial term was included to allow for potentially
non-linear relationships (note that the first and second degree effects
of \emph{raw} polynomials can be interpreted independently as the linear
part and the ``curvy'' part of the relationship). For the CHOCO model,
mildly informative and effect-agnostic (i.e., centered at zero) priors
were used. The same formula was used for all parameters, except that
(based on the reliability analysis of Study 1), items were only included
as random effects for \emph{p}, and participants were not included for
\emph{bex}.

\textbf{TODO}. For sample 2, the analysis was based on that of Sample 1.
The main differences are 1) the inclusion of the ``Condition'' (whether
the picture was presented as ``Real'' of ``Fake'' in the first phase of
the experiment) as an additional predictor (it was entered as the only
random slope for all participant random effects); 2) the inclusion of an
additional parameters, \emph{pmid}, modelling the probability of
answering the middle-point of the Likert scale (representing an
``undecided'' option).

The same analysis was applied for Beauty.

\subsection{Results}\label{results-1}

As the complete model parameters tables are available at
https://github.com/RealityBending/FictionChoco, we will focus on
reporting noteworthy findings below.

In sample 1, the traditional approach suggested a significant linear
relationship between the mean level of ``Reality'' and attractiveness
(\(\beta_{poly1}=3.63, ~95\%~CI [2.71, 4.56], pd = ...\%\)) for males
only. The CHOCO model revealed that attractiveness had a significantly
positive linear relationship with the probability \emph{p} of judging
faces as real (STATS), but a quadratic relationship with the confidence
in real judgments (\emph{confright}; STATS). Attractiveness was also
associated with less confidence in fake judgments (\emph{confleft};
STATS), and related linearly with a stronger bias towards extreme Real
responses (\emph{bex}; STATS). These effects were present for males
only, and no relationship was found for females. The reliability of the
effect of attractiveness (the random slopes) was very low for all
parameters (D-vour \textless{} 0.1).

The effect of Beauty was the same as above, with some additional
evidence in favour of an effect in women, in particular for a linear
relationship with \emph{p} for females as well
(\(\beta_{poly1}=7.02, ~95\%~CI [-0.90, 14.97], pd = 95.91\%\)), as well
as a quadradic relationship with the confidence in fake judgments
(\emph{confleft}; STATS).

In Sample 2, the traditional approach suggested a significant linear
relationship between the mean level of ``Reality'' and attractiveness
for males (STATS) and women (STATS), with no effect of the Condition.
The CHOCO model revealed that attractiveness had a significantly
positive linear relationship with the probability \emph{p} of judging
faces as real for males (STATS) and females (STATS). It also had a
linear relationship with the confidence in real judgments
(\emph{confright}) for males (STATS) and females (STATS), as well as a
significant quadratic curvature for males only (STATS). Attractiveness
also linearly decreased the confidence in fake judgments only for males
(\emph{confleft}; STATS), but increased the probability of an undecided
response (\emph{pmid}) in women (STATS), and displayed in females only a
quadratic relationship with the amount of extreme answers (\emph{pex};
STATS) and its bias in favour of right-side extreme response
(\emph{bex}; STATS). No consistent and reliable effect of condition was
found.

The effect of Beauty was\ldots{}

\subsection{Discussion}\label{discussion-1}

The fact that the effect was less obvious in women in the sample 1 might
be explained with power (less relevant items were included for women).

\section{General Discussion}\label{general-discussion}

Little difference between Beauty and Attractiveness, aside that Beauty
revealed more effects in women. Why?

Likert-scales (with a limited number of discrete options) are commonly
used in psychology, and often modeled as continuous. While this fact has
spurred its own polarized debate (REFS), our position is that the model
used should ideally reflect the data \emph{generating} mechanism rather
than necessarily focus on the \emph{observed} data\footnote{Ideally, the
  data recording method should be aligned with the assumed generating
  mechanism, which is an experiment design issue rather than a
  statistical one.} - although the former cannot always be easily
inferred or known. For the data of Sample 2, we held the assumption that
the underlying latent distribution was CHOCO distributed (as there is no
reason to assume it would be different from Sample 1), treated as
continuous CHOCO-distributed data, despite the fact that it was
collected on a 7-point Likert scale. The motivation was to see if the
model could extrapolate and reliably estimate all the parameters based
on a limited number of data points but means in our case that,
effectively, each Beta distribution was extrapolated from the frequency
of only two unique values (or 3 for the no-extreme version). While this
is far from ideal, the CHOCO model still provided a better fit to the
data than alternative approaches. \textbf{TODO} However, it might have
lead to a lesser ability to detect finer grain changes, including
U-shaped relationships. Having 7-points (or 6 with no mid-value) is the
strict minimum to use the CHOCO model.

Future studies should assess the ability of CHOCO models to also fit
unimodal distributions (so it can be used more generally), as well as
the benefits of more parsimonious parametrization (for instance, a
unique precision parameter controlling both the left and right sides, or
a linked confidence parameter e.g.~where \emph{confleft} is expressed as
a function of \emph{confright}).

\section{Data Availability}\label{data-availability}

Data, code and everything is available at
https://github.com/RealityBending/FictionChoco. The CHOCO model is
implemented in the \emph{cogmod} R package
(https://github.com/DominiqueMakowski/cogmod).

\section{Acknowledgements}\label{acknowledgements}

We would like to thank the dissertation students from the University of
Sussex for their help in data collection.

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-burkner2017brms}
Bürkner, P.-C. (2017). Brms: An r package for bayesian multilevel models
using stan. \emph{Journal of Statistical Software}, \emph{80}, 1--28.

\bibitem[\citeproctext]{ref-chen2021broadening}
Chen, J. M., Norman, J. B., \& Nam, Y. (2021). Broadening the stimulus
set: Introducing the american multiracial faces database. \emph{Behavior
Research Methods}, \emph{53}, 371--389.

\bibitem[\citeproctext]{ref-open2015estimating}
Collaboration, O. S. (2015). Estimating the reproducibility of
psychological science. \emph{Science}, \emph{349}(6251), aac4716.

\bibitem[\citeproctext]{ref-gulati2024beautiful}
Gulati, A., Martı́nez-Garcia, M., Fernández, D., Lozano, M. A., Lepri,
B., \& Oliver, N. (2024). What is beautiful is still good: The
attractiveness halo effect in the era of beauty filters. \emph{Royal
Society Open Science}, \emph{11}(11), 240882.

\bibitem[\citeproctext]{ref-hou2023neural}
Hou, X., Shang, J., \& Tong, S. (2023). Neural mechanisms of the
conscious and subliminal processing of facial attractiveness.
\emph{Brain Sciences}, \emph{13}(6), 855.

\bibitem[\citeproctext]{ref-hung2016unconscious}
Hung, S.-M., Nieh, C.-H., \& Hsieh, P.-J. (2016). Unconscious processing
of facial attractiveness: Invisible attractive faces orient visual
attention. \emph{Scientific Reports}, \emph{6}(1), 37117.

\bibitem[\citeproctext]{ref-kubinec2023ordered}
Kubinec, R. (2023). Ordered beta regression: A parsimonious,
well-fitting model for continuous data with lower and upper bounds.
\emph{Political Analysis}, \emph{31}(4), 519--536.

\bibitem[\citeproctext]{ref-little2021facial}
Little, A. C. (2021). Facial attractiveness. \emph{Encyclopedia of
Evolutionary Psychological Science}, 2887--2891.

\bibitem[\citeproctext]{ref-ludecke2021performance}
Lüdecke, D., Ben-Shachar, M. S., Patil, I., Waggoner, P., \& Makowski,
D. (2021). Performance: An r package for assessment, comparison and
testing of statistical models. \emph{Journal of Open Source Software},
\emph{6}(60).

\bibitem[\citeproctext]{ref-luo2019robust}
Luo, Q., Rossion, B., \& Dzhelyova, M. (2019). A robust implicit measure
of facial attractiveness discrimination. \emph{Social Cognitive and
Affective Neuroscience}, \emph{14}(7), 737--746.

\bibitem[\citeproctext]{ref-makowski2019indices}
Makowski, D., Ben-Shachar, M. S., Chen, S. A., \& Lüdecke, D. (2019).
Indices of effect existence and significance in the bayesian framework.
\emph{Frontiers in Psychology}, \emph{10}, 2767.

\bibitem[\citeproctext]{ref-makowski2025too}
Makowski, D., Te, A. S., Neves, A., Kirk, S., Liang, N. Z., Mavros, P.,
\& Chen, S. A. (2025). Too beautiful to be fake: Attractive faces are
less likely to be judged as artificially generated. \emph{Acta
Psychologica}, \emph{252}, 104670.

\bibitem[\citeproctext]{ref-makowski2023we}
Makowski, D., \& Waggoner, P. D. (2023). Where are we going with
statistical computing? From mathematical statistics to collaborative
data science. In \emph{Mathematics} (8; Vol. 11, p. 1821). MDPI.

\bibitem[\citeproctext]{ref-monk2021beholding}
Monk Jr, E. P., Esposito, M. H., \& Lee, H. (2021). Beholding
inequality: Race, gender, and returns to physical attractiveness in the
united states. \emph{American Journal of Sociology}, \emph{127}(1),
194--241.

\bibitem[\citeproctext]{ref-ospina2012general}
Ospina, R., \& Ferrari, S. L. (2012). A general class of zero-or-one
inflated beta regression models. \emph{Computational Statistics \& Data
Analysis}, \emph{56}(6), 1609--1623.

\bibitem[\citeproctext]{ref-rouder2024hierarchical}
Rouder, J. N., \& Mehrvarz, M. (2024). Hierarchical-model insights for
planning and interpreting individual-difference studies of cognitive
abilities. \emph{Current Directions in Psychological Science},
\emph{33}(2), 128--135.

\bibitem[\citeproctext]{ref-vehtari2017practical}
Vehtari, A., Gelman, A., \& Gabry, J. (2017). Practical bayesian model
evaluation using leave-one-out cross-validation and WAIC.
\emph{Statistics and Computing}, \emph{27}, 1413--1432.

\bibitem[\citeproctext]{ref-williams2021beneath}
Williams, D. R., Mulder, J., Rouder, J. N., \& Rast, P. (2021). Beneath
the surface: Unearthing within-person variability and mean relations
with bayesian mixed models. \emph{Psychological Methods}, \emph{26}(1),
74.

\end{CSLReferences}

\end{document}

### 2025-04-27 analysis/combine_models.R

@@ -1,23 +1,0 @@
-library(brms)
-library(cmdstanr)
-library(rstan)
-
-combine_and_save <- function(name = "sample1_zoib") {
-  files <- list.files('./models/', pattern = paste0(name, '_.*rds$'), full.names = TRUE)
-  m <- brms::combine_models(mlist = lapply(files, readRDS))
-
-  # m <- brms::add_criterion(m, criterion="waic", model_name = name)
-  # loo1 <- loo::loo_subsample(m, observations = 10, loo_approximation_draws = 100, cores = 4)
-  # fit_laplace <- rstan::optimizing(m, data = insight::get_data(m), draws = 1000, importance_resampling = TRUE)
-  # parameter_draws_laplace <- fit_laplace$theta_tilde # draws from approximate posterior
-  # log_p <- fit_laplace$log_p # log density of the posterior
-  # log_g <- fit_laplace$log_g # log density of the approximation
-
-
-  saveRDS(m, paste0("models/", name, ".rds"))
-}
-
-combine_and_save("sample1_zoib")
-combine_and_save("sample1_bext")
-combine_and_save("sample1_choco")
-combine_and_save("sample1_chocoattractiveness")

### 2025-04-27 analysis/make_models.R

@@ -1,278 +1,0 @@
-# Fit model with 16 chains per task
-
-library(brms)
-library(cmdstanr)
-library(cogmod)
-
-options(brms.backend = "cmdstanr")
-
-# Get array task ID from Slurm
-task_id <- as.numeric(commandArgs(trailingOnly = TRUE)[1])
-
-# Number of chains per task from Slurm (fixed at 16)
-chains_per_task <- as.numeric(Sys.getenv("SLURM_CPUS_PER_TASK"))
-
-# Calculate starting chain ID
-start_chain <- (task_id - 1) * chains_per_task + 1
-
-iter <- 1000 # 50% of that is warmup
-
-# Test --------------------------------------------------------------------
-
-# model <- brm(
-#   formula = mpg ~ wt,
-#   data = mtcars,
-#   family = gaussian(),
-#   chains = chains_per_task,    # 16 chains per task
-#   cores = chains_per_task,     # Use all 16 CPUs
-#   iter = 500,
-#   seed = 1234 + start_chain,   # Unique seed per task
-#   file = paste0("./modeltoy_task_", task_id, ".rds")
-# )
-
-
-
-# Sample 1 ----------------------------------------------------------------
-
-df <- read.csv("https://raw.githubusercontent.com/RealityBending/FictionChoco/refs/heads/main/data/sample1.csv")
-df$Sex <- factor(df$Sex, levels = c("Male", "Female"))
-# df <- df[df$Participant %in% unique(df$Participant)[1:10],]
-
-# ZOIB
-print("===== ZOIB =====")
-print(Sys.time())
-
-f <- bf(
-  Real ~ 0 + Intercept + Sex + (0 + Intercept | Participant) + (0 + Intercept | Item),
-  phi ~ 0 + Intercept + Sex + (0 + Intercept | Participant) + (0 + Intercept | Item),
-  zoi ~ 0 + Intercept + Sex + (0 + Intercept | Participant) + (0 + Intercept | Item),
-  coi ~ 0 + Intercept + Sex + (0 + Intercept | Participant) + (0 + Intercept | Item)
-)
-
-model <- brm(
-  formula = f,
-  data = df,
-  family = zero_one_inflated_beta(),
-  init = 0,
-  chains = chains_per_task,    # 16 chains per task
-  cores = chains_per_task,     # Use all 16 CPUs
[…]

### 2025-04-27 analysis/server/combine_models.R

@@ -1,0 +1,31 @@
+library(brms)
+library(cmdstanr)
+library(rstan)
+library(loo)
+
+
+# Get the number of cores and task ID from the environment variables
+options(mc.cores = as.numeric(Sys.getenv("SLURM_CPUS_PER_TASK")))
+task_id <- as.numeric(Sys.getenv("SLURM_ARRAY_TASK_ID"))
+
+# List of models
+model_names <- c("sample1_zoib", "sample1_betagate", "sample1_choco",
+                 "sample1_chocoattractiveness", "sample1_chocobeauty",
+                 "sample2_chocoattractiveness")[6]
+
+# Select the model name based on the task ID
+model_name <- model_names[task_id]
+
+combine_and_save <- function(name) {
+  print(paste0(name, ": ", Sys.time()))
+  files <- list.files(".", pattern = paste0(name, '_.*rds$'), full.names = TRUE)
+  m <- brms::combine_models(mlist = lapply(files, readRDS))
+
+  # Add criterion and save
+  m <- brms::add_criterion(m, "waic", ndraws = 1500, file = name)
+
+  # saveRDS(m, paste0(name, ".rds"))
+}
+
+combine_and_save(model_name)
+print(paste0("Completed: ", Sys.time()))

### 2025-04-27 analysis/server/make_models.R

@@ -1,0 +1,366 @@
+# Fit model with 16 chains per task
+
+library(brms)
+library(cmdstanr)
+library(cogmod)
+
+options(brms.backend = "cmdstanr")
+
+# Get array task ID from Slurm
+task_id <- as.numeric(commandArgs(trailingOnly = TRUE)[1])
+
+# Number of chains per task from Slurm (fixed at 16)
+chains_per_task <- as.numeric(Sys.getenv("SLURM_CPUS_PER_TASK"))
+
+# Calculate starting chain ID
+start_chain <- (task_id - 1) * chains_per_task + 1
+
+iter <- 1500 # 50% of that is warmup
+
+# Test --------------------------------------------------------------------
+
+# model <- brm(
+#   formula = mpg ~ wt,
+#   data = mtcars,
+#   family = gaussian(),
+#   chains = chains_per_task,    # 16 chains per task
+#   cores = chains_per_task,     # Use all 16 CPUs
+#   iter = 500,
+#   seed = 1234 + start_chain,   # Unique seed per task
+#   file = paste0("./modeltoy_task_", task_id, ".rds")
+# )
+
+
+
+# Sample 1 ----------------------------------------------------------------
+
+df <- read.csv("https://raw.githubusercontent.com/RealityBending/FictionChoco/refs/heads/main/data/sample1.csv")
+df$Sex <- factor(df$Sex, levels = c("Male", "Female"))
+# df <- df[df$Participant %in% unique(df$Participant)[1:10],]
+
+
+
+
+# ZOIB
+print("===== ZOIB =====")
+print(Sys.time())
+
+f <- bf(
+  Real ~ 0 + Intercept + Sex + (1 | Participant) + (1 | Item),
+  phi ~ 0 + Intercept + Sex + (1 | Participant) + (1 | Item),
+  zoi ~ 0 + Intercept + Sex + (1 | Participant) + (1 | Item),
+  coi ~ 0 + Intercept + Sex + (1 | Participant) + (1 | Item)
+)
+
+model <- brm(
+  formula = f,
+  data = df,
+  family = zero_one_inflated_beta(),
+  init = 0,
+  chains = chains_per_task,    # 16 chains per task
+  cores = chains_per_task,     # Use all 16 CPUs
+  iter = iter,  # Number of iterations and warmup must be equal (warmup is 50% iter)
+  seed = 1234 + start_chain,   # Unique seed per task
[…]

### 2025-05-06 analysis/1_study1.qmd
update what I have
@AnafNeves I'm pushing what I have if you want to take a look at the ms draft

@@
 library(rstan)
-
-set.seed(123)
 ```
@@
 
+set.seed(123)
+
 dfraw <- read.csv("https://raw.githubusercontent.com/RealityBending/FakeFace/refs/heads/main/data/data.csv") |> 
@@
          Item = gsub(".jpg", "", Stimulus)) |> 
-  filter(Sex %in% c("Male", "Female"), Orientation %in% c("Heterosexual", "Bisexual", "Homosexual")) |> 
+  filter(Sex %in% c("Male", "Female"), Orientation %in% c("Heterosexual", "Bisexual")) |> 
   mutate(Relevance = case_when(
-    Sex == "Male" & Orientation == "Heterosexual" & grepl("NF", Stimulus) ~ "Relevant",
-    Sex == "Female" & Orientation == "Heterosexual" & grepl("NM", Stimulus) ~ "Relevant",
-    Sex == "Male" & Orientation == "Bisexual" & grepl("NF", Stimulus) ~ "Relevant",
-    Sex == "Female" & Orientation == "Bisexual" & grepl("NM", Stimulus) ~ "Relevant",
-    Sex == "Male" & Orientation == "Homosexual" & grepl("NM", Stimulus) ~ "Relevant",
-    Sex == "Female" & Orientation == "Homosexual" & grepl("NF", Stimulus) ~ "Relevant",
+    Sex == "Male" & grepl("NF", Stimulus) ~ "Relevant",
+    Sex == "Female" & grepl("NM", Stimulus) ~ "Relevant",
     .default = "Irrelevant"
@@
   mutate(SexOrientation = paste(Sex, "-", Orientation)) |> 
-  filter(Relevance == "Relevant", Orientation == "Heterosexual")
+  filter(Relevance == "Relevant")
 

### 2025-05-06 analysis/2_study2.qmd
update what I have
@AnafNeves I'm pushing what I have if you want to take a look at the ms draft

@@
   precleft <- modelbased::estimate_prediction(m, data = newdata, predict = "precleft", iterations = 50)
+  if ("pex" %in% names(insight::find_parameters(m))) {
     pex <- modelbased::estimate_prediction(m, data = newdata, predict = "pex", iterations = 50)
+  } else {
+    pex <- list(Predicted = rep(0, nrow(newdata)))
+  }
+  if ("bex" %in% names(insight::find_parameters(m))) {
     bex <- modelbased::estimate_prediction(m, data = newdata, predict = "bex", iterations = 50)
+  } else {
+    bex <- list(Predicted = rep(0.5, nrow(newdata)))
+  }
   if ("pmid" %in% names(insight::find_parameters(m))) {
@@
 
-df <- read.csv("https://raw.githubusercontent.com/RealityBending/FakeFace2/refs/heads/main/data/data_task.csv") |> 
+set.seed(123)
+
+dfraw <- read.csv("https://raw.githubusercontent.com/RealityBending/FakeFace2/refs/heads/main/data/data_task.csv") |> 
+  full_join(read.csv("https://raw.githubusercontent.com/RealityBending/FakeFace2/refs/heads/main/data/data_participants.csv") |> 
+              select(Participant, Age), by = "Participant") |> 
+  rename(Sex = Gender, Orientation = SexualOrientation, Item=Stimulus) |> 
+  filter(Sex %in% c("Male", "Female"), Orientation %in% c("Heterosexual", "Bisexual")) |> 
+  mutate(Relevance = case_when(
+    Sex == "Male" & grepl("NF", Item) ~ "Relevant",
+    Sex == "Female" & grepl("NM", Item) ~ "Relevant",
+    .default = "Irrelevant"
+  ))  |> 
   filter(Relevance == "Relevant") |> 
@@
          Beauty = Beauty / 6,
-         RealNoex = datawizard::change_scale(Real, to = c(0.125, 0.875))) |> 
-  select(Participant, Sex=Gender, Condition, Item=Stimulus, Real, RealNoex, Attractive, Beauty)
+         RealNoex = datawizard::change_scale(Real, to = c(0.125, 0.875)),
+         SexOrientation = paste(Sex, "-", Orientation))
 
-write.csv(df, "../data/sample2.csv", row.names = FALSE)
+dfraw |> 
+  select(Participant, Sex, Condition, Item, Real, RealNoex, Attractive, Beauty) |> 
+  write.csv("../data/sample2.csv", row.names = FALSE)
+```
+
+```{r}
+#| eval: false
+
+# For demographic section
+dfraw |> 
+  mutate(n_stims = n(), .by = "Participant") |> 
+  select(Participant, Age, Sex, n_stims) |> 
+  slice(1, .by = "Participant") |> 
+  report::report_participants()
+
+report::report(dfraw$Sex)
 ```
@@
 ```{r}
-m <- lme4::lmer(Real ~ Sex / poly(Attractive, 2) * Condition + (1|Participant) + (1|Item), data = df)
+ggplot(df, aes(x=Real)) +
+  geom_bar() +
+  facet_wrap(~Sex)
+```
+
+```{r}
+m <- lme4::lmer(RealNoex ~ Sex / poly(Attractive, 2) * Condition + (1|Participant) + (1|Item), data = df)
 display(parameters::parameters(m, effects = "fixed"))
@@
   geom_line(aes(color = Condition), linewidth = 1) +
-  geom_smooth(data=df, aes(color = Condition, y = Real), method = "loess", se = FALSE, linetype = "dotted", formula = 'y ~ x') +
+  geom_smooth(data=dfraw, aes(color = Condition, y = Real), method = "loess", se = FALSE, linetype = "dotted", formula = 'y ~ x') +
   scale_color_manual(values = c("AI-Generated"="#3F51B5", "Photograph"="#FF5722")) +
@@
 
-#### Posterior Predictive Check
+#### Model Comparison
 
 ```{r}
+# m2_betagate <- readRDS("models/sample2_betagate.rds")
+m2_betagate <- readRDS("models/sample2noex_betagate.rds")
 m2_chocoat <- readRDS("models/sample2_chocoattractiveness.rds")
+m2_chocoat <- readRDS("models/sample2noex_chocoattractiveness.rds")
+```
 
-m <- m2_chocoat
+```{r}
+loocomp <- loo::loo_compare(m2_betagate, m2_chocoat, criterion = "waic")
 
+parameters(loocomp, include_ENP = TRUE)
+
+report::report(loocomp)
+```
+
+##### Posterior Predictive Check
+
+```{r}
 make_ppcheck  <- function(m, fill = "#FF9800") {
@@
     ggplot(aes(x = iter_value)) +
-    stat_bin(data = df, aes(x = Real, y = after_stat(density)),
+    stat_bin(data = df, aes(x = RealNoex, y = after_stat(density)),
              geom="bar", breaks = seq(0, 1.0, by = 0.04), fill = "#424242") +
@@
 
-make_ppcheck(m2_chocoat) +
+p_ppc1 <- make_ppcheck(m2_chocoat, fill = "brown") +
   labs("Real") 
+p_ppc2 <- make_ppcheck(m2_betagate, fill = "#FF9800") +
+  labs("Real")
+
+p_ppc1 / p_ppc2 
 ```
@@
 ```{r}
-# c <- estimate_contrasts(m2_chocoat, contrast = "Condition", by = "Sex", predict = "mu")
-# c <- estimate_contrasts(m2_chocoat, contrast = "Condition", by = "Sex", predict = "mu", backend = "emmeans")
-# c2 <- estimate_contrasts(m2_chocoat, contrast = "Condition", by = "Sex", predict = "confright", backend = "emmeans")
+c <- data.frame()
+for(param in c("mu", "confright", "confleft", "precright", "precleft", "pmid")) {
+  c <- estimate_contrasts(m2_chocoat, contrast = "Condition", by = "Sex", predict = param, backend = "emmeans") |>
+    mutate(Parameter = ifelse(param == "mu", "p", param)) |> 
+    rbind(c)
+}
+c
 ```
@@
                          newdata=insight::get_datagrid(m2_chocoat, length = 5, preserve_range = FALSE),
-                         x = seq(0, 1, length.out=7))
+                         x = seq(0, 1, length.out=9))
 
@@
          y = ifelse(x %in% c(0, 1), y * 3, y),  # Inflate extremes for visibility
-         x = ifelse(Condition == "AI-Generated", x + 0.1, x)) |>  
+         x = ifelse(Condition == "AI-Generated", x + 0.01, x)) |>  
   filter(Condition == "Photograph") |>
   ggplot(aes(x=x, y=y)) +
-  geom_line(aes(color=Attractive), linewidth = 1.25) +
-  geom_path(data = filter(pred$mu, Condition == "Photograph"), aes(group = side), linewidth=0, lineend = 'round', 
-            arrow = arrow(length = unit(0.025, "npc"), type = "closed")) +
+  geom_line(aes(color=Attractive, linetype = Condition), linewidth = 1.25) +
+  # geom_path(data = filter(pred$mu, Condition == "Photograph"), aes(group = side), linewidth=0, lineend = 'round', 
+  #           arrow = arrow(length = unit(0.025, "npc"), type = "closed")) +
   facet_wrap(~Sex) +
@@
 
+#### Model Description
+
+::: {.panel-tabset}
+
+##### Effect of Attractiveness
+
+```{r}
+m2_chocobeauty <- readRDS("models/sample2noex_chocobeauty.rds")
+params_table(m2_chocobeauty, keep = "polyBeauty|Condition")
+```
+
+:::
 ## Figures

### 2025-05-06 paper/manuscript.qmd
update what I have
@AnafNeves I'm pushing what I have if you want to take a look at the ms draft

@@
 The sample is the same as in [@makowski2025too].
-We included all heterosexual male and female participants, for a final sample of 125 participants (Mean age = 28.6, SD = 9.2, range: [19, 66]; Sex: 44.8% females).
-We included only sexually relevant stimuli for men and women (i.e., all 89 female faces for men and 20 male faces for women).
+We included all heterosexual and bisexual (these two groups did not seem to differ based on preliminary analyses and were thus grouped to maximize power) male and female participants, for a final sample of 141 participants (Mean age = 28.4, SD = 9.0, range: [19, 66]; Sex: 47.5% females).
+For each participant, we included only stimuli of the opposite gender (i.e., all 89 female faces for men and 20 male faces for women).
 
@@
 The difference in predictive accuracy, as indexed by Expected Log Predictive Density (ELPD-WAIC), suggests that the *CHOCO* is the best
-model ($ELPD = -138.45$), followed by *Beta-Gate* ($\Delta_{ELPD} = -1875.31 \pm 61.18, p < .001)$ and *ZOIB* ($\Delta_{ELPD} = -1916.11 \pm 61.75, p < .001$).
+model ($ELPD = -203.54$), followed by *Beta-Gate* ($\Delta_{ELPD} = -1794.57 \pm 63.12, p < .001)$ and *ZOIB* ($\Delta_{ELPD} = -1833.59 \pm 63.52, p < .001$).
 See @fig-three for the posterior predictive checks, showing that only the CHOCO model managed to capture the bimodal distribution of data.
@@
 
-The ZOIB model suggested that women had higher mean (*mu* parameter) scores of reality beliefs ($\Delta_{\text{Female}} = 0.16,~95\%~ CI~[0.01, 0.32], pd = 98.10\%$), less extreme values (*zoi* parameter; STATS) but more ones relative to zeros (*coi* parameter; STATS).
+The ZOIB model suggested that women had higher mean (*mu* parameter) scores of reality beliefs ($\Delta_{\text{Female}} = 0.20,~95\%~ CI~[0.03, 0.37], pd = 98.77\%$), less extreme values (*zoi* parameter; STATS) but more ones relative to zeros (*coi* parameter; STATS).
 
@@
 
-The ZOIB model estimated reliable variability in the participant's *phi* parameter (D-vour = 0.88) and *zoi* parameter (D-vour = 0.84), as well as in the *mu* parameter related to individual items (D-vour = 0.82).
+The ZOIB model estimated reliable variability in the participant's *phi* parameter (D-vour = 0.88) and *zoi* parameter (D-vour = 0.85), as well as in the *mu* parameter related to individual items (D-vour = 0.82).
 Moderate reliability was also observed for items in the *coi* parameter (D-vour = 0.71) and for participants in the *mu* parameter (D-vour = 0.69).
-The Beta-Gate model yielded similar results: a high reliability of participant's *phi* parameter (D-vour = 0.88), *pex* parameter (D-vour = 0.84).
+The Beta-Gate model yielded similar results: a high reliability of participant's *phi* parameter (D-vour = 0.88), *pex* parameter (D-vour = 0.85).
 The *mu* parameter's variability was reliably captured for items (D-vour = 0.85) and moderately for participants (D-vour = 0.72).
-Of note that the parameters of Beta-Gate and ZOIB correlate almost perfectly, underlining their empirical similarity despite a different parametrization.
 
-The CHOCO model yielded reliable estimates for all parameters except *bex* for participants (*confright* D-vour = 0.94, *confleft* D-vour = 0.91, *pex* D-vour = 0.80, *p* D-vour = 0.73, *precright* D-vour = 0.79, *precleft* D-vour = 0.73).
+The CHOCO model yielded reliable estimates for all parameters except *bex* for participants (*confright* D-vour = 0.94, *confleft* D-vour = 0.91, *pex* D-vour = 0.79, *p* D-vour = 0.73, *precright* D-vour = 0.77, *precleft* D-vour = 0.67).
 Item's variability was primarily reflected through the *p* parameter (D-vour = 0.86).
 
-Finally, the empirical average correlated the most strongly with CHOCO's *p* (r = .94), rather than ZOIB's *mu* (r = .74) or
+Finally, the empirical average correlated the most strongly with CHOCO's *p* (r = .86), rather than ZOIB's *mu* (r = .77) or
 Beta-Gate's *mu* (r = .82).
@@
 The empirical *pex* correlated the strongest with ZOIB's *zoi* (r = .90), and Beta-Gate's *pex* (r = .90), and CHOCO's *pex* (r = .88).
+Of note that the parameters of Beta-Gate and ZOIB correlate almost perfectly, underlining their empirical similarity despite a different parametrization.
 
@@
 The first sample includes the same participants as study 1 (see above).
+
 **TODO** for sample 2.
+
+The final sample included 189 participants (Mean age = 28.4, SD = 14.0, range: [18, 69]; Sex: 76.2% females).
 
@@
 
-In sample 1, the traditional approach suggested a significant linear relationship between the mean level of "Reality" and attractiveness ($\beta_{poly1}=3.63, ~95\%~CI [2.71, 4.56], pd = ...\%$) for males only.
+### Sample 1
+
+In sample 1, the traditional approach suggested a significant linear relationship between the mean level of "Reality" and attractiveness ($\beta_{poly1}=3.42, ~95\%~CI [2.50, 4.34], p < .001\%$) for males only.
 The CHOCO model revealed that attractiveness had a significantly positive linear relationship with the probability *p* of judging faces as real (STATS), but a quadratic relationship with the confidence in real judgments (*confright*; STATS).
-Attractiveness was also associated with less confidence in fake judgments (*confleft*; STATS), and related linearly with a stronger bias towards extreme Real responses (*bex*; STATS).
+Attractiveness was also associated with less confidence in fake judgments (*confleft*; STATS), a quadratic relationship with left precision (*precleft*; STATS), and related linearly with a stronger bias towards extreme Real responses (*bex*; STATS).
 These effects were present for males only, and no relationship was found for females.
-The reliability of the effect of attractiveness (the random slopes) was very low for all parameters (D-vour < 0.1).
+The reliability of the effect of attractiveness (the random slopes) was very low for all parameters (D-vour < 0.01).
+The effect of Beauty was the same as of attractiveness.
 
-The effect of Beauty was the same as above, with some additional evidence in favour of an effect in women, in particular for a linear relationship with *p* for females as well ($\beta_{poly1}=7.02, ~95\%~CI [-0.90, 14.97], pd = 95.91\%$), as well as a quadradic relationship with the confidence in fake judgments (*confleft*; STATS).
+<!-- , with some additional evidence in favour of an effect in women, in particular for a linear relationship with *p* for females as well ($\beta_{poly1}=7.02, ~95\%~CI [-0.90, 14.97], pd = 95.91\%$), as well as a quadratic relationship with the confidence in fake judgments (*confleft*; STATS).
+-->
+
+### Sample 2
 
@@
 
+This study introduces a new statistical model to analyze bimodal data, which can be observed for instance when using subjective rating scales in which the two sides correspond to a different "choice" (e.g., "false" vs.
+"true", "real" vs "fake").
+The Choice-Confidence (CHOCO) model conceptualizes the data as a mixture of two distributions, one for each choice, and models the probability of choosing one or the other as well as the degree of confidence in each choice.
+We validated this model on real-life data by appling it to the reality beliefs of @makowski2025too, and showed that it was able to capture the bimodal nature of the data and provides a more accurate representation of the underlying processes than other alternative approaches, including Zero-or-One inflated Beta (ZOIB) models.
+
 Little difference between Beauty and Attractiveness, aside that Beauty revealed more effects in women. Why?
+
+We demonstrated the flexibility of the CHOCO model, being usable in analog scales as well as discrete ratings with a mid-point.
 
@@
 Having 7-points (or 6 with no mid-value) is the strict minimum to use the CHOCO model.
+
+Future studies should assess psychometric property liek the minimum amount of data required to estimate all parameters reliably.
 

### 2025-05-06 analysis/server/combine_models.R

@@ -8,10 +8,13 @@ library(loo)
 options(mc.cores = as.numeric(Sys.getenv("SLURM_CPUS_PER_TASK")))
 task_id <- as.numeric(Sys.getenv("SLURM_ARRAY_TASK_ID"))
 
+setwd("/mnt/lustre/users/psych/dmm56/FictionChoco/")
+
 # List of models
 model_names <- c("sample1_zoib", "sample1_betagate", "sample1_choco",
                  "sample1_chocoattractiveness", "sample1_chocobeauty",
-                 "sample2_chocoattractiveness")[6]
+                 "sample2_chocoattractiveness", "sample2_betagate",
+                 "sample2noex_betagate", "sample2noex_chocoattractiveness", "sample2noex_chocobeauty")[4:10]
 
 # Select the model name based on the task ID
 model_name <- model_names[task_id]

### 2025-05-06 analysis/server/make_models.R

@@ -16,6 +16,7 @@ chains_per_task <- as.numeric(Sys.getenv("SLURM_CPUS_PER_TASK"))
 start_chain <- (task_id - 1) * chains_per_task + 1
 
 iter <- 1500 # 50% of that is warmup
+path <- "/mnt/lustre/users/psych/dmm56/FictionChoco/"
 
 # Test --------------------------------------------------------------------
 
@@ -27,16 +28,15 @@ iter <- 1500 # 50% of that is warmup
 #   cores = chains_per_task,     # Use all 16 CPUs
 #   iter = 500,
 #   seed = 1234 + start_chain,   # Unique seed per task
-#   file = paste0("./modeltoy_task_", task_id, ".rds")
+#   file = paste0(path, "modeltoy_task_", task_id, ".rds")
 # )
-
 
 
 # Sample 1 ----------------------------------------------------------------
 
+
 df <- read.csv("https://raw.githubusercontent.com/RealityBending/FictionChoco/refs/heads/main/data/sample1.csv")
 df$Sex <- factor(df$Sex, levels = c("Male", "Female"))
-# df <- df[df$Participant %in% unique(df$Participant)[1:10],]
 
 
 
@@ -61,7 +61,7 @@ model <- brm(
   cores = chains_per_task,     # Use all 16 CPUs
   iter = iter,  # Number of iterations and warmup must be equal (warmup is 50% iter)
   seed = 1234 + start_chain,   # Unique seed per task
-  file = paste0("./sample1_zoib_task_", task_id, ".rds")
+  file = paste0(path, "sample1_zoib_task_", task_id, ".rds")
 )
 
 # # # # # DO NOT DO XBX
@@ -83,7 +83,7 @@ model <- brm(
 # # # # #   cores = chains_per_task,     # Use all 16 CPUs
 # # # # #   iter = iter,  # Number of iterations and warmup must be equal (warmup is 50% iter)
 # # # # #   seed = 1234 + start_chain,   # Unique seed per task
-# # # # #   file = paste0("./sample1_xbx_task_", task_id, ".rds")
+# # # # #   file = paste0(path, "sample1_xbx_task_", task_id, ".rds")
 # # # # # )
 # # #
 # # # # #
@@ -110,7 +110,7 @@ model <- brm(
   cores = chains_per_task,     # Use all 16 CPUs
   iter = iter,  # Number of iterations and warmup must be equal (warmup is 50% iter)
[…]

### 2025-05-23 paper/manuscript.qmd
typo

@@
 
-For the population-level effects, we will consider significant and report (using the median of the posterior distribution) effects for which the 95% Credible Interval (CI) does not include zero [and when the probability of direction *pd* is > \~97\%; @makowski2019indices].
+For the population-level effects, we will consider significant and report (using the median of the posterior distribution) effects for which the 95% Credible Interval (CI) does not include zero [and when the probability of direction *pd* is > \~97\%, @makowski2019indices].
 For the individual-level parameters (i.e., the random intercepts of each parameter for each participant and each item), we will first analyze their reliability using the Variance-Over-Uncertainty Ratio index (*D-vour*).
@@
 One such heuristic might be facial attractiveness.
-Attractiveness appraisals are known to be automatic and unconscious [@luo2019robust; @hung2016unconscious; @hou2023neural], and carry strong real-life consequences, as demonstrated by the large body of literature on the "beauty premium" [@little2021facial; @gulati2024beautiful, **MORE**]^[As an eloquent example, @monk2021beholding reported in a large representative US sample that the magnitude of the earnings disparities among white women along the perceived attractiveness continuum exceeds in magnitude the canonical black-white race gap.]
+Attractiveness appraisals are known to be automatic and unconscious [@luo2019robust; @hung2016unconscious; @hou2023neural], and carry strong real-life consequences, as demonstrated by the large body of literature on the "beauty premium" [@little2021facial; @gulati2024beautiful, **MORE**]^[As an eloquent example, @monk2021beholding reported in a large representative US sample that the magnitude of earnings disparities among white women along the perceived attractiveness continuum exceeds in magnitude the canonical black-white race gap.]
 

### 2025-05-28 paper/manuscript.qmd
changes to script

@@
 
+In a post-truth era [@lewandowsky2017beyond], where near-perfect simulations are increasingly feasible due to rapid technological advancements, distinguishing between what is real and what is fake poses a growing challenge [**REF**].
+As these technologies advance and physical cues become indistinguishable, individuals are likely to rely more heavily on alternative sources of information -such as contextual cues and cognitive heuristic -when making reality judgments [i.e., simulation monitoring; @makowski2018cognitive, @makowski2019phenomenal].
+This reliance is particularly pronounced under conditions of high ambiguity, where processing demands are higher and the atomization and decontextualization of information (especially prevalent in online environments) further complicate the discernment of authenticity.
+
 Another unknown is whether these beliefs are fully stimuli-driven or whether there are some stable individual tendencies that are at stake. 
+For instance, in the absence of clear contextual information, individual-level characteristics may increasingly guide beliefs about reality.
+Traits such as narcissism have been shown to influence such judgments [@makowski2025too].
+Moreover, it has been proposed that the interaction between stimulus characteritics and individual differences plays a crucial role in shaping reality beliefs.
+For example, self-reported arousal has been found to predict the likelihood of perceiving an image as real [@azevedo2020body], while judging a stimulus as fake has been associated with emotional downregulation [@makowski2019phenomenal].
+[**should probably add other examples**]
+
+These findings underscore that what drives beliefs and judgments of reality remains unclear.
+Yet, understanding these mechanisms is increasingly vital given the proliferation of AI-generated content and the growing risk of misjudging the authenticity of stimuli, particularly in contexts with significant societal consequences (e.g., the influence of misinformation on democratic processes).
 
@@
 In the second phase, participants were informed that "about half of the previously seen images were AI-generated".
-The same faces were presented again in a new random order (same display time), followed by ratings of "reality" (whether they believed the image was AI-generated - left anchor - or a real photograph - right anchor).
+The same faces were presented again in a new random order (same display time), followed by ratings of "reality" (whether they believed the image was fake - left anchor - or a real - right anchor).
 @makowski2025too showed that most stimuli yielded highly variable reality ratings, suggesting that idiosyncratic cognitive processes were at play.
@@
 For all models and each parameter, the full formula was entered: $Real\sim Sex + (1 | Participant) + (1 | Item)$ (with Sex as the main predictor and participants and items entered as random intercepts).
-The models were run using *brms* [@burkner2017brms] R package, and analyzed using the *easystats* collection of packages [**REFS**].
+The models were run using *brms* [@burkner2017brms] R package, and analyzed using the *easystats* collection of packages [@easystats2022].
 To maximize the comparability across models We used the default priors (uniform) for all models, and we ran 16 chains of 1400 iterations each on the University of Sussex High-Performance Computing (HPC) cluster.
@@
 
-The explosion of accessibility of state-of-the-art AI tools has made it effortless to generate realistic images, including human 
-faces that are often indistinguishable from real ones **[REF]**.
+The explosion of accessibility of state-of-the-art AI tools has made it effortless to generate realistic images, including human faces that are often indistinguishable from real ones [@miller2023ai; @nightingale2022ai; @bozkir2024can].
 These synthetic visuals are flooding the cyberspace across various domains, such as art, advertising and entertainment, to education and information.
@@
 One such heuristic might be facial attractiveness.
-Attractiveness appraisals are known to be automatic and unconscious [@luo2019robust; @hung2016unconscious; @hou2023neural], and carry strong real-life consequences, as demonstrated by the large body of literature on the "beauty premium" [@little2021facial; @gulati2024beautiful, **MORE**]^[As an eloquent example, @monk2021beholding reported in a large representative US sample that the magnitude of earnings disparities among white women along the perceived attractiveness continuum exceeds in magnitude the canonical black-white race gap.]
+Attractiveness appraisals are known to be automatic and unconscious [@luo2019robust; @hung2016unconscious; @hou2023neural], and carry strong real-life consequences, as demonstrated by the large body of literature on the "beauty premium" [@little2021facial; @gulati2024beautiful;@pandey2021face; @kukkonen2024beauty]^[As an eloquent example, @monk2021beholding reported in a large representative US sample that the magnitude of earnings disparities among white women along the perceived attractiveness continuum exceeds in magnitude the canonical black-white race gap.]
 
 **TODO:** However, it is important to allow for the possibility of a U-shaped relationship.
-For instance, Peng et al.
-(2020) reported that both attractive and unattractive people sold more (underlining a "plainness penalty").
-<!-- The Faces of Success: Beauty and Ugliness Premiums in e-Commerce Platforms -->
+For instance, @peng2020faces reported that both attractive and unattractive people sold more products (underlining a "plainness penalty").
 
-Although its role as a potential modulator of reality beliefs remains underexplored, Makowski et al... **TODO: detail findings of that study + other studies**.
+Although its role as a potential modulator of reality beliefs remains underexplored, @makowski2025too found significant associations between participants’ realness ratings and facial salience.
+Female participants judged both highly attractive and highly unattractive faces as more likely to be real, whereas for male participants, this effect was limited to highly attractive faces.
+These findings offer a complementary perspective to those of @miller2023ai, who reported that AI-generated faces tend to be rated as more attractive, and that participants used attractiveness as a distinguishing cue between real and AI-generated faces.
+Together, these results highlight a possible bidirectional and context-dependent role of attractiveness in shaping reality judgments.
 
@@
 For the second sample, the procedure was relatively similar with a few key differences.
-**TODO** Describe  fake Face 2. 
-The main difference is that the subjective ratings were done on a X-point Likert scale, rather than analog, with a mid-point value.
+Most importantly, subjective ratings were collected using a 7-point Likert scale ranging from 0 to 6, rather than a visual analog scale, which included a clear midpoint. 
+Additionally, the familiarity rating was removed from the set of facial evaluations. 
+In the second phase, realness judgments were also recorded on a 7-point Likert scale, this time ranging from -3 to +3, again with a defined midpoint, replacing the visual analog format used in the first study.
+For this scale, the endpoints were labeled 'AI-generated' and 'Photograph', replacing the 'Fake' and 'Real' anchors used in Study 1.
+
+There were also modifications to the instructions provided to participants. 
+In Phase 1, an additional prompt was introduced, stating that the study was conducted in partnership with a young AI startup. 
+This was intended to strengthen participants’ belief that the faces they were shown could plausibly be either real or AI-generated. 
+Additionally, participants were explicitly asked to respond based on their immediate, first impressions—an instruction not included in the first study.
+
+In Phase 2, participants were not informed about the proportion of AI-generated versus real faces encountered during Phase 1. 
+This omission was deliberate, aimed at reducing demand characteristics and preventing participants from mechanically selecting half the faces as real and half as AI-generated. 
+Instead, they were encouraged to base their responses solely on their level of confidence regarding the origin of each image.
 

### 2025-05-28 paper/manuscript.tex
changes to script

@@
 \makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\newsavebox\pandoc@box
+\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
+  \sbox\pandoc@box{#1}%
+  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
+  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
+  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
+  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
+  \else\usebox{\pandoc@box}%
+  \fi%
+}
 % Set default figure placement to htbp
-\makeatletter
 \def\fps@figure{htbp}
@@
 \par{ }
-\par{     \begin{tcolorbox}[enhanced jigsaw, left=2mm, arc=.35mm, bottomrule=.15mm, colframe=quarto-callout-note-color-frame, leftrule=.75mm, breakable, rightrule=.15mm, toprule=.15mm, colback=white, opacityback=0]
+\par{     \begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, bottomrule=.15mm, colback=white, colframe=quarto-callout-note-color-frame, opacityback=0, breakable, arc=.35mm, leftrule=.75mm, toprule=.15mm, left=2mm]
 
@@
 \begin{center}
-\includegraphics[width=0.2\textwidth,height=\textheight]{manuscript_files/mediabag/ReBeL_LogoOnly_hu114.png}
+\includegraphics[width=0.2\linewidth,height=\textheight,keepaspectratio]{manuscript_files/mediabag/ReBeL_LogoOnly_hu114.png}
 \end{center}
@@
 misleading conclusions when applied indiscriminately. It is thus
-important that psychologists use models that can best describe the data
-they collect, to fully exploit them and bring more nuance and accuracy
-to their conclusions.
+important that psychologists use models that can best describe (or
+generate) the data they collect, to fully exploit them and bring more
+nuance and accuracy to their conclusions.
 
@@
 
-\includegraphics[width=0.5\textwidth,height=\textheight]{./figures/fig1.png}
+\includegraphics[width=0.5\linewidth,height=\textheight,keepaspectratio]{./figures/fig1.png}
 
@@
 
-\includegraphics[width=0.5\textwidth,height=\textheight]{./figures/fig2.png}
+\includegraphics[width=0.5\linewidth,height=\textheight,keepaspectratio]{./figures/fig2.png}
 
@@
 
+In a post-truth era (\citeproc{ref-lewandowsky2017beyond}{Lewandowsky et
+al., 2017}), where near-perfect simulations are increasingly feasible
+due to rapid technological advancements, distinguishing between what is
+real and what is fake poses a growing challenge {[}\textbf{REF}{]}. As
+these technologies advance and physical cues become indistinguishable,
+individuals are likely to rely more heavily on alternative sources of
+information -such as contextual cues and cognitive heuristic -when
+making reality judgments {[}i.e., simulation monitoring; Makowski
+(\citeproc{ref-makowski2018cognitive}{2018}), Makowski, Sperduti, et al.
+(\citeproc{ref-makowski2019phenomenal}{2019}){]}. This reliance is
+particularly pronounced under conditions of high ambiguity, where
+processing demands are higher and the atomization and
+decontextualization of information (especially prevalent in online
+environments) further complicate the discernment of authenticity.
+
 Another unknown is whether these beliefs are fully stimuli-driven or
 whether there are some stable individual tendencies that are at stake.
+For instance, in the absence of clear contextual information,
+individual-level characteristics may increasingly guide beliefs about
+reality. Traits such as narcissism have been shown to influence such
+judgments (\citeproc{ref-makowski2025too}{Makowski et al., 2025}).
+Moreover, it has been proposed that the interaction between stimulus
+characteritics and individual differences plays a crucial role in
+shaping reality beliefs. For example, self-reported arousal has been
+found to predict the likelihood of perceiving an image as real
+(\citeproc{ref-azevedo2020body}{Azevedo et al., 2020}), while judging a
+stimulus as fake has been associated with emotional downregulation
+(\citeproc{ref-makowski2019phenomenal}{Makowski, Sperduti, et al.,
+2019}). {[}\textbf{should probably add other examples}{]}
+
+These findings underscore that what drives beliefs and judgments of
+reality remains unclear. Yet, understanding these mechanisms is
+increasingly vital given the proliferation of AI-generated content and
+the growing risk of misjudging the authenticity of stimuli, particularly
+in contexts with significant societal consequences (e.g., the influence
+of misinformation on democratic processes).
 
@@
 The sample is the same as in (\citeproc{ref-makowski2025too}{Makowski et
-al., 2025}). We included all heterosexual male and female participants,
-for a final sample of 125 participants (Mean age = 28.6, SD = 9.2,
-range: {[}19, 66{]}; Sex: 44.8\% females). We included only sexually
-relevant stimuli for men and women (i.e., all 89 female faces for men
-and 20 male faces for women).
+al., 2025}). We included all heterosexual and bisexual (these two groups
+did not seem to differ based on preliminary analyses and were thus
+grouped to maximize power) male and female participants, for a final
+sample of 141 participants (Mean age = 28.4, SD = 9.0, range: {[}19,
+66{]}; Sex: 47.5\% females). For each participant, we included only
+stimuli of the opposite gender (i.e., all 89 female faces for men and 20
+male faces for women).
 
@@
 a new random order (same display time), followed by ratings of
-``reality'' (whether they believed the image was AI-generated - left
-anchor - or a real photograph - right anchor). Makowski et al.
+``reality'' (whether they believed the image was fake - left anchor - or
+a real - right anchor). Makowski et al.
 (\citeproc{ref-makowski2025too}{2025}) showed that most stimuli yielded
@@
 (\citeproc{ref-burkner2017brms}{Bürkner, 2017}) R package, and analyzed
-using the \emph{easystats} collection of packages {[}\textbf{REFS}{]}.
-To maximize the comparability across models We used the default priors
-(uniform) for all models, and we ran 16 chains of 1400 iterations each
-on the University of Sussex High-Performance Computing (HPC) cluster.
+using the \emph{easystats} collection of packages
+(\citeproc{ref-easystats2022}{Lüdecke et al., 2022}). To maximize the
+comparability across models We used the default priors (uniform) for all
+models, and we ran 16 chains of 1400 iterations each on the University
+of Sussex High-Performance Computing (HPC) cluster.
 
@@
 report (using the median of the posterior distribution) effects for
-which the 95\% Credible Interval (CI) does not include zero {[}and when
+which the 95\% Credible Interval (CI) does not include zero (and when
 the probability of direction \emph{pd} is \textgreater{}
-\textasciitilde97\%; Makowski et al.
-(\citeproc{ref-makowski2019indices}{2019}){]}. For the individual-level
-parameters (i.e., the random intercepts of each parameter for each
-participant and each item), we will first analyze their reliability
-using the Variance-Over-Uncertainty Ratio index (\emph{D-vour}). This
-index, implemented in the \emph{performance} package
+\textasciitilde97\%, \citeproc{ref-makowski2019indices}{Makowski,
+Ben-Shachar, et al., 2019}). For the individual-level parameters (i.e.,
+the random intercepts of each parameter for each participant and each
+item), we will first analyze their reliability using the
+Variance-Over-Uncertainty Ratio index (\emph{D-vour}). This index,
+implemented in the \emph{performance} package
 (\citeproc{ref-ludecke2021performance}{Lüdecke et al., 2021}), is
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{./figures/fig3.png}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{./figures/fig3.png}
 
@@
 indexed by Expected Log Predictive Density (ELPD-WAIC), suggests that
-the \emph{CHOCO} is the best model (\(ELPD = -138.45\)), followed by
-\emph{Beta-Gate} (\(\Delta_{ELPD} = -1875.31 \pm 61.18, p < .001)\) and
-\emph{ZOIB} (\(\Delta_{ELPD} = -1916.11 \pm 61.75, p < .001\)). See
+the \emph{CHOCO} is the best model (\(ELPD = -203.54\)), followed by
+\emph{Beta-Gate} (\(\Delta_{ELPD} = -1794.57 \pm 63.12, p < .001)\) and
+\emph{ZOIB} (\(\Delta_{ELPD} = -1833.59 \pm 63.52, p < .001\)). See
 Figure~\ref{fig-three} for the posterior predictive checks, showing that
@@
 parameter) scores of reality beliefs
-(\(\Delta_{\text{Female}} = 0.16,~95\%~ CI~[0.01, 0.32], pd = 98.10\%\)),
+(\(\Delta_{\text{Female}} = 0.20,~95\%~ CI~[0.03, 0.37], pd = 98.77\%\)),
 less extreme values (\emph{zoi} parameter; STATS) but more ones relative
@@
 
-Across all these models, no effect of Sex on the prediction parameter
-was observed.
+Across all these models, no effect of Sex on the precision parameter was
+observed.
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{./figures/fig4.png}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{./figures/fig4.png}
 
@@
 \emph{phi} parameter (D-vour = 0.88) and \emph{zoi} parameter (D-vour =
-0.84), as well as in the \emph{mu} parameter related to individual items
+0.85), as well as in the \emph{mu} parameter related to individual items
 (D-vour = 0.82). Moderate reliability was also observed for items in the
@@
 results: a high reliability of participant's \emph{phi} parameter
-(D-vour = 0.88), \emph{pex} parameter (D-vour = 0.84). The \emph{mu}
+(D-vour = 0.88), \emph{pex} parameter (D-vour = 0.85). The \emph{mu}
 parameter's variability was reliably captured for items (D-vour = 0.85)
-and moderately for participants (D-vour = 0.72). Of note that the
-parameters of Beta-Gate and ZOIB correlate almost perfectly, underlining
-their empirical similarity despite a different parametrization.
+and moderately for participants (D-vour = 0.72).
 
@@
 \emph{bex} for participants (\emph{confright} D-vour = 0.94,
-\emph{confleft} D-vour = 0.91, \emph{pex} D-vour = 0.80, \emph{p} D-vour
-= 0.73, \emph{precright} D-vour = 0.79, \emph{precleft} D-vour = 0.73).
+\emph{confleft} D-vour = 0.91, \emph{pex} D-vour = 0.79, \emph{p} D-vour
+= 0.73, \emph{precright} D-vour = 0.77, \emph{precleft} D-vour = 0.67).
 Item's variability was primarily reflected through the \emph{p}
@@
 Finally, the empirical average correlated the most strongly with CHOCO's
-\emph{p} (r = .94), rather than ZOIB's \emph{mu} (r = .74) or
+\emph{p} (r = .86), rather than ZOIB's \emph{mu} (r = .77) or
 Beta-Gate's \emph{mu} (r = .82). The empirical overall confidence was
@@
 \emph{zoi} (r = .90), and Beta-Gate's \emph{pex} (r = .90), and CHOCO's
-\emph{pex} (r = .88).
+\emph{pex} (r = .88). Of note that the parameters of Beta-Gate and ZOIB
+correlate almost perfectly, underlining their empirical similarity
+despite a different parametrization.
 
@@
 effortless to generate realistic images, including human faces that are
-often indistinguishable from real ones \textbf{{[}REF{]}}. These
+often indistinguishable from real ones
+(\citeproc{ref-bozkir2024can}{Bozkir et al., 2024};
+\citeproc{ref-miller2023ai}{Miller et al., 2023};
+\citeproc{ref-nightingale2022ai}{Nightingale \& Farid, 2022}). These
 synthetic visuals are flooding the cyberspace across various domains,
@@
 on the ``beauty premium'' (\citeproc{ref-gulati2024beautiful}{Gulati et
-al., 2024}, \textbf{MORE}; \citeproc{ref-little2021facial}{Little,
-2021})\footnote{As an eloquent example, Monk Jr et al.
-  (\citeproc{ref-monk2021beholding}{2021}) reported that the the
-  magnitude of the earnings disparities among white women along the
-  perceived attractiveness continuum exceeds in magnitude the canonical
-  black-white race gap in a large representative US sample.}
+al., 2024}; \citeproc{ref-kukkonen2024beauty}{Kukkonen et al., 2024};
+\citeproc{ref-little2021facial}{Little, 2021};
+\citeproc{ref-pandey2021face}{Pandey \& Zayas, 2021})\footnote{As an
+  eloquent example, Monk Jr et al.
+  (\citeproc{ref-monk2021beholding}{2021}) reported in a large
+  representative US sample that the magnitude of earnings disparities
+  among white women along the perceived attractiveness continuum exceeds
+  in magnitude the canonical black-white race gap.}
 
 \textbf{TODO:} However, it is important to allow for the possibility of
-a U-shaped relationship. For instance, Peng et al.~(2020) reported that
-both attractive and unattractive people sold more (underlining a
-``plainness penalty'').
+a U-shaped relationship. For instance, Peng et al.
+(\citeproc{ref-peng2020faces}{2020}) reported that both attractive and
+unattractive people sold more products (underlining a ``plainness
+penalty'').
 
 Although its role as a potential modulator of reality beliefs remains
-underexplored, Makowski et al\ldots{} \textbf{TODO: detail findings of
-that study + other studies}.
+underexplored, Makowski et al. (\citeproc{ref-makowski2025too}{2025})
+found significant associations between participants' realness ratings
+and facial salience. Female participants judged both highly attractive
+and highly unattractive faces as more likely to be real, whereas for
+male participants, this effect was limited to highly attractive faces.
+These findings offer a complementary perspective to those of Miller et
+al. (\citeproc{ref-miller2023ai}{2023}), who reported that AI-generated
+faces tend to be rated as more attractive, and that participants used
+attractiveness as a distinguishing cue between real and AI-generated
+faces. Together, these results highlight a possible bidirectional and
+context-dependent role of attractiveness in shaping reality judgments.
 
@@
 The first sample includes the same participants as study 1 (see above).
+
 \textbf{TODO} for sample 2.
 
+The final sample included 189 participants (Mean age = 28.4, SD = 14.0,
+range: {[}18, 69{]}; Sex: 76.2\% females).
+
 \subsubsection{Procedure}\label{procedure-1}
@@
 For the second sample, the procedure was relatively similar with a few
-key differences. \textbf{TODO} Describe fake Face 2. The main difference
-is that the subjective ratings were done on a X-point Likert scale,
-rather than analog, with a mid-point value.
+key differences. Most importantly, subjective ratings were collected
+using a 7-point Likert scale ranging from 0 to 6, rather than a visual
+analog scale, which included a clear midpoint. Additionally, the
+familiarity rating was removed from the set of facial evaluations. In
+the second phase, realness judgments were also recorded on a 7-point
+Likert scale, this time ranging from -3 to +3, again with a defined
+midpoint, replacing the visual analog format used in the first study.
+For this scale, the endpoints were labeled `AI-generated' and
+`Photograph', replacing the `Fake' and `Real' anchors used in Study 1.
+
+There were also modifications to the instructions provided to
+participants. In Phase 1, an additional prompt was introduced, stating
+that the study was conducted in partnership with a young AI startup.
+This was intended to strengthen participants' belief that the faces they
+were shown could plausibly be either real or AI-generated. Additionally,
+participants were explicitly asked to respond based on their immediate,
+first impressions---an instruction not included in the first study.
+
+In Phase 2, participants were not informed about the proportion of
+AI-generated versus real faces encountered during Phase 1. This omission
+was deliberate, aimed at reducing demand characteristics and preventing
+participants from mechanically selecting half the faces as real and half
+as AI-generated. Instead, they were encouraged to base their responses
+solely on their level of confidence regarding the origin of each image.
 
@@
 
+\subsubsection{Sample 1}\label{sample-1}
+
 In sample 1, the traditional approach suggested a significant linear
 relationship between the mean level of ``Reality'' and attractiveness
-(\(\beta_{poly1}=3.63, ~95\%~CI [2.71, 4.56], pd = ...\%\)) for males
+(\(\beta_{poly1}=3.42, ~95\%~CI [2.50, 4.34], p < .001\%\)) for males
 only. The CHOCO model revealed that attractiveness had a significantly
@@
 associated with less confidence in fake judgments (\emph{confleft};
+STATS), a quadratic relationship with left precision (\emph{precleft};
 STATS), and related linearly with a stronger bias towards extreme Real
@@
 effect of attractiveness (the random slopes) was very low for all
-parameters (D-vour \textless{} 0.1).
-
-The effect of Beauty was the same as above, with some additional
-evidence in favour of an effect in women, in particular for a linear
-relationship with \emph{p} for females as well
-(\(\beta_{poly1}=7.02, ~95\%~CI [-0.90, 14.97], pd = 95.91\%\)), as well
-as a quadradic relationship with the confidence in fake judgments
-(\emph{confleft}; STATS).
+parameters (D-vour \textless{} 0.01). The effect of Beauty was the same
+as of attractiveness.
+
+\subsubsection{Sample 2}\label{sample-2}
 
@@
 
+This study introduces a new statistical model to analyze bimodal data,
+which can be observed for instance when using subjective rating scales
+in which the two sides correspond to a different ``choice'' (e.g.,
+``false'' vs.~``true'', ``real'' vs ``fake''). The Choice-Confidence
+(CHOCO) model conceptualizes the data as a mixture of two distributions,
+one for each choice, and models the probability of choosing one or the
+other as well as the degree of confidence in each choice. We validated
+this model on real-life data by appling it to the reality beliefs of
+Makowski et al. (\citeproc{ref-makowski2025too}{2025}), and showed that
+it was able to capture the bimodal nature of the data and provides a
+more accurate representation of the underlying processes than other
+alternative approaches, including Zero-or-One inflated Beta (ZOIB)
+models.
+
 Little difference between Beauty and Attractiveness, aside that Beauty
 revealed more effects in women. Why?
+
+We demonstrated the flexibility of the CHOCO model, being usable in
+analog scales as well as discrete ratings with a mid-point.
 
@@
 
+Future studies should assess psychometric property liek the minimum
+amount of data required to estimate all parameters reliably.
+
 Future studies should assess the ability of CHOCO models to also fit
@@
 We would like to thank the dissertation students from the University of
-Sussex for their help in data collection.
+Sussex for their help in data collection. DM would also like to thank
+\ldots{} for the motivation provided to write this paper.
 
@@
 \begin{CSLReferences}{1}{0}
+\bibitem[\citeproctext]{ref-azevedo2020body}
+Azevedo, R., Tucciarelli, R., De Beukelaer, S., Ambroziak, K., Jones,
+I., \& Tsakiris, M. (2020). \emph{A body of evidence:`feeling in
+seeing'predicts realness judgments for photojournalistic images.}
+
+\bibitem[\citeproctext]{ref-bozkir2024can}
+Bozkir, E., Riedmiller, C., Skodras, A. N., Kasneci, G., \& Kasneci, E.
+(2024). Can you tell real from fake face images? Perception of
+computer-generated faces by humans. \emph{ACM Transactions on Applied
+Perception}, \emph{22}(2), 1--23.
+
 \bibitem[\citeproctext]{ref-burkner2017brms}
@@
 
+\bibitem[\citeproctext]{ref-kukkonen2024beauty}
+Kukkonen, I., Pajunen, T., Sarpila, O., \& Åberg, E. (2024). Is
+beauty-based inequality gendered? A systematic review of gender
+differences in socioeconomic outcomes of physical attractiveness in
+labor markets. \emph{European Societies}, \emph{26}(1), 117--148.
+
+\bibitem[\citeproctext]{ref-lewandowsky2017beyond}
+Lewandowsky, S., Ecker, U. K., \& Cook, J. (2017). Beyond
+misinformation: Understanding and coping with the {``post-truth''} era.
+\emph{Journal of Applied Research in Memory and Cognition}, \emph{6}(4),
+353--369.
+
 \bibitem[\citeproctext]{ref-little2021facial}
@@
 
+\bibitem[\citeproctext]{ref-easystats2022}
+Lüdecke, D., Ben-Shachar, M. S., Patil, I., Wiernik, B. M., Bacher, E.,
+Thériault, R., \& Makowski, D. (2022). Easystats: Framework for easy
+statistical modeling, visualization, and reporting. \emph{CRAN}.
+\url{https://doi.org/10.32614/CRAN.package.easystats}
+
 \bibitem[\citeproctext]{ref-luo2019robust}
@@
 
+\bibitem[\citeproctext]{ref-makowski2018cognitive}
+Makowski, D. (2018). \emph{Cognitive neuropsychology of implicit emotion
+regulation through fictional reappraisal} {[}PhD thesis{]}. Sorbonne
+Paris Cit{é}.
+
 \bibitem[\citeproctext]{ref-makowski2019indices}
@@
 \emph{Frontiers in Psychology}, \emph{10}, 2767.
+
+\bibitem[\citeproctext]{ref-makowski2019phenomenal}
+Makowski, D., Sperduti, M., Pelletier, J., Blondé, P., La Corte, V.,
+Arcangeli, M., Zalla, T., Lemaire, S., Dokic, J., Nicolas, S., et al.
+(2019). Phenomenal, bodily and brain correlates of fictional reappraisal
+as an implicit emotion regulation strategy. \emph{Cognitive, Affective,
+\& Behavioral Neuroscience}, \emph{19}, 877--897.
 
@@
 
+\bibitem[\citeproctext]{ref-miller2023ai}
+Miller, E. J., Steward, B. A., Witkower, Z., Sutherland, C. A.,
+Krumhuber, E. G., \& Dawel, A. (2023). AI hyperrealism: Why AI faces are
+perceived as more real than human ones. \emph{Psychological Science},
+\emph{34}(12), 1390--1403.
+
 \bibitem[\citeproctext]{ref-monk2021beholding}
@@
 
+\bibitem[\citeproctext]{ref-nightingale2022ai}
+Nightingale, S. J., \& Farid, H. (2022). AI-synthesized faces are
+indistinguishable from real faces and more trustworthy.
+\emph{Proceedings of the National Academy of Sciences}, \emph{119}(8),
+e2120481119.
+
 \bibitem[\citeproctext]{ref-ospina2012general}
@@
 Analysis}, \emph{56}(6), 1609--1623.
+
+\bibitem[\citeproctext]{ref-pandey2021face}
+Pandey, G., \& Zayas, V. (2021). What is a face worth? Facial
+attractiveness biases experience-based monetary decision-making.
+\emph{British Journal of Psychology}, \emph{112}(4), 934--963.
+
+\bibitem[\citeproctext]{ref-peng2020faces}
+Peng, L., Cui, G., Chung, Y., \& Zheng, W. (2020). The faces of success:
+Beauty and ugliness premiums in e-commerce platforms. \emph{Journal of
+Marketing}, \emph{84}(4), 67--85.
 

### 2025-05-29 paper/manuscript.qmd
updated manuscript

@@
 
-There were also modifications to the instructions provided to participants. 
+There were also modifications to the instructions provided to participants and in the paradigm. 
 In Phase 1, an additional prompt was introduced, stating that the study was conducted in partnership with a young AI startup. 
 This was intended to strengthen participants’ belief that the faces they were shown could plausibly be either real or AI-generated.
-Additionally, participants were explicitly asked to respond based on their immediate, first impressions—an instruction not included in the first study.
+In this phase, a prompt appeared for 1–2 seconds before each face, informing participants whether the upcoming image was "AI-generated" or "Photograph", notably, all faces were in fact real photograph.
 

### 2025-05-29 paper/manuscript.tex
updated manuscript

@@
 \par{ }
-\par{     \begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, bottomrule=.15mm, colback=white, colframe=quarto-callout-note-color-frame, opacityback=0, breakable, arc=.35mm, leftrule=.75mm, toprule=.15mm, left=2mm]
+\par{     \begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, arc=.35mm, left=2mm, colback=white, rightrule=.15mm, bottomrule=.15mm, leftrule=.75mm, breakable, toprule=.15mm, opacityback=0]
 
@@
 There were also modifications to the instructions provided to
-participants. In Phase 1, an additional prompt was introduced, stating
-that the study was conducted in partnership with a young AI startup.
-This was intended to strengthen participants' belief that the faces they
-were shown could plausibly be either real or AI-generated. Additionally,
-participants were explicitly asked to respond based on their immediate,
-first impressions---an instruction not included in the first study.
+participants and in the paradigm. In Phase 1, an additional prompt was
+introduced, stating that the study was conducted in partnership with a
+young AI startup. This was intended to strengthen participants' belief
+that the faces they were shown could plausibly be either real or
+AI-generated. In this phase, a prompt appeared for 1--2 seconds before
+each face, informing participants whether the upcoming image was
+``AI-generated'' or ``Photograph'', notably, all faces were in fact real
+photograph.
 

### 2025-06-07 paper/manuscript.qmd
edits

--- 2025-05-23 paper/manuscript.qmd
@@
 
-**TODO**: Write some introduction about reality beliefs in the context of ambiguity.
-Underline that what drives judgments and beliefs of reality remains unclear.
-And this is important in today's post-truth era.
+In today's post-truth era, the proliferation of advanced AI technologies has made it increasingly challenging to distinguish between authentic and synthetic media, bearing significant implications for information integrity and public trust [@lewandowsky2017beyond].
+As traditional cues become less reliable (e.g., visual glitches and artefacts in generated images, formulaic generated text), people increasingly depend on contextual information and cognitive heuristics to assess authenticity, a process dubbed simulation monitoring [@makowski2019phenomenal].
 
-Another unknown is whether these beliefs are fully stimuli-driven or whether there are some stable individual tendencies that are at stake.
+This reliance is particularly pronounced under conditions of high ambiguity, where the decontextualization of information, especially prevalent in online environments, complicates authenticity assessments.
+An open question in this domain is the extent to which reality judgments are influenced by the stimuli themselves versus stable individual characteristics like personality, expectations or expertise - or transient psychophysiological states [@makowski2025too].
+
+Faces - socially and perceptually rich stimuli for which AI-generation has been particularly successful - have frequently been used to investigate reality judgments [@azevedo2020body; @makowski2025too; @tucciarelli2022realness; @nightingale2022ai].
+Studies asking participants to judge whether face images are real or artificially generated reveal that such judgments can be shaped by low-level features (e.g., clarity, symmetry), higher-level attributes (e.g., attractiveness, trustworthiness), and interindividual variability.
+In the present study, we apply the CHOCO model to data from such a paradigms to evaluate its capacity to recover interpretable parameters related to individual-level determinants of reality beliefs.
 
@@
 
-In the first phase, participants viewed 109 neutral-expression photographs of faces (random order, display time of 500 ms) from the American Multiracial Face Database [@chen2021broadening].
+In the first phase, participants viewed 109 neutral-expression photographs of faces (random order, display time of 500 ms) from the American Multiracial Face Database [AMFD; @chen2021broadening].
 After each image, participants rated the face on trustworthiness, familiarity, attractiveness, and beauty using visual analog scales.
 In the second phase, participants were informed that "about half of the previously seen images were AI-generated".
-The same faces were presented again in a new random order (same display time), followed by ratings of "reality" (whether they believed the image was AI-generated - left anchor - or a real photograph - right anchor).
+The same faces were presented again in a new random order (same display time), followed by ratings of "reality" (whether they believed the image was fake - left anchor - or real - right anchor).
 @makowski2025too showed that most stimuli yielded highly variable reality ratings, suggesting that idiosyncratic cognitive processes were at play.
@@
 For all models and each parameter, the full formula was entered: $Real\sim Sex + (1 | Participant) + (1 | Item)$ (with Sex as the main predictor and participants and items entered as random intercepts).
-The models were run using *brms* [@burkner2017brms] R package, and analyzed using the *easystats* collection of packages [**REFS**].
+The models were run using *brms* [@burkner2017brms] R package, and analyzed using the *easystats* collection of packages [@makowski2025modelbased; @ludecke2020extracting; @patil2022datawizard].
 To maximize the comparability across models We used the default priors (uniform) for all models, and we ran 16 chains of 1400 iterations each on the University of Sussex High-Performance Computing (HPC) cluster.
@@
 One such heuristic might be facial attractiveness.
-Attractiveness appraisals are known to be automatic and unconscious [@luo2019robust; @hung2016unconscious; @hou2023neural], and carry strong real-life consequences, as demonstrated by the large body of literature on the "beauty premium" [@little2021facial; @gulati2024beautiful, **MORE**]^[As an eloquent example, @monk2021beholding reported in a large representative US sample that the magnitude of earnings disparities among white women along the perceived attractiveness continuum exceeds in magnitude the canonical black-white race gap.]
+Attractiveness appraisals are known to be automatic and unconscious [@luo2019robust; @hung2016unconscious; @hou2023neural], and carry strong real-life consequences, as demonstrated by the large body of literature on the "beauty premium" [@little2021facial; @gulati2024beautiful;@pandey2021face; @kukkonen2024beauty]^[As an eloquent example, @monk2021beholding reported in a large representative US sample that the magnitude of earnings disparities among white women along the perceived attractiveness continuum exceeds in magnitude the canonical black-white race gap.]
 
-**TODO:** However, it is important to allow for the possibility of a U-shaped relationship.
-For instance, Peng et al.
-(2020) reported that both attractive and unattractive people sold more (underlining a "plainness penalty").
-<!-- The Faces of Success: Beauty and Ugliness Premiums in e-Commerce Platforms -->
+Although its role as a potential modulator of reality beliefs remains under-explored, @makowski2025too found significant associations between participants' realness ratings and facial attractiveness, with a potential sexual dymorphism: male participants judged attractive faces as being likely more real, whereas evidence suggested a midler and quadratic (U-shaped) relationship between attractiveness and reality beliefs.
+These findings offer a complementary perspective to those of @miller2023ai, who reported that participants used attractiveness as a distinguishing cue between real and AI-generated faces.
+These results highlight a possible bidirectional and context-dependent influence of attractiveness on reality judgments, of which the exact shape needs further investigation.
 
-Although its role as a potential modulator of reality beliefs remains underexplored, Makowski et al... **TODO: detail findings of that study + other studies**.
-
-In this study, we will apply the CHOCO model to the data of MAKOWSKI, and see whether different analysis approaches lead to different conclusions.
-We will then try to replicate these findings in a new sample of participants.
+<!-- However, it is important to allow for the possibility of a U-shaped relationship.
+For instance, @peng2020faces reported that both attractive and unattractive people sold more products (underlining a "plainness penalty").
+-->
 
@@
 
-For the second sample, the procedure was relatively similar with a few key differences.
-**TODO** Describe  fake Face 2. 
-The main difference is that the subjective ratings were done on a X-point Likert scale, rather than analog, with a mid-point value.
+For the second sample, the procedure was relatively similar to that of the first sample (described in Study 1) with a few key differences.
+
+The main difference was the introduction of an experimental manipulation: while in the first study, participants were simply informed of the presence of AI-generated or face pictures, the second sample's paradigm included (randomized) cues before each stimuli, indicating whether the upcoming image was "AI-generated" or "Photograph" (note that again, all images were in reality photographs).
+Specifically, at the beginning of the experiment, a cover story was presented, presenting the study as a partnership with an AI startup aimed at testing the quality of a new face AI-generation algorithm.
+Following that, they would see the 109 neutral-expression photographs from the AMFD database (1 s), each preceded by a textual cue indicating whether the image was "AI-generated" or "Photograph" (2 s).
+Ratings of attractiveness, beauty and trustworthiness were collected after each image.
+
+**Ana can you verify the timings?**
+
+This phase of the experiment would conclude with multi-choice questions asking participants to indicate whether they believed in the cover story.
+The second phase would start with a new set of instructions revealing (misleadingly)...
+
+<!-- Most importantly, subjective ratings (of reality and attractiveness) were collected using a 7-point Likert scale ranging from 0 to 6 (which included a clear midpoint option), rather than a visual analog scale.
+-->
+<!-- For this scale, the endpoints were labeled 'AI-generated' and 'Photograph', replacing the 'Fake' and 'Real' anchors used in Study 1. -->
+
+<!-- There were also modifications to the instructions provided to participants and in the paradigm.  -->
+
+<!-- This was intended to strengthen participants’ belief that the faces they were shown could plausibly be either real or AI-generated.
+In this phase, a prompt appeared for 1–2 seconds before each face, informing participants whether the upcoming image was "AI-generated" or "Photograph", notably, all faces were in fact real photograph.
+-->
+
+<!-- In Phase 2, participants were not informed about the proportion of AI-generated versus real faces encountered during Phase 1.  -->
+<!-- This omission was deliberate, aimed at reducing demand characteristics and preventing participants from mechanically selecting half the faces as real and half as AI-generated.  -->
+<!-- Instead, they were encouraged to base their responses solely on their level of confidence regarding the origin of each image. -->
 

### 2025-06-11 paper/manuscript.qmd
update

@@
 Despite significant advancements in psychological science following the replication crisis [@open2015estimating], its progress is still hindered by its sub-optimal (or inappropriate) usage of statistical tools [@makowski2023we].
-A prevalent issue is the continued reliance on linear models that assume normally distributed (Gaussian) data - as this assumption often does not hold true for many types of psychological outcomes. 
+A prevalent issue is the continued reliance on linear models that assume normally distributed (Gaussian) data^[More specifically, that the outcome is distributed according to a Normal distribution which parameters are expressed as a linear function of the predictors.] - as this assumption often does not hold true for many types of psychological outcomes.
 For instance, reaction times typically exhibit skewed distributions, choices can be represented as binary variables, and count data consists of strictly positive integers. 
@@
 #| apa-twocolumn: false  # A Figure Spanning Two Columns When in Journal Mode
-#| out-width: "50%"
+#| out-width: "100%"
 
@@
 Specifically, we will assess whether 1) CHOCO provides improved model fit, 2) yields deeper insights into population-level effects than traditional approaches (gender differences in reality beliefs), and 3) allows for the reliable estimation of interpretable individual-level parameters through random effects.
-
+**Study 2** will apply this model to more subtle effects, such as the effect perceived facial attractiveness on reality judgments, and test the ability to fit alternative data structures (scales with ordinal response options and mid-points).
 To this end, we analyze data from two separate studies in which participants judged whether a face image was AI-generated ("fake") or a real photograph.
-**Study 2** will apply this model to more subtle effects, such as the effect perceived facial attractiveness and beauty on reality judgments.
 
@@
 In today's post-truth era, the proliferation of advanced AI technologies has made it increasingly challenging to distinguish between authentic and synthetic media, bearing significant implications for information integrity and public trust [@lewandowsky2017beyond].
-As traditional cues become less reliable (e.g., visual glitches and artefacts in generated images, formulaic generated text), people increasingly depend on contextual information and cognitive heuristics to assess authenticity, a process dubbed simulation monitoring [@makowski2019phenomenal].
+As traditional cues become less reliable (e.g., visual glitches and artefacts in generated images; formulaic generated text, etc.), people increasingly depend on contextual information and cognitive heuristics to assess authenticity, a process referred to as "simulation monitoring" [@makowski2019phenomenal].
 
-This reliance is particularly pronounced under conditions of high ambiguity, where the decontextualization of information, especially prevalent in online environments, complicates authenticity assessments.
+This reliance on alternative epistemological sources is particularly pronounced under conditions of high ambiguity, where the decontextualization of information, especially prevalent in online environments, complicates authenticity assessments.
 An open question in this domain is the extent to which reality judgments are influenced by the stimuli themselves versus stable individual characteristics like personality, expectations or expertise - or transient psychophysiological states [@makowski2025too].
 
-Faces - socially and perceptually rich stimuli for which AI-generation has been particularly successful - have frequently been used to investigate reality judgments [@azevedo2020body; @makowski2025too; @tucciarelli2022realness; @nightingale2022ai].
+Images of faces - socially and perceptually rich stimuli for which AI-generation has been particularly successful - are a paradigmatic example that have been used to investigate reality judgments [@azevedo2020body; @makowski2025too; @tucciarelli2022realness; @nightingale2022ai].
 Studies asking participants to judge whether face images are real or artificially generated reveal that such judgments can be shaped by low-level features (e.g., clarity, symmetry), higher-level attributes (e.g., attractiveness, trustworthiness), and interindividual variability.
-In the present study, we apply the CHOCO model to data from such a paradigms to evaluate its capacity to recover interpretable parameters related to individual-level determinants of reality beliefs.
+In the present study, we apply the CHOCO model to such data to evaluate its capacity to recover interpretable parameters related to individual-level determinants of reality beliefs.
 
@@
 
-The sample is the same as in [@makowski2025too].
-We included all heterosexual and bisexual (these two groups did not seem to differ based on preliminary analyses and were thus grouped to maximize power) male and female participants, for a final sample of 141 participants (Mean age = 28.4, SD = 9.0, range: [19, 66]; Sex: 47.5% females).
+Using the open-accss data from @makowski2025too, we included all heterosexual and bisexual (as these two groups did not seem to differ based on preliminary analyses and were thus grouped to maximize power) male and female participants, for a final sample of 141 participants (Mean age = 28.4, SD = 9.0, range: [19, 66]; Sex: 47.5% females).
 For each participant, we included only stimuli of the opposite gender (i.e., all 89 female faces for men and 20 male faces for women).
@@
 
-In the first phase, participants viewed 109 neutral-expression photographs of faces (random order, display time of 500 ms) from the American Multiracial Face Database [AMFD; @chen2021broadening].
+In the first phase, participants viewed 109 neutral-expression photographs of faces (random order, display time of 3 s) from the American Multiracial Face Database [AMFD, @chen2021broadening].
 After each image, participants rated the face on trustworthiness, familiarity, attractiveness, and beauty using visual analog scales.
@@
 The same faces were presented again in a new random order (same display time), followed by ratings of "reality" (whether they believed the image was fake - left anchor - or real - right anchor).
-@makowski2025too showed that most stimuli yielded highly variable reality ratings, suggesting that idiosyncratic cognitive processes were at play.
+<!-- @makowski2025too showed that most stimuli yielded highly variable reality ratings, suggesting that idiosyncratic cognitive processes were at play. -->
 
@@
 
-Finally, we will run a correlation analysis of the models' individual-level estimates against "empirical" (indices computed directly on the observed data), including the empirical *p* ($P(y > 0.5)$), the overall *conf* ($mean(|y - 0.5|)$), *pex* ($P(y \in [0, 1])$) and *bex* ($P(y == 1) / P(y \in [0, 1])$), assessing whether the model's estimate are in-line with easily interpretable indices.
+Finally, we will run a correlation analysis of the models' individual-level estimates against "empirical" (indices computed directly on the observed data), including the empirical *p*, the overall *conf*, *pex* and *bex* (respectively calculated as $P(y > 0.5)$; $mean(|y - 0.5|)$; $P(y \in [0, 1])$; and $P(y == 1) / P(y \in [0, 1])$), assessing whether the model's estimate are in-line with easily interpretable indices.
 
@@
 #| label: fig-three
-#| fig-cap: "TODO."
+#| fig-cap: "Top: Model comparison revealed that the CHOCO model was a significantly better fit for the data (raw distribution in purple) compared to a ZOIB or Beta-Gate models, capturing its bimodal distribution.
+Bottom: the CHOCO model can be set to estimate effect on any of its parameters, such as the overall probability of responding on one side as well as the confidence in both choices.
+We illustrate this by showing the effect of Sex on the distribution of reality beliefs based on a CHOCO model."
 #| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
@@
 The difference in predictive accuracy, as indexed by Expected Log Predictive Density (ELPD-WAIC), suggests that the *CHOCO* is the best
-model ($ELPD = -203.54$), followed by *Beta-Gate* ($\Delta_{ELPD} = -1794.57 \pm 63.12, p < .001)$ and *ZOIB* ($\Delta_{ELPD} = -1833.59 \pm 63.52, p < .001$).
+model ($ELPD = -203.54$), followed by *Beta-Gate* ($\Delta_{ELPD} = -1794.57 \pm 63.12, ~p < .001)$ and *ZOIB* ($\Delta_{ELPD} = -1833.59 \pm 63.52, ~p < .001$).
 See @fig-three for the posterior predictive checks, showing that only the CHOCO model managed to capture the bimodal distribution of data.
@@
 
-The ZOIB model suggested that women had higher mean (*mu* parameter) scores of reality beliefs ($\Delta_{\text{Female}} = 0.20,~95\%~ CI~[0.03, 0.37], pd = 98.77\%$), less extreme values (*zoi* parameter; STATS) but more ones relative to zeros (*coi* parameter; STATS).
+The ZOIB model suggested that women had higher mean scores of reality beliefs ($\mu_{~\text{Female}} = 0.20,~95\%~ CI~[0.03, 0.37], ~pd = 98.77\%$), less extreme values ($zoi_{~\text{Female}} = -1.24,~95\%~ CI~[-2.46, -0.07], ~pd = 98.25\%$) but more ones relative to zeros ($coi_{~\text{Female}} = 1.63,~95\%~ CI~[0.70, 2.56], ~pd = 99.97\%$).
 
-The Beta-Gate model similarly suggested that women had higher mean (*mu* parameter) scores of reality beliefs (STATS), less extreme values (*pex* parameter; STATS), and a greater tendency to answer one relative to zero (*bex* parameter; STATS).
+The Beta-Gate model similarly suggested that women had higher mean scores of reality beliefs ($\mu_{~\text{Female}} = 0.21,~95\%~ CI~[0.02, 0.40], ~pd = 98.52\%$), less extreme values ($pex_{~\text{Female}} = -1.34,~95\%~ CI~[-2.56, -0.17], ~pd = 98.75\%$), and a greater tendency to answer one relative to zero ($bex_{~\text{Female}} = 1.13,~95\%~ CI~[0.34, 1.91], ~pd = 99.76\%$).
 
-The CHOCO model shows that women had a higher probability *p* of judging faces as real (STATS), but are not more confident when doing so (*confright* parameter; STATS).
-However, they were less confident when answering that an image was AI-generated (*confleft* parameter; STATS).
-There were also less likely to produce extreme answers (*pex* parameter; STATS), but no strong evidence supporting a directional bias was observed (*bex* parameter; STATS).
+The CHOCO model shows that women had a higher probability *p* of judging faces as real ($P_{~\text{Female}} = 0.45,~95\%~ CI~[0.05, 0.86], ~pd = 98.54\%$), but are not more confident when doing so ($confright_{~\text{Female}} = -0.13,~95\%~ CI~[-0.47, 0.21], ~pd = 77.73\%$).
+However, they were less confident when answering that an image was AI-generated ($confleft_{~\text{Female}} = -0.53,~95\%~ CI~[-0.89, -0.17], ~pd = 99.84\%$).
+There were also less likely to produce extreme answers ($pex_{~\text{Female}} = -1.19,~95\%~ CI~[-2.40, 0.01], ~pd = 97.64\%$), but no strong evidence supporting a directional bias was observed ($bex_{~\text{Female}} = 0.48,~95\%~ CI~[-0.12, 1.09], ~pd = 94.32\%$).
 
@@
 #| label: fig-four
-#| fig-cap: "TODO."
+#| fig-cap: "Top: the participant and item-level estimates for the CHOCO parameters, under the distribution of their point-estimates.
+Reliable effects are characterised by a higher between-group variability relative (the dispersion of the distribution of point-estimates) to its within-group variability (the average uncertainty of individual estimates represented by the error bar).
+For example, inter-individual variability in the parameters reflecting the confidence in left and right choices is reliably capture, as opposed to the inter-individual variability in the 'bex' parameter.
+Bottom: correlation matrix between participant-level estimates from different models and empirical indices (e.g., the raw proportion of responses on the right per participant).
+The CHOCO parameters are easily interpretable, strongly correlating with their empirical counterparts."
 #| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
-#| out-width: "100%"
+#| out-width: "95%"
 
@@
 
-The CHOCO model yielded reliable estimates for all parameters except *bex* for participants (*confright* D-vour = 0.94, *confleft* D-vour = 0.91, *pex* D-vour = 0.79, *p* D-vour = 0.73, *precright* D-vour = 0.77, *precleft* D-vour = 0.67).
+The CHOCO model yielded reliable estimates (@fig-four) for all parameters except *bex* for participants (*confright* D-vour = 0.94, *confleft* D-vour = 0.91, *pex* D-vour = 0.79, *p* D-vour = 0.73, *precright* D-vour = 0.77, *precleft* D-vour = 0.67).
 Item's variability was primarily reflected through the *p* parameter (D-vour = 0.86).
@@
 An interesting dissociation emerged between participant- and item-level variability: the latter seemed mostly to be represented in the *p* parameter, while participants reliably varied in most of the components (aside from *bex*).
-This could be interpreted as that external item characteristics primarily influence the probability of being judged as real vs.
+This could suggest that external item characteristics primarily influence the probability of being judged as real vs.
 fake, while the expressed confidence is first and foremost an individual characteristic.
@@
 One such heuristic might be facial attractiveness.
-Attractiveness appraisals are known to be automatic and unconscious [@luo2019robust; @hung2016unconscious; @hou2023neural], and carry strong real-life consequences, as demonstrated by the large body of literature on the "beauty premium" [@little2021facial; @gulati2024beautiful;@pandey2021face; @kukkonen2024beauty]^[As an eloquent example, @monk2021beholding reported in a large representative US sample that the magnitude of earnings disparities among white women along the perceived attractiveness continuum exceeds in magnitude the canonical black-white race gap.]
+Attractiveness appraisals are known to be automatic and unconscious [@luo2019robust; @hung2016unconscious; @hou2023neural], and carry strong real-life consequences, as demonstrated by the large body of literature on the "beauty premium"^[As an eloquent example, @monk2021beholding reported in a large representative US sample that the magnitude of earnings disparities among white women along the perceived attractiveness continuum exceeds in magnitude the canonical black-white race gap.] [@little2021facial; @gulati2024beautiful;@pandey2021face; @kukkonen2024beauty].
 
-Although its role as a potential modulator of reality beliefs remains under-explored, @makowski2025too found significant associations between participants' realness ratings and facial attractiveness, with a potential sexual dymorphism: male participants judged attractive faces as being likely more real, whereas evidence suggested a midler and quadratic (U-shaped) relationship between attractiveness and reality beliefs.
+Although its role as a potential modulator of reality beliefs remains under-explored, @makowski2025too found significant associations between participants' realness ratings and facial attractiveness, with a potential sexual dymorphism: male participants judged attractive faces as being likely more real, whereas evidence suggested a milder and quadratic (U-shaped) relationship between attractiveness and reality beliefs for females.
 These findings offer a complementary perspective to those of @miller2023ai, who reported that participants used attractiveness as a distinguishing cue between real and AI-generated faces.
@@
 
-The first sample includes the same participants as study 1 (see above).
-
-**TODO** for sample 2.
-
-The final sample included 189 participants (Mean age = 28.4, SD = 14.0, range: [18, 69]; Sex: 76.2% females).
+The first sample includes the same participants as study 1 (N = 141, see above). 
+For the second sample (data and preprocessing information available at https://github.com/RealityBending/FakeFace2), 248 participants were initially recruited through academic platforms (SONA and SurveySwap).
+We removed 15 participants with data suggesting low-effort responding as well as those (N=16) that did not believe in the experimental manipulation and were fully confident that all images were real or fake.
+<!-- 8 (3.23%) participants with no variation in their responses and 7 (2.92%) participants with no coherence (r < 0.1) b -->
+Similarly to the first sample, we included all hetero and bisexual participants and stimuli from the opposite gender, resulting in a final sample included 189 participants (Mean age = 28.4, SD = 14.0, range: [18, 69]; Sex: 76.2% females).
+The study was approved by the University of Sussex Ethics Board (ER/ST633/1).
 
@@
 
-For the second sample, the procedure was relatively similar to that of the first sample (described in Study 1) with a few key differences.
+For the second sample, the procedure was relatively similar to that of the first sample (described above in Study 1) with a few key differences.
 
-The main difference was the introduction of an experimental manipulation: while in the first study, participants were simply informed of the presence of AI-generated or face pictures, the second sample's paradigm included (randomized) cues before each stimuli, indicating whether the upcoming image was "AI-generated" or "Photograph" (note that again, all images were in reality photographs).
-Specifically, at the beginning of the experiment, a cover story was presented, presenting the study as a partnership with an AI startup aimed at testing the quality of a new face AI-generation algorithm.
-Following that, they would see the 109 neutral-expression photographs from the AMFD database (1 s), each preceded by a textual cue indicating whether the image was "AI-generated" or "Photograph" (2 s).
+The main difference was the introduction of an experimental manipulation: while for sample 1, participants were simply informed of the presence of AI-generated stimuli among photographs (not providing information as to specifically which image), the reality beliefs were directly manipulated in sample 2.
+At the beginning of the experiment, a cover story presented the study as a partnership with an AI startup aimed at testing the quality of a new face AI-generation algorithm.
+Following that, participants would see the 109 neutral-expression photographs from the AMFD database (1 s), each preceded by a randomly assigned textual cue indicating whether the image was "AI-generated" or "Photograph" (2 s).
 Ratings of attractiveness, beauty and trustworthiness were collected after each image.
 
-**Ana can you verify the timings?**
+This phase of the experiment concluded with multi-choice questions asking participants to indicate whether they believed in the cover story.
+The second phase would start with a new set of instructions (falsely) revealing that the cues were "mixed up" (shuffled randomly), and that they would now be presented with the faces again (1 s) followed by an assessment of their own beliefs about whether the image was real or fake.
 
-This phase of the experiment would conclude with multi-choice questions asking participants to indicate whether they believed in the cover story.
-The second phase would start with a new set of instructions revealing (misleadingly)...
-
-<!-- Most importantly, subjective ratings (of reality and attractiveness) were collected using a 7-point Likert scale ranging from 0 to 6 (which included a clear midpoint option), rather than a visual analog scale.
--->
+The second main difference was the subjective ratings' format, collected using a 7-point Likert scale ranging from 0 to 6 (which included a clear midpoint option), rather than a visual analog scale for sample 1.
+The 3 buttons on the left side (0, 1, 2) were colored in red and corresponded to the more-or-less pronounced belief that the image was "AI-generated", the 3 values on the right (4, 5, 6) for "Photograph" were colored in green, and the middle value (3), representing an undecided option, was colored in yellow.
 <!-- For this scale, the endpoints were labeled 'AI-generated' and 'Photograph', replacing the 'Fake' and 'Real' anchors used in Study 1. -->
-
-<!-- There were also modifications to the instructions provided to participants and in the paradigm.  -->
-
-<!-- This was intended to strengthen participants’ belief that the faces they were shown could plausibly be either real or AI-generated.
-In this phase, a prompt appeared for 1–2 seconds before each face, informing participants whether the upcoming image was "AI-generated" or "Photograph", notably, all faces were in fact real photograph.
--->
-
-<!-- In Phase 2, participants were not informed about the proportion of AI-generated versus real faces encountered during Phase 1.  -->
-<!-- This omission was deliberate, aimed at reducing demand characteristics and preventing participants from mechanically selecting half the faces as real and half as AI-generated.  -->
-<!-- Instead, they were encouraged to base their responses solely on their level of confidence regarding the origin of each image. -->
 
@@
 To compare the benefits of CHOCO models to a "traditional" analytic approach, we started by fitting a frequentist linear mixed model to predict reality beliefs with the formula $Real\sim Sex / poly(Attractive, 2) + (poly(Attractive, 2)|Participant) + (1|Item)$.
-The 2nd degree raw polynomial term was included to allow for potentially non-linear relationships (note that the first and second degree effects of *raw* polynomials can be interpreted independently as the linear part and the "curvy" part of the relationship).
+The second degree orthogonal polynomial term was included to allow for potentially non-linear relationships (note that the first and second degree effects of orthogonal polynomials can be interpreted independently as the linear part and the "curvy" part of the relationship).
 For the CHOCO model, mildly informative and effect-agnostic (i.e., centered at zero) priors were used.
-The same formula was used for all parameters, except that (based on the reliability analysis of Study 1), items were only included as random effects for *p*, and participants were not included for *bex*.
+The same formula was used for all parameters, except that items were only included as random effects for *p*, and participants were not included for *bex* (based on the reliability analysis of Study 1).
 
-**TODO**.
 For sample 2, the analysis was based on that of Sample 1.
-The main differences are 1) the inclusion of the "Condition" (whether the picture was presented as "Real" of "Fake" in the first phase of the experiment) as an additional predictor (it was entered as the only random slope for all participant random effects); 2) the inclusion of an additional parameters, *pmid*, modelling the probability of answering the middle-point of the Likert scale (representing an "undecided" option).
+The main differences are 1) the inclusion of the "Condition" (whether the picture was presented as "Real" of "Fake" in the first phase of the experiment) as an additional predictor (entered as the only random slope for all participant random effects); 2) the inclusion of an additional parameters, *pmid*, modelling the probability of answering the middle-point of the Likert scale (representing an "undecided" option); 3) as they were effectively only 3 distinct values on each side of the scale, and to showcase the flexibility of the CHOCO model, we decided to *not* treat the extreme values as 0 and 1 (and model them via the separate parameters *pex* and *bex*, which would distributions to be estimated from only 2 values), but instead to treat them as part of the continuous distribution.
+The 7 response options were rescaled to be evenly spaced between 0 and 1 excluded (i.e., [0.125, 0.875]).
+The *pex* and *bex* parameters were fixed to 0, making for a slightly more parsimonious model.
 
-The same analysis was applied for Beauty.
+<!-- The same analysis was applied for Beauty.  -->
 
@@
 
-In sample 1, the traditional approach suggested a significant linear relationship between the mean level of "Reality" and attractiveness ($\beta_{poly1}=3.42, ~95\%~CI [2.50, 4.34], p < .001\%$) for males only.
-The CHOCO model revealed that attractiveness had a significantly positive linear relationship with the probability *p* of judging faces as real (STATS), but a quadratic relationship with the confidence in real judgments (*confright*; STATS).
-Attractiveness was also associated with less confidence in fake judgments (*confleft*; STATS), a quadratic relationship with left precision (*precleft*; STATS), and related linearly with a stronger bias towards extreme Real responses (*bex*; STATS).
-These effects were present for males only, and no relationship was found for females.
+```{r}
+#| label: fig-five
+#| fig-cap: "The effect of attractiveness on reality beliefs.
+Top: the effect of different levels of attractiveness (shades of red) on the CHOCO distribution of the reality ratings in both samples, showing primarily that, for male participants, more attractive faces were judged more likely as a photograph rather than an AI-generated image.
+For the second sample, stimuli that were during the initial viewing presented as 'AI-generated' are represented via a dotted line.
+Bottom: The impact of attractiveness (x-axis) on different CHOCO distribution parameters, for male and female participants (the AI-generated condition is added as a transparent line for sample 2).
+Given the presence of polynomial terms, significant effects are denoted by a symbol representing the shape of the relationship (/ or \\\\ for linear, U or inverse-U for quadratic links)."
+#| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
+#| out-width: "100%"
+
+knitr::include_graphics("./figures/fig5.png")
+```
+
+In sample 1, the traditional approach suggested a significant linear relationship between the mean level of "Reality" and attractiveness for males only ($\beta_{~\text{poly1}}=3.42, ~95\%~CI [2.50, 4.34], ~p < .001$), the second largest effect being that of a quadratic link for women ($\beta_{~\text{poly2}}=1.82, ~95\%~CI [-0.27, 3.92], ~p = .09$).
+The CHOCO model revealed that, for males only, attractiveness had a significantly positive linear relationship with the probability *p* of judging faces as real ($P_{~\text{poly1}}=13.64, ~95\%~CI [8.42, 18.99], ~pd = 100\%$), but a quadratic relationship with the confidence in real judgments ($confright_{~\text{poly2}}=3.70, ~95\%~CI [0.80, 6.40], ~pd = 99.53\%$).
+Attractiveness was also associated with less confidence in fake judgments ($confleft_{~\text{poly1}}=-5.49, ~95\%~CI [-8.84, -2.11], ~pd =  99.92\%$), a quadratic relationship with left precision ($precleft_{~\text{poly2}}=-12.65	, ~95\%~CI [-24.59, -0.36], ~pd = 97.86\%$), and related linearly with a stronger bias towards extreme Real responses ($bex_{~\text{poly1}}=9.56	, ~95\%~CI [2.44, 16.73], ~pd = 99.55\%$).
+No significant relationship was found for females.
 The reliability of the effect of attractiveness (the random slopes) was very low for all parameters (D-vour < 0.01).
-The effect of Beauty was the same as of attractiveness.
+<!-- The effect of Beauty was the same as of attractiveness. -->
 
-<!-- , with some additional evidence in favour of an effect in women, in particular for a linear relationship with *p* for females as well ($\beta_{poly1}=7.02, ~95\%~CI [-0.90, 14.97], pd = 95.91\%$), as well as a quadratic relationship with the confidence in fake judgments (*confleft*; STATS).
+<!-- , with some additional evidence in favour of an effect in women, in particular for a linear relationship with *p* for females as well ($\beta_{poly1}=7.02, ~95\%~CI [-0.90, 14.97], ~pd = 95.91\%$), as well as a quadratic relationship with the confidence in fake judgments (*confleft*; STATS).
 -->
@@
 
-In Sample 2, the traditional approach suggested a significant linear relationship between the mean level of "Reality" and attractiveness for males (STATS) and women (STATS), with no effect of the Condition.
-The CHOCO model revealed that attractiveness had a significantly positive linear relationship with the probability *p* of judging faces as real for males (STATS) and females (STATS).
-It also had a linear relationship with the confidence in real judgments (*confright*) for males (STATS) and females (STATS), as well as a significant quadratic curvature for males only (STATS).
-Attractiveness also linearly decreased the confidence in fake judgments only for males (*confleft*; STATS), but increased the probability of an undecided response (*pmid*) in women (STATS), and displayed in females only a quadratic relationship with the amount of extreme answers (*pex*; STATS) and its bias in favour of right-side extreme response (*bex*; STATS).
-No consistent and reliable effect of condition was found.
+In Sample 2, the traditional approach suggested a significant linear relationship between the mean level of "Reality" and attractiveness for males ($\beta_{~\text{poly1}}=3.87, ~95\%~CI [2.91, 4.83], ~p < .001$) and females ($\beta_{~\text{poly1}}=1.88, ~95\%~CI [0.74, 3.01], ~p < .001$), with no effect of the Condition.
+The CHOCO model revealed that attractiveness had a significantly positive linear relationship with the probability *p* of judging faces as real for males ($P_{~\text{poly1}}=18.44, ~95\%~CI [11.36, 25.31], ~pd = 100\%$) and females ($P_{~\text{poly1}}=9.08, ~95\%~CI [1.90, 16.35], ~pd = 99.27\%$).
+It also had a linear relationship with the confidence in real judgments for males ($confright_{~\text{poly1}}=7.51, ~95\%~CI [3.90, 11.30], ~pd = 99.99\%$), as well as a significant quadratic relationship for females ($confright_{~\text{poly2}}=4.48, ~95\%~CI [0.74, 8.18], ~pd = 99.09\%$).
+Attractiveness also linearly decreased the confidence in fake judgments only for males ($confleft_{~\text{poly1}}=-6.67, ~95\%~CI [-10.04, -3.38], ~pd = 100\%$).
+Marginal contrasts suggested that stimuli previously labelled as photographs increased the probability *p* of judging faces as real, only for females ($P \Delta_{~\text{real - fake}}=0.06, ~95\%~CI [0.00, 0.11], ~pd = 98.36\%$).
 
-The effect of Beauty was...
+### Beauty
+
+The same analysis was run for Beauty, which indicated the following differences: 
+**ANA: can you report the main differences?**
 
 ## Discussion
+
+**WIP**
 
@@
 # General Discussion
+
+**WIP**
 
@@
 
-Future studies should assess psychometric property liek the minimum amount of data required to estimate all parameters reliably.
+While our study was pertaining to *simulation monitoring*, the process of assessing and forming beliefs about the nature of external stimuli (REF makowski phenomenological), we believe that it is part of the mechanisms at play in the fabric of our sense of reality (REF makowski thesis), alongside others, such as presence (the embodied feeling of being physically "in" an experience) or *reality monitoring* (the process of distinguishing between internally generated and externally perceived events, i.e., imagination vs.
+perception).
+Interestingly, recent studies about the latter support a possible distinction at a neural level between ...
+and (binary) decision making.
+(https://www.cell.com/neuron/fulltext/S0896-6273(25)00362-9).
+
+Future studies should assess psychometric property liek the minimum amount of data required to estimate and recover all parameters reliably.
 
@@
 We would like to thank the dissertation students from the University of Sussex for their help in data collection.
-DM would also like to thank ... for the motivation provided to write this paper.
+DM would also like to thank F. Rocher for inspiring some aspects of this paper.
 

### 2025-06-11 paper/manuscript.tex
update

--- 2025-05-28 paper/manuscript.tex
@@
 \par{ }
-\par{     \begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, bottomrule=.15mm, colback=white, colframe=quarto-callout-note-color-frame, opacityback=0, breakable, arc=.35mm, leftrule=.75mm, toprule=.15mm, left=2mm]
+\par{     \begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, rightrule=.15mm, bottomrule=.15mm, colback=white, arc=.35mm, breakable, toprule=.15mm, leftrule=.75mm, opacityback=0, left=2mm]
 
@@
 issue is the continued reliance on linear models that assume normally
-distributed (Gaussian) data - as this assumption often does not hold
-true for many types of psychological outcomes. For instance, reaction
-times typically exhibit skewed distributions, choices can be represented
-as binary variables, and count data consists of strictly positive
-integers. Applying models that presume normality and model the ``mean''
-of the outcome variable can lead to misinterpretation and potentially
-misleading conclusions when applied indiscriminately. It is thus
-important that psychologists use models that can best describe (or
-generate) the data they collect, to fully exploit them and bring more
-nuance and accuracy to their conclusions.
+distributed (Gaussian) data\footnote{More specifically, that the outcome
+  is distributed according to a Normal distribution which parameters are
+  expressed as a linear function of the predictors.} - as this
+assumption often does not hold true for many types of psychological
+outcomes. For instance, reaction times typically exhibit skewed
+distributions, choices can be represented as binary variables, and count
+data consists of strictly positive integers. Applying models that
+presume normality and model the ``mean'' of the outcome variable can
+lead to misinterpretation and potentially misleading conclusions when
+applied indiscriminately. It is thus important that psychologists use
+models that can best describe (or generate) the data they collect, to
+fully exploit them and bring more nuance and accuracy to their
+conclusions.
 
@@
 
-\includegraphics[width=0.5\linewidth,height=\textheight,keepaspectratio]{./figures/fig1.png}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{./figures/fig1.png}
 
@@
 choice and beta regression for confidence (see
-\citeproc{ref-makowski2025too}{Makowski et al., 2025} for an example).
-While this approach can provide additional insights into underlying
-mechanisms compared to a unique model, it assumes psychological and
-statistical independence between choice and confidence, which may not
-hold true in practice.
+\citeproc{ref-makowski2025too}{Makowski, Te, et al., 2025} for an
+example). While this approach can provide additional insights into
+underlying mechanisms compared to a unique model, it assumes
+psychological and statistical independence between choice and
+confidence, which may not hold true in practice.
 
@@
 
-\includegraphics[width=0.5\linewidth,height=\textheight,keepaspectratio]{./figures/fig2.png}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{./figures/fig2.png}
 
@@
 interpretable individual-level parameters through random effects.
-
-To this end, we analyze data from two separate studies in which
-participants judged whether a face image was AI-generated (``fake'') or
-a real photograph. \textbf{Study 2} will apply this model to more subtle
-effects, such as the effect perceived facial attractiveness and beauty
-on reality judgments.
+\textbf{Study 2} will apply this model to more subtle effects, such as
+the effect perceived facial attractiveness on reality judgments, and
+test the ability to fit alternative data structures (scales with ordinal
+response options and mid-points). To this end, we analyze data from two
+separate studies in which participants judged whether a face image was
+AI-generated (``fake'') or a real photograph.
 
@@
 
-\textbf{TODO}: Write some introduction about reality beliefs in the
-context of ambiguity. Underline that what drives judgments and beliefs
-of reality remains unclear. And this is important in today's post-truth
-era.
-
-In a post-truth era (\citeproc{ref-lewandowsky2017beyond}{Lewandowsky et
-al., 2017}), where near-perfect simulations are increasingly feasible
-due to rapid technological advancements, distinguishing between what is
-real and what is fake poses a growing challenge {[}\textbf{REF}{]}. As
-these technologies advance and physical cues become indistinguishable,
-individuals are likely to rely more heavily on alternative sources of
-information -such as contextual cues and cognitive heuristic -when
-making reality judgments {[}i.e., simulation monitoring; Makowski
-(\citeproc{ref-makowski2018cognitive}{2018}), Makowski, Sperduti, et al.
-(\citeproc{ref-makowski2019phenomenal}{2019}){]}. This reliance is
-particularly pronounced under conditions of high ambiguity, where
-processing demands are higher and the atomization and
-decontextualization of information (especially prevalent in online
-environments) further complicate the discernment of authenticity.
-
-Another unknown is whether these beliefs are fully stimuli-driven or
-whether there are some stable individual tendencies that are at stake.
-For instance, in the absence of clear contextual information,
-individual-level characteristics may increasingly guide beliefs about
-reality. Traits such as narcissism have been shown to influence such
-judgments (\citeproc{ref-makowski2025too}{Makowski et al., 2025}).
-Moreover, it has been proposed that the interaction between stimulus
-characteritics and individual differences plays a crucial role in
-shaping reality beliefs. For example, self-reported arousal has been
-found to predict the likelihood of perceiving an image as real
-(\citeproc{ref-azevedo2020body}{Azevedo et al., 2020}), while judging a
-stimulus as fake has been associated with emotional downregulation
-(\citeproc{ref-makowski2019phenomenal}{Makowski, Sperduti, et al.,
-2019}). {[}\textbf{should probably add other examples}{]}
-
-These findings underscore that what drives beliefs and judgments of
-reality remains unclear. Yet, understanding these mechanisms is
-increasingly vital given the proliferation of AI-generated content and
-the growing risk of misjudging the authenticity of stimuli, particularly
-in contexts with significant societal consequences (e.g., the influence
-of misinformation on democratic processes).
+In today's post-truth era, the proliferation of advanced AI technologies
+has made it increasingly challenging to distinguish between authentic
+and synthetic media, bearing significant implications for information
+integrity and public trust
+(\citeproc{ref-lewandowsky2017beyond}{Lewandowsky et al., 2017}). As
+traditional cues become less reliable (e.g., visual glitches and
+artefacts in generated images; formulaic generated text, etc.), people
+increasingly depend on contextual information and cognitive heuristics
+to assess authenticity, a process referred to as ``simulation
+monitoring'' (\citeproc{ref-makowski2019phenomenal}{Makowski, Sperduti,
+et al., 2019}).
+
+This reliance on alternative epistemological sources is particularly
+pronounced under conditions of high ambiguity, where the
+decontextualization of information, especially prevalent in online
+environments, complicates authenticity assessments. An open question in
+this domain is the extent to which reality judgments are influenced by
+the stimuli themselves versus stable individual characteristics like
+personality, expectations or expertise - or transient
+psychophysiological states (\citeproc{ref-makowski2025too}{Makowski, Te,
+et al., 2025}).
+
+Images of faces - socially and perceptually rich stimuli for which
+AI-generation has been particularly successful - are a paradigmatic
+example that have been used to investigate reality judgments
+(\citeproc{ref-azevedo2020body}{Azevedo et al., 2020};
+\citeproc{ref-makowski2025too}{Makowski, Te, et al., 2025};
+\citeproc{ref-nightingale2022ai}{Nightingale \& Farid, 2022};
+\citeproc{ref-tucciarelli2022realness}{Tucciarelli et al., 2022}).
+Studies asking participants to judge whether face images are real or
+artificially generated reveal that such judgments can be shaped by
+low-level features (e.g., clarity, symmetry), higher-level attributes
+(e.g., attractiveness, trustworthiness), and interindividual
+variability. In the present study, we apply the CHOCO model to such data
+to evaluate its capacity to recover interpretable parameters related to
+individual-level determinants of reality beliefs.
 
@@
 
-The sample is the same as in (\citeproc{ref-makowski2025too}{Makowski et
-al., 2025}). We included all heterosexual and bisexual (these two groups
-did not seem to differ based on preliminary analyses and were thus
-grouped to maximize power) male and female participants, for a final
-sample of 141 participants (Mean age = 28.4, SD = 9.0, range: {[}19,
-66{]}; Sex: 47.5\% females). For each participant, we included only
-stimuli of the opposite gender (i.e., all 89 female faces for men and 20
-male faces for women).
+Using the open-accss data from Makowski, Te, et al.
+(\citeproc{ref-makowski2025too}{2025}), we included all heterosexual and
+bisexual (as these two groups did not seem to differ based on
+preliminary analyses and were thus grouped to maximize power) male and
+female participants, for a final sample of 141 participants (Mean age =
+28.4, SD = 9.0, range: {[}19, 66{]}; Sex: 47.5\% females). For each
+participant, we included only stimuli of the opposite gender (i.e., all
+89 female faces for men and 20 male faces for women).
 
@@
 In the first phase, participants viewed 109 neutral-expression
-photographs of faces (random order, display time of 500 ms) from the
-American Multiracial Face Database
-(\citeproc{ref-chen2021broadening}{Chen et al., 2021}). After each
-image, participants rated the face on trustworthiness, familiarity,
+photographs of faces (random order, display time of 3 s) from the
+American Multiracial Face Database (AMFD,
+\citeproc{ref-chen2021broadening}{Chen et al., 2021}). After each image,
+participants rated the face on trustworthiness, familiarity,
 attractiveness, and beauty using visual analog scales. In the second
@@
 ``reality'' (whether they believed the image was fake - left anchor - or
-a real - right anchor). Makowski et al.
-(\citeproc{ref-makowski2025too}{2025}) showed that most stimuli yielded
-highly variable reality ratings, suggesting that idiosyncratic cognitive
-processes were at play.
+real - right anchor).
 
@@
 using the \emph{easystats} collection of packages
-(\citeproc{ref-easystats2022}{Lüdecke et al., 2022}). To maximize the
-comparability across models We used the default priors (uniform) for all
-models, and we ran 16 chains of 1400 iterations each on the University
-of Sussex High-Performance Computing (HPC) cluster.
+(\citeproc{ref-ludecke2020extracting}{Lüdecke et al., 2020};
+\citeproc{ref-makowski2025modelbased}{Makowski, Ben-Shachar, et al.,
+2025}; \citeproc{ref-patil2022datawizard}{Patil et al., 2022}). To
+maximize the comparability across models We used the default priors
+(uniform) for all models, and we ran 16 chains of 1400 iterations each
+on the University of Sussex High-Performance Computing (HPC) cluster.
 
@@
 individual-level estimates against ``empirical'' (indices computed
-directly on the observed data), including the empirical \emph{p}
-(\(P(y > 0.5)\)), the overall \emph{conf} (\(mean(|y - 0.5|)\)),
-\emph{pex} (\(P(y \in [0, 1])\)) and \emph{bex}
-(\(P(y == 1) / P(y \in [0, 1])\)), assessing whether the model's
-estimate are in-line with easily interpretable indices.
+directly on the observed data), including the empirical \emph{p}, the
+overall \emph{conf}, \emph{pex} and \emph{bex} (respectively calculated
+as \(P(y > 0.5)\); \(mean(|y - 0.5|)\); \(P(y \in [0, 1])\); and
+\(P(y == 1) / P(y \in [0, 1])\)), assessing whether the model's estimate
+are in-line with easily interpretable indices.
 
@@
 
-{\caption{{TODO.}{\label{fig-three}}}}
+{\caption{{Top: Model comparison revealed that the CHOCO model was a
+significantly better fit for the data (raw distribution in purple)
+compared to a ZOIB or Beta-Gate models, capturing its bimodal
+distribution. Bottom: the CHOCO model can be set to estimate effect on
+any of its parameters, such as the overall probability of responding on
+one side as well as the confidence in both choices. We illustrate this
+by showing the effect of Sex on the distribution of reality beliefs
+based on a CHOCO model.}{\label{fig-three}}}}
 
@@
 the \emph{CHOCO} is the best model (\(ELPD = -203.54\)), followed by
-\emph{Beta-Gate} (\(\Delta_{ELPD} = -1794.57 \pm 63.12, p < .001)\) and
-\emph{ZOIB} (\(\Delta_{ELPD} = -1833.59 \pm 63.52, p < .001\)). See
+\emph{Beta-Gate} (\(\Delta_{ELPD} = -1794.57 \pm 63.12, ~p < .001)\) and
+\emph{ZOIB} (\(\Delta_{ELPD} = -1833.59 \pm 63.52, ~p < .001\)). See
 Figure~\ref{fig-three} for the posterior predictive checks, showing that
@@
 
-The ZOIB model suggested that women had higher mean (\emph{mu}
-parameter) scores of reality beliefs
-(\(\Delta_{\text{Female}} = 0.20,~95\%~ CI~[0.03, 0.37], pd = 98.77\%\)),
-less extreme values (\emph{zoi} parameter; STATS) but more ones relative
-to zeros (\emph{coi} parameter; STATS).
+The ZOIB model suggested that women had higher mean scores of reality
+beliefs
+(\(\mu_{~\text{Female}} = 0.20,~95\%~ CI~[0.03, 0.37], ~pd = 98.77\%\)),
+less extreme values
+(\(zoi_{~\text{Female}} = -1.24,~95\%~ CI~[-2.46, -0.07], ~pd = 98.25\%\))
+but more ones relative to zeros
+(\(coi_{~\text{Female}} = 1.63,~95\%~ CI~[0.70, 2.56], ~pd = 99.97\%\)).
 
 The Beta-Gate model similarly suggested that women had higher mean
-(\emph{mu} parameter) scores of reality beliefs (STATS), less extreme
-values (\emph{pex} parameter; STATS), and a greater tendency to answer
-one relative to zero (\emph{bex} parameter; STATS).
+scores of reality beliefs
+(\(\mu_{~\text{Female}} = 0.21,~95\%~ CI~[0.02, 0.40], ~pd = 98.52\%\)),
+less extreme values
+(\(pex_{~\text{Female}} = -1.34,~95\%~ CI~[-2.56, -0.17], ~pd = 98.75\%\)),
+and a greater tendency to answer one relative to zero
+(\(bex_{~\text{Female}} = 1.13,~95\%~ CI~[0.34, 1.91], ~pd = 99.76\%\)).
 
 The CHOCO model shows that women had a higher probability \emph{p} of
-judging faces as real (STATS), but are not more confident when doing so
-(\emph{confright} parameter; STATS). However, they were less confident
-when answering that an image was AI-generated (\emph{confleft}
-parameter; STATS). There were also less likely to produce extreme
-answers (\emph{pex} parameter; STATS), but no strong evidence supporting
-a directional bias was observed (\emph{bex} parameter; STATS).
+judging faces as real
+(\(P_{~\text{Female}} = 0.45,~95\%~ CI~[0.05, 0.86], ~pd = 98.54\%\)),
+but are not more confident when doing so
+(\(confright_{~\text{Female}} = -0.13,~95\%~ CI~[-0.47, 0.21], ~pd = 77.73\%\)).
+However, they were less confident when answering that an image was
+AI-generated
+(\(confleft_{~\text{Female}} = -0.53,~95\%~ CI~[-0.89, -0.17], ~pd = 99.84\%\)).
+There were also less likely to produce extreme answers
+(\(pex_{~\text{Female}} = -1.19,~95\%~ CI~[-2.40, 0.01], ~pd = 97.64\%\)),
+but no strong evidence supporting a directional bias was observed
+(\(bex_{~\text{Female}} = 0.48,~95\%~ CI~[-0.12, 1.09], ~pd = 94.32\%\)).
 
@@
 
-{\caption{{TODO.}{\label{fig-four}}}}
-
-\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{./figures/fig4.png}
+{\caption{{Top: the participant and item-level estimates for the CHOCO
+parameters, under the distribution of their point-estimates. Reliable
+effects are characterised by a higher between-group variability relative
+(the dispersion of the distribution of point-estimates) to its
+within-group variability (the average uncertainty of individual
+estimates represented by the error bar). For example, inter-individual
+variability in the parameters reflecting the confidence in left and
+right choices is reliably capture, as opposed to the inter-individual
+variability in the `bex' parameter. Bottom: correlation matrix between
+participant-level estimates from different models and empirical indices
+(e.g., the raw proportion of responses on the right per participant).
+The CHOCO parameters are easily interpretable, strongly correlating with
+their empirical counterparts.}{\label{fig-four}}}}
+
+\includegraphics[width=0.95\linewidth,height=\textheight,keepaspectratio]{./figures/fig4.png}
 
@@
 
-The CHOCO model yielded reliable estimates for all parameters except
-\emph{bex} for participants (\emph{confright} D-vour = 0.94,
-\emph{confleft} D-vour = 0.91, \emph{pex} D-vour = 0.79, \emph{p} D-vour
-= 0.73, \emph{precright} D-vour = 0.77, \emph{precleft} D-vour = 0.67).
-Item's variability was primarily reflected through the \emph{p}
-parameter (D-vour = 0.86).
+The CHOCO model yielded reliable estimates (Figure~\ref{fig-four}) for
+all parameters except \emph{bex} for participants (\emph{confright}
+D-vour = 0.94, \emph{confleft} D-vour = 0.91, \emph{pex} D-vour = 0.79,
+\emph{p} D-vour = 0.73, \emph{precright} D-vour = 0.77, \emph{precleft}
+D-vour = 0.67). Item's variability was primarily reflected through the
+\emph{p} parameter (D-vour = 0.86).
 
@@
 participants reliably varied in most of the components (aside from
-\emph{bex}). This could be interpreted as that external item
-characteristics primarily influence the probability of being judged as
-real vs.~fake, while the expressed confidence is first and foremost an
-individual characteristic.
+\emph{bex}). This could suggest that external item characteristics
+primarily influence the probability of being judged as real vs.~fake,
+while the expressed confidence is first and foremost an individual
+characteristic.
 
@@
 real-life consequences, as demonstrated by the large body of literature
-on the ``beauty premium'' (\citeproc{ref-gulati2024beautiful}{Gulati et
-al., 2024}; \citeproc{ref-kukkonen2024beauty}{Kukkonen et al., 2024};
-\citeproc{ref-little2021facial}{Little, 2021};
-\citeproc{ref-pandey2021face}{Pandey \& Zayas, 2021})\footnote{As an
-  eloquent example, Monk Jr et al.
-  (\citeproc{ref-monk2021beholding}{2021}) reported in a large
+on the ``beauty premium''\footnote{As an eloquent example, Monk Jr et
+  al. (\citeproc{ref-monk2021beholding}{2021}) reported in a large
   representative US sample that the magnitude of earnings disparities
@@
   in magnitude the canonical black-white race gap.}
-
-\textbf{TODO:} However, it is important to allow for the possibility of
-a U-shaped relationship. For instance, Peng et al.
-(\citeproc{ref-peng2020faces}{2020}) reported that both attractive and
-unattractive people sold more products (underlining a ``plainness
-penalty'').
+(\citeproc{ref-gulati2024beautiful}{Gulati et al., 2024};
+\citeproc{ref-kukkonen2024beauty}{Kukkonen et al., 2024};
+\citeproc{ref-little2021facial}{Little, 2021};
+\citeproc{ref-pandey2021face}{Pandey \& Zayas, 2021}).
 
 Although its role as a potential modulator of reality beliefs remains
-underexplored, Makowski et al. (\citeproc{ref-makowski2025too}{2025})
-found significant associations between participants' realness ratings
-and facial salience. Female participants judged both highly attractive
-and highly unattractive faces as more likely to be real, whereas for
-male participants, this effect was limited to highly attractive faces.
-These findings offer a complementary perspective to those of Miller et
-al. (\citeproc{ref-miller2023ai}{2023}), who reported that AI-generated
-faces tend to be rated as more attractive, and that participants used
-attractiveness as a distinguishing cue between real and AI-generated
-faces. Together, these results highlight a possible bidirectional and
-context-dependent role of attractiveness in shaping reality judgments.
-
-In this study, we will apply the CHOCO model to the data of MAKOWSKI,
-and see whether different analysis approaches lead to different
-conclusions. We will then try to replicate these findings in a new
-sample of participants.
+under-explored, Makowski, Te, et al.
+(\citeproc{ref-makowski2025too}{2025}) found significant associations
+between participants' realness ratings and facial attractiveness, with a
+potential sexual dymorphism: male participants judged attractive faces
+as being likely more real, whereas evidence suggested a milder and
+quadratic (U-shaped) relationship between attractiveness and reality
+beliefs for females. These findings offer a complementary perspective to
+those of Miller et al. (\citeproc{ref-miller2023ai}{2023}), who reported
+that participants used attractiveness as a distinguishing cue between
+real and AI-generated faces. These results highlight a possible
+bidirectional and context-dependent influence of attractiveness on
+reality judgments, of which the exact shape needs further investigation.
 
@@
 
-The first sample includes the same participants as study 1 (see above).
-
-\textbf{TODO} for sample 2.
-
-The final sample included 189 participants (Mean age = 28.4, SD = 14.0,
-range: {[}18, 69{]}; Sex: 76.2\% females).
+The first sample includes the same participants as study 1 (N = 141, see
+above). For the second sample (data and preprocessing information
+available at https://github.com/RealityBending/FakeFace2), 248
+participants were initially recruited through academic platforms (SONA
+and SurveySwap). We removed 15 participants with data suggesting
+low-effort responding as well as those (N=16) that did not believe in
+the experimental manipulation and were fully confident that all images
+were real or fake. Similarly to the first sample, we included all hetero
+and bisexual participants and stimuli from the opposite gender,
+resulting in a final sample included 189 participants (Mean age = 28.4,
+SD = 14.0, range: {[}18, 69{]}; Sex: 76.2\% females). The study was
+approved by the University of Sussex Ethics Board (ER/ST633/1).
 
@@
 
-For the second sample, the procedure was relatively similar with a few
-key differences. Most importantly, subjective ratings were collected
-using a 7-point Likert scale ranging from 0 to 6, rather than a visual
-analog scale, which included a clear midpoint. Additionally, the
-familiarity rating was removed from the set of facial evaluations. In
-the second phase, realness judgments were also recorded on a 7-point
-Likert scale, this time ranging from -3 to +3, again with a defined
-midpoint, replacing the visual analog format used in the first study.
-For this scale, the endpoints were labeled `AI-generated' and
-`Photograph', replacing the `Fake' and `Real' anchors used in Study 1.
-
-There were also modifications to the instructions provided to
-participants. In Phase 1, an additional prompt was introduced, stating
-that the study was conducted in partnership with a young AI startup.
-This was intended to strengthen participants' belief that the faces they
-were shown could plausibly be either real or AI-generated. Additionally,
-participants were explicitly asked to respond based on their immediate,
-first impressions---an instruction not included in the first study.
-
-In Phase 2, participants were not informed about the proportion of
-AI-generated versus real faces encountered during Phase 1. This omission
-was deliberate, aimed at reducing demand characteristics and preventing
-participants from mechanically selecting half the faces as real and half
-as AI-generated. Instead, they were encouraged to base their responses
-solely on their level of confidence regarding the origin of each image.
+For the second sample, the procedure was relatively similar to that of
+the first sample (described above in Study 1) with a few key
+differences.
+
+The main difference was the introduction of an experimental
+manipulation: while for sample 1, participants were simply informed of
+the presence of AI-generated stimuli among photographs (not providing
+information as to specifically which image), the reality beliefs were
+directly manipulated in sample 2. At the beginning of the experiment, a
+cover story presented the study as a partnership with an AI startup
+aimed at testing the quality of a new face AI-generation algorithm.
+Following that, participants would see the 109 neutral-expression
+photographs from the AMFD database (1 s), each preceded by a randomly
+assigned textual cue indicating whether the image was ``AI-generated''
+or ``Photograph'' (2 s). Ratings of attractiveness, beauty and
+trustworthiness were collected after each image.
+
+This phase of the experiment concluded with multi-choice questions
+asking participants to indicate whether they believed in the cover
+story. The second phase would start with a new set of instructions
+(falsely) revealing that the cues were ``mixed up'' (shuffled randomly),
+and that they would now be presented with the faces again (1 s) followed
+by an assessment of their own beliefs about whether the image was real
+or fake.
+
+The second main difference was the subjective ratings' format, collected
+using a 7-point Likert scale ranging from 0 to 6 (which included a clear
+midpoint option), rather than a visual analog scale for sample 1. The 3
+buttons on the left side (0, 1, 2) were colored in red and corresponded
+to the more-or-less pronounced belief that the image was
+``AI-generated'', the 3 values on the right (4, 5, 6) for ``Photograph''
+were colored in green, and the middle value (3), representing an
+undecided option, was colored in yellow.
 
@@
 \(Real\sim Sex / poly(Attractive, 2) + (poly(Attractive, 2)|Participant) + (1|Item)\).
-The 2nd degree raw polynomial term was included to allow for potentially
-non-linear relationships (note that the first and second degree effects
-of \emph{raw} polynomials can be interpreted independently as the linear
-part and the ``curvy'' part of the relationship). For the CHOCO model,
-mildly informative and effect-agnostic (i.e., centered at zero) priors
-were used. The same formula was used for all parameters, except that
-(based on the reliability analysis of Study 1), items were only included
-as random effects for \emph{p}, and participants were not included for
-\emph{bex}.
-
-\textbf{TODO}. For sample 2, the analysis was based on that of Sample 1.
-The main differences are 1) the inclusion of the ``Condition'' (whether
-the picture was presented as ``Real'' of ``Fake'' in the first phase of
-the experiment) as an additional predictor (it was entered as the only
-random slope for all participant random effects); 2) the inclusion of an
-additional parameters, \emph{pmid}, modelling the probability of
-answering the middle-point of the Likert scale (representing an
-``undecided'' option).
-
-The same analysis was applied for Beauty.
+The second degree orthogonal polynomial term was included to allow for
+potentially non-linear relationships (note that the first and second
+degree effects of orthogonal polynomials can be interpreted
+independently as the linear part and the ``curvy'' part of the
+relationship). For the CHOCO model, mildly informative and
+effect-agnostic (i.e., centered at zero) priors were used. The same
+formula was used for all parameters, except that items were only
+included as random effects for \emph{p}, and participants were not
+included for \emph{bex} (based on the reliability analysis of Study 1).
+
+For sample 2, the analysis was based on that of Sample 1. The main
+differences are 1) the inclusion of the ``Condition'' (whether the
+picture was presented as ``Real'' of ``Fake'' in the first phase of the
+experiment) as an additional predictor (entered as the only random slope
+for all participant random effects); 2) the inclusion of an additional
+parameters, \emph{pmid}, modelling the probability of answering the
+middle-point of the Likert scale (representing an ``undecided'' option);
+3) as they were effectively only 3 distinct values on each side of the
+scale, and to showcase the flexibility of the CHOCO model, we decided to
+\emph{not} treat the extreme values as 0 and 1 (and model them via the
+separate parameters \emph{pex} and \emph{bex}, which would distributions
+to be estimated from only 2 values), but instead to treat them as part
+of the continuous distribution. The 7 response options were rescaled to
+be evenly spaced between 0 and 1 excluded (i.e., {[}0.125, 0.875{]}).
+The \emph{pex} and \emph{bex} parameters were fixed to 0, making for a
+slightly more parsimonious model.
 
@@
 
+\begin{figure*}[!htbp]
+
+{\caption{{The effect of attractiveness on reality beliefs. Top: the
+effect of different levels of attractiveness (shades of red) on the
+CHOCO distribution of the reality ratings in both samples, showing
+primarily that, for male participants, more attractive faces were judged
+more likely as a photograph rather than an AI-generated image. For the
+second sample, stimuli that were during the initial viewing presented as
+`AI-generated' are represented via a dotted line. Bottom: The impact of
+attractiveness (x-axis) on different CHOCO distribution parameters, for
+male and female participants (the AI-generated condition is added as a
+transparent line for sample 2). Given the presence of polynomial terms,
+significant effects are denoted by a symbol representing the shape of
+the relationship (/ or \textbackslash{} for linear, U or inverse-U for
+quadratic links).}{\label{fig-five}}}}
+
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{./figures/fig5.png}
+
+\end{figure*}
+
 In sample 1, the traditional approach suggested a significant linear
 relationship between the mean level of ``Reality'' and attractiveness
-(\(\beta_{poly1}=3.42, ~95\%~CI [2.50, 4.34], p < .001\%\)) for males
-only. The CHOCO model revealed that attractiveness had a significantly
-positive linear relationship with the probability \emph{p} of judging
-faces as real (STATS), but a quadratic relationship with the confidence
-in real judgments (\emph{confright}; STATS). Attractiveness was also
-associated with less confidence in fake judgments (\emph{confleft};
-STATS), a quadratic relationship with left precision (\emph{precleft};
-STATS), and related linearly with a stronger bias towards extreme Real
-responses (\emph{bex}; STATS). These effects were present for males
-only, and no relationship was found for females. The reliability of the
-effect of attractiveness (the random slopes) was very low for all
-parameters (D-vour \textless{} 0.01). The effect of Beauty was the same
-as of attractiveness.
+for males only
+(\(\beta_{~\text{poly1}}=3.42, ~95\%~CI [2.50, 4.34], ~p < .001\)), the
+second largest effect being that of a quadratic link for women
+(\(\beta_{~\text{poly2}}=1.82, ~95\%~CI [-0.27, 3.92], ~p = .09\)). The
+CHOCO model revealed that, for males only, attractiveness had a
+significantly positive linear relationship with the probability \emph{p}
+of judging faces as real
+(\(P_{~\text{poly1}}=13.64, ~95\%~CI [8.42, 18.99], ~pd = 100\%\)), but
+a quadratic relationship with the confidence in real judgments
+(\(confright_{~\text{poly2}}=3.70, ~95\%~CI [0.80, 6.40], ~pd = 99.53\%\)).
+Attractiveness was also associated with less confidence in fake
+judgments
+(\(confleft_{~\text{poly1}}=-5.49, ~95\%~CI [-8.84, -2.11], ~pd =  99.92\%\)),
+a quadratic relationship with left precision
+(\(precleft_{~\text{poly2}}=-12.65    , ~95\%~CI [-24.59, -0.36], ~pd = 97.86\%\)),
+and related linearly with a stronger bias towards extreme Real responses
+(\(bex_{~\text{poly1}}=9.56    , ~95\%~CI [2.44, 16.73], ~pd = 99.55\%\)).
+No significant relationship was found for females. The reliability of
+the effect of attractiveness (the random slopes) was very low for all
+parameters (D-vour \textless{} 0.01).
 
@@
 relationship between the mean level of ``Reality'' and attractiveness
-for males (STATS) and women (STATS), with no effect of the Condition.
-The CHOCO model revealed that attractiveness had a significantly
-positive linear relationship with the probability \emph{p} of judging
-faces as real for males (STATS) and females (STATS). It also had a
-linear relationship with the confidence in real judgments
-(\emph{confright}) for males (STATS) and females (STATS), as well as a
-significant quadratic curvature for males only (STATS). Attractiveness
-also linearly decreased the confidence in fake judgments only for males
-(\emph{confleft}; STATS), but increased the probability of an undecided
-response (\emph{pmid}) in women (STATS), and displayed in females only a
-quadratic relationship with the amount of extreme answers (\emph{pex};
-STATS) and its bias in favour of right-side extreme response
-(\emph{bex}; STATS). No consistent and reliable effect of condition was
-found.
-
-The effect of Beauty was\ldots{}
+for males
+(\(\beta_{~\text{poly1}}=3.87, ~95\%~CI [2.91, 4.83], ~p < .001\)) and
+females
+(\(\beta_{~\text{poly1}}=1.88, ~95\%~CI [0.74, 3.01], ~p < .001\)), with
+no effect of the Condition. The CHOCO model revealed that attractiveness
+had a significantly positive linear relationship with the probability
+\emph{p} of judging faces as real for males
+(\(P_{~\text{poly1}}=18.44, ~95\%~CI [11.36, 25.31], ~pd = 100\%\)) and
+females
+(\(P_{~\text{poly1}}=9.08, ~95\%~CI [1.90, 16.35], ~pd = 99.27\%\)). It
+also had a linear relationship with the confidence in real judgments for
+males
+(\(confright_{~\text{poly1}}=7.51, ~95\%~CI [3.90, 11.30], ~pd = 99.99\%\)),
+as well as a significant quadratic relationship for females
+(\(confright_{~\text{poly2}}=4.48, ~95\%~CI [0.74, 8.18], ~pd = 99.09\%\)).
+Attractiveness also linearly decreased the confidence in fake judgments
+only for males
+(\(confleft_{~\text{poly1}}=-6.67, ~95\%~CI [-10.04, -3.38], ~pd = 100\%\)).
+Marginal contrasts suggested that stimuli previously labelled as
+photographs increased the probability \emph{p} of judging faces as real,
+only for females
+(\(P \Delta_{~\text{real - fake}}=0.06, ~95\%~CI [0.00, 0.11], ~pd = 98.36\%\)).
+
+\subsubsection{Beauty}\label{beauty}
+
+The same analysis was run for Beauty, which indicated the following
+differences: \textbf{ANA: can you report the main differences?}
 
 \subsection{Discussion}\label{discussion-1}
+
+\textbf{WIP}
 
@@
 \section{General Discussion}\label{general-discussion}
+
+\textbf{WIP}
 
@@
 this model on real-life data by appling it to the reality beliefs of
-Makowski et al. (\citeproc{ref-makowski2025too}{2025}), and showed that
-it was able to capture the bimodal nature of the data and provides a
-more accurate representation of the underlying processes than other
+Makowski, Te, et al. (\citeproc{ref-makowski2025too}{2025}), and showed
+that it was able to capture the bimodal nature of the data and provides
+a more accurate representation of the underlying processes than other
 alternative approaches, including Zero-or-One inflated Beta (ZOIB)
@@
 
+While our study was pertaining to \emph{simulation monitoring}, the
+process of assessing and forming beliefs about the nature of external
+stimuli (REF makowski phenomenological), we believe that it is part of
+the mechanisms at play in the fabric of our sense of reality (REF
+makowski thesis), alongside others, such as presence (the embodied
+feeling of being physically ``in'' an experience) or \emph{reality
+monitoring} (the process of distinguishing between internally generated
+and externally perceived events, i.e., imagination vs.~perception).
+Interestingly, recent studies about the latter support a possible
+distinction at a neural level between \ldots{} and (binary) decision
+making. (https://www.cell.com/neuron/fulltext/S0896-6273(25)00362-9).
+
 Future studies should assess psychometric property liek the minimum
-amount of data required to estimate all parameters reliably.
+amount of data required to estimate and recover all parameters reliably.
 
@@
 We would like to thank the dissertation students from the University of
-Sussex for their help in data collection. DM would also like to thank
-\ldots{} for the motivation provided to write this paper.
+Sussex for their help in data collection. DM would also like to thank F.
+Rocher for inspiring some aspects of this paper.
 
@@
 
+\bibitem[\citeproctext]{ref-ludecke2020extracting}
+Lüdecke, D., Ben-Shachar, M. S., Patil, I., \& Makowski, D. (2020).
+Extracting, computing and exploring the parameters of statistical models
+using r. \emph{Journal of Open Source Software}, \emph{5}(53), 2445.
+
 \bibitem[\citeproctext]{ref-ludecke2021performance}
@@
 
-\bibitem[\citeproctext]{ref-easystats2022}
-Lüdecke, D., Ben-Shachar, M. S., Patil, I., Wiernik, B. M., Bacher, E.,
-Thériault, R., \& Makowski, D. (2022). Easystats: Framework for easy
-statistical modeling, visualization, and reporting. \emph{CRAN}.
-\url{https://doi.org/10.32614/CRAN.package.easystats}
-
 \bibitem[\citeproctext]{ref-luo2019robust}
@@
 
-\bibitem[\citeproctext]{ref-makowski2018cognitive}
-Makowski, D. (2018). \emph{Cognitive neuropsychology of implicit emotion
-regulation through fictional reappraisal} {[}PhD thesis{]}. Sorbonne
-Paris Cit{é}.
-
 \bibitem[\citeproctext]{ref-makowski2019indices}
@@
 \emph{Frontiers in Psychology}, \emph{10}, 2767.
+
+\bibitem[\citeproctext]{ref-makowski2025modelbased}
+Makowski, D., Ben-Shachar, M. S., Wiernik, B. M., Patil, I., Thériault,
+R., \& Lüdecke, D. (2025). Modelbased: An r package to make the most out
+of your statistical models through marginal means, marginal effects, and
+model predictions. \emph{Journal of Open Source Software},
+\emph{10}(109), 7969.
 
@@
 
-\bibitem[\citeproctext]{ref-peng2020faces}
-Peng, L., Cui, G., Chung, Y., \& Zheng, W. (2020). The faces of success:
-Beauty and ugliness premiums in e-commerce platforms. \emph{Journal of
-Marketing}, \emph{84}(4), 67--85.
+\bibitem[\citeproctext]{ref-patil2022datawizard}
+Patil, I., Makowski, D., Ben-Shachar, M. S., Wiernik, B. M., Bacher, E.,
+\& Lüdecke, D. (2022). Datawizard: An r package for easy data
+preparation and statistical transformations. \emph{Journal of Open
+Source Software}, \emph{7}(78), 4684.
 
@@
 
+\bibitem[\citeproctext]{ref-tucciarelli2022realness}
+Tucciarelli, R., Vehar, N., Chandaria, S., \& Tsakiris, M. (2022). On
+the realness of people who do not exist: The social processing of
+artificial faces. \emph{Iscience}, \emph{25}(12).
+
 \bibitem[\citeproctext]{ref-vehtari2017practical}

### 2025-06-11 functions/choco.R

@@ -1,242 +1,0 @@
-# Choice-Confidence (CHOCO) distribution to model bimodal analog/Likert scale data (with zeros and ones).
-# The idea is to have two separate ordered beta distributions (see Kubinec, 2023) modeling the left and the right hand side of the scale.
-
-# Regarding Ordered Beta distributions, see:
-# - Stan code: https://github.com/saudiwin/ordbetareg/blob/master/beta_logit.stan
-# - Python code: https://github.com/saudiwin/ordbetareg_py/blob/main/ordbetareg/model.py
-
-# Simulation ---------------------------------------------------------
-
-#' Random simulation from the CHOCO distribution:
-#' @param n Number of random draws.
-#' @param mu Probability of choosing the right side (relative to the left side).
-#' @param muleft The center of the left side beta distribution.
-#' @param phileft The shape parameter of the left side beta distribution.
-#' @param kleft The cutoff parameter for the left side beta distribution.
-#' @param muright The center of the right side beta distribution (overridden by mud if NULL).
-#' @param phiright The shape parameter of the right side beta distribution (overridden by phid if NULL).
-#' @param kright The cutoff parameter for the right side beta distribution (overridden by kd if NULL).
-#' @param mud Deviation for muright on the logit scale relative to muleft (used if muright is NULL).
-#' @param phid Deviation for phiright as a log-multiplier relative to phileft (used if phiright is NULL).
-#' @param kd Deviation for kright on the logit scale relative to kleft (used if kright is NULL).
-#'
-#' @examples
-#' hist(rchoco(3000, mu=0.5, muleft=0.5, phileft=3, kleft=0.95), breaks=100)
-#' hist(rchoco(3000, mu=0.6, muleft=0.5, phileft=5, kleft=0.99), breaks=100)
-#' hist(rchoco(3000, mu=0.4, muleft=0.5, phileft=3, kleft=0.85), breaks=100)
-rchoco <- function(n, mu = 0.5, muleft = 0.5, phileft = 3, kleft = 0.95,
-                   muright = NULL, phiright = NULL, kright = NULL,
-                   mud = 0, phid = 0, kd = 0) {
-  # Overall probabilities for choosing left or right side
[…]

### 2025-06-11 paper/manuscript.qmd
line number

@@
 keywords: "subjective scales, choice confidence model, reality beliefs, AI-generated faces, attractiveness"
-numbered-lines: false
+numbered-lines: true
 floatsintext: true
@@
 #| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
-#| out-width: "95%"
+#| out-width: "80%"
 

### 2025-06-11 paper/manuscript.tex
line number

@@
 
+\usepackage[nolongtablepatch]{lineno}
+\linenumbers
+
 \usepackage{newtx}
@@
 \par{ }
-\par{     \begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, rightrule=.15mm, bottomrule=.15mm, colback=white, arc=.35mm, breakable, toprule=.15mm, leftrule=.75mm, opacityback=0, left=2mm]
+\par{     \begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, leftrule=.75mm, breakable, rightrule=.15mm, bottomrule=.15mm, toprule=.15mm, left=2mm, arc=.35mm, opacityback=0, colback=white]
 
@@
 
+\resetlinenumber[1]
+
 \section{Introduction}\label{introduction}
@@
 
-\includegraphics[width=0.95\linewidth,height=\textheight,keepaspectratio]{./figures/fig4.png}
+\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{./figures/fig4.png}
 

### 2025-06-11 paper/manuscript.qmd
minor

@@
 In experimental paradigms, these can be somewhat disentangled by prompting participants to make a discrete choice selection (e.g., "True" vs. "False"), followed by a separate confidence rating. 
-However, this artificial separation makes its joint analysis difficulty, and may not reflect real-world decision-making, where individuals often express both choice and confidence simultaneously using a single, continuous scale.
-In such scales, each side can represent a distinct latent category, and the distance from the midpoint can indicates the level of confidence or certainty.
+However, this artificial separation makes its joint analysis difficult, and may not reflect real-world scenarios, where individuals often express both choice and confidence simultaneously using a single, continuous scale.
+In such scales, each side can represent a distinct latent category, and the distance from the midpoint can indicate the level of confidence or certainty.
 This integrated response format typically results in bimodal distributions, with peaks corresponding to the mean confidence on either side.
@@
 One alternative is to transform the data into two variables a posteriori: binarizing the side to represent choice and calculating the absolute distance from the midpoint to represent confidence.
-These can then be modeled separately, for instance, using logistic regression for choice and beta regression for confidence [see @makowski2025too for an example]. 
+These can then be modeled separately, for instance, using logistic regressions for choice and beta regressions for confidence [see @makowski2025too for an example]. 
 While this approach can provide additional insights into underlying mechanisms compared to a unique model, it assumes psychological and statistical independence between choice and confidence, which may not hold true in practice.

### 2025-06-11 paper/manuscript.tex
minor

@@
 \par{ }
-\par{     \begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, leftrule=.75mm, breakable, rightrule=.15mm, bottomrule=.15mm, toprule=.15mm, left=2mm, arc=.35mm, opacityback=0, colback=white]
+\par{     \begin{tcolorbox}[enhanced jigsaw, breakable, toprule=.15mm, leftrule=.75mm, bottomrule=.15mm, opacityback=0, arc=.35mm, colback=white, rightrule=.15mm, left=2mm, colframe=quarto-callout-note-color-frame]
 
@@
 vs.~``False''), followed by a separate confidence rating. However, this
-artificial separation makes its joint analysis difficulty, and may not
-reflect real-world decision-making, where individuals often express both
+artificial separation makes its joint analysis difficult, and may not
+reflect real-world scenarios, where individuals often express both
 choice and confidence simultaneously using a single, continuous scale.
 In such scales, each side can represent a distinct latent category, and
-the distance from the midpoint can indicates the level of confidence or
+the distance from the midpoint can indicate the level of confidence or
 certainty. This integrated response format typically results in bimodal
@@
 absolute distance from the midpoint to represent confidence. These can
-then be modeled separately, for instance, using logistic regression for
-choice and beta regression for confidence (see
+then be modeled separately, for instance, using logistic regressions for
+choice and beta regressions for confidence (see
 \citeproc{ref-makowski2025too}{Makowski, Te, et al., 2025} for an

### 2025-06-12 analysis/2_study2.qmd
update

@@
 
-- p: poly1 male
-- confright: poly2 male
-- confleft: poly1 male (neg)
-- bex: poly1 male
-- Suggests that (related to the original FakeFace1 hypothesis) reality would be driven by extremes degrees of attractiveness.
-
 #### Model Visualization
@@
 
-  p <- modelbased::estimate_prediction(m, data = newdata, predict = "mu", iterations = 50)
-  confright <- modelbased::estimate_prediction(m, data = newdata, predict = "confright", iterations = 50)
-  confleft <- modelbased::estimate_prediction(m, data = newdata, predict = "confleft", iterations = 50)
-  precright <- modelbased::estimate_prediction(m, data = newdata, predict = "precright", iterations = 50)
-  precleft <- modelbased::estimate_prediction(m, data = newdata, predict = "precleft", iterations = 50)
+  p <- modelbased::estimate_prediction(m, data = newdata, predict = "mu", iterations = 100)
+  confright <- modelbased::estimate_prediction(m, data = newdata, predict = "confright", iterations = 100)
+  confleft <- modelbased::estimate_prediction(m, data = newdata, predict = "confleft", iterations = 100)
+  precright <- modelbased::estimate_prediction(m, data = newdata, predict = "precright", iterations = 100)
+  precleft <- modelbased::estimate_prediction(m, data = newdata, predict = "precleft", iterations = 100)
   if ("pex" %in% names(insight::find_parameters(m))) {
-    pex <- modelbased::estimate_prediction(m, data = newdata, predict = "pex", iterations = 50)
+    pex <- modelbased::estimate_prediction(m, data = newdata, predict = "pex", iterations = 100)
   } else {
@@
   if ("bex" %in% names(insight::find_parameters(m))) {
-    bex <- modelbased::estimate_prediction(m, data = newdata, predict = "bex", iterations = 50)
+    bex <- modelbased::estimate_prediction(m, data = newdata, predict = "bex", iterations = 100)
   } else {
@@
   if ("pmid" %in% names(insight::find_parameters(m))) {
-    pmid <- modelbased::estimate_prediction(m, data = newdata, predict = "pmid", iterations = 50)
+    pmid <- modelbased::estimate_prediction(m, data = newdata, predict = "pmid", iterations = 100)
   } else {
@@
 
-pred <- visualize_dchoco(m1_chocoat,
+pred1 <- visualize_dchoco(m1_chocoat,
                           newdata=insight::get_datagrid(m1_chocoat, length = 5, preserve_range = FALSE))
 
-p3_density <- pred$densities |> 
+p3_density <- pred1$densities |> 
   mutate(Attractive = as.factor(Attractive),
@@
   geom_line(aes(color=Attractive), linewidth = 1.25) +
-  geom_path(data = pred$mu, aes(group = side), linewidth=0, lineend = 'round', 
-            arrow = arrow(length = unit(0.025, "npc"), type = "closed")) +
+  geom_path(data = pred1$mu, aes(group = side), linewidth=0.3, lineend = 'round', 
+            arrow = arrow(length = unit(0.03, "npc"), type = "closed")) +
   facet_wrap(~Sex) +
-  labs(y = "Distribution", x = "Do you think this image is real?", color = "Attractiveness") +
-  scale_colour_manual(values=scales::seq_gradient_pal("#FFCDD2", "#B71C1C", "Lab")(seq(0,1,length.out=5))) +
+  labs(y = "Distribution", x = "Do you think this image is real?", color = "Attractiveness",
+       title = "Effect on CHOCO distribution") +
+  scale_colour_manual(values=scales::seq_gradient_pal("#FFCDD2", "#B71C1C", "Lab")(seq(0, 1, length.out=5))) +
   scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.25, 0.5, 0.75, 1), labels = c("      0%", "25%", "50%", "75%", "100%      ")) +
-  scale_linewidth_continuous(range = c(0.5, 2), guide = "none") +
+  # scale_linewidth_continuous(range = c(0.5, 2), guide = "none") +
   theme_minimal() +
@@
         axis.text.y = element_blank(), 
-        # axis.title.y = element_blank(), 
+        plot.title = element_text(face = "bold"),
         strip.background = element_rect(fill = "lightgrey", color = NA)) 
@@
 colors <- c("p" = "#9C27B0", "confright" = "#795548", "confleft" = "#795548",
-            "precright" = "#607D8B","precleft" = "#607D8B", "pex" = "#FF9800", "bex" = "#F44336", "pmid" = "black")
+            "precright" = "#607D8B","precleft" = "#607D8B", "pex" = "#FF9800", "bex" = "#FF7811", "pmid" = "#4CAF50")
 
@@
     pred_params <- m |> 
-      modelbased::estimate_prediction(data = "grid", length = 20, preserve_range = FALSE, predict = param, iterations = 10) |>
+      modelbased::estimate_prediction(data = "grid", length = 40, preserve_range = FALSE, 
+                                      predict = param, iterations = 200) |>
       as.data.frame() |> 
@@
     mutate(Parameter = ifelse(Parameter == "mu", "p", Parameter),
-           Parameter = fct_relevel(Parameter, "p", "confright", "precright", "pex", "confleft", "precleft", "bex")) |> 
-    filter(Parameter != "bex") |>
+           Parameter = fct_relevel(Parameter, "p", "confright", "precright", "bex", "confleft", "precleft", "pex")) |> 
+    filter(Parameter != "pex") |>
     ggplot(aes(x = x, y = Predicted)) +
@@
     facet_wrap(~Parameter, scales = "free_y", ncol=3, strip.position = "left") +
-    labs(x = xlabel) +
+    labs(x = xlabel, title = "Effect on model parameters") +
     scale_fill_manual(values = colors, guide = "none") +
     scale_color_manual(values = colors, guide = "none") +
-    # scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c("      0%", "50%", "100%      ")) +
+    scale_x_continuous(breaks = c(0, 0.25, 0.5, 0.75, 1), labels = c(" 0", "0.25", "0.5", "0.75", "1 ")) +
     theme_minimal() +
@@
           axis.title.x = element_text(face = "bold"),
+          plot.title = element_text(face = "bold"),
           strip.text = element_text(size = rel(1), face = "italic"),
@@
 
-plot_params1(m1_chocoat, xlabel = "Attractiveness")
+p3_params <- plot_params1(m1_chocoat, xlabel = "Attractiveness")
+
+# Add significance
+sig <- data.frame(
+  Sex = c("Male", "Male", "Male", "Male", "Male"),
+  Label = c("/", "U", "\\", "∩", "/"),
+  Parameter = c("p", "confright", "confleft", "precleft", "bex"),
+  x = 0.5,
+  y = c(0.51, 0.6, 0.63, 3.75, 0.42)
+) |> 
+  mutate(Parameter =  fct_relevel(Parameter, "p", "confright", "bex", "confleft", "precleft"))
+
+p3_paramssig <- p3_params + 
+  geom_text(data = sig, aes(x = x, y = y, label = Label), 
+            size = 5, color = "black", hjust = 0.5) 
+p3_paramssig
 ```
@@
 
-```{r}
-#| cache-lazy: false
-
-plot_params1(m1_chocobeauty, xlabel = "Beauty")
-```
-
 ## Sample 2
@@
   geom_line(aes(color = Condition), linewidth = 1) +
-  geom_smooth(data=dfraw, aes(color = Condition, y = Real), method = "loess", se = FALSE, linetype = "dotted", formula = 'y ~ x') +
+  geom_smooth(data=df, aes(color = Condition, y = RealNoex), method = "loess", se = FALSE, linetype = "dotted", formula = 'y ~ x') +
   scale_color_manual(values = c("AI-Generated"="#3F51B5", "Photograph"="#FF5722")) +
@@
 
+Interestingly, Beta-gate appears as a better fit, but this is likely related to the nature of the index used, WAIC, which heavily penalizes for the number of parameters (which is *much* larger for CHOCO).
+Probably not a very meaningful result, especially since the posterior checks clearly show its misfit.
+
 ```{r}
@@
 m2_betagate <- readRDS("models/sample2noex_betagate.rds")
-m2_chocoat <- readRDS("models/sample2_chocoattractiveness.rds")
+# m2_chocoat <- readRDS("models/sample2_chocoattractiveness.rds")
 m2_chocoat <- readRDS("models/sample2noex_chocoattractiveness.rds")
@@
 for(param in c("mu", "confright", "confleft", "precright", "precleft", "pmid")) {
-  c <- estimate_contrasts(m2_chocoat, contrast = "Condition", by = "Sex", predict = param, backend = "emmeans") |>
+  c <- estimate_contrasts(m2_chocoat, contrast = "Condition", by = "Sex", predict = param, 
+                          backend = "emmeans", test = "pd") |>
     mutate(Parameter = ifelse(param == "mu", "p", param)) |> 
@@
 c
+```
+
+```{r}
+estimate_contrasts(m2_chocoat, contrast = "Condition", by = c("Attractive", "Sex"), 
+                   predict = "mu", length = 5, backend = "emmeans", test = "pd") 
 ```
@@
 ```{r}
-pred <- visualize_dchoco(m2_chocoat,
-                         newdata=insight::get_datagrid(m2_chocoat, length = 5, preserve_range = FALSE),
-                         x = seq(0, 1, length.out=9))
+pred2 <- visualize_dchoco(m2_chocoat,
+                          newdata=insight::get_datagrid(m2_chocoat, length = 5, preserve_range = FALSE))
 
-p4_density <- pred$densities |> 
-  mutate(Attractive = as.factor(Attractive),
-         y = ifelse(x %in% c(0, 1), y * 3, y),  # Inflate extremes for visibility
-         x = ifelse(Condition == "AI-Generated", x + 0.01, x)) |>  
-  filter(Condition == "Photograph") |>
+p4_density <- pred2$densities |> 
+  mutate(Attractive = as.factor(Attractive)) |> 
+         # y = ifelse(x %in% c(0, 1), y * 3, y),  # Inflate extremes for visibility
+         # x = ifelse(Condition == "AI-Generated", x + 0.01, x)) |>  
+  # filter(Condition == "Photograph") |>
   ggplot(aes(x=x, y=y)) +
-  geom_line(aes(color=Attractive, linetype = Condition), linewidth = 1.25) +
-  # geom_path(data = filter(pred$mu, Condition == "Photograph"), aes(group = side), linewidth=0, lineend = 'round', 
+  geom_line(aes(color=Attractive, linetype = Condition, linewidth = Condition)) +
+  # geom_path(data = filter(pred2$mu, Condition == "Photograph"), aes(group = side), linewidth=0, lineend = 'round', 
   #           arrow = arrow(length = unit(0.025, "npc"), type = "closed")) +
   facet_wrap(~Sex) +
-  labs(y = "Distribution", x = "Do you think this image is real?", color = "Attractiveness") +
+  labs(y = "Distribution", x = "Do you think this image is real?", color = "Attractiveness",
+       title = "Effect on CHOCO distribution") +
   scale_colour_manual(values=scales::seq_gradient_pal("#FFCDD2", "#B71C1C", "Lab")(seq(0,1,length.out=5))) +
   scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.25, 0.5, 0.75, 1), labels = c("      0%", "25%", "50%", "75%", "100%      ")) +
-  scale_linewidth_continuous(range = c(0.5, 2), guide = "none") +
+  scale_linetype_manual(values = c("Photograph" = "solid", "AI-Generated" = "dotted")) +
+  scale_linewidth_manual(values = c("Photograph" = 1.25, "AI-Generated" = 0.8)) +
   theme_minimal() +
@@
         axis.text.y = element_blank(),  
+        plot.title = element_text(face = "bold"),
         # axis.title.y = element_blank(), 
@@
 
+```{r}
+#| cache-lazy: false
+
+plot_params2 <- function(m, xlabel = "Attractiveness") {
+  pred_params <- data.frame()
+  for(param in c("mu", "confright", "confleft", "precright", "precleft", "pmid")) {
+    pred_params <- m |> 
+      modelbased::estimate_prediction(data = "grid", length = 40, preserve_range = FALSE, predict = param, iterations = 200) |>
+      as.data.frame() |> 
+      dplyr::mutate(Parameter = param) |> 
+      rbind(pred_params)
+  }
+
+  if(xlabel == "Attractiveness") {
+    pred_params$x <- pred_params$Attractive
+  } else {
+    pred_params$x <- pred_params$Beauty
+  }
+
+  pred_params |> 
+    mutate(Parameter = ifelse(Parameter == "mu", "p", Parameter),
+           Parameter = fct_relevel(Parameter, "p", "confright", "precright", "pmid", "confleft", "precleft")) |> 
+    ggplot(aes(x = x, y = Predicted)) +
+    geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = Parameter, 
+                    group = interaction(Sex, Condition), alpha = Condition)) +
+    geom_line(aes(color = Parameter, linetype = Sex, 
+                  alpha = interaction(Sex, Condition), group = interaction(Sex, Condition)), linewidth = 1) +
+    facet_wrap(~Parameter, scales = "free_y", ncol=3, strip.position = "left") +
+    labs(x = xlabel, title = "Effect on model parameters") +
+    scale_fill_manual(values = colors, guide = "none") +
+    scale_color_manual(values = colors, guide = "none") +
+    scale_alpha_manual(values = c("Photograph" = 0.2, "AI-Generated" = 0, 
+                                  "Male.Photograph" = 1, "Female.Photograph" = 1,
+                                  "Male.AI-Generated" = 0.2, "Female.AI-Generated"=0.2), guide = "none") +
+    scale_x_continuous(breaks = c(0, 0.25, 0.5, 0.75, 1), labels = c(" 0", "0.25", "0.5", "0.75", "1 ")) +
+    theme_minimal() +
+    theme(strip.placement = "outside",
+          axis.title.y = element_blank(),
+          axis.title.x = element_text(face = "bold"),
+          plot.title = element_text(face = "bold"),
+          strip.text = element_text(size = rel(1), face = "italic"),
+          panel.grid.minor.x = element_blank())
+}
+
+p4_params <- plot_params2(m2_chocoat, xlabel = "Attractiveness")
+
+# Add significance
+sig <- data.frame(
+  Sex = c("Male", "Male", "Female", "Male"),
+  Label = c("/", "/", "U", "\\"),
+  Parameter = c("p", "confright", "confright", "confleft"),
+  x = 0.5,
+  y = c(0.48, 0.425, 0.49, 0.48)
+) |> 
+  mutate(Parameter =  fct_relevel(Parameter, "p", "confright", "confleft"))
+
+p4_paramssig <- p4_params + 
+  geom_text(data = sig, aes(x = x, y = y, label = Label), 
+            size = 5, color = "black", hjust = 0.5) 
+p4_paramssig
+```
+
 ### Beauty
@@
 ```{r}
-m <- lme4::lmer(Real ~ Sex / poly(Beauty, 2) * Condition + (1|Participant) + (1|Item), data = df)
+m <- lme4::lmer(RealNoex ~ Sex / poly(Beauty, 2) * Condition + (1|Participant) + (1|Item), data = df)
 display(parameters::parameters(m, effects = "fixed"))
@@
 
-##### Effect of Attractiveness
+##### Effect of Beauty
 
@@
 
+##### Random Effects
+
+```{r}
+table_dvour(modelbased::estimate_grouplevel(m2_chocobeauty))
+```
+
+##### Full
+
+```{r}
+summary(m2_chocobeauty, robust = TRUE)
+```
+
 :::
+
+#### Model Visualization
+
+```{r}
+pred <- visualize_dchoco(m2_chocobeauty,
+                         newdata=insight::get_datagrid(m2_chocobeauty, length = 5, preserve_range = FALSE))
+
+p5_density <- pred$densities |> 
+  mutate(Beauty = as.factor(Beauty)) |> 
+         # y = ifelse(x %in% c(0, 1), y * 3, y),  # Inflate extremes for visibility
+         # x = ifelse(Condition == "AI-Generated", x + 0.01, x)) |>  
+  # filter(Condition == "Photograph") |>
+  ggplot(aes(x=x, y=y)) +
+  geom_line(aes(color=Beauty, linetype = Condition, linewidth = Condition)) +
+  # geom_path(data = filter(pred$mu, Condition == "Photograph"), aes(group = side), linewidth=0, lineend = 'round', 
+  #           arrow = arrow(length = unit(0.025, "npc"), type = "closed")) +
+  facet_wrap(~Sex) +
+  labs(y = "Distribution", x = "Do you think this image is real?", color = "Attractiveness",
+       title = "Effect on CHOCO distribution") +
+  scale_colour_manual(values=scales::seq_gradient_pal("#C5CAE9", "#283593", "Lab")(seq(0,1,length.out=5))) +
+  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.25, 0.5, 0.75, 1), labels = c("      0%", "25%", "50%", "75%", "100%      ")) +
+  scale_linetype_manual(values = c("Photograph" = "solid", "AI-Generated" = "dotted")) +
+  scale_linewidth_manual(values = c("Photograph" = 1.25, "AI-Generated" = 0.8)) +
+  theme_minimal() +
+  theme(axis.title.x = element_text(face = "bold"), 
+        axis.text.y = element_blank(),  
+        plot.title = element_text(face = "bold"),
+        # axis.title.y = element_blank(), 
+        strip.background = element_rect(fill = "lightgrey", color = NA)) 
+p5_density
+```
+
 ## Figures
@@
 ```{r}
-#| eval: false
+#| eval: true
+#| cache-lazy: false
 
-p3_a <- wrap_elements(p3_density / p3_params + 
-         patchwork::plot_annotation(title = "Effect of Attractiveness", theme = theme(plot.title = element_text(face = "bold"))))
+p3_a <- wrap_elements((p3_density + theme(legend.position = "none")) / 
+                        (p3_paramssig + theme(legend.position = "none")) + 
+         patchwork::plot_annotation(title = "Sample 1", 
+                                    theme = theme(plot.title = element_text(face = "bold.italic", hjust = 0.5))))
+
+p4_a <- wrap_elements((p4_density + theme(axis.title.y = element_blank())) / 
+                        p4_paramssig + 
+         patchwork::plot_annotation(title = "Sample 2", 
+                                    theme = theme(plot.title = element_text(face = "bold.italic", hjust = 0.5))))
+
+p5 <- (p3_a | p4_a) + 
+  plot_layout(guides = "collect", widths = c(0.45, 0.55))
+
+ggsave("../paper/figures/fig5.png", p5, width = 13, height = 11, dpi = 300)
 ```

### 2025-06-12 paper/manuscript.qmd
update

@@
 ---
-title: "**Introducing the Choice-Confidence (CHOCO) Model for Bimodal Scales Data: Application to the Relationship between Reality Beliefs about AI-Generated Faces and Beauty**"
+title: "**Introducing the Choice-Confidence (CHOCO) Model for Bimodal Data from Subjective Ratings: Application to the Effect of Attractiveness on Reality Beliefs about AI-Generated Faces**"
 shorttitle: "Choice-Confidence (CHOCO) Model"
@@
 keywords: "subjective scales, choice confidence model, reality beliefs, AI-generated faces, attractiveness"
+# Add more keywords if possible: beauty, Ordered Beta model, Bimodal data, decision-making
 numbered-lines: true
@@
 <!-- Problem statement -->
-Despite significant advancements in psychological science following the replication crisis [@open2015estimating], its progress is still hindered by its sub-optimal (or inappropriate) usage of statistical tools [@makowski2023we].
+Despite significant advancements in psychological science following the replication crisis [@open2015estimating], its progress is still hindered by its sub-optimal (or inappropriate) usage of statistical tools [@makowski2023we; @cumming2014new; @blanca2018current].
 A prevalent issue is the continued reliance on linear models that assume normally distributed (Gaussian) data^[More specifically, that the outcome is distributed according to a Normal distribution which parameters are expressed as a linear function of the predictors.] - as this assumption often does not hold true for many types of psychological outcomes.
@@
 
-Using the open-accss data from @makowski2025too, we included all heterosexual and bisexual (as these two groups did not seem to differ based on preliminary analyses and were thus grouped to maximize power) male and female participants, for a final sample of 141 participants (Mean age = 28.4, SD = 9.0, range: [19, 66]; Sex: 47.5% females).
+Using the open-access data from @makowski2025too, we included all heterosexual and bisexual (as these two groups did not seem to differ based on preliminary analyses and were thus grouped to maximize power) male and female participants, for a final sample of 141 participants (Mean age = 28.4, SD = 9.0, range: [19, 66]; Sex: 47.5% females).
 For each participant, we included only stimuli of the opposite gender (i.e., all 89 female faces for men and 20 male faces for women).
@@
 
+The findings of Study 2 replicate and extend those of Study 1, supporting the idea that facial attractiveness significantly shapes participants' beliefs about reality beliefs pertaining to human faces.
+This effect was particularly strong and robust among male participants.
+For females, the effect was weaker in Study 1 but more pronounced in Study 2, which might partly reflect increased statistical power.
+
+Another possibility, however, is that the second sample's experimental manipulation introduced conflicting cues (e.g., faces labelled "AI-generated") that could have disrupted intuitive heuristics, thereby increasing reliance on attractiveness as a diagnostic cue, particularly in ambiguous trials.
+<!-- This could be tested by analyzing believed vs. non-believed items -->
+
 The fact that the effect was less obvious in women in the sample 1 might be explained with power (less relevant items were included for women).
@@
 
+Also: https://www.cell.com/neuron/fulltext/S0896-6273(16)30016-2
+
 Future studies should assess psychometric property liek the minimum amount of data required to estimate and recover all parameters reliably.

### 2025-06-12 paper/manuscript.tex
update

@@
 \title{\textbf{Introducing the Choice-Confidence (CHOCO) Model for
-Bimodal Scales Data: Application to the Relationship between Reality
-Beliefs about AI-Generated Faces and Beauty}}
+Bimodal Data from Subjective Ratings: Application to the Effect of
+Attractiveness on Reality Beliefs about AI-Generated Faces}}
 
@@
 \par{ }
-\par{     \begin{tcolorbox}[enhanced jigsaw, breakable, toprule=.15mm, leftrule=.75mm, bottomrule=.15mm, opacityback=0, arc=.35mm, colback=white, rightrule=.15mm, left=2mm, colframe=quarto-callout-note-color-frame]
+\par{     \begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, opacityback=0, colback=white, bottomrule=.15mm, breakable, toprule=.15mm, rightrule=.15mm, arc=.35mm, left=2mm, leftrule=.75mm]
 
@@
 inappropriate) usage of statistical tools
-(\citeproc{ref-makowski2023we}{Makowski \& Waggoner, 2023}). A prevalent
+(\citeproc{ref-blanca2018current}{Blanca et al., 2018};
+\citeproc{ref-cumming2014new}{Cumming, 2014};
+\citeproc{ref-makowski2023we}{Makowski \& Waggoner, 2023}). A prevalent
 issue is the continued reliance on linear models that assume normally
@@
 
-Using the open-accss data from Makowski, Te, et al.
+Using the open-access data from Makowski, Te, et al.
 (\citeproc{ref-makowski2025too}{2025}), we included all heterosexual and
@@
 
+The findings of Study 2 replicate and extend those of Study 1,
+supporting the idea that facial attractiveness significantly shapes
+participants' beliefs about reality beliefs pertaining to human faces.
+This effect was particularly strong and robust among male participants.
+For females, the effect was weaker in Study 1 but more pronounced in
+Study 2, which might partly reflect increased statistical power.
+
+Another possibility, however, is that the second sample's experimental
+manipulation introduced conflicting cues (e.g., faces labelled
+``AI-generated'') that could have disrupted intuitive heuristics,
+thereby increasing reliance on attractiveness as a diagnostic cue,
+particularly in ambiguous trials.
+
 The fact that the effect was less obvious in women in the sample 1 might
@@
 
+Also: https://www.cell.com/neuron/fulltext/S0896-6273(16)30016-2
+
 Future studies should assess psychometric property liek the minimum
@@
 
+\bibitem[\citeproctext]{ref-blanca2018current}
+Blanca, M. J., Alarcón, R., \& Bono, R. (2018). Current practices in
+data analysis procedures in psychology: What has changed?
+\emph{Frontiers in Psychology}, \emph{9}, 2558.
+
 \bibitem[\citeproctext]{ref-bozkir2024can}
@@
 psychological science. \emph{Science}, \emph{349}(6251), aac4716.
+
+\bibitem[\citeproctext]{ref-cumming2014new}
+Cumming, G. (2014). The new statistics: Why and how. \emph{Psychological
+Science}, \emph{25}(1), 7--29.
 

### 2025-06-13 paper/manuscript.qmd
elements of discussion

@@
 The same analysis was run for Beauty, which indicated the following differences: 
-**ANA: can you report the main differences?**
+**ANA: can you check and write down the main differences?**
 
@@
 
-**WIP**
-
-The findings of Study 2 replicate and extend those of Study 1, supporting the idea that facial attractiveness significantly shapes participants' beliefs about reality beliefs pertaining to human faces.
-This effect was particularly strong and robust among male participants.
-For females, the effect was weaker in Study 1 but more pronounced in Study 2, which might partly reflect increased statistical power.
-
-Another possibility, however, is that the second sample's experimental manipulation introduced conflicting cues (e.g., faces labelled "AI-generated") that could have disrupted intuitive heuristics, thereby increasing reliance on attractiveness as a diagnostic cue, particularly in ambiguous trials.
+This study provides further evidence that attractive faces are more likely to be judged as real (*vs.* AI-generated), This effect was particularly strong and robust among male participants.
+For females, the effect was weaker in sample 1 but more pronounced in sample 2, which might partly reflect increased statistical power.
+An alternative possibility is that the second sample's experimental manipulation introduced conflicting cues (e.g., faces labelled "AI-generated") that could have disrupted intuitive heuristics, thereby increasing reliance on attractiveness as a diagnostic cue, particularly in ambiguous trials.
 <!-- This could be tested by analyzing believed vs. non-believed items -->
 
-The fact that the effect was less obvious in women in the sample 1 might be explained with power (less relevant items were included for women).
+<!-- Paragraph about potential mechanisms -->
+However, the mechanisms that underlie why attractiveness predicts reality judgments remains unclear.
+One possibility is that this positive link reflects a broader bias toward attributing positive qualities to attractive individuals, in-line with the substantial body of literature on the "beauty-is-good" stereotype, whereby attractive faces are judged as more trustworthy [@eagly1991beautiful], warmer and more approachable [@fiske2002model], and perceived as more authentic and sincere [@little2021facial].
+These traits may implicitly overlap with qualities ascribed to "real" humans.
+In this sense, reality judgments could be an extension of existing social heuristics that conflate visual appeal with genuineness.
+
+Alternatively, the effect might be primarily driven by attention.
+Attractive faces tend to capture and hold visual attention more effectively [@nakamura2014attractive], which could lead to a deeper processing.
+This greater accumulation of evidence could influence the bias systematically towards a particular category ("reality") or perhaps, towards the "true" category of the stimulus (which we cannot delineate from our data as all our stimuli were actually real photographs).
+Alternatively, such attentional bias may boost perceptual fluency and subjective certainty, that would be predominantly reflected in the confidence parameter of the CHOCO model.
+The attention hypothesis could be formally tested via incorporating a reality judgment task of real and actual AI-generated faces within attentional paradigms to see whether the confidence and bias changes as a function of attentional engagement.
+<!-- This could be tested via an attention (ANT) paradigm. -->
+
+Importantly for the scope of this study, the CHOCO model allowed us to dissect this effect in greater detail compared to traditional analyses, revealing that attractiveness not only shifted the choice toward "real" judgments but also modulated the confidence and extremity of those judgments.
+With the second sample, we demonstrated the flexibility of this model by fitting it to discrete ratings with a mid-point - providing potential insight into epistemic uncertainty in decision-making tasks.
+
+However, one might be concerned about the application of the CHOCO model (a mixture of two continuous Beta distributions) to ordinal data in Sample 2 (where a 7-point Likert scale was used).
+Ordinal ratings (limited number of discrete options represented numerically by integers) are commonly used in psychology, and often modeled as continuous.
+While this fact has spurred its own polarized debate (REFS), our position is that the model used should ideally reflect the data *generating* mechanism rather than necessarily focus on the *observed* data^[Ideally, the data recording method should be aligned with the assumed generating mechanism, which is an experiment design issue rather than a statistical one.] - although the former cannot always be easily inferred or known.
+For the data of Sample 2, we held the assumption that the underlying latent distribution was CHOCO distributed (as there is no reason to assume it would be different from Sample 1), treated as continuous CHOCO-distributed data, despite the fact that it was collected on a 7-point scale.
+The motivation was to see if the model could extrapolate and reliably estimate all the parameters based on a limited number of data points but means in our case that, effectively, each Beta distribution was extrapolated from the frequency of only three unique values.
+While this is far from ideal, the CHOCO model still proved insightful, although it might have impacted its ability detect finer-grain changes, including quadratic relationships.
+
+<!-- Interestingly, in Sample 2, prior labels (“AI-generated” vs.
+“Photograph”) had a modest but significant influence on beliefs—particularly for females—suggesting that top-down expectations can modulate reality perception.
+This effect was small but detectable, and illustrates how pre-existing contextual information can shape subjective reality beliefs, possibly interacting with heuristic features such as attractiveness.
+-->
 
@@
 
-We demonstrated the flexibility of the CHOCO model, being usable in analog scales as well as discrete ratings with a mid-point.
-
-Likert-scales (with a limited number of discrete options) are commonly used in psychology, and often modeled as continuous.
-While this fact has spurred its own polarized debate (REFS), our position is that the model used should ideally reflect the data *generating* mechanism rather than necessarily focus on the *observed* data^[Ideally, the data recording method should be aligned with the assumed generating mechanism, which is an experiment design issue rather than a statistical one.] - although the former cannot always be easily inferred or known.
-For the data of Sample 2, we held the assumption that the underlying latent distribution was CHOCO distributed (as there is no reason to assume it would be different from Sample 1), treated as continuous CHOCO-distributed data, despite the fact that it was collected on a 7-point Likert scale.
-The motivation was to see if the model could extrapolate and reliably estimate all the parameters based on a limited number of data points but means in our case that, effectively, each Beta distribution was extrapolated from the frequency of only two unique values (or 3 for the no-extreme version).
-While this is far from ideal, the CHOCO model still provided a better fit to the data than alternative approaches. 
-**TODO** However, it might have lead to a lesser ability to detect finer grain changes, including U-shaped relationships.
-Having 7-points (or 6 with no mid-value) is the strict minimum to use the CHOCO model.
-
 While our study was pertaining to *simulation monitoring*, the process of assessing and forming beliefs about the nature of external stimuli (REF makowski phenomenological), we believe that it is part of the mechanisms at play in the fabric of our sense of reality (REF makowski thesis), alongside others, such as presence (the embodied feeling of being physically "in" an experience) or *reality monitoring* (the process of distinguishing between internally generated and externally perceived events, i.e., imagination vs.
@@
 
-Future studies should assess psychometric property liek the minimum amount of data required to estimate and recover all parameters reliably.
+Future studies should assess psychometric property like the minimum amount of data required to estimate and recover all parameters reliably.
+As well as the minimum amount of unique value (for Likert scales) requires to detect weak or non-linearly shaped effects.
 

### 2025-06-13 paper/manuscript.tex
elements of discussion

@@
 \par{ }
-\par{     \begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, opacityback=0, colback=white, bottomrule=.15mm, breakable, toprule=.15mm, rightrule=.15mm, arc=.35mm, left=2mm, leftrule=.75mm]
+\par{     \begin{tcolorbox}[enhanced jigsaw, breakable, colback=white, opacityback=0, rightrule=.15mm, arc=.35mm, bottomrule=.15mm, left=2mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, leftrule=.75mm]
 
@@
 
-The findings of Study 2 replicate and extend those of Study 1,
-supporting the idea that facial attractiveness significantly shapes
-participants' beliefs about reality beliefs pertaining to human faces.
-This effect was particularly strong and robust among male participants.
-For females, the effect was weaker in Study 1 but more pronounced in
-Study 2, which might partly reflect increased statistical power.
-
-Another possibility, however, is that the second sample's experimental
-manipulation introduced conflicting cues (e.g., faces labelled
-``AI-generated'') that could have disrupted intuitive heuristics,
-thereby increasing reliance on attractiveness as a diagnostic cue,
-particularly in ambiguous trials.
-
-The fact that the effect was less obvious in women in the sample 1 might
-be explained with power (less relevant items were included for women).
+This study provides further evidence that attractive faces are more
+likely to be judged as real (vs.~AI-generated), This effect was
+particularly strong and robust among male participants. For females, the
+effect was weaker in sample 1 but more pronounced in sample 2, which
+might partly reflect increased statistical power. An alternative
+possibility is that the second sample's experimental manipulation
+introduced conflicting cues (e.g., faces labelled ``AI-generated'') that
+could have disrupted intuitive heuristics, thereby increasing reliance
+on attractiveness as a diagnostic cue, particularly in ambiguous trials.
+
+However, the mechanisms that underlie why attractiveness predicts
+reality judgments remains unclear. One possibility is that this positive
+link reflects a broader bias toward attributing positive qualities to
+attractive individuals, in-line with the substantial body of literature
+on the ``beauty-is-good'' stereotype, whereby attractive faces are
+judged as more trustworthy (\citeproc{ref-eagly1991beautiful}{Eagly et
+al., 1991}), warmer and more approachable
+(\citeproc{ref-fiske2002model}{Fiske et al., 2002}), and perceived as
+more authentic and sincere (\citeproc{ref-little2021facial}{Little,
+2021}). These traits may implicitly overlap with qualities ascribed to
+``real'' humans. In this sense, reality judgments could be an extension
+of existing social heuristics that conflate visual appeal with
+genuineness.
+
+Alternatively, the effect might be primarily driven by attention.
+Attractive faces tend to capture and hold visual attention more
+effectively (\citeproc{ref-nakamura2014attractive}{Nakamura \& Kawabata,
+2014}), which could lead to a deeper processing. This greater
+accumulation of evidence could influence the bias systematically towards
+a particular category (``reality'') or perhaps, towards the ``true''
+category of the stimulus (which we cannot delineate from our data as all
+our stimuli were actually real photographs). Alternatively, such
+attentional bias may boost perceptual fluency and subjective certainty,
+that would be predominantly reflected in the confidence parameter of the
+CHOCO model. The attention hypothesis could be formally tested via
+incorporating a reality judgment task of real and actual AI-generated
+faces within attentional paradigms to see whether the confidence and
+bias changes as a function of attentional engagement.
+
+Importantly for the scope of this study, the CHOCO model allowed us to
+dissect this effect in greater detail compared to traditional analyses,
+revealing that attractiveness not only shifted the choice toward
+``real'' judgments but also modulated the confidence and extremity of
+those judgments. With the second sample, we demonstrated the flexibility
+of this model by fitting it to discrete ratings with a mid-point -
+providing potential insight into epistemic uncertainty in
+decision-making tasks.
+
+However, one might be concerned about the application of a continuous
+distribution (the CHOCO mixture of two Beta distributions) to ordinal
+data (in Sample 2, where a 7-point Likert scale was used), as it
+translated effectively to only three unique values used to estimate a
+Beta distribution. Ordinal ratings (limited number of discrete options
+represented numerically by integers) are commonly used in psychology,
+and often modeled as continuous. While this fact has spurred its own
+polarized debate (REFS), our position is that the model used should
+ideally reflect the data \emph{generating} mechanism rather than
+necessarily focus on the \emph{observed} data\footnote{Ideally, the data
+  recording method should be aligned with the assumed generating
+  mechanism, which is an experiment design issue rather than a
+  statistical one.} - although the former cannot always be easily
+inferred or known. For the data of Sample 2, we held the assumption that
+the underlying latent distribution was CHOCO distributed (as there is no
+reason to assume it would be different from Sample 1), treated as
+continuous CHOCO-distributed data, despite the fact that it was
+collected on a 7-point scale. The motivation was to see if the model
+could extrapolate and reliably estimate all the parameters based on a
+limited number of data points but means in our case that, effectively,
+each Beta distribution was extrapolated from the frequency of only three
+unique values. While this is far from ideal, the CHOCO model still
+proved insightful, although it might have impacted its ability detect
+finer-grain changes, including quadratic relationships.
 
@@
 
-We demonstrated the flexibility of the CHOCO model, being usable in
-analog scales as well as discrete ratings with a mid-point.
-
-Likert-scales (with a limited number of discrete options) are commonly
-used in psychology, and often modeled as continuous. While this fact has
-spurred its own polarized debate (REFS), our position is that the model
-used should ideally reflect the data \emph{generating} mechanism rather
-than necessarily focus on the \emph{observed} data\footnote{Ideally, the
-  data recording method should be aligned with the assumed generating
-  mechanism, which is an experiment design issue rather than a
-  statistical one.} - although the former cannot always be easily
-inferred or known. For the data of Sample 2, we held the assumption that
-the underlying latent distribution was CHOCO distributed (as there is no
-reason to assume it would be different from Sample 1), treated as
-continuous CHOCO-distributed data, despite the fact that it was
-collected on a 7-point Likert scale. The motivation was to see if the
-model could extrapolate and reliably estimate all the parameters based
-on a limited number of data points but means in our case that,
-effectively, each Beta distribution was extrapolated from the frequency
-of only two unique values (or 3 for the no-extreme version). While this
-is far from ideal, the CHOCO model still provided a better fit to the
-data than alternative approaches. \textbf{TODO} However, it might have
-lead to a lesser ability to detect finer grain changes, including
-U-shaped relationships. Having 7-points (or 6 with no mid-value) is the
-strict minimum to use the CHOCO model.
-
 While our study was pertaining to \emph{simulation monitoring}, the
@@
 
-Future studies should assess psychometric property liek the minimum
+Future studies should assess psychometric property like the minimum
 amount of data required to estimate and recover all parameters reliably.
+As well as the minimum amount of unique value (for Likert scales)
+requires to detect weak or non-linearly shaped effects.
 
@@
 
+\bibitem[\citeproctext]{ref-eagly1991beautiful}
+Eagly, A. H., Ashmore, R. D., Makhijani, M. G., \& Longo, L. C. (1991).
+What is beautiful is good, but\ldots: A meta-analytic review of research
+on the physical attractiveness stereotype. \emph{Psychological
+Bulletin}, \emph{110}(1), 109.
+
+\bibitem[\citeproctext]{ref-fiske2002model}
+Fiske, S. T., Cuddy, A. J., Glick, P., \& Xu, J. (2002). A model of
+(often mixed) stereotype content: Competence and warmth respectively
+follow from perceived status and competition. \emph{Journal of
+Personality and Social Psychology}, \emph{82}(6), 878.
+
 \bibitem[\citeproctext]{ref-gulati2024beautiful}
@@
 
+\bibitem[\citeproctext]{ref-nakamura2014attractive}
+Nakamura, K., \& Kawabata, H. (2014). Attractive faces temporally
+modulate visual attention. \emph{Frontiers in Psychology}, \emph{5},
+620.
+
 \bibitem[\citeproctext]{ref-nightingale2022ai}

### 2025-06-14 paper/manuscript.qmd
oki

@@
 abstract: |
-  TO DO.
+  As AI-generated content becomes increasingly indistinguishable from real stimuli, understanding how individuals form beliefs about the reality of what they perceive is both theoretically and practically urgent.
+This study introduces the Choice-Confidence (CHOCO) model as a tool to dissect the cognitive mechanisms underlying decision-making judgments.
+Conceptualized as a mixture of two Beta distributions, CHOCO simultaneously estimates the probability of selecting one category (e.g., "real" vs.
+"AI-generated") and the confidence associated with each choice.
+We apply this model to two datasets (N=141 and N=189) in which participants judged whether faces were real photographs or AI-generated images, finding that facial attractiveness systematically increased the likelihood of being judged as "real", particularly for male participants.
+We discuss how these findings relate to the broader effect of facial attractiveness, and suggest future directions for both the cognitive and psychometric application of the CHOCO framework.
 #   **Significance Statement**. TODO.
-keywords: "subjective scales, choice confidence model, reality beliefs, AI-generated faces, attractiveness"
+keywords: "subjective scales, choice confidence model, reality beliefs, AI-generated faces, attractiveness, simulation monitoring"
 # Add more keywords if possible: beauty, Ordered Beta model, Bimodal data, decision-making
@@
 ---
+
+<!-- https://royalsocietypublishing.org/journal/rsos -->
+<!-- Scientific reports -->
+<!-- BRM -->
 
@@
 
-Decision-making is often conceptualized as involving distinct processes: the choice itself, and the confidence associated with that choice. 
-In experimental paradigms, these can be somewhat disentangled by prompting participants to make a discrete choice selection (e.g., "True" vs. "False"), followed by a separate confidence rating. 
+Decision-making is often conceptualized as involving distinct processes: the choice itself, and the confidence associated with that choice [@sanders2016signatures]. 
+In experimental paradigms, these can be somewhat disentangled by prompting participants to make a discrete choice selection (e.g., "True" *vs.* "False"), followed by a separate confidence rating. 
 However, this artificial separation makes its joint analysis difficult, and may not reflect real-world scenarios, where individuals often express both choice and confidence simultaneously using a single, continuous scale.
@@
 Study 1 revealed that the Choice-Confidence (CHOCO) model was a much better fit for bimodal bounded data, compared to other alternatives like the Zero-and-One Inflated Beta (ZOIB) and Beta-Gate (Ordered Beta) models.
-It also allowed for a deeper understanding through its interpretable parameters, offering insights into possibly distinct cognitive mechanisms, such as the probability of answering real vs.
-fake, and the associated confidence in these two choices.
+It also allowed for a deeper understanding through its interpretable parameters, offering insights into possibly distinct cognitive mechanisms, such as the probability of answering real *vs.* fake, and the associated confidence in these two choices.
 This was illustrated by modelling the effect of sex on all the CHOCO parameters.
@@
 An interesting dissociation emerged between participant- and item-level variability: the latter seemed mostly to be represented in the *p* parameter, while participants reliably varied in most of the components (aside from *bex*).
-This could suggest that external item characteristics primarily influence the probability of being judged as real vs.
-fake, while the expressed confidence is first and foremost an individual characteristic.
+This could suggest that external item characteristics primarily influence the probability of being judged as real *vs.* fake, while the expressed confidence is first and foremost an individual characteristic.
 
@@
 
+<!-- Difference with Makowski 2025? Explaiend by different model but also the fact that we only included Hetero participants with sexually-relevant items. -->
+
 <!-- Paragraph about potential mechanisms -->
@@
 Ordinal ratings (limited number of discrete options represented numerically by integers) are commonly used in psychology, and often modeled as continuous.
-While this fact has spurred its own polarized debate (REFS), our position is that the model used should ideally reflect the data *generating* mechanism rather than necessarily focus on the *observed* data^[Ideally, the data recording method should be aligned with the assumed generating mechanism, which is an experiment design issue rather than a statistical one.] - although the former cannot always be easily inferred or known.
-For the data of Sample 2, we held the assumption that the underlying latent distribution was CHOCO distributed (as there is no reason to assume it would be different from Sample 1), treated as continuous CHOCO-distributed data, despite the fact that it was collected on a 7-point scale.
+While this fact has spurred its own polarized debate [@norman2010likert; @sullivan2013analyzing; @liddell2018analyzing], our position is that the model used should ideally reflect the data *generating* mechanism rather than necessarily focus on the *observed* data^[Ideally, the data recording method should be aligned with the assumed generating mechanism, which is an experiment design issue rather than a statistical one.] - although the former cannot always be easily inferred or known.
+For the data of Sample 2, we held the assumption that the underlying latent distribution was CHOCO distributed (as there is no reason to assume it would be different from Sample 1), and thus treated as continuous CHOCO-distributed data, despite the fact that it was collected on a 7-point scale.
 The motivation was to see if the model could extrapolate and reliably estimate all the parameters based on a limited number of data points but means in our case that, effectively, each Beta distribution was extrapolated from the frequency of only three unique values.
@@
 
-**WIP**
+<!-- Summary of s tudy 1-->
+This study introduces a new statistical model to analyze bimodal data, which can be observed when using subjective rating scales in which the two sides correspond to a different category (e.g., "false" vs.
+"true", "real" vs "fake", "positive" vs.
+"negative).
+The Choice-Confidence (CHOCO) model conceptualizes the data as a mixture of two distributions, one for each choice, and models the probability of choosing one or the other as well as the degree of confidence in each choice.
+We validated this model on real-life data by applying it to the reality beliefs of @makowski2025too as well as to a new sample, showing that it captures the underlying cognitive processes more accurately than alternative models (e.g., linear or ZOIB models), and provides interpretable psychological parameters.
+These parameters, such as the choice probability and respective confidence, offer a cognitively meaningful decomposition of subjective decisions, applicable to a wide range of tasks.
 
-This study introduces a new statistical model to analyze bimodal data, which can be observed for instance when using subjective rating scales in which the two sides correspond to a different "choice" (e.g., "false" vs.
-"true", "real" vs "fake").
-The Choice-Confidence (CHOCO) model conceptualizes the data as a mixture of two distributions, one for each choice, and models the probability of choosing one or the other as well as the degree of confidence in each choice.
-We validated this model on real-life data by appling it to the reality beliefs of @makowski2025too, and showed that it was able to capture the bimodal nature of the data and provides a more accurate representation of the underlying processes than other alternative approaches, including Zero-or-One inflated Beta (ZOIB) models.
+<!-- Link with sense of reality -->
+Beyond the statistical contribution, the study also replicates and expands findings on the mechanisms that underpin simulation monitoring, or judgments pertaining to beliefs about the real *vs.* simulated (or synthetic) nature of external stimuli.
+This process, important for navigating and making sense of our environment, is likely contributing to our sense of reality [the feeling and belief of being real in a real environment, @makowski2018cognitive], together with other mechanisms such as presence (the embodied feeling of being physically "in" an experience) and reality monitoring (the process of distinguishing between internally generated and externally perceived events, i.e., imagination *vs.* perception).
 
-Little difference between Beauty and Attractiveness, aside that Beauty revealed more effects in women. Why?
+<!-- Possible Neuroscientific support -->
+Interestingly, recent studies about the latter support a possible distinction at a neural level between a graded "reality signal" strength, primarily tracked by the fusiform gyrus, and its categorical thresholding supporting the binary classification of reality *vs.* imagination, supported by a frontal network of brain regions, including dorsomedial prefrontal cortex and the anterior insula [@dijkstra2025neural].
+Although tangential, these findings might provide potential neuroscientific empirical grounding for the CHOCO's model usefulness in investigating dimensions of the sense of reality, opening the doors for future research exploring the neural correlates of the CHOCO parameters.
 
-While our study was pertaining to *simulation monitoring*, the process of assessing and forming beliefs about the nature of external stimuli (REF makowski phenomenological), we believe that it is part of the mechanisms at play in the fabric of our sense of reality (REF makowski thesis), alongside others, such as presence (the embodied feeling of being physically "in" an experience) or *reality monitoring* (the process of distinguishing between internally generated and externally perceived events, i.e., imagination vs.
-perception).
-Interestingly, recent studies about the latter support a possible distinction at a neural level between ...
-and (binary) decision making.
-(https://www.cell.com/neuron/fulltext/S0896-6273(25)00362-9).
+<!-- Attractiveness and faces -->
+The main empirical finding replicated in our study is that perceived facial attractiveness predicts the likelihood of believing that a face is real (*vs.* AI-generated).
+In both samples, males were more likely to identify highly attractive faces as real.
+The effect was weaker for females, possibly clouded by statistical power and/or the presence of other interacting moderators.
+Our work highlights how this key social heuristic not only biases trait perceptions (i.e., ratings of trustworthiness, competence, etc.) but also extends into other types of judgments, including reality beliefs.
 
-Also: https://www.cell.com/neuron/fulltext/S0896-6273(16)30016-2
+<!-- Limitations -->
+Despite these promising findings, several limitations must be acknowledged.
+First, the stimuli shown to male and female participants were not identical, each group rated faces of the opposite sex, limiting our ability to make direct sex-based comparisons and fully answer the potential sexual dimorphism existing in the formation of reality beliefs.
+Second, all faces presented were real photographs, which could raise concerns about the validity of the procedure.
+This however was a deliberate design feature: our goal was not to assess the detectability of current AI-generated images and study true discrimination abilities, but rather to study the cognitive mechanisms that shape reality judgments assuming the simulation is perceptually flawless.
+Nevertheless, future work could leverage generative AI to systematically manipulate facial attributes, such as attractiveness, while holding other features constant, offering greater experimental control and the ability to test causal hypotheses about the role of attractiveness for simulation monitoring.
 
-Future studies should assess psychometric property like the minimum amount of data required to estimate and recover all parameters reliably.
-As well as the minimum amount of unique value (for Likert scales) requires to detect weak or non-linearly shaped effects.
+<!-- Future directions -->
+While our application focused on facial judgments, the CHOCO model has broader potential.
+Many real-world rating tasks - whether about emotions, morality, authenticity, political veracity, or aesthetic appeal—produce bimodal or skewed response distributions that traditional models fail to capture well.
+CHOCO offers a flexible, interpretable approach that accommodates both the binary outcome of decision-making and its related continuous evaluations.
+However, some methodological and practical questions remain to be answered.
 
-Future studies should assess the ability of CHOCO models to also fit unimodal distributions (so it can be used more generally), as well as the benefits of more parsimonious parametrization (for instance, a unique precision parameter controlling both the left and right sides, or a linked confidence parameter e.g.
-where *confleft* is expressed as a function of *confright*).
+First, future studies should investigate the psychometric quality of the CHOCO model to derive practical guidelines for its optimal use.
+These include assessing the minimum amount of data required to reliably estimate and recover the different CHOCO parameters.
+When applying the model to ordinal data (such as Likert scales), another key question is the minimum number of unique response categories needed to detect subtle or nonlinear effects, such as quadratic modulation of confidence.
+Second, assessing the model’s robustness when applied to unimodal or skewed data would extend its utility beyond strictly bimodal contexts and support its general applicability.
+Third, more parsimonious variants of the model could be explored, such as versions with a single shared precision parameter for both sides or formulations where the confidence on one side is modeled as a function of the other.
+Such simplifications may retain interpretability while reducing parameter overhead and improving model stability.
+
+In conclusion, we developed and validated the CHOCO model, a theoretically grounded and practically flexible tool for modeling subjective judgments and delineating choice from confidence processes.
+It not only advances statistical modeling of decision-making data but opens new avenues for dissecting the cognitive mechanisms of belief, confidence, and reality perception.
+
+<!-- Little difference between Beauty and Attractiveness, aside that Beauty revealed more effects in women. Why? -->
 

### 2025-06-14 paper/manuscript.tex
oki

@@
 \usepackage{subcaption}
+
 \renewcommand\thesubfigure{\Alph{subfigure}}
@@
 
-\abstract{TO DO. }
+\abstract{As AI-generated content becomes increasingly indistinguishable
+from real stimuli, understanding how individuals form beliefs about the
+reality of what they perceive is both theoretically and practically
+urgent. This study introduces the Choice-Confidence (CHOCO) model as a
+tool to dissect the cognitive mechanisms underlying decision-making
+judgments. Conceptualized as a mixture of two Beta distributions, CHOCO
+simultaneously estimates the probability of selecting one category
+(e.g., ``real'' vs.~``AI-generated'') and the confidence associated with
+each choice. We apply this model to two datasets (N=141 and N=189) in
+which participants judged whether faces were real photographs or
+AI-generated images, finding that facial attractiveness systematically
+increased the likelihood of being judged as ``real'', particularly for
+male participants. We discuss how these findings relate to the broader
+effect of facial attractiveness, and suggest future directions for both
+the cognitive and psychometric application of the CHOCO framework. }
 
 \keywords{subjective scales, choice confidence model, reality beliefs,
-AI-generated faces, attractiveness}
+AI-generated faces, attractiveness, simulation monitoring}
 
@@
 \par{ }
-\par{     \begin{tcolorbox}[enhanced jigsaw, breakable, colback=white, opacityback=0, rightrule=.15mm, arc=.35mm, bottomrule=.15mm, left=2mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, leftrule=.75mm]
+\par{     \begin{tcolorbox}[enhanced jigsaw, breakable, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, left=2mm, rightrule=.15mm, arc=.35mm, opacityback=0, colback=white, leftrule=.75mm, toprule=.15mm]
 
@@
 \usepackage{pbalance}
-\usepackage{float}
+% \usepackage{float}
 \makeatletter
@@
 \onecolumn
-\begin{minipage}{0.5\textwidth}
+\begin{minipage}{0.485\textwidth}
 \oldtpt[#1]
@@
 \onecolumn
-\begin{minipage}{0.5\textwidth}
+\begin{minipage}{0.48\textwidth}
 \oldtpt
@@
 Decision-making is often conceptualized as involving distinct processes:
-the choice itself, and the confidence associated with that choice. In
+the choice itself, and the confidence associated with that choice
+(\citeproc{ref-sanders2016signatures}{Sanders et al., 2016}). In
 experimental paradigms, these can be somewhat disentangled by prompting
 participants to make a discrete choice selection (e.g., ``True''
-vs.~``False''), followed by a separate confidence rating. However, this
-artificial separation makes its joint analysis difficult, and may not
-reflect real-world scenarios, where individuals often express both
-choice and confidence simultaneously using a single, continuous scale.
-In such scales, each side can represent a distinct latent category, and
-the distance from the midpoint can indicate the level of confidence or
-certainty. This integrated response format typically results in bimodal
-distributions, with peaks corresponding to the mean confidence on either
-side. Traditional beta regression models, which assume unimodal
-distributions within the (0,1) interval, are ill-suited for such data.
-One alternative is to transform the data into two variables a
-posteriori: binarizing the side to represent choice and calculating the
-absolute distance from the midpoint to represent confidence. These can
-then be modeled separately, for instance, using logistic regressions for
-choice and beta regressions for confidence (see
+\emph{vs.} ``False''), followed by a separate confidence rating.
+However, this artificial separation makes its joint analysis difficult,
+and may not reflect real-world scenarios, where individuals often
+express both choice and confidence simultaneously using a single,
+continuous scale. In such scales, each side can represent a distinct
+latent category, and the distance from the midpoint can indicate the
+level of confidence or certainty. This integrated response format
+typically results in bimodal distributions, with peaks corresponding to
+the mean confidence on either side. Traditional beta regression models,
+which assume unimodal distributions within the (0,1) interval, are
+ill-suited for such data. One alternative is to transform the data into
+two variables a posteriori: binarizing the side to represent choice and
+calculating the absolute distance from the midpoint to represent
+confidence. These can then be modeled separately, for instance, using
+logistic regressions for choice and beta regressions for confidence (see
 \citeproc{ref-makowski2025too}{Makowski, Te, et al., 2025} for an
@@
 cognitive mechanisms, such as the probability of answering real
-vs.~fake, and the associated confidence in these two choices. This was
-illustrated by modelling the effect of sex on all the CHOCO parameters.
+\emph{vs.} fake, and the associated confidence in these two choices.
+This was illustrated by modelling the effect of sex on all the CHOCO
+parameters.
 
@@
 \emph{bex}). This could suggest that external item characteristics
-primarily influence the probability of being judged as real vs.~fake,
-while the expressed confidence is first and foremost an individual
+primarily influence the probability of being judged as real \emph{vs.}
+fake, while the expressed confidence is first and foremost an individual
 characteristic.
@@
 The same analysis was run for Beauty, which indicated the following
-differences: \textbf{ANA: can you report the main differences?}
+differences: \textbf{ANA: can you check and write down the main
+differences?}
 
@@
 
-\textbf{WIP}
-
 This study provides further evidence that attractive faces are more
-likely to be judged as real (vs.~AI-generated), This effect was
+likely to be judged as real (\emph{vs.} AI-generated), This effect was
 particularly strong and robust among male participants. For females, the
@@
 
-However, one might be concerned about the application of a continuous
-distribution (the CHOCO mixture of two Beta distributions) to ordinal
-data (in Sample 2, where a 7-point Likert scale was used), as it
-translated effectively to only three unique values used to estimate a
-Beta distribution. Ordinal ratings (limited number of discrete options
-represented numerically by integers) are commonly used in psychology,
-and often modeled as continuous. While this fact has spurred its own
-polarized debate (REFS), our position is that the model used should
-ideally reflect the data \emph{generating} mechanism rather than
-necessarily focus on the \emph{observed} data\footnote{Ideally, the data
-  recording method should be aligned with the assumed generating
-  mechanism, which is an experiment design issue rather than a
-  statistical one.} - although the former cannot always be easily
-inferred or known. For the data of Sample 2, we held the assumption that
-the underlying latent distribution was CHOCO distributed (as there is no
-reason to assume it would be different from Sample 1), treated as
-continuous CHOCO-distributed data, despite the fact that it was
-collected on a 7-point scale. The motivation was to see if the model
-could extrapolate and reliably estimate all the parameters based on a
-limited number of data points but means in our case that, effectively,
-each Beta distribution was extrapolated from the frequency of only three
-unique values. While this is far from ideal, the CHOCO model still
-proved insightful, although it might have impacted its ability detect
-finer-grain changes, including quadratic relationships.
+However, one might be concerned about the application of the CHOCO model
+(a mixture of two continuous Beta distributions) to ordinal data in
+Sample 2 (where a 7-point Likert scale was used). Ordinal ratings
+(limited number of discrete options represented numerically by integers)
+are commonly used in psychology, and often modeled as continuous. While
+this fact has spurred its own polarized debate
+(\citeproc{ref-liddell2018analyzing}{Liddell \& Kruschke, 2018};
+\citeproc{ref-norman2010likert}{Norman, 2010};
+\citeproc{ref-sullivan2013analyzing}{Sullivan \& Artino Jr, 2013}), our
+position is that the model used should ideally reflect the data
+\emph{generating} mechanism rather than necessarily focus on the
+\emph{observed} data\footnote{Ideally, the data recording method should
+  be aligned with the assumed generating mechanism, which is an
+  experiment design issue rather than a statistical one.} - although the
+former cannot always be easily inferred or known. For the data of Sample
+2, we held the assumption that the underlying latent distribution was
+CHOCO distributed (as there is no reason to assume it would be different
+from Sample 1), and thus treated as continuous CHOCO-distributed data,
+despite the fact that it was collected on a 7-point scale. The
+motivation was to see if the model could extrapolate and reliably
+estimate all the parameters based on a limited number of data points but
+means in our case that, effectively, each Beta distribution was
+extrapolated from the frequency of only three unique values. While this
+is far from ideal, the CHOCO model still proved insightful, although it
+might have impacted its ability detect finer-grain changes, including
+quadratic relationships.
 
@@
 
-\textbf{WIP}
-
 This study introduces a new statistical model to analyze bimodal data,
-which can be observed for instance when using subjective rating scales
-in which the two sides correspond to a different ``choice'' (e.g.,
-``false'' vs.~``true'', ``real'' vs ``fake''). The Choice-Confidence
-(CHOCO) model conceptualizes the data as a mixture of two distributions,
-one for each choice, and models the probability of choosing one or the
-other as well as the degree of confidence in each choice. We validated
-this model on real-life data by appling it to the reality beliefs of
-Makowski, Te, et al. (\citeproc{ref-makowski2025too}{2025}), and showed
-that it was able to capture the bimodal nature of the data and provides
-a more accurate representation of the underlying processes than other
-alternative approaches, including Zero-or-One inflated Beta (ZOIB)
-models.
-
-Little difference between Beauty and Attractiveness, aside that Beauty
-revealed more effects in women. Why?
-
-While our study was pertaining to \emph{simulation monitoring}, the
-process of assessing and forming beliefs about the nature of external
-stimuli (REF makowski phenomenological), we believe that it is part of
-the mechanisms at play in the fabric of our sense of reality (REF
-makowski thesis), alongside others, such as presence (the embodied
-feeling of being physically ``in'' an experience) or \emph{reality
-monitoring} (the process of distinguishing between internally generated
-and externally perceived events, i.e., imagination vs.~perception).
+which can be observed when using subjective rating scales in which the
+two sides correspond to a different category (e.g., ``false''
+vs.~``true'', ``real'' vs ``fake'', ``positive'' vs.~``negative). The
+Choice-Confidence (CHOCO) model conceptualizes the data as a mixture of
+two distributions, one for each choice, and models the probability of
+choosing one or the other as well as the degree of confidence in each
+choice. We validated this model on real-life data by applying it to the
+reality beliefs of Makowski, Te, et al.
+(\citeproc{ref-makowski2025too}{2025}) as well as to a new sample,
+showing that it captures the underlying cognitive processes more
+accurately than alternative models (e.g., linear or ZOIB models), and
+provides interpretable psychological parameters. These parameters, such
+as the choice probability and respective confidence, offer a cognitively
+meaningful decomposition of subjective decisions, applicable to a wide
+range of tasks.
+
+Beyond the statistical contribution, the study also replicates and
+expands findings on the mechanisms that underpin simulation monitoring,
+or judgments pertaining to beliefs about the real \emph{vs.} simulated
+(or synthetic) nature of external stimuli. This process, important for
+navigating and making sense of our environment, is likely contributing
+to our sense of reality (the feeling and belief of being real in a real
+environment, \citeproc{ref-makowski2018cognitive}{Makowski, 2018}),
+together with other mechanisms such as presence (the embodied feeling of
+being physically ``in'' an experience) and reality monitoring (the
+process of distinguishing between internally generated and externally
+perceived events, i.e., imagination \emph{vs.} perception).
+
 Interestingly, recent studies about the latter support a possible
-distinction at a neural level between \ldots{} and (binary) decision
-making. (https://www.cell.com/neuron/fulltext/S0896-6273(25)00362-9).
-
-Also: https://www.cell.com/neuron/fulltext/S0896-6273(16)30016-2
-
-Future studies should assess psychometric property like the minimum
-amount of data required to estimate and recover all parameters reliably.
-As well as the minimum amount of unique value (for Likert scales)
-requires to detect weak or non-linearly shaped effects.
-
-Future studies should assess the ability of CHOCO models to also fit
-unimodal distributions (so it can be used more generally), as well as
-the benefits of more parsimonious parametrization (for instance, a
-unique precision parameter controlling both the left and right sides, or
-a linked confidence parameter e.g.~where \emph{confleft} is expressed as
-a function of \emph{confright}).
+distinction at a neural level between a graded ``reality signal''
+strength, primarily tracked by the fusiform gyrus, and its categorical
+thresholding supporting the binary classification of reality \emph{vs.}
+imagination, supported by a frontal network of brain regions, including
+dorsomedial prefrontal cortex and the anterior insula
+(\citeproc{ref-dijkstra2025neural}{Dijkstra et al., 2025}). Although
+tangential, these findings might provide potential neuroscientific
+empirical grounding for the CHOCO's model usefulness in investigating
+dimensions of the sense of reality, opening the doors for future
+research exploring the neural correlates of the CHOCO parameters.
+
+The main empirical finding replicated in our study is that perceived
+facial attractiveness predicts the likelihood of believing that a face
+is real (\emph{vs.} AI-generated). In both samples, males were more
+likely to identify highly attractive faces as real. The effect was
+weaker for females, possibly clouded by statistical power and/or the
+presence of other interacting moderators. Our work highlights how this
+key social heuristic not only biases trait perceptions (i.e., ratings of
+trustworthiness, competence, etc.) but also extends into other types of
+judgments, including reality beliefs.
+
+Despite these promising findings, several limitations must be
+acknowledged. First, the stimuli shown to male and female participants
+were not identical, each group rated faces of the opposite sex, limiting
+our ability to make direct sex-based comparisons and fully answer the
+potential sexual dimorphism existing in the formation of reality
+beliefs. Second, all faces presented were real photographs, which could
+raise concerns about the validity of the procedure. This however was a
+deliberate design feature: our goal was not to assess the detectability
+of current AI-generated images and study true discrimination abilities,
+but rather to study the cognitive mechanisms that shape reality
+judgments assuming the simulation is perceptually flawless.
+Nevertheless, future work could leverage generative AI to systematically
+manipulate facial attributes, such as attractiveness, while holding
+other features constant, offering greater experimental control and the
+ability to test causal hypotheses about the role of attractiveness for
+simulation monitoring.
+
+While our application focused on facial judgments, the CHOCO model has
+broader potential. Many real-world rating tasks - whether about
+emotions, morality, authenticity, political veracity, or aesthetic
+appeal---produce bimodal or skewed response distributions that
+traditional models fail to capture well. CHOCO offers a flexible,
+interpretable approach that accommodates both the binary outcome of
+decision-making and its related continuous evaluations. However, some
+methodological and practical questions remain to be answered.
+
+First, future studies should investigate the psychometric quality of the
+CHOCO model to derive practical guidelines for its optimal use. These
+include assessing the minimum amount of data required to reliably
+estimate and recover the different CHOCO parameters. When applying the
+model to ordinal data (such as Likert scales), another key question is
+the minimum number of unique response categories needed to detect subtle
+or nonlinear effects, such as quadratic modulation of confidence.
+Second, assessing the model's robustness when applied to unimodal or
+skewed data would extend its utility beyond strictly bimodal contexts
+and support its general applicability. Third, more parsimonious variants
+of the model could be explored, such as versions with a single shared
+precision parameter for both sides or formulations where the confidence
+on one side is modeled as a function of the other. Such simplifications
+may retain interpretability while reducing parameter overhead and
+improving model stability.
+
+In conclusion, we developed and validated the CHOCO model, a
+theoretically grounded and practically flexible tool for modeling
+subjective judgments and delineating choice from confidence processes.
+It not only advances statistical modeling of decision-making data but
+opens new avenues for dissecting the cognitive mechanisms of belief,
+confidence, and reality perception.
 
@@
 
+\bibitem[\citeproctext]{ref-dijkstra2025neural}
+Dijkstra, N., Rein, T. von, Kok, P., \& Fleming, S. M. (2025). A neural
+basis for distinguishing imagination from reality. \emph{Neuron}.
+
 \bibitem[\citeproctext]{ref-eagly1991beautiful}
@@
 
+\bibitem[\citeproctext]{ref-liddell2018analyzing}
+Liddell, T. M., \& Kruschke, J. K. (2018). Analyzing ordinal data with
+metric models: What could possibly go wrong? \emph{Journal of
+Experimental Social Psychology}, \emph{79}, 328--348.
+
 \bibitem[\citeproctext]{ref-little2021facial}
@@
 Affective Neuroscience}, \emph{14}(7), 737--746.
+
+\bibitem[\citeproctext]{ref-makowski2018cognitive}
+Makowski, D. (2018). \emph{Cognitive neuropsychology of implicit emotion
+regulation through fictional reappraisal} {[}PhD thesis{]}. Sorbonne
+Paris Cit{é}.
 
@@
 
+\bibitem[\citeproctext]{ref-norman2010likert}
+Norman, G. (2010). Likert scales, levels of measurement and the
+{``laws''} of statistics. \emph{Advances in Health Sciences Education},
+\emph{15}, 625--632.
+
 \bibitem[\citeproctext]{ref-ospina2012general}
@@
 
+\bibitem[\citeproctext]{ref-sanders2016signatures}
+Sanders, J. I., Hangya, B., \& Kepecs, A. (2016). Signatures of a
+statistical computation in the human sense of confidence. \emph{Neuron},
+\emph{90}(3), 499--506.
+
+\bibitem[\citeproctext]{ref-sullivan2013analyzing}
+Sullivan, G. M., \& Artino Jr, A. R. (2013). Analyzing and interpreting
+data from likert-type scales. \emph{Journal of Graduate Medical
+Education}, \emph{5}(4), 541.
+
 \bibitem[\citeproctext]{ref-tucciarelli2022realness}

### 2025-06-14 paper/manuscript.qmd
rename

@@
 $$
-D_{\text{vour}} = \frac{\sigma_B^2}{\sigma_B^2 + \mu_{\text{SE}}^2}
+D_{\text{vour}} = \frac{\sigma_{\text{between}}^2}{\sigma_{\text{between}}^2 + \mu_{\text{within}}^2}
 $$
 
-Where $\sigma_B^2$ is the between-group variability (computed as the SD of the random effect point-estimates) and $\mu_{\text{SE}}^2$ is the mean squared uncertainty in random effect estimates (i.e., the average uncertainty).
+Where $\sigma_{between}^2$ is the between-group variability (computed as the SD of the random effect point-estimates) and $\mu_{\text{within}}^2$ is the mean squared uncertainty (SE) in random effect estimates (i.e., the average uncertainty).
 We use as *D-vour* = 0.666 as the threshold for moderately reliable random effect estimates, which corresponds to a 2:1 ratio of between-group variance to uncertainty.

### 2025-06-14 paper/manuscript.tex
rename

@@
 \par{ }
-\par{     \begin{tcolorbox}[enhanced jigsaw, breakable, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, left=2mm, rightrule=.15mm, arc=.35mm, opacityback=0, colback=white, leftrule=.75mm, toprule=.15mm]
+\par{     \begin{tcolorbox}[enhanced jigsaw, breakable, arc=.35mm, colframe=quarto-callout-note-color-frame, colback=white, leftrule=.75mm, opacityback=0, rightrule=.15mm, bottomrule=.15mm, toprule=.15mm, left=2mm]
 
@@
 \[
-D_{\text{vour}} = \frac{\sigma_B^2}{\sigma_B^2 + \mu_{\text{SE}}^2}
+D_{\text{vour}} = \frac{\sigma_{\text{between}}^2}{\sigma_{\text{between}}^2 + \mu_{\text{within}}^2}
 \]
 
-Where \(\sigma_B^2\) is the between-group variability (computed as the
-SD of the random effect point-estimates) and \(\mu_{\text{SE}}^2\) is
-the mean squared uncertainty in random effect estimates (i.e., the
-average uncertainty). We use as \emph{D-vour} = 0.666 as the threshold
-for moderately reliable random effect estimates, which corresponds to a
-2:1 ratio of between-group variance to uncertainty.
+Where \(\sigma_{between}^2\) is the between-group variability (computed
+as the SD of the random effect point-estimates) and
+\(\mu_{\text{within}}^2\) is the mean squared uncertainty (SE) in random
+effect estimates (i.e., the average uncertainty). We use as
+\emph{D-vour} = 0.666 as the threshold for moderately reliable random
+effect estimates, which corresponds to a 2:1 ratio of between-group
+variance to uncertainty.
 

### 2025-06-18 paper/manuscript.qmd
add paragraph on metacognitive stuff

@@
 Ordinal ratings (limited number of discrete options represented numerically by integers) are commonly used in psychology, and often modeled as continuous.
-While this fact has spurred its own polarized debate [@norman2010likert; @sullivan2013analyzing; @liddell2018analyzing], our position is that the model used should ideally reflect the data *generating* mechanism rather than necessarily focus on the *observed* data^[Ideally, the data recording method should be aligned with the assumed generating mechanism, which is an experiment design issue rather than a statistical one.] - although the former cannot always be easily inferred or known.
+While this fact has spurred its own polarized debate [@norman2010likert; @sullivan2013analyzing; @liddell2018analyzing; @rausch2014comparison], our position is that the model used should ideally reflect the data *generating* mechanism rather than necessarily focus on the *observed* data^[Ideally, the data recording method should be aligned with the assumed generating mechanism, which is an experiment design issue rather than a statistical one.] - although the former cannot always be easily inferred or known.
 For the data of Sample 2, we held the assumption that the underlying latent distribution was CHOCO distributed (as there is no reason to assume it would be different from Sample 1), and thus treated as continuous CHOCO-distributed data, despite the fact that it was collected on a 7-point scale.
@@
 
-First, future studies should investigate the psychometric quality of the CHOCO model to derive practical guidelines for its optimal use.
+Firstly, future studies should investigate the psychometric quality of the CHOCO model to derive practical guidelines for its optimal use.
 These include assessing the minimum amount of data required to reliably estimate and recover the different CHOCO parameters.
@@
 
-In conclusion, we developed and validated the CHOCO model, a theoretically grounded and practically flexible tool for modeling subjective judgments and delineating choice from confidence processes.
+In the domain of reaction times, researchers distinguish between "descriptive" models (e.g., ex-Gaussian or log-normal distributions) that capture characteristic distributional shapes, and "data-generating" models like the Drift Diffusion Model [DDM, @ratcliff2016diffusion] or Linear Ballistic Accumulator [LBA, @brown2008simplest], which simulate an assumed generative process (typically, sequential evidence accumulation).
+Drawing a comparison with this distinction, CHOCO was designed as a descriptive model to flexibly fit bimodal confidence data from subjective scales where responses implicitly reflect two categories.
+However, CHOCO remains agnostic about the underlying cognitive dynamics that generate those confidence ratings.
+In contrast, the field of metacognition has developed cognitive models that conceptualize confidence generation as a second-order inference.
+These include approaches derived from Signal Detection Theory, such as the meta-d' method which quantifies metacognitive sensitivity by assessing how well confidence discriminates between correct and incorrect decisions [@maniscalco2012signal; @fleming2017hmeta], as well as dynamic models like the Weighted Evidence and Visibility model or the Post-Decisional Accumulation model, which propose that confidence arises from mechanisms like post-decisional evidence sampling [@rausch2021modelling].
+A critical distinction is that these models are typically applied to two-step paradigms, where participants first make a categorical decision and then separately report their confidence.
+Future studies could investigate the differences and potential overlap in the applicability area and insights provided by these two classes of models.
+
+In conclusion, we developed and validated the CHOCO model, a mathematically grounded and practically flexible tool for modeling subjective judgments and delineating choice from confidence processes.
 It not only advances statistical modeling of decision-making data but opens new avenues for dissecting the cognitive mechanisms of belief, confidence, and reality perception.

### 2025-06-18 paper/manuscript.tex
add paragraph on metacognitive stuff

@@
 \par{ }
-\par{     \begin{tcolorbox}[enhanced jigsaw, breakable, arc=.35mm, colframe=quarto-callout-note-color-frame, colback=white, leftrule=.75mm, opacityback=0, rightrule=.15mm, bottomrule=.15mm, toprule=.15mm, left=2mm]
+\par{     \begin{tcolorbox}[enhanced jigsaw, breakable, opacityback=0, colframe=quarto-callout-note-color-frame, arc=.35mm, colback=white, left=2mm, rightrule=.15mm, leftrule=.75mm, bottomrule=.15mm, toprule=.15mm]
 
@@
 \citeproc{ref-norman2010likert}{Norman, 2010};
+\citeproc{ref-rausch2014comparison}{Rausch \& Zehetleitner, 2014};
 \citeproc{ref-sullivan2013analyzing}{Sullivan \& Artino Jr, 2013}), our
@@
 
-First, future studies should investigate the psychometric quality of the
-CHOCO model to derive practical guidelines for its optimal use. These
-include assessing the minimum amount of data required to reliably
+Firstly, future studies should investigate the psychometric quality of
+the CHOCO model to derive practical guidelines for its optimal use.
+These include assessing the minimum amount of data required to reliably
 estimate and recover the different CHOCO parameters. When applying the
@@
 
+In the domain of reaction times, researchers distinguish between
+``descriptive'' models (e.g., ex-Gaussian or log-normal distributions)
+that capture characteristic distributional shapes, and
+``data-generating'' models like the Drift Diffusion Model (DDM,
+\citeproc{ref-ratcliff2016diffusion}{Ratcliff et al., 2016}) or Linear
+Ballistic Accumulator (LBA, \citeproc{ref-brown2008simplest}{Brown \&
+Heathcote, 2008}), which simulate an assumed generative process
+(typically, sequential evidence accumulation). Drawing a comparison with
+this distinction, CHOCO was designed as a descriptive model to flexibly
+fit bimodal confidence data from subjective scales where responses
+implicitly reflect two categories. However, CHOCO remains agnostic about
+the underlying cognitive dynamics that generate those confidence
+ratings. In contrast, the field of metacognition has developed cognitive
+models that conceptualize confidence generation as a second-order
+inference. These include approaches derived from Signal Detection
+Theory, such as the meta-d' method which quantifies metacognitive
+sensitivity by assessing how well confidence discriminates between
+correct and incorrect decisions
+(\citeproc{ref-fleming2017hmeta}{Fleming, 2017};
+\citeproc{ref-maniscalco2012signal}{Maniscalco \& Lau, 2012}), as well
+as dynamic models like the Weighted Evidence and Visibility model or the
+Post-Decisional Accumulation model, which propose that confidence arises
+from mechanisms like post-decisional evidence sampling
+(\citeproc{ref-rausch2021modelling}{Rausch et al., 2021}). A critical
+distinction is that these models are typically applied to two-step
+paradigms, where participants first make a categorical decision and then
+separately report their confidence. Future studies could investigate the
+differences and potential overlap in the applicability area and insights
+provided by these two classes of models.
+
 In conclusion, we developed and validated the CHOCO model, a
-theoretically grounded and practically flexible tool for modeling
+mathematically grounded and practically flexible tool for modeling
 subjective judgments and delineating choice from confidence processes.
@@
 
+\bibitem[\citeproctext]{ref-brown2008simplest}
+Brown, S. D., \& Heathcote, A. (2008). The simplest complete model of
+choice response time: Linear ballistic accumulation. \emph{Cognitive
+Psychology}, \emph{57}(3), 153--178.
+
 \bibitem[\citeproctext]{ref-burkner2017brms}
@@
 Personality and Social Psychology}, \emph{82}(6), 878.
+
+\bibitem[\citeproctext]{ref-fleming2017hmeta}
+Fleming, S. M. (2017). HMeta-d: Hierarchical bayesian estimation of
+metacognitive efficiency from confidence ratings. \emph{Neuroscience of
+Consciousness}, \emph{2017}(1), nix007.
 
@@
 
+\bibitem[\citeproctext]{ref-maniscalco2012signal}
+Maniscalco, B., \& Lau, H. (2012). A signal detection theoretic approach
+for estimating metacognitive sensitivity from confidence ratings.
+\emph{Consciousness and Cognition}, \emph{21}(1), 422--430.
+
 \bibitem[\citeproctext]{ref-miller2023ai}
@@
 
+\bibitem[\citeproctext]{ref-ratcliff2016diffusion}
+Ratcliff, R., Smith, P. L., Brown, S. D., \& McKoon, G. (2016).
+Diffusion decision model: Current issues and history. \emph{Trends in
+Cognitive Sciences}, \emph{20}(4), 260--281.
+
+\bibitem[\citeproctext]{ref-rausch2021modelling}
+Rausch, M., Hellmann, S., \& Zehetleitner, M. (2021). Modelling
+visibility judgments using models of decision confidence.
+\emph{Attention, Perception, \& Psychophysics}, \emph{83}(8),
+3311--3336.
+
+\bibitem[\citeproctext]{ref-rausch2014comparison}
+Rausch, M., \& Zehetleitner, M. (2014). A comparison between a visual
+analogue scale and a four point scale as measures of conscious
+experience of motion. \emph{Consciousness and Cognition}, \emph{28},
+126--140.
+
 \bibitem[\citeproctext]{ref-rouder2024hierarchical}

### 2025-06-18 paper/manuscript.qmd
mention metad in intro

@@
 Decision-making is often conceptualized as involving distinct processes: the choice itself, and the confidence associated with that choice [@sanders2016signatures]. 
-In experimental paradigms, these can be somewhat disentangled by prompting participants to make a discrete choice selection (e.g., "True" *vs.* "False"), followed by a separate confidence rating. 
-However, this artificial separation makes its joint analysis difficult, and may not reflect real-world scenarios, where individuals often express both choice and confidence simultaneously using a single, continuous scale.
-In such scales, each side can represent a distinct latent category, and the distance from the midpoint can indicate the level of confidence or certainty.
+In experimental paradigms, these can be somewhat disentangled by prompting participants to make a discrete choice selection (e.g., correct *vs.* error), followed by a separate confidence rating, which has led to the development of bespoke approaches to model the underlying data-generation process [e.g., @fleming2017hmeta; @rausch2021modelling].
+However, this artificial separation may not reflect real-world scenarios, where individuals express simultaneously through a single scale both a discrete choice and an analog judgment of confidence, certainty, or degree of agreement (note that we will use the term "confidence" hereafter as a general term, without assuming a specific cognitive process).
+In instances where each side represents a distinct latent category (e.g., True/False, Agree/Disagree, Positive/Negative) - as opposed to a truly unimodal dimension, the absolute distance from the midpoint could be conceptualized as a indicating a continuous degree of "confidence".
 This integrated response format typically results in bimodal distributions, with peaks corresponding to the mean confidence on either side.
@@
 
-To model data of subjective scales in which the left and right sides can be conceptualized as two different choices (e.g., True/False, Agree/Disagree, etc.) and the magnitude of the response (how much the cursor is set away from the mid-point) as the confidence, we introduce the Choice-Confidence (CHOCO) model (@fig-two).
+To model data of subjective scales in which the left and right sides can be conceptualized as two different choices and the magnitude of the response (how much the cursor is set away from the mid-point) as the confidence, we introduce the Choice-Confidence (CHOCO) model (@fig-two).
 It consists of a three‐part mixture on $x \in [0, 1]$:

### 2025-06-18 paper/manuscript.tex
mention metad in intro

@@
 \par{ }
-\par{     \begin{tcolorbox}[enhanced jigsaw, breakable, opacityback=0, colframe=quarto-callout-note-color-frame, arc=.35mm, colback=white, left=2mm, rightrule=.15mm, leftrule=.75mm, bottomrule=.15mm, toprule=.15mm]
+\par{     \begin{tcolorbox}[enhanced jigsaw, bottomrule=.15mm, arc=.35mm, colback=white, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, colframe=quarto-callout-note-color-frame, opacityback=0, toprule=.15mm]
 
@@
 experimental paradigms, these can be somewhat disentangled by prompting
-participants to make a discrete choice selection (e.g., ``True''
-\emph{vs.} ``False''), followed by a separate confidence rating.
-However, this artificial separation makes its joint analysis difficult,
-and may not reflect real-world scenarios, where individuals often
-express both choice and confidence simultaneously using a single,
-continuous scale. In such scales, each side can represent a distinct
-latent category, and the distance from the midpoint can indicate the
-level of confidence or certainty. This integrated response format
-typically results in bimodal distributions, with peaks corresponding to
-the mean confidence on either side. Traditional beta regression models,
-which assume unimodal distributions within the (0,1) interval, are
-ill-suited for such data. One alternative is to transform the data into
-two variables a posteriori: binarizing the side to represent choice and
-calculating the absolute distance from the midpoint to represent
-confidence. These can then be modeled separately, for instance, using
-logistic regressions for choice and beta regressions for confidence (see
+participants to make a discrete choice selection (e.g., correct
+\emph{vs.} error), followed by a separate confidence rating, which has
+led to the development of bespoke approaches to model the underlying
+data-generation process (e.g., \citeproc{ref-fleming2017hmeta}{Fleming,
+2017}; \citeproc{ref-rausch2021modelling}{Rausch et al., 2021}).
+However, this artificial separation may not reflect real-world
+scenarios, where individuals express simultaneously through a single
+scale both a discrete choice and an analog judgment of confidence,
+certainty, or degree of agreement (note that we will use the term
+``confidence'' hereafter as a general term, without assuming a specific
+cognitive process). In instances where each side represents a distinct
+latent category (e.g., True/False, Agree/Disagree, Positive/Negative) -
+as opposed to a truly unimodal dimension, the absolute distance from the
+midpoint could be conceptualized as a indicating a continuous degree of
+``confidence''. This integrated response format typically results in
+bimodal distributions, with peaks corresponding to the mean confidence
+on either side. Traditional beta regression models, which assume
+unimodal distributions within the (0,1) interval, are ill-suited for
+such data. One alternative is to transform the data into two variables a
+posteriori: binarizing the side to represent choice and calculating the
+absolute distance from the midpoint to represent confidence. These can
+then be modeled separately, for instance, using logistic regressions for
+choice and beta regressions for confidence (see
 \citeproc{ref-makowski2025too}{Makowski, Te, et al., 2025} for an
@@
 To model data of subjective scales in which the left and right sides can
-be conceptualized as two different choices (e.g., True/False,
-Agree/Disagree, etc.) and the magnitude of the response (how much the
-cursor is set away from the mid-point) as the confidence, we introduce
-the Choice-Confidence (CHOCO) model (Figure~\ref{fig-two}). It consists
-of a three‐part mixture on \(x \in [0, 1]\):
+be conceptualized as two different choices and the magnitude of the
+response (how much the cursor is set away from the mid-point) as the
+confidence, we introduce the Choice-Confidence (CHOCO) model
+(Figure~\ref{fig-two}). It consists of a three‐part mixture on
+\(x \in [0, 1]\):
 

### 2025-06-21 analysis/0_proof.qmd
update based on comments

---
title: "Proof of Concept"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    toc: true
    code-fold: true
    self-contained: false
execute: 
  cache: true
  cache-lazy: false
---

```{r}
#| message: false
#| warning: false

options(mc.cores = parallel::detectCores() - 2)

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
library(ggdist)
library(brms)
library(cmdstanr)
library(rstan)
library(cogmod)
library(gt)
library(webshot2)
Sys.setenv(CHROMOTE_CHROME = "C:/Program Files (x86)/Microsoft/Edge/Application/msedge.exe")
```

## Data

```{r}
set.seed(33)

c <- c("Condition A: CHOCO(p = 0.5, confleft = 0.3, confright = 0.3)",
       "Condition B: CHOCO(p = 0.3, confleft = 0.3, confright = 0.7)")

df <- data.frame(
  y = c(
    rchoco(1000, p = 0.5, confleft = 0.3, confright = 0.3, pex = 0),
    rchoco(1000, p = 0.3, confleft = 0.3, confright = 0.7, pex = 0)
    ),
  Condition = c(rep(c("A", "B"), each = 1000))
)

p1 <- data.frame(
  x = c(seq(0, 1, length.out = 1000), seq(0, 1, length.out = 1000)),
  y = c(
    dchoco(seq(0, 1, length.out = 1000), p = 0.5, confleft = 0.3, confright = 0.3, pex = 0),
    dchoco(seq(0, 1, length.out = 1000), p = 0.3, confleft = 0.3, confright = 0.7, pex = 0)
  ),
  Condition = rep(c, each = 1000)
) |> 
  ggplot(aes(x = x, y = y)) +
  geom_histogram(data = df, aes(x = y, y = after_stat(density), fill = Condition),
                bins = 35, alpha = 0.5, position = "identity") +
  geom_line(aes(color = Condition), linewidth = 2, alpha = 0.95) +
  scale_fill_manual(values = c("#0072B2", "#FF5722"), guide = "none") +
  scale_color_manual(values = c("#0072B2", "#FF5722")) +
  labs(x = "Confidence", y = "Distribution", title = "Usage Example") +
  theme_minimal() +
  theme(legend.position = "top",
        legend.direction='vertical',
        legend.title = element_blank(),
        legend.justification = c(0, 0.5),
        axis.text.y = element_blank(),
        plot.title = element_text(face = "bold"))
p1
```

## Linear model

```{r}

model1 <- lm(y ~ Condition, data = df)

t1 <- parameters::parameters(model1) |> 
  format_table(zap_small = TRUE) |> 
  data_select(exclude = "SE") |> 
  data_rename(select = "Parameter", "Effect") |> 
  gt() |> 
  data_color(
    columns = "Effect",
    palette = c("#0072B2", "#FF5722")) |>
  data_color(
    columns = "p",
    rows = 2,
    palette = "red") |> 
  tab_header(
    title = html("<strong>Linear Model</strong>"),
    subtitle = html("<p style='text-align: left;'><code>lm(y ~ Condition)</code><br></p>")) |> 
  cols_align(columns = -Effect, align = "right")|>
  cols_label(Effect = md("***Effect***"),
             `95% CI` = md("***95% CI***"),
             Coefficient = md("***Coefficient***"),
             `t(1998)` = md("***t(1998)***"),
             p = md("***p***")) 
t1

gtsave(t1, "table1.png", path="./", vwidth = 1280)
p_t1 <- png::readPNG("table1.png") |> 
  grid::rasterGrob()
file.remove("table1.png")
```

## CHOCO

```{r}
model2 <- brm(bf(
    y ~ Condition,
    confleft ~ Condition,
    confright ~ Condition,
    precright ~ 1,
    precleft ~ 1,
    pex = 0,
    bex = 0.5,
    pmid = 0),
  data = df,
  family = choco(link_mu = "identity", 
                 link_confleft = "identity", 
                 link_confright = "identity"),
  stanvars = cogmod::choco_stanvars(),
  init = 0.5,
  iter = 2000,
  warmup = 1000,
  chains = 4,
  refresh = 0,
  cores = parallel::detectCores() - 2,
  backend = "cmdstanr"
)

t2 <- parameters::parameters(model2, dispersion = TRUE) |> 
  as.data.frame() |> 
  mutate(t = Median / MAD, p = bayestestR::pd_to_p(pd)) |> 
  head(6) |> 
  select(Parameter = Component, Effect = Parameter, Coefficient = Median, CI, CI_low, CI_high, t, p) |> 
  mutate(Effect = rep(c("(Intercept)", "Condition [B]"), 3),
         Parameter = ifelse(Parameter == "conditional", "p", Parameter)) |> 
  format_table(zap_small = TRUE) |> 
  gt() |> 
  data_color(
    columns = "Effect",
    palette = rep(c("#0072B2", "#FF5722"), 3)) |> 
  data_color(
    columns = "Parameter",
    palette = c("#9a7b78", "#c09569", "#9C27B0")) |> 
  data_color(
    columns = "p",
    rows = c(2, 6),
    palette = "green") |>
   data_color(
    columns = "p",
    rows = c(4),
    palette = "red") |>
  tab_header(
    title = html("<strong>CHOCO Model</strong>"),
    subtitle = html("<p style='text-align: left;'><code>brm(bf(<br>&emsp;&emsp;y ~ Condition, <br>&emsp;&emsp;confleft ~ Condition, <br>&emsp;&emsp;confright ~ Condition), <br>family = cogmod::choco())</code><br></p>")) |>
  cols_align(columns = -c(Effect, Parameter), align = "right") |>
  cols_label(Effect = md("***Effect***"),
             Parameter = md("***Parameter***"),
             Coefficient = md("***Coefficient***"),
             `95% CI` = md("***95% CI***"),
             `t` = md("***t***"),
             `p` = md("***p***"))

t2

gtsave(t2, "table2.png", path="./", vwidth = 1280)
p_t2 <- png::readPNG("table2.png") |> 
  grid::rasterGrob()
file.remove("table2.png")
```

## Figure

```{r}
# p2 <- wrap_elements(wrap_elements(p_t1) / wrap_elements(p_t2) + 
#          patchwork::plot_annotation(title = "Effect of Condition", theme = theme(plot.title = element_text(face = "bold"))) +
#          patchwork::plot_layout(heights = c(0.35, 0.65)))
p2 <- gridExtra::grid.arrange(p_t1, p_t2, heights = c(0.35, 0.65))

ggsave("table3.png", p2, width = 10, height = 10, dpi = 300)
p2 <- png::readPNG("table3.png") |> 
  grid::rasterGrob()
file.remove("table3.png")

fig <- gridExtra::grid.arrange(p1, p2, ncol = 2, widths = c(0.45, 0.55))
ggsave("../paper/figures/fig2b.png", fig, dpi = 300, width = 9, height=6)

# ----
fig2a <- grid::rasterGrob(png::readPNG("../paper/figures/fig2.png"))
fig2b <- grid::rasterGrob(png::readPNG("../paper/figures/fig2b.png"))

figfinal <- gridExtra::grid.arrange(fig2a, fig2b,
  ncol = 1, heights = c(0.5, 0.5))
figfinal
ggsave("../paper/figures/fig2final.png", figfinal, dpi = 600, 
       width = grid::resolveRasterSize(fig2a)$width,
       height = grid::resolveRasterSize(fig2a)$height + grid::resolveRasterSize(fig2b)$height)
```

### 2025-06-21 paper/manuscript.qmd
update based on comments

@@
     role:
-      - Writing – original draft
       - Writing – review & editing
@@
 A prevalent issue is the continued reliance on linear models that assume normally distributed (Gaussian) data^[More specifically, that the outcome is distributed according to a Normal distribution which parameters are expressed as a linear function of the predictors.] - as this assumption often does not hold true for many types of psychological outcomes.
-For instance, reaction times typically exhibit skewed distributions, choices can be represented as binary variables, and count data consists of strictly positive integers. 
+For instance, reaction times typically exhibit skewed distributions, choices can be represented as binary variables, and count data consist of strictly positive integers. 
 Applying models that presume normality and model the "mean" of the outcome variable can lead to misinterpretation and potentially misleading conclusions when applied indiscriminately.
-It is thus important that psychologists use models that can best describe (or generate) the data they collect, to fully exploit them and bring more nuance and accuracy to their conclusions.
+It is thus important that psychologists use models that can best describe (or generate) the data they collect, to exploit them fully and bring more nuance and accuracy to their conclusions.
 
@@
 Among the most commonly collected data in psychology are responses on subjective scales, such as Likert-type items or visual analog scales, which exhibit some fundamental properties: these responses are bounded (and can be rescaled to a 0-1 range) and frequently display clustering at the extremes.
-Traditional linear models being ill-suited for such data, researchers have turned to using Beta distributions to model this data (instead of Gaussian), suited for continuous data within the (0,1) interval (i.e., excluding extreme responses).
+Traditional linear models being ill-suited for such data, researchers have turned to using Beta distributions to model these data (instead of Gaussian), suited for continuous data within the (0,1) interval (i.e., excluding extreme responses).
 To address the frequent occurrence of exact zeros and ones (i.e., extreme values), zero-one inflated beta (ZOIB) models have been developed [@ospina2012general] to accommodate the excess of boundary values by incorporating additional components that model the probabilities of responses at 0 and 1 as a separate, independent process.
@@
 The Beta-Gate model is a reparametrized Ordered Beta model [@kubinec2023ordered]^[In the Ordered Beta model, the cutpoints on the log-scale are directly used as parameters, instead of being derived from *pex* and *bex*.] available in the *cogmod* package in R (https://github.com/DominiqueMakowski/cogmod), in which participants' answers on bounded scales are conceptualized as latent responses that can fall past a pair of probabilistic "gates" (or cutpoints) that control whether the response is recorded as an extreme (0 or 1) or as a nuanced, continuous value in between (@fig-one).
-These distance of these gates from the edges of the scale varies based on two interpretable parameters: *pex* (the propensity by which people are likely to answer extreme values), and *bex*, a bias toward the upper extreme (1) versus the lower (0).
+The distance of these gates from the edges of the scale varies based on two interpretable parameters: *pex* (the propensity by which people are likely to answer extreme values), and *bex*, a bias toward the upper extreme (1) versus the lower (0).
 A person's internal response that lies close to the edge might be "caught" by a gate and recorded as an extreme, while others pass through to express a continuous response (Beta-distributed with $\mu$ (*mu*) and $\phi$ (*phi*) as its mean and precision parameters).
@@
 'Mu' and 'phi' correspond to the mean and precision of the continuous part of the distribution (between 0 and 1), and 'pex' (propensity of extremes) and 'bex' (balance of extremes) indirectly control the proportion of zeros and ones by specifying the location of the \"gates\", past which the latent response process is likely to generate extreme values.
-Specifically, 'pex' defines the total distance of both gates from the extremes (in yellow), and 'bex' determines the proportion of the right gate distance relative to the left.
+Specifically, 'pex' defines the total distance of both gates from the extremes (i.e., the sum length of the two orange arrows), and 'bex' determines the proportion of the right gate distance relative to the left (i.e., the length of the right arrow relative to the left one).
 In this example, the total distance from the extremes is 'pex' = 0.1, with 40% ('bex' = 0.4) of that distance being on the right (and 60% on the left).
 The left gate is thus located at 0.6, and the right at 1-0.04 = 0.96."
+#| fig-align: "center"
 #| apa-twocolumn: false  # A Figure Spanning Two Columns When in Journal Mode
@@
 
-- $P(x = 0) = \text{logistic}\left( \text{logit}(pex \cdot (1 - bex)) - \text{logit}(\mu) \right)$
-- $P(x = 1) = 1 - \text{logistic}\left( \text{logit}(1 - pex \cdot bex) - \text{logit}(\mu) \right)$
-- $P(x \in (0,1)) = 1 - P(x = 0) - P(x = 1)$
+$$
+\begin{cases}
+\mathrm{logistic}\left( \mathrm{logit}(pex \cdot (1 - bex)) - \mathrm{logit}(\mu) \right), & x = 0, \\[6pt]
+1 - \mathrm{logistic}\left( \mathrm{logit}(1 - pex \cdot bex) - \mathrm{logit}(\mu) \right), & x = 1, \\[6pt]
+1 - P(x = 0) - P(x = 1), & 0 < x < 1. \\
+\end{cases}
+$$
 
@@
 #| label: fig-two
-#| fig-cap: "The CHOCO Model uses a mixture of Beta-Gate distributions to model separately the right and left sides of the scale (e.g., a rating of whether a statement was 'Truth' vs.
+#| fig-cap: "Top: The CHOCO Model uses a mixture of Beta-Gate distributions to model separately the right and left sides of the scale (e.g., a rating of whether a statement was 'Truth' vs.
 'Lie'), as well as their relative proportion.
@@
 confleft = 0.3).
-Extreme values (zeros and ones) are governed by the same mechanism as for Beta-Gate models."
-#| apa-twocolumn: false  # A Figure Spanning Two Columns When in Journal Mode
-#| out-width: "100%"
+Extreme values (zeros and ones) are governed by the same mechanism as for Beta-Gate models.
+Below: An illustrative dataset in which participants, in condition A (blue), respond as often on the left and right side of the scale, with low confidence (the confidence is concentrated towards the centre).
+In condition B (orange), they answer more often on the left (but also with low confidence), but their responses on the right are more extreme.
+As shown in the top table, a typical linear model could lead the imprudent researcher to misleadingly conclude that \"there is no effect of Condition\" (as a linear regression models changes in the mean, which is actually equal for the two conditions).
+The CHOCO model correctly reveals that condition B is associated with a lower probability but higher confidence of answers on the right side of the scale.
+Note that the CHOCO model is currently only available through Bayesian modeling, hence we report for consistency pseudo t- and p-values in the table, which correspond respectively to the coefficients divided by their SE, and the values obtained via the pd-to-p conversion [@makowski2019indices]."
+#| fig-align: "center"
+#| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
+#| out-width: "70%"
 
-knitr::include_graphics("./figures/fig2.png")
+knitr::include_graphics("./figures/fig2final.png")
 ```
@@
 **Study 1** aims to evaluate the CHOCO model’s ability to better capture subjective scale responses that (potentially) reflect an underlying discrete choice, in comparison to existing models such as the ZOIB and Beta-Gate.
-Specifically, we will assess whether 1) CHOCO provides improved model fit, 2) yields deeper insights into population-level effects than traditional approaches (gender differences in reality beliefs), and 3) allows for the reliable estimation of interpretable individual-level parameters through random effects.
-**Study 2** will apply this model to more subtle effects, such as the effect perceived facial attractiveness on reality judgments, and test the ability to fit alternative data structures (scales with ordinal response options and mid-points).
+Specifically, we will assess whether 1) CHOCO provides improved model fit, 2) yields deeper insights into population-level effects than traditional approaches (sex differences in reality beliefs), and 3) allows for the reliable estimation of interpretable individual-level parameters through random effects.
+**Study 2** will apply this model to more subtle effects, such as the effect of perceived facial attractiveness on reality judgments, and test the ability to fit alternative data structures (scales with ordinal response options and mid-points).
 To this end, we analyze data from two separate studies in which participants judged whether a face image was AI-generated ("fake") or a real photograph.
@@
 
-In today's post-truth era, the proliferation of advanced AI technologies has made it increasingly challenging to distinguish between authentic and synthetic media, bearing significant implications for information integrity and public trust [@lewandowsky2017beyond].
-As traditional cues become less reliable (e.g., visual glitches and artefacts in generated images; formulaic generated text, etc.), people increasingly depend on contextual information and cognitive heuristics to assess authenticity, a process referred to as "simulation monitoring" [@makowski2019phenomenal].
+The question of "what is real" has become increasingly relevant in the context of recent technological developments, particularly with the rise of artificial intelligence (AI), capable of mass producing highly realistic and relevant content, such as images, videos or text.
+These synthetic productions are flooding the cyberspace across various domains, such as art, advertising and entertainment, to education and information. 
+The democratization of such technology carries an important potential for misuse, such as in disinformation campaigns and scams (e.g., AI-bots, identity theft), and abuse [e.g., deepfakes, @viola2023designed], and raises pressing concerns about the value of authenticity and the potential erosion of media trust in our increasingly *post-truth* society [@lewandowsky2017beyond].
+Critically, as technology improves, traditional cues for detecting what is fake become less reliable (e.g., visual glitches and artefacts in generated images; formulaic generated text, etc.), leading people to increasingly depend on contextual information and cognitive heuristics to assess authenticity, a process referred to as "simulation monitoring" [@makowski2019phenomenal].
 
@@
 Using the open-access data from @makowski2025too, we included all heterosexual and bisexual (as these two groups did not seem to differ based on preliminary analyses and were thus grouped to maximize power) male and female participants, for a final sample of 141 participants (Mean age = 28.4, SD = 9.0, range: [19, 66]; Sex: 47.5% females).
-For each participant, we included only stimuli of the opposite gender (i.e., all 89 female faces for men and 20 male faces for women).
+For each participant, we included only stimuli of the opposite sex (i.e., all 89 female faces for males and 20 male faces for females).
 
@@
 After each image, participants rated the face on trustworthiness, familiarity, attractiveness, and beauty using visual analog scales.
-In the second phase, participants were informed that "about half of the previously seen images were AI-generated".
+In the second phase, participants were informed that "about half of the previously seen images were AI-generated" (in actuality, all faces were photographs).
 The same faces were presented again in a new random order (same display time), followed by ratings of "reality" (whether they believed the image was fake - left anchor - or real - right anchor).
@@
 The models were run using *brms* [@burkner2017brms] R package, and analyzed using the *easystats* collection of packages [@makowski2025modelbased; @ludecke2020extracting; @patil2022datawizard].
-To maximize the comparability across models We used the default priors (uniform) for all models, and we ran 16 chains of 1400 iterations each on the University of Sussex High-Performance Computing (HPC) cluster.
+To maximize comparability across models, we used the default priors (uniform) for all models, and we ran 16 chains of 1400 iterations each on the University of Sussex High-Performance Computing (HPC) cluster.
 
@@
 
-For the population-level effects, we will consider significant and report (using the median of the posterior distribution) effects for which the 95% Credible Interval (CI) does not include zero [and when the probability of direction *pd* is > \~97\%, @makowski2019indices].
+For the population-level effects, we will consider "significant" (noteworthy) and report (using the median of the posterior distribution) effects for which the 95% Credible Interval (CI) does not include zero [and when the probability of direction *pd* is > \~97\%, @makowski2019indices].
 For the individual-level parameters (i.e., the random intercepts of each parameter for each participant and each item), we will first analyze their reliability using the Variance-Over-Uncertainty Ratio index (*D-vour*).
@@
 We use as *D-vour* = 0.666 as the threshold for moderately reliable random effect estimates, which corresponds to a 2:1 ratio of between-group variance to uncertainty.
+Reliable in this context means that the variability between groups (e.g., participants) is greater than the uncertainty in the estimates, suggesting that the parameter can be used to assess inter-individual differences.
 
@@
 We illustrate this by showing the effect of Sex on the distribution of reality beliefs based on a CHOCO model."
+#| fig-align: "center"
 #| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
@@
 
-The ZOIB model suggested that women had higher mean scores of reality beliefs ($\mu_{~\text{Female}} = 0.20,~95\%~ CI~[0.03, 0.37], ~pd = 98.77\%$), less extreme values ($zoi_{~\text{Female}} = -1.24,~95\%~ CI~[-2.46, -0.07], ~pd = 98.25\%$) but more ones relative to zeros ($coi_{~\text{Female}} = 1.63,~95\%~ CI~[0.70, 2.56], ~pd = 99.97\%$).
+The ZOIB model suggested that females had higher mean scores of reality beliefs ($\mu_{~\text{Female}} = 0.20,~95\%~ CI~[0.03, 0.37], ~pd = 98.77\%$), less extreme values ($zoi_{~\text{Female}} = -1.24,~95\%~ CI~[-2.46, -0.07], ~pd = 98.25\%$) but more ones relative to zeros ($coi_{~\text{Female}} = 1.63,~95\%~ CI~[0.70, 2.56], ~pd = 99.97\%$).
 
-The Beta-Gate model similarly suggested that women had higher mean scores of reality beliefs ($\mu_{~\text{Female}} = 0.21,~95\%~ CI~[0.02, 0.40], ~pd = 98.52\%$), less extreme values ($pex_{~\text{Female}} = -1.34,~95\%~ CI~[-2.56, -0.17], ~pd = 98.75\%$), and a greater tendency to answer one relative to zero ($bex_{~\text{Female}} = 1.13,~95\%~ CI~[0.34, 1.91], ~pd = 99.76\%$).
+The Beta-Gate model similarly suggested that females had higher mean scores of reality beliefs ($\mu_{~\text{Female}} = 0.21,~95\%~ CI~[0.02, 0.40], ~pd = 98.52\%$), less extreme values ($pex_{~\text{Female}} = -1.34,~95\%~ CI~[-2.56, -0.17], ~pd = 98.75\%$), and a greater tendency to answer one relative to zero ($bex_{~\text{Female}} = 1.13,~95\%~ CI~[0.34, 1.91], ~pd = 99.76\%$).
 
-The CHOCO model shows that women had a higher probability *p* of judging faces as real ($P_{~\text{Female}} = 0.45,~95\%~ CI~[0.05, 0.86], ~pd = 98.54\%$), but are not more confident when doing so ($confright_{~\text{Female}} = -0.13,~95\%~ CI~[-0.47, 0.21], ~pd = 77.73\%$).
+The CHOCO model shows that females had a higher probability *p* of judging faces as real ($P_{~\text{Female}} = 0.45,~95\%~ CI~[0.05, 0.86], ~pd = 98.54\%$), but are not more confident when doing so ($confright_{~\text{Female}} = -0.13,~95\%~ CI~[-0.47, 0.21], ~pd = 77.73\%$).
 However, they were less confident when answering that an image was AI-generated ($confleft_{~\text{Female}} = -0.53,~95\%~ CI~[-0.89, -0.17], ~pd = 99.84\%$).
@@
 Reliable effects are characterised by a higher between-group variability relative (the dispersion of the distribution of point-estimates) to its within-group variability (the average uncertainty of individual estimates represented by the error bar).
-For example, inter-individual variability in the parameters reflecting the confidence in left and right choices is reliably capture, as opposed to the inter-individual variability in the 'bex' parameter.
+For example, inter-individual variability in the parameters reflecting the confidence in left and right choices is reliably captured, as opposed to the inter-individual variability in the 'bex' parameter.
 Bottom: correlation matrix between participant-level estimates from different models and empirical indices (e.g., the raw proportion of responses on the right per participant).
 The CHOCO parameters are easily interpretable, strongly correlating with their empirical counterparts."
+#| fig-align: "center"
 #| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
@@
 The explosion of accessibility of state-of-the-art AI tools has made it effortless to generate realistic images, including human faces that are often indistinguishable from real ones [@miller2023ai; @nightingale2022ai; @bozkir2024can].
-These synthetic visuals are flooding the cyberspace across various domains, such as art, advertising and entertainment, to education and information.
-This technological advancement carries an important potential for misuse, such as in disinformation campaigns, scams (e.g., AI-bots, identity theft), and abuse.
-The democratization of such technology raises pressing concerns about the value of authenticity and the potential erosion of media trust in our increasingly *post-truth* society.
-
 In this evolving landscape, understanding the cognitive mechanisms that underpin our judgments of reality becomes paramount.
@@
 
-Although its role as a potential modulator of reality beliefs remains under-explored, @makowski2025too found significant associations between participants' realness ratings and facial attractiveness, with a potential sexual dymorphism: male participants judged attractive faces as being likely more real, whereas evidence suggested a milder and quadratic (U-shaped) relationship between attractiveness and reality beliefs for females.
+Although the role of attractiveness as a potential modulator of reality beliefs remains under-explored, @makowski2025too found significant associations between participants' realness ratings and facial attractiveness, with a potential sexual dymorphism: male participants judged attractive faces as being likely more real, whereas evidence suggested a milder and quadratic (U-shaped) relationship between attractiveness and reality beliefs for females.
 These findings offer a complementary perspective to those of @miller2023ai, who reported that participants used attractiveness as a distinguishing cue between real and AI-generated faces.
@@
 For the second sample (data and preprocessing information available at https://github.com/RealityBending/FakeFace2), 248 participants were initially recruited through academic platforms (SONA and SurveySwap).
-We removed 15 participants with data suggesting low-effort responding as well as those (N=16) that did not believe in the experimental manipulation and were fully confident that all images were real or fake.
+We removed 15 participants with data suggesting low-effort responding (no variation in their responses indicative of straightlining) as well as those (N=16) that did not believe in the experimental manipulation and were fully confident that all images were real or fake.
 <!-- 8 (3.23%) participants with no variation in their responses and 7 (2.92%) participants with no coherence (r < 0.1) b -->
@@
 At the beginning of the experiment, a cover story presented the study as a partnership with an AI startup aimed at testing the quality of a new face AI-generation algorithm.
-Following that, participants would see the 109 neutral-expression photographs from the AMFD database (1 s), each preceded by a randomly assigned textual cue indicating whether the image was "AI-generated" or "Photograph" (2 s).
+Following that, participants would see the 109 neutral-expression pictures (all real photographs) from the AMFD database (1 s), each preceded by a randomly assigned textual cue indicating whether the image was "AI-generated" or "Photograph" (2 s).
 Ratings of attractiveness, beauty and trustworthiness were collected after each image.
@@
 For sample 2, the analysis was based on that of Sample 1.
-The main differences are 1) the inclusion of the "Condition" (whether the picture was presented as "Real" of "Fake" in the first phase of the experiment) as an additional predictor (entered as the only random slope for all participant random effects); 2) the inclusion of an additional parameters, *pmid*, modelling the probability of answering the middle-point of the Likert scale (representing an "undecided" option); 3) as they were effectively only 3 distinct values on each side of the scale, and to showcase the flexibility of the CHOCO model, we decided to *not* treat the extreme values as 0 and 1 (and model them via the separate parameters *pex* and *bex*, which would distributions to be estimated from only 2 values), but instead to treat them as part of the continuous distribution.
+The main differences are 1) the inclusion of the "Condition" (whether the picture was presented as "Real" of "Fake" in the first phase of the experiment) as an additional predictor (entered as the only random slope for all participant random effects); 2) the inclusion of an additional parameters, *pmid*, modelling the probability of answering the middle-point of the Likert scale (representing an "undecided" option); 3) as they were effectively only 3 distinct values on each side of the scale, and to showcase the flexibility of the CHOCO model, we decided to *not* treat the extreme values as 0 and 1 (and model them via the separate parameters *pex* and *bex*, which would imply estimating the distributions from only 2 values), but instead to treat them as part of the continuous distribution.
 The 7 response options were rescaled to be evenly spaced between 0 and 1 excluded (i.e., [0.125, 0.875]).
@@
 Given the presence of polynomial terms, significant effects are denoted by a symbol representing the shape of the relationship (/ or \\\\ for linear, U or inverse-U for quadratic links)."
+#| fig-align: "center"
 #| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
@@
 In sample 1, the traditional approach suggested a significant linear relationship between the mean level of "Reality" and attractiveness for males only ($\beta_{~\text{poly1}}=3.42, ~95\%~CI [2.50, 4.34], ~p < .001$), the second largest effect being that of a quadratic link for women ($\beta_{~\text{poly2}}=1.82, ~95\%~CI [-0.27, 3.92], ~p = .09$).
-The CHOCO model revealed that, for males only, attractiveness had a significantly positive linear relationship with the probability *p* of judging faces as real ($P_{~\text{poly1}}=13.64, ~95\%~CI [8.42, 18.99], ~pd = 100\%$), but a quadratic relationship with the confidence in real judgments ($confright_{~\text{poly2}}=3.70, ~95\%~CI [0.80, 6.40], ~pd = 99.53\%$).
+The CHOCO model revealed that, for males only, attractiveness had a significantly positive linear relationship with the probability *p* of judging faces as real ($P_{~\text{poly1}}=13.64, ~95\%~CI [8.42, 18.99], ~pd = 100\%$), and a quadratic relationship with the confidence in real judgments ($confright_{~\text{poly2}}=3.70, ~95\%~CI [0.80, 6.40], ~pd = 99.53\%$).
 Attractiveness was also associated with less confidence in fake judgments ($confleft_{~\text{poly1}}=-5.49, ~95\%~CI [-8.84, -2.11], ~pd =  99.92\%$), a quadratic relationship with left precision ($precleft_{~\text{poly2}}=-12.65	, ~95\%~CI [-24.59, -0.36], ~pd = 97.86\%$), and related linearly with a stronger bias towards extreme Real responses ($bex_{~\text{poly1}}=9.56	, ~95\%~CI [2.44, 16.73], ~pd = 99.55\%$).
@@
 
-### Beauty
+<!-- ### Beauty -->
 
-The same analysis was run for Beauty, which indicated the following differences: 
-**ANA: can you check and write down the main differences?**
+<!-- The same analysis was run for Beauty, which indicated the following differences:  -->
+<!-- **ANA: can you check and write down the main differences?** -->
 
@@
 The motivation was to see if the model could extrapolate and reliably estimate all the parameters based on a limited number of data points but means in our case that, effectively, each Beta distribution was extrapolated from the frequency of only three unique values.
-While this is far from ideal, the CHOCO model still proved insightful, although it might have impacted its ability detect finer-grain changes, including quadratic relationships.
+While this is far from ideal, the CHOCO model still proved insightful, although it might have restricted its ability to detect finer-grain changes, including quadratic relationships.
 
@@
 
-We would like to thank the dissertation students from the University of Sussex for their help in data collection.
+We would like to thank the dissertation students from the University of Sussex for their help in data collection, as well as T. Karapanagiotidis for feedback.
 DM would also like to thank F. Rocher for inspiring some aspects of this paper.

### 2025-06-21 paper/manuscript.tex
update based on comments

@@
 \par{ }
-\par{     \begin{tcolorbox}[enhanced jigsaw, bottomrule=.15mm, arc=.35mm, colback=white, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, colframe=quarto-callout-note-color-frame, opacityback=0, toprule=.15mm]
+\par{     \begin{tcolorbox}[enhanced jigsaw, opacityback=0, left=2mm, leftrule=.75mm, colback=white, breakable, bottomrule=.15mm, colframe=quarto-callout-note-color-frame, toprule=.15mm, arc=.35mm, rightrule=.15mm]
 
@@
 -- original draft; Ana Neves:   Data curation, Writing -- original
-draft, Writing -- review \& editing; Andy Field:   Writing -- original
-draft, Writing -- review \& editing}
+draft, Writing -- review \& editing; Andy Field:   Writing -- review \&
+editing}
 \par{Correspondence concerning this article should be addressed
@@
 distributions, choices can be represented as binary variables, and count
-data consists of strictly positive integers. Applying models that
-presume normality and model the ``mean'' of the outcome variable can
-lead to misinterpretation and potentially misleading conclusions when
-applied indiscriminately. It is thus important that psychologists use
-models that can best describe (or generate) the data they collect, to
-fully exploit them and bring more nuance and accuracy to their
-conclusions.
+data consist of strictly positive integers. Applying models that presume
+normality and model the ``mean'' of the outcome variable can lead to
+misinterpretation and potentially misleading conclusions when applied
+indiscriminately. It is thus important that psychologists use models
+that can best describe (or generate) the data they collect, to exploit
+them fully and bring more nuance and accuracy to their conclusions.
 
@@
 at the extremes. Traditional linear models being ill-suited for such
-data, researchers have turned to using Beta distributions to model this
+data, researchers have turned to using Beta distributions to model these
 data (instead of Gaussian), suited for continuous data within the (0,1)
@@
 control whether the response is recorded as an extreme (0 or 1) or as a
-nuanced, continuous value in between (Figure~\ref{fig-one}). These
+nuanced, continuous value in between (Figure~\ref{fig-one}). The
 distance of these gates from the edges of the scale varies based on two
@@
 extreme values. Specifically, `pex' defines the total distance of both
-gates from the extremes (in yellow), and `bex' determines the proportion
-of the right gate distance relative to the left. In this example, the
-total distance from the extremes is `pex' = 0.1, with 40\% (`bex' = 0.4)
-of that distance being on the right (and 60\% on the left). The left
-gate is thus located at 0.6, and the right at 1-0.04 = 0.96.}
+gates from the extremes (i.e., the sum length of the two orange arrows),
+and `bex' determines the proportion of the right gate distance relative
+to the left (i.e., the length of the right arrow relative to the left
+one). In this example, the total distance from the extremes is `pex' =
+0.1, with 40\% (`bex' = 0.4) of that distance being on the right (and
+60\% on the left). The left gate is thus located at 0.6, and the right
+at 1-0.04 = 0.96.}
 
@@
 
-\begin{itemize}
-\tightlist
-\item
-  \(P(x = 0) = \text{logistic}\left( \text{logit}(pex \cdot (1 - bex)) - \text{logit}(\mu) \right)\)
-\item
-  \(P(x = 1) = 1 - \text{logistic}\left( \text{logit}(1 - pex \cdot bex) - \text{logit}(\mu) \right)\)
-\item
-  \(P(x \in (0,1)) = 1 - P(x = 0) - P(x = 1)\)
-\end{itemize}
+\[
+\begin{cases}
+\mathrm{logistic}\left( \mathrm{logit}(pex \cdot (1 - bex)) - \mathrm{logit}(\mu) \right), & x = 0, \\[6pt]
+1 - \mathrm{logistic}\left( \mathrm{logit}(1 - pex \cdot bex) - \mathrm{logit}(\mu) \right), & x = 1, \\[6pt]
+1 - P(x = 0) - P(x = 1), & 0 < x < 1. \\
+\end{cases}
+\]
 
@@
 
-\begin{figure}
-
-\caption{\label{fig-two}The CHOCO Model uses a mixture of Beta-Gate
+\begin{figure*}[!htbp]
+
+{\caption{{Top: The CHOCO Model uses a mixture of Beta-Gate
 distributions to model separately the right and left sides of the scale
@@
 (confright = 0.6 vs.~confleft = 0.3). Extreme values (zeros and ones)
-are governed by the same mechanism as for Beta-Gate models.}
-
-\centering{
-
-\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{./figures/fig2.png}
-
-}
-
-\end{figure}%
+are governed by the same mechanism as for Beta-Gate models. Below: An
+illustrative dataset in which participants, in condition A (blue),
+respond as often on the left and right side of the scale, with low
+confidence (the confidence is concentrated towards the centre). In
+condition B (orange), they answer more often on the left (but also with
+low confidence), but their responses on the right are more extreme. As
+shown in the top table, a typical linear model could lead the imprudent
+researcher to misleadingly conclude that ``there is no effect of
+Condition'' (as a linear regression models changes in the mean, which is
+actually equal for the two conditions). The CHOCO model correctly
+reveals that condition B is associated with a lower probability but
+higher confidence of answers on the right side of the scale. Note that
+the CHOCO model is currently only available through Bayesian modeling,
+hence we report for consistency pseudo t- and p-values in the table,
+which correspond respectively to the coefficients divided by their SE,
+and the values obtained via the pd-to-p conversion
+(\citeproc{ref-makowski2019indices}{Makowski, Ben-Shachar, et al.,
+2019}).}{\label{fig-two}}}}
+
+\begin{center}
+\includegraphics[width=0.7\linewidth,height=\textheight,keepaspectratio]{./figures/fig2final.png}
+\end{center}
+
+\end{figure*}
 
@@
 provides improved model fit, 2) yields deeper insights into
-population-level effects than traditional approaches (gender differences
-in reality beliefs), and 3) allows for the reliable estimation of
+population-level effects than traditional approaches (sex differences in
+reality beliefs), and 3) allows for the reliable estimation of
 interpretable individual-level parameters through random effects.
 \textbf{Study 2} will apply this model to more subtle effects, such as
-the effect perceived facial attractiveness on reality judgments, and
+the effect of perceived facial attractiveness on reality judgments, and
 test the ability to fit alternative data structures (scales with ordinal
@@
 
-In today's post-truth era, the proliferation of advanced AI technologies
-has made it increasingly challenging to distinguish between authentic
-and synthetic media, bearing significant implications for information
-integrity and public trust
-(\citeproc{ref-lewandowsky2017beyond}{Lewandowsky et al., 2017}). As
-traditional cues become less reliable (e.g., visual glitches and
-artefacts in generated images; formulaic generated text, etc.), people
+The question of ``what is real'' has become increasingly relevant in the
+context of recent technological developments, particularly with the rise
+of artificial intelligence (AI), capable of mass producing highly
+realistic and relevant content, such as images, videos or text. These
+synthetic productions are flooding the cyberspace across various
+domains, such as art, advertising and entertainment, to education and
+information. The democratization of such technology carries an important
+potential for misuse, such as in disinformation campaigns and scams
+(e.g., AI-bots, identity theft), and abuse (e.g., deepfakes,
+\citeproc{ref-viola2023designed}{Viola \& Voto, 2023}), and raises
+pressing concerns about the value of authenticity and the potential
+erosion of media trust in our increasingly \emph{post-truth} society
+(\citeproc{ref-lewandowsky2017beyond}{Lewandowsky et al., 2017}).
+Critically, as technology improves, traditional cues for detecting what
+is fake become less reliable (e.g., visual glitches and artefacts in
+generated images; formulaic generated text, etc.), leading people to
 increasingly depend on contextual information and cognitive heuristics
@@
 28.4, SD = 9.0, range: {[}19, 66{]}; Sex: 47.5\% females). For each
-participant, we included only stimuli of the opposite gender (i.e., all
-89 female faces for men and 20 male faces for women).
+participant, we included only stimuli of the opposite sex (i.e., all 89
+female faces for males and 20 male faces for females).
 
@@
 phase, participants were informed that ``about half of the previously
-seen images were AI-generated''. The same faces were presented again in
-a new random order (same display time), followed by ratings of
-``reality'' (whether they believed the image was fake - left anchor - or
-real - right anchor).
+seen images were AI-generated'' (in actuality, all faces were
+photographs). The same faces were presented again in a new random order
+(same display time), followed by ratings of ``reality'' (whether they
+believed the image was fake - left anchor - or real - right anchor).
 
@@
 2025}; \citeproc{ref-patil2022datawizard}{Patil et al., 2022}). To
-maximize the comparability across models We used the default priors
+maximize comparability across models, we used the default priors
 (uniform) for all models, and we ran 16 chains of 1400 iterations each
@@
 
-For the population-level effects, we will consider significant and
-report (using the median of the posterior distribution) effects for
-which the 95\% Credible Interval (CI) does not include zero (and when
-the probability of direction \emph{pd} is \textgreater{}
+For the population-level effects, we will consider ``significant''
+(noteworthy) and report (using the median of the posterior distribution)
+effects for which the 95\% Credible Interval (CI) does not include zero
+(and when the probability of direction \emph{pd} is \textgreater{}
 \textasciitilde97\%, \citeproc{ref-makowski2019indices}{Makowski,
@@
 effect estimates, which corresponds to a 2:1 ratio of between-group
-variance to uncertainty.
+variance to uncertainty. Reliable in this context means that the
+variability between groups (e.g., participants) is greater than the
+uncertainty in the estimates, suggesting that the parameter can be used
+to assess inter-individual differences.
 
@@
 
+\begin{center}
 \includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{./figures/fig3.png}
+\end{center}
 
@@
 
-The ZOIB model suggested that women had higher mean scores of reality
+The ZOIB model suggested that females had higher mean scores of reality
 beliefs
@@
 
-The Beta-Gate model similarly suggested that women had higher mean
+The Beta-Gate model similarly suggested that females had higher mean
 scores of reality beliefs
@@
 
-The CHOCO model shows that women had a higher probability \emph{p} of
+The CHOCO model shows that females had a higher probability \emph{p} of
 judging faces as real
@@
 variability in the parameters reflecting the confidence in left and
-right choices is reliably capture, as opposed to the inter-individual
+right choices is reliably captured, as opposed to the inter-individual
 variability in the `bex' parameter. Bottom: correlation matrix between
@@
 
+\begin{center}
 \includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{./figures/fig4.png}
+\end{center}
 
@@
 \citeproc{ref-miller2023ai}{Miller et al., 2023};
-\citeproc{ref-nightingale2022ai}{Nightingale \& Farid, 2022}). These
-synthetic visuals are flooding the cyberspace across various domains,
-such as art, advertising and entertainment, to education and
-information. This technological advancement carries an important
-potential for misuse, such as in disinformation campaigns, scams (e.g.,
-AI-bots, identity theft), and abuse. The democratization of such
-technology raises pressing concerns about the value of authenticity and
-the potential erosion of media trust in our increasingly
-\emph{post-truth} society.
-
-In this evolving landscape, understanding the cognitive mechanisms that
-underpin our judgments of reality becomes paramount. Despite the
-increasing prevalence of digitally altered or AI-generated content,
-humans still rely on certain heuristics to assess authenticity. One such
-heuristic might be facial attractiveness. Attractiveness appraisals are
-known to be automatic and unconscious (\citeproc{ref-hou2023neural}{Hou
-et al., 2023}; \citeproc{ref-hung2016unconscious}{Hung et al., 2016};
+\citeproc{ref-nightingale2022ai}{Nightingale \& Farid, 2022}). In this
+evolving landscape, understanding the cognitive mechanisms that underpin
+our judgments of reality becomes paramount. Despite the increasing
+prevalence of digitally altered or AI-generated content, humans still
+rely on certain heuristics to assess authenticity. One such heuristic
+might be facial attractiveness. Attractiveness appraisals are known to
+be automatic and unconscious (\citeproc{ref-hou2023neural}{Hou et al.,
+2023}; \citeproc{ref-hung2016unconscious}{Hung et al., 2016};
 \citeproc{ref-luo2019robust}{Luo et al., 2019}), and carry strong
@@
 
-Although its role as a potential modulator of reality beliefs remains
-under-explored, Makowski, Te, et al.
+Although the role of attractiveness as a potential modulator of reality
+beliefs remains under-explored, Makowski, Te, et al.
 (\citeproc{ref-makowski2025too}{2025}) found significant associations
@@
 and SurveySwap). We removed 15 participants with data suggesting
-low-effort responding as well as those (N=16) that did not believe in
-the experimental manipulation and were fully confident that all images
-were real or fake. Similarly to the first sample, we included all hetero
-and bisexual participants and stimuli from the opposite gender,
-resulting in a final sample included 189 participants (Mean age = 28.4,
-SD = 14.0, range: {[}18, 69{]}; Sex: 76.2\% females). The study was
-approved by the University of Sussex Ethics Board (ER/ST633/1).
+low-effort responding (no variation in their responses indicative of
+straightlining) as well as those (N=16) that did not believe in the
+experimental manipulation and were fully confident that all images were
+real or fake. Similarly to the first sample, we included all hetero and
+bisexual participants and stimuli from the opposite gender, resulting in
+a final sample included 189 participants (Mean age = 28.4, SD = 14.0,
+range: {[}18, 69{]}; Sex: 76.2\% females). The study was approved by the
+University of Sussex Ethics Board (ER/ST633/1).
 
@@
 Following that, participants would see the 109 neutral-expression
-photographs from the AMFD database (1 s), each preceded by a randomly
-assigned textual cue indicating whether the image was ``AI-generated''
-or ``Photograph'' (2 s). Ratings of attractiveness, beauty and
-trustworthiness were collected after each image.
+pictures (all real photographs) from the AMFD database (1 s), each
+preceded by a randomly assigned textual cue indicating whether the image
+was ``AI-generated'' or ``Photograph'' (2 s). Ratings of attractiveness,
+beauty and trustworthiness were collected after each image.
 
@@
 \emph{not} treat the extreme values as 0 and 1 (and model them via the
-separate parameters \emph{pex} and \emph{bex}, which would distributions
-to be estimated from only 2 values), but instead to treat them as part
-of the continuous distribution. The 7 response options were rescaled to
-be evenly spaced between 0 and 1 excluded (i.e., {[}0.125, 0.875{]}).
-The \emph{pex} and \emph{bex} parameters were fixed to 0, making for a
-slightly more parsimonious model.
+separate parameters \emph{pex} and \emph{bex}, which would imply
+estimating the distributions from only 2 values), but instead to treat
+them as part of the continuous distribution. The 7 response options were
+rescaled to be evenly spaced between 0 and 1 excluded (i.e., {[}0.125,
+0.875{]}). The \emph{pex} and \emph{bex} parameters were fixed to 0,
+making for a slightly more parsimonious model.
 
@@
 
+\begin{center}
 \includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{./figures/fig5.png}
+\end{center}
 
@@
 of judging faces as real
-(\(P_{~\text{poly1}}=13.64, ~95\%~CI [8.42, 18.99], ~pd = 100\%\)), but
+(\(P_{~\text{poly1}}=13.64, ~95\%~CI [8.42, 18.99], ~pd = 100\%\)), and
 a quadratic relationship with the confidence in real judgments
@@
 (\(P \Delta_{~\text{real - fake}}=0.06, ~95\%~CI [0.00, 0.11], ~pd = 98.36\%\)).
-
-\subsubsection{Beauty}\label{beauty}
-
-The same analysis was run for Beauty, which indicated the following
-differences: \textbf{ANA: can you check and write down the main
-differences?}
 
@@
 is far from ideal, the CHOCO model still proved insightful, although it
-might have impacted its ability detect finer-grain changes, including
-quadratic relationships.
+might have restricted its ability to detect finer-grain changes,
+including quadratic relationships.
 
@@
 We would like to thank the dissertation students from the University of
-Sussex for their help in data collection. DM would also like to thank F.
-Rocher for inspiring some aspects of this paper.
+Sussex for their help in data collection, as well as T. Karapanagiotidis
+for feedback. DM would also like to thank F. Rocher for inspiring some
+aspects of this paper.
 
@@
 
+\bibitem[\citeproctext]{ref-viola2023designed}
+Viola, M., \& Voto, C. (2023). Designed to abuse? Deepfakes and the
+non-consensual diffusion of intimate images. \emph{Synthese},
+\emph{201}(1), 30.
+
 \bibitem[\citeproctext]{ref-williams2021beneath}

### 2025-06-25 paper/manuscript.qmd
abstract

@@
 abstract: |
+  We introduce the Choice-Confidence (CHOCO) model as a tool to dissect the cognitive mechanisms underlying decision-making judgments, offering a better fit for data from subjective rating scales for which the two extremes correspond to discrete choices (e.g., "positive" *vs.* "negative", "true" *vs.* false", "real" *vs.* "AI-generated", ...), and for which the distance from the center might reflect a gradual process, such as confidence or certainty in that choice.
+Conceptualized as a mixture of two Beta distributions, CHOCO simultaneously estimates the probability of selecting one category (the right side of the scale over the left side) and the confidence associated with each choice.
+To demonstrate its usage, we apply this model to two datasets (N=141 and N=189) in which participants judged whether face images were real photographs or AI-generated images, finding that facial attractiveness increased the likelihood of being judged as "real", particularly for male participants.
 As AI-generated content becomes increasingly indistinguishable from real stimuli, understanding how individuals form beliefs about the reality of what they perceive is both theoretically and practically urgent.
-This study introduces the Choice-Confidence (CHOCO) model as a tool to dissect the cognitive mechanisms underlying decision-making judgments.
-Conceptualized as a mixture of two Beta distributions, CHOCO simultaneously estimates the probability of selecting one category (e.g., "real" vs.
-"AI-generated") and the confidence associated with each choice.
-We apply this model to two datasets (N=141 and N=189) in which participants judged whether faces were real photographs or AI-generated images, finding that facial attractiveness systematically increased the likelihood of being judged as "real", particularly for male participants.
 We discuss how these findings relate to the broader effect of facial attractiveness, and suggest future directions for both the cognitive and psychometric application of the CHOCO framework.
 #   **Significance Statement**. TODO.
-keywords: "subjective scales, choice confidence model, reality beliefs, AI-generated faces, attractiveness, simulation monitoring"
+keywords: "subjective scales, choice confidence model, reality beliefs, AI-generated faces, attractiveness, simulation monitoring, zero-and-one inflated Beta regression"
 # Add more keywords if possible: beauty, Ordered Beta model, Bimodal data, decision-making
@@
 **Study 2** will apply this model to more subtle effects, such as the effect of perceived facial attractiveness on reality judgments, and test the ability to fit alternative data structures (scales with ordinal response options and mid-points).
-To this end, we analyze data from two separate studies in which participants judged whether a face image was AI-generated ("fake") or a real photograph.
+As a case-study illustration, we analyze data from two separate studies in which participants judged whether a face image was AI-generated ("fake") or a real photograph.
 
@@
 
-This study provides further evidence that attractive faces are more likely to be judged as real (*vs.* AI-generated), This effect was particularly strong and robust among male participants.
+This study provides further evidence that attractive faces are more likely to be judged as real (*vs.* AI-generated).
+This effect was particularly strong and robust among male participants.
 For females, the effect was weaker in sample 1 but more pronounced in sample 2, which might partly reflect increased statistical power.

### 2025-06-25 paper/manuscript.tex
abstract

@@
 
-\abstract{As AI-generated content becomes increasingly indistinguishable
-from real stimuli, understanding how individuals form beliefs about the
-reality of what they perceive is both theoretically and practically
-urgent. This study introduces the Choice-Confidence (CHOCO) model as a
-tool to dissect the cognitive mechanisms underlying decision-making
-judgments. Conceptualized as a mixture of two Beta distributions, CHOCO
-simultaneously estimates the probability of selecting one category
-(e.g., ``real'' vs.~``AI-generated'') and the confidence associated with
-each choice. We apply this model to two datasets (N=141 and N=189) in
-which participants judged whether faces were real photographs or
-AI-generated images, finding that facial attractiveness systematically
-increased the likelihood of being judged as ``real'', particularly for
-male participants. We discuss how these findings relate to the broader
-effect of facial attractiveness, and suggest future directions for both
-the cognitive and psychometric application of the CHOCO framework. }
+\abstract{We introduce the Choice-Confidence (CHOCO) model as a tool to
+dissect the cognitive mechanisms underlying decision-making judgments,
+offering a better fit for data from subjective rating scales for which
+the two extremes correspond to discrete choices (e.g., ``positive''
+\emph{vs.} ``negative'', ``true'' \emph{vs.} false'', ``real''
+\emph{vs.} ``AI-generated'', \ldots), and for which the distance from
+the center might reflect a gradual process, such as confidence or
+certainty in that choice. Conceptualized as a mixture of two Beta
+distributions, CHOCO simultaneously estimates the probability of
+selecting one category (the right side of the scale over the left side)
+and the confidence associated with each choice. To demonstrate its
+usage, we apply this model to two datasets (N=141 and N=189) in which
+participants judged whether face images were real photographs or
+AI-generated images, finding that facial attractiveness increased the
+likelihood of being judged as ``real'', particularly for male
+participants. As AI-generated content becomes increasingly
+indistinguishable from real stimuli, understanding how individuals form
+beliefs about the reality of what they perceive is both theoretically
+and practically urgent. We discuss how these findings relate to the
+broader effect of facial attractiveness, and suggest future directions
+for both the cognitive and psychometric application of the CHOCO
+framework. }
 
 \keywords{subjective scales, choice confidence model, reality beliefs,
-AI-generated faces, attractiveness, simulation monitoring}
+AI-generated faces, attractiveness, simulation monitoring, zero-and-one
+inflated Beta regression}
 
@@
 \par{ }
-\par{     \begin{tcolorbox}[enhanced jigsaw, opacityback=0, left=2mm, leftrule=.75mm, colback=white, breakable, bottomrule=.15mm, colframe=quarto-callout-note-color-frame, toprule=.15mm, arc=.35mm, rightrule=.15mm]
+\par{     \begin{tcolorbox}[enhanced jigsaw, toprule=.15mm, leftrule=.75mm, left=2mm, rightrule=.15mm, colback=white, colframe=quarto-callout-note-color-frame, breakable, bottomrule=.15mm, opacityback=0, arc=.35mm]
 
@@
 test the ability to fit alternative data structures (scales with ordinal
-response options and mid-points). To this end, we analyze data from two
-separate studies in which participants judged whether a face image was
-AI-generated (``fake'') or a real photograph.
+response options and mid-points). As a case-study illustration, we
+analyze data from two separate studies in which participants judged
+whether a face image was AI-generated (``fake'') or a real photograph.
 
@@
 This study provides further evidence that attractive faces are more
-likely to be judged as real (\emph{vs.} AI-generated), This effect was
+likely to be judged as real (\emph{vs.} AI-generated). This effect was
 particularly strong and robust among male participants. For females, the

### 2025-06-25 paper/manuscript.qmd
submit

@@
   - name: Andy Field 
-    orcid: ""
+    orcid: "0000-0003-3306-4695"
     role:
@@
 abstract: |
-  We introduce the Choice-Confidence (CHOCO) model as a tool to dissect the cognitive mechanisms underlying decision-making judgments, offering a better fit for data from subjective rating scales for which the two extremes correspond to discrete choices (e.g., "positive" *vs.* "negative", "true" *vs.* false", "real" *vs.* "AI-generated", ...), and for which the distance from the center might reflect a gradual process, such as confidence or certainty in that choice.
+  We introduce the Choice-Confidence (CHOCO) model as a tool to dissect the cognitive mechanisms underlying decision-making judgments, offering a better fit for data from subjective rating scales for which the two extremes correspond to discrete choices (e.g., "positive" *vs.* "negative", "true" *vs.* false", "real" *vs.* "AI-generated"), and for which the distance from the center might reflect a gradual process, such as confidence or certainty in that choice.
 Conceptualized as a mixture of two Beta distributions, CHOCO simultaneously estimates the probability of selecting one category (the right side of the scale over the left side) and the confidence associated with each choice.
 To demonstrate its usage, we apply this model to two datasets (N=141 and N=189) in which participants judged whether face images were real photographs or AI-generated images, finding that facial attractiveness increased the likelihood of being judged as "real", particularly for male participants.
-As AI-generated content becomes increasingly indistinguishable from real stimuli, understanding how individuals form beliefs about the reality of what they perceive is both theoretically and practically urgent.
+Given that AI-generated content is becoming increasingly indistinguishable from real stimuli, understanding how individuals form beliefs about the reality of what they perceive is both theoretically important and practically urgent.
 We discuss how these findings relate to the broader effect of facial attractiveness, and suggest future directions for both the cognitive and psychometric application of the CHOCO framework.
@@
 Similarly to the first sample, we included all hetero and bisexual participants and stimuli from the opposite gender, resulting in a final sample included 189 participants (Mean age = 28.4, SD = 14.0, range: [18, 69]; Sex: 76.2% females).
-The study was approved by the University of Sussex Ethics Board (ER/ST633/1).
+The study was approved by the University of Sussex Ethics Board (ER/ST633/1), but not preregistered.
 
@@
 
+The procedure for the first sample is described in study 1.
 For the second sample, the procedure was relatively similar to that of the first sample (described above in Study 1) with a few key differences.

### 2025-06-25 paper/manuscript.tex
submit

@@
 \emph{vs.} ``negative'', ``true'' \emph{vs.} false'', ``real''
-\emph{vs.} ``AI-generated'', \ldots), and for which the distance from
-the center might reflect a gradual process, such as confidence or
-certainty in that choice. Conceptualized as a mixture of two Beta
-distributions, CHOCO simultaneously estimates the probability of
-selecting one category (the right side of the scale over the left side)
-and the confidence associated with each choice. To demonstrate its
-usage, we apply this model to two datasets (N=141 and N=189) in which
-participants judged whether face images were real photographs or
-AI-generated images, finding that facial attractiveness increased the
-likelihood of being judged as ``real'', particularly for male
-participants. As AI-generated content becomes increasingly
-indistinguishable from real stimuli, understanding how individuals form
-beliefs about the reality of what they perceive is both theoretically
-and practically urgent. We discuss how these findings relate to the
-broader effect of facial attractiveness, and suggest future directions
-for both the cognitive and psychometric application of the CHOCO
-framework. }
+\emph{vs.} ``AI-generated''), and for which the distance from the center
+might reflect a gradual process, such as confidence or certainty in that
+choice. Conceptualized as a mixture of two Beta distributions, CHOCO
+simultaneously estimates the probability of selecting one category (the
+right side of the scale over the left side) and the confidence
+associated with each choice. To demonstrate its usage, we apply this
+model to two datasets (N=141 and N=189) in which participants judged
+whether face images were real photographs or AI-generated images,
+finding that facial attractiveness increased the likelihood of being
+judged as ``real'', particularly for male participants. Given that
+AI-generated content is becoming increasingly indistinguishable from
+real stimuli, understanding how individuals form beliefs about the
+reality of what they perceive is both theoretically important and
+practically urgent. We discuss how these findings relate to the broader
+effect of facial attractiveness, and suggest future directions for both
+the cognitive and psychometric application of the CHOCO framework. }
 
@@
 Makowski}{0000-0001-5375-9967}}\par{\addORCIDlink{Ana
-Neves}{0009-0006-0020-7599}}\par{\addORCIDlink{Andy Field}{true}} 
+Neves}{0009-0006-0020-7599}}\par{\addORCIDlink{Andy
+Field}{0000-0003-3306-4695}} 
 \par{ }
-\par{     \begin{tcolorbox}[enhanced jigsaw, toprule=.15mm, leftrule=.75mm, left=2mm, rightrule=.15mm, colback=white, colframe=quarto-callout-note-color-frame, breakable, bottomrule=.15mm, opacityback=0, arc=.35mm]
+\par{     \begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, leftrule=.75mm, colframe=quarto-callout-note-color-frame, colback=white, breakable, arc=.35mm, bottomrule=.15mm, left=2mm, toprule=.15mm, opacityback=0]
 
@@
 range: {[}18, 69{]}; Sex: 76.2\% females). The study was approved by the
-University of Sussex Ethics Board (ER/ST633/1).
+University of Sussex Ethics Board (ER/ST633/1), but not preregistered.
 
@@
 
-For the second sample, the procedure was relatively similar to that of
-the first sample (described above in Study 1) with a few key
-differences.
+The procedure for the first sample is described in study 1. For the
+second sample, the procedure was relatively similar to that of the first
+sample (described above in Study 1) with a few key differences.
 
